### 2021-09-24 lin-vs-cen.tex
First commit

\documentclass{patmorin}
\listfiles
\usepackage{pat}
\usepackage{paralist}
\usepackage{dsfont}  % for \mathds{A}
\usepackage[utf8x]{inputenc}
\usepackage{skull}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage[noend]{algorithmic}

\usepackage[normalem]{ulem}
\usepackage{cancel}
\usepackage{enumitem}

\usepackage{todonotes}

\usepackage[longnamesfirst,numbers,sort&compress]{natbib}

\usepackage[mathlines]{lineno}
\setlength{\linenumbersep}{2em}
\linenumbers
% \rightlinenumbers
\linenumbers
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
 \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
 \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
 \renewenvironment{#1}%
    {\linenomath\csname old#1\endcsname}%
    {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
 \patchAmsMathEnvironmentForLineno{#1}%
 \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\newcommand{\coloured}[2]{{\color{#1}{#2}}}
\newenvironment{colourblock}[1]{\color{#1}}{}

\newcommand{\condref}[1]{(C\ref{#1})}

% Taken from
% https://tex.stackexchange.com/questions/42726/align-but-show-one-equation-number-at-the-end
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\setlength{\parskip}{1ex}

\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\tw}{tw}
\DeclareMathOperator{\stw}{stw}
\DeclareMathOperator{\ltw}{ltw}
\DeclareMathOperator{\pw}{pw}
\DeclareMathOperator{\lpw}{lpw}
\DeclareMathOperator{\lhptw}{lhp-tw}
\DeclareMathOperator{\lhppw}{lhp-pw}

\DeclareMathOperator{\x}{x}
\DeclareMathOperator{\depth}{d}
\DeclareMathOperator{\sh}{cbt}
\DeclareMathOperator{\cbt}{cbt}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\dc}{dc}

\DeclareMathOperator{\afci}{\overline{\chi}_\pi}
\DeclareMathOperator{\afcn}{\dot{\chi}_\pi}

\newcommand{\ellt}{{\lfloor\ell/2\rfloor}}

\title{\MakeUppercase{Linear versus Centered Colouring Numbers}\thanks{This research was partly funded by NSERC.}}
\author{Prosenjit Bose, Vida Dujmović, Mehrnoosh Javarsineh, and Pat Morin}

\date{}

\DeclareMathOperator{\ddiv}{div}
\DeclareMathOperator{\hist}{p}

\newcommand{\colored}[2]{{\color{#1}#2}}

\usepackage{tabularx}

\DeclareMathOperator{\ci}{\overline{\pi}}

\DeclareMathOperator{\chicen}{\chi_{\mathrm{cen}}}
\DeclareMathOperator{\chilin}{\chi_{\mathrm{lin}}}

\begin{document}

% \begin{titlepage}
\maketitle

\begin{abstract}
    These are some notes on the relationships betwee linear colouring numbers and centered colouring numbers.
\end{abstract}
% \end{titlepage}

% \pagenumbering{roman}
% \tableofcontents
%
% \newpage
% \pagenumbering{arabic}

\section{Introduction}

Let $G$ be a simple undirected graph.  A \emph{$k$-colouring} of $G$ is any function $\varphi:V(G)\to S$ where $S$ is a set of size $k$.  A vertex $v$ of $G$ is a \emph{center} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.  The colouring $\varphi$ is \emph{centered} if every connected sugraph $H\subseteq G$ has a center.  The colouring $\varphi$ is \emph{linear} if every simple path $P\subseteq G$ has a center.

The \emph{centered chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centered $c$-colouring.  The \emph{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.  Since every path in $G$ is a connected subgraph of $G$, every centered colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.

The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet[Theorem~1]{kun.obrien.ea:polynomial} who showed the following result:

\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun-obrien-general}
  For any graph $G$, $\chicen(G)\le \chilin^{190}(G)\cdot\log^{O(1)}(\chilin(G))$.  
\end{thm}

Their proof of \cref{kun-obrien-general} has three steps:
\begin{enumerate}
  \item A theorem of \citet{kawarabayashi.rossman:polynomial} shows that, if $\chicen(G)\ge k^{190}\log^{O(1)} k$ then $\chilin(G)\ge k^{38}$ or $G$ has treewidth $\tw(G)\ge k^{38}$.  In the former case there is nothing left to prove.
  \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy:improved} shows that, if $\tw(G)\ge k^{38}$, then $G$ contains an $\Omega(k^2\times k^2)$ grid minor.
  \item A technical lemma \cite[Lemma~5]{kun.obrien.ea:polynomial} shows that, if $G$ contains a $k^2\times k^2$ grid minor, then $\chilin(G)\in\Omega(k)$.
\end{enumerate}

These notes are an attempt to improve the bound in \cref{kun-obrien-general} in the general case as well as for some special classes of graphs.

As a first step, we attempt to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.  This, by itself, reduces the exponent in \cref{kun.obrien.general} from $190$ to $95$.  Next, we observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.  For example, if we consider only planar graphs, then it is well known that any planar graph $G$ of treewidth $k$ contains an $\Omega(k\times k)$ grid minor.  This already reduces the exponent further (for planar graphs) to $5$.  If one could prove a similar result for the first ``heavy hammer'' then this would show that, for any planar graph $G$, $\chicen(G)\in O(\chilin(G))$.

\section{The Linear Colouring Number of Pseudogrids}

A \emph{subdivision} $G'$ of a graph $G$ is any graph that can be obtained from $G$ by repeatedly replacing some edge edge $uw$ of $G$ with a path $uvw$, where $v\not\in V(G)$ is a newly introduced degree-$2$ vertex.  Alternatively, $G'$ is obtained from $G$ by replacing some of the edges of $G$ with paths whose internal vertices have degree $2$.

The $k\times k$ grid graph $G_k$ is the graph with vertex set $V(G_k):=\{1,\ldots,k\}^2$ and that contain an edge with endpoints $(v,w)$ and $x,y$ if and only $|v-x|+|w-y|=1$.  A graph $H$ is a $k\times k$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $G$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_k$.  Since $G_k$ has maximum degree $4$ any graph $G$ that contains a $G_k$-minor contains a subgraph $H$ that is a $k\times k$ pseudogrid.

\begin{lem}
  For any $k\times k$ pseudogrid $H$, $\chilin(G)\in\Omega(k)$.
\end{lem}

\begin{proof}
  Let $\varphi:V(H)\to\{1,\ldots,c\}$ be an arbitrarily $c$-colouring of some $k\times k$ pseudogrid $H$ for some $c\le \epsilon k$, where $\epsilon$ is a small positive constant.  We will show that $H$ contains a path $P$ which has no center with respect to $c$.  First we introduce some notations:

  For each vertex $v$ of $G_k$ there is a path $P_v$ in $\mathcal{P}$ (possibly consisting of one vertex) such that the vertex $v$ appears in $H'$ as a result of contracting $P_v$ in $H$.  In this way, each vertex $v$ of $G_k$ is associated with a non-empty \emph{colour set} $\phi(v):=\{\varphi(w):w\in V(P_v)\}$.  Similarly, for each edge $e$ of $G_k$ there is a path $P_e$ in $H'$ that comes from subdividing $e$.  Thus, each edge $e$ of $G_k$ is associated with (possibly empty) \emph{colour set} $\phi(e):=\{\varphi(w):w\in P^-_e\}$, where $P^-_e$ denotes the (possibly empty) set of internal vertices on the path $P_e$.

  For each colour $\alpha\in\{1,\ldots,c\}$, define the \emph{multiplicity} of $\alpha$ as the number of vertices and edges of $G_k$ that include $\alpha$ in their colour set.  Fix some large integer constant $A$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity at most $A$.

  If there exists some infrequent colour $\alpha\in\{1,\ldots,c\}$, we identify every row and column of $G_k$ that contains an edge or vertex with $\alpha$ in its colour set.  We then \emph{remove} each of these (at most $A$) rows and columns from $H$ by removing all edges of $H$ in $P_e$ for each $e$ in the relevant row (column) of $G_k$ and then discarding any isolated vertices.  After doing this for each row and column containing $\alpha$, the graph $H$ may be disconnected.  We discard any isolated vertices and any path $P_{v}$ that is disconnected from the rest of $H$.

  The resulting graph $\tilde{H}$ contains a $k'\times k'$ pseudogrid for some $k' \le k-A$ and contains no vertex of colour $\alpha$.  We then set $H\gets \tilde{H}$, $k\gets k'$ and repeat this process until eventually obtain a new $k'\times k'$ pseudogrid $\tilde{H}$ for some $k'\ge k-cA\ge (1-\epsilon A)k \ge k/2$, for $\epsilon \le 1/2A$.  From this point onward that that there is no infrequent colour, i.e., every colour $\alpha\in\{1,\ldots,c\}$ has multiplicity at least $A$.

  Next we run a two stage process that, for each $\alpha\in\{1,\ldots,c\}$, will identify a pair of vertices $v_\alpha,w_\alpha$ in $H$ whose colour is $\alpha$.  These pairs are carefully chosen so that we can construct a simple path $P$ in $H$ that contains $\bigcup_{\alpha\in\{1,\ldots,c\}} \{v_\alpha,w_\alpha\}$.

  \begin{compactenum}[Stage 1]
    \item We begin this stage by defining each edge and vertex of $G_k$ as being initially \emph{unmarked}.  Fix some constant $B>0$.  We iterate over each $\alpha\in\{1,\ldots,c\}$ in turn.  If $G_k$ contain a pair $(x_\alpha,y_\alpha)$ where each of $x_\alpha$ and $y_\alpha$ is an unmarked vertex or unmarked edge of $G_k$ with $\alpha$ in its colour set and the distance in $G_k$ between $x_\alpha$ and $y_\alpha$ is at least $B$ then Stage~1 \emph{succeeds} for $\alpha$.  In this case, we take $v_\alpha$ to be any vertex in $P_{x_\alpha}$ of colour $\alpha$ and $w_\alpha$ to be any vertex in $P_{y_\alpha}$ of colour $\alpha$.  We then \emph{mark} every vertex and edge of $G_k$ whose distance from $x_\alpha$ or $y_\alpha$ is less than $B$.

    If $G_k$ contains no such pair $(x_\alpha,y_\alpha)$, then we leave $v_\alpha$ and $w_\alpha$ undefined and say that Stage~1 \emph{failed} for colour $\alpha$.

    \item Let $S\subset\{1,\ldots,c\}$ be the set of colours $\alpha$ for which Stage~1 succeeded and let $\overline{S}:=\{1,\ldots,c\}\setminus S$ be the set of colours for which Stage~1 failed.  We construct a bipartite graph $K$ with parts $L(K):=\overline{S}$ and $R(K):=\bigcup_{\alpha\in S}\{x_\alpha,y_\alpha\}$.   For each marked vertex or edge $z$ of $G_k$, let $s(z)$ be the vertex in $R(k)$ that is closest to $z$ in $G_k$.\todo{Gotta be really precise here to avoid degree 3 issue.}  The edge set of $K$ contains an edge $\alpha\beta$ precisely when there is a marked edge or vertex $z$ of $G_k$ such that $\alpha\in\phi(z)$ and $s(z)=\beta$.

    \begin{clm}
      $K$ contains two matchings $M_1$ and $M_2$ such that each vertex in $L(K)$ appears in each of $M_1$ and $M_2$ and each vertex in $R(k)$ appears in at most one of $M_1$ or $M_2$.
    \end{clm}

    \begin{proof}
      Use Hall's Mathcing Theorem/Lemma.
    \end{proof}

    Finally, for each $\alpha\in\overline{S}$, we define $x_\alpha$ to be the the unique element of $R(K)$ such that $M_1$ includes the edge $\alphax_\alpha$ and we define $y_\alpha$ to be the unique element of $R(K)$ such that $M_2$ includes the edge $\alpha y_\alpha$.  We define $v_\alpha$ and $w_\alpha$ as before.

    Now, let $R:=\bigcup_{\alpha\in\{1,\ldots,c\}} \{x_\alpha,y_\alpha\}$.

    \begin{clm}
        There is no vertex of $G_k$ that has three or more incident edges in $R$.
    \end{clm}

    \begin{clm}
        Each $B/2\times B_/2$ subgrid of $G_k$ contains at most four elements of $R$.
    \end{clm}

    Now classify each vertex of $G_k$ as either being \emph{straight} or \emph{bent}. Define what it means for a path $P$ to go straight or bend at a vertex.  Now we should be able to finish with this lemma:

    \begin{lem}
        $G_k$ contain a simple path $P$ that contains every edge and vertex in $R$.  Furthermore, $P$ bends at each bent vertex in $R$ and $P$ goes straight at each straight vertex in $R$. 
    \end{lem}

\end{proof}

\bibliographystyle{plainurlnat}
\bibliography{lin-vs-cen}

\end{document}

Two words $s$ and $t$ are \emph{anagrams} of each other if $s$ is a permutation of $t$.  A single word $w:=w_1,\ldots,w_{2r}$ is \emph{anagramish} if its first half $w_1,\ldots,w_r$ and its second half $w_{r+1},\ldots,w_{2r}$ are anagrams of each other.  A word is \emph{anagram-free} if it does not contain an anagramish factor.  Anagramish words are also known as \emph{abelian squares} and anagram-free words are also known as \emph{abelian square free} words \cite{keranen:abelian,keranen:powerful}, \emph{strongly nonrepetitive} words \cite{pleasants:non-repetitive}, or \emph{strongly asymmetric} sequences \cite{evdokimov:strongly,evdokimov:strongly2}.

In 1961, \citet{erdos:some} asked if there exist arbitrarily long words over an alphabet of size $4$ that are anagram-free.\footnote{This was an incredibly prescient question since it is not at all obvious that there exist arbitrarily long anagram-free words over \emph{any} finite alphabet. The only justification for choosing the constant $4$ is that a short case analysis rules out the possibility of length-$8$ anagram-free words over an alphabet of size $3$.}  In 1968 \citet{evdokimov:strongly,evdokimov:strongly2} showed the existence of arbitrarily long anagram-free words over an alphabet of size $25$ and in 1971 \citet{pleasants:non-repetitive} showed an alphabet of size $5$ is sufficient.  Erd\H{o}s's question was not fully resolved until 1992, when \citet{keranen:abelian} answered it in the affirmative.

A path $v_1,\ldots,v_{2r}$ in a graph $G$ is \emph{anagramish} under a vertex $c$-colouring $\phi:V(G)\to\{1,\ldots,c\}$ if $\phi(v_1),\ldots,\phi(v_{2r})$ is an anagramish word.  The colouring $\phi$ is an \emph{anagram-free colouring} of $G$ if no path in $G$ is anagramish under $\phi$.  The minimum integer $c$ for which $G$ has an anagram-free vertex $c$-colouring is called the \emph{anagram-free chromatic number} of $G$ and is denoted by $\afcn(G)$.

For a graph family $\mathcal{G}$, $\afcn(\mathcal{G}):=\max\{\afcn(G):G\in\mathcal{G}\}$ or $\afcn(\mathcal{G}):=\infty$ if the maximum is undefined. The results on anagram-free words discussed in the preceding paragraph can be interpreted in terms of $\afcn(\mathcal{P})$ where $\mathcal{P}$ is the family of all paths. Indeed, Ker\"anen's result shows that $\afcn(\mathcal{P})\le 4$. Slightly more complicated than paths are trees.  \citet{wilson.wood:anagram-free} showed that $\afcn(\mathcal{T})=\infty$ for the family $\mathcal{T}$ of trees and \citet{kamcev.luczak.ea:anagram-free} showed that $\afcn(\mathcal{T}_2)=\infty$ even for the family $\mathcal{T}_2$ of binary trees.

One positive result in this context is that of
\citet{wilson.wood:anagram-free}, who showed that every tree $T$ of pathwidth $p$ has $\afcn(T)\le 4p+1$. Since trees are graphs of treewidth $1$ it is natural to ask if this result can be extended to show that every graph $G$ of treewidth $t$ and pathwidth $p$ has $\afcn(G)\le f(t,p)$ for some function $f:\N^2\to\N$.  \citet{carmi.dujmovic.ea:anagram-free} showed that such a generalization is not possible for any $t\ge 3$ by giving examples of $n$-vertex graphs of pathwidth $3$ (and treewidth $3$) with $\afcn(G)\in\Omega(\log n)$.  The obvious remaining gap left by these two works is graphs of treewidth $2$. Our main result is to show that $G_n$, the $2\times n$ square grid has $\afcn(G_n)\in\omega_n(1)$. Since $G_n$ has pathwidth 2, we have:

\begin{thm}\label{main_vertex}
    For every $c\in\N$, there exists a graph of pathwidth $2$ that has no anagram-free vertex $c$-colouring.
\end{thm}

\citet[Section 7.1]{wilson:anagram-free} conjectured that $\afcn(G_n)\in\omega_n(1)$, so this work confirms this conjecture.  Prior to the current work, it was not even known if the family of $n\times n$ square grids had anagram-free colourings using a bounded number of colours.

In a larger context, this lower bound gives more evidence that, except for a few special cases (paths \cite{evdokimov:strongly,pleasants:non-repetitive,keranen:abelian}, trees of bounded pathwidth \cite{wilson.wood:anagram-free}, and highly subdivided graphs \cite{wilson.wood:anagram-free2}), the qualitative behaviour of anagram-free chromatic number is not much different than that of treedepth/centered colouring \cite{nesetril.ossona:tree-depth}.  Very roughly: For most graph classes, every graph in the class has an anagram-free colouring using a bounded number of colors precisely when every graph in the class has a colouring using a bounded number of colours in which every path contains a colour that appears only once in the path.

% \todo[inline]{Statement of vertex colouring result (hopefully)}
%
% An \emph{edge $c$-colouring} $\varphi:E(G)\to\{1,\ldots,c\}$ is \emph{anagram-free} if $G$ has no path $v_0,\ldots,v_{2r}$ such that $\varphi(v_0v_1),\ldots,\varphi(v_{2r-1}v_{2r})$ is anagramish.  We also give an edge colouring version of \cref{main_vertex}.
%
% \begin{thm}\label{main_edge}
%     For every $c\in\N$, there exists a graph of $G$ of pathwidth $2$ that has no anagram-free edge $c$-colouring.
% \end{thm}

% \todo[inline]{Relation to Existing Work}

The remainder of this paper is organized as follows: \Cref{near_anagram_statement} gives some definitions and states a key lemma which shows, under a certain periodicity condition, that every sufficiently long word contains a factor that is $\epsilon$-close to being anagramish.
% In \cref{edge_colourings} we prove \cref{main_edge}.
In \cref{vertex_colourings} we use this key lemma to prove \cref{main_vertex}.  In \cref{near_anagram_proof} we prove the key lemma. \cref{reflections} concludes with some final remarks about the (non-)constructiveness of our proof technique.

\section{Periodicity in words}
\label{near_anagram_statement}

An \emph{alphabet} $\Sigma$ is a finite non-empty set and each element of $\Sigma$ is called a \emph{letter}. A \emph{word} over $\Sigma$ is a (possibly empty) sequence $w:=w_1,\ldots,w_n$ with $w_i\in\Sigma$ for each $i\in\{1,\ldots,n\}$. The \emph{length} $|w|$ of $w$ is the length, $n$, of the sequence.  A word of length at least $1$ is \emph{non-empty}.  The unique word of length $0$ is denoted by $\varepsilon$. For a word $w:=w_1,\ldots,w_n$, we let $\Sigma_w:=\{w_i:i\in\{1,\ldots,n\}\}$.  For each $1 \le i \le j\le n+1$, $w_i,w_{i+1},\ldots,w_{j-1}$ is called a \emph{factor} of $w$.\footnote{Note that this definition includes the empty factor $\varepsilon$ obtained when $i=j$.}  The set of all factors of $w$ is denoted by $F(w)$.

A \emph{language} $L$ over $\Sigma$ is a set of words over $\Sigma$.  A language $L$ is \emph{infinite} if $|L|=\infty$.  We let $\Sigma_L:=\bigcup_{w\in L}\Sigma_w$ and $F(L):=\bigcup_{w\in L}F(w)$.  The language $L$ is \emph{factor-closed} if $F(w)\subseteq L$ for each $w\in L$.  For any alphabet $\Sigma$ and any $k\in\N$, $\Sigma^k$ (the $k$-fold cartesian product of $\Sigma$ with itself) is the language consisting of all length-$k$ words over $\Sigma$.  The \emph{Kleene closure} $\Sigma^*:=\bigcup_{k=0}^\infty \Sigma^k$ is the language consisting of all words over $\Sigma$.

 % (Note the convention that $w[i\mathbin{:}j]$ does not include $w_j$, so $w[i\mathbin{:}j]$ has length $j-i$.)\todo{Is this notation still needed after revisions?}

Let $w:=w_1,\ldots,w_n$ be a word over an alphabet $\Sigma$ and, for each $a\in \Sigma$, define $|w|_a:=|\{i\in\{1,\ldots,n\}:w_i=a\}|$.  The \emph{Parikh vector} of $w$ is the integer-valued $|\Sigma|$-vector $\hist(w):=(|w|_{a}:a\in\Sigma)$ that is indexed by elements of $\Sigma$.  Observe that a word $w_1,\ldots,w_{2r}$ is anagramish if and only if $\hist(w_1,\ldots,w_r)=\hist(w_{r+1},\ldots,w_{2r})$ or, equivalently, $\hist(w_1,\ldots,w_r)-\hist(w_{r+1},\ldots,w_{2r})=\boldsymbol{0}$.
% Let $\|\cdot\|_1$ denote the L1 norm of a vector.\footnote{The L1 norm of a real-valued $d$-vector $\boldsymbol{v}:=(v_1,\ldots,v_d)$, $\|\boldsymbol{v}\|_1=\sum_{i=1}^d|v_i|$.}
For each $a\in\Sigma$, let $\delta_a(w):=|w_1,\ldots,w_r|_{a}-|w_{r+1},\ldots,w_{2r}|_{a}$ and let $\tau_a(w):=|\delta_a|$.
Then $\tau(w):=\sum_{a\in\Sigma}\tau_a(w)$ is a useful measure of how far a word is from being anagramish and that $\tau(w)=0$ if and only if $w$ is anagramish.

A word $w:=w_1,\ldots,w_n$ is \emph{$\ell$-periodic} over an alphabet $\Sigma\supseteq\Sigma_w$ if each length-$\ell$ factor of $w$ contains every letter in $\Sigma$.  A language $L$ is \emph{$\ell$-periodic}  if each of its words is $\ell$-periodic over $\Sigma_L$.  A language $L$ is \emph{periodic} if it is $\ell$-periodic for some finite $\ell$.  We make use of the following key lemma, which states that every sufficiently long $\ell$-periodic word contains arbitrarily long factors that are $\epsilon$-close to being anagramish.

\begin{lem}\label{near_anagram_fourier}
    For each $r_0,\ell\in\N$ and each $\epsilon>0$, there exists a positive integer $n$ such that each $\ell$-periodic word $w_1,\ldots,w_n$ contains a factor $w:=w_{i+1},\ldots,w_{i+2r}$ of length $2r \ge 2r_0$ such that $\tau(w)\le \epsilon r$.
\end{lem}

The proof of \cref{near_anagram_fourier} is deferred to \cref{near_anagram_proof}.  We now give some intuition as to how it is used.  The process of checking if a word is anagramish can be viewed as finding common letters in the first and second halves and crossing them both out.  If this results in a complete cancellation of all letters, then the word is an anagram.  \cref{near_anagram_fourier} tells us that we can always find a long factor $w$ where, after exhaustive cancellation, only an $\epsilon$-fraction of the original letters remain.
% Informally, the factor $w$ is $\epsilon$-close to being anagramish.

\cref{near_anagram_fourier} says that, if up to $\epsilon r$ letters in each half of $w$ were each allowed to cancel two letters of the same letters in the other half of $w$, then it would be possible to complete the cancellation process.  To achieve this type of one-versus-two cancellation in our setting, we decompose our coloured pathwidth-$2$ graph into pieces of constant size.  The vertices in each piece can be covered with one path or partitioned into two paths (see \cref{paths}).  In this way an occurrence $H_z$ of a particular coloured piece in one half can be matched with two like-coloured pieces $H_x$ and $H_y$ in the other half. We construct a single path $P$ that contains all vertices in $H_z$ and only half the vertices in each of $H_x$ and $H_y$.  In this way, the colours of vertices $P\cap H_z$ can cancel the colours of the vertices in $P\cap(H_x\cup H_y)$.

Since \cref{near_anagram_fourier} requires that the word $w$ be $\ell$-periodic, the following lemma will be helpful in obtaining words that can be used with \cref{near_anagram_fourier}. (Although \cref{periodicity} follows immediately from considerably stronger results in combinatorics on words---for example \citet[Proposition~1.5.12]{lothaire:algebraic}---we provide a proof here for the sake of completeness.)

\begin{lem}\label{periodicity}
    Let $L$ be an infinite factor-closed language over an alphabet $\Sigma$.  Then there exists an infinite periodic factor-closed language $M\subseteq L$.
\end{lem}

\begin{proof}
    Let $\Xi\subseteq\Sigma$ be any minimal subset of $\Sigma$ with the property that $M:=\Xi^* \cap L$ is infinite.  Observe that, since $L$ is factor closed, $M$ is also factor-closed.  Therefore $M$ is infinite and factor-closed.  All that remains is to show that $M$ is $\ell$-periodic for some integer $\ell$.

    Since $\Xi$ is minimal, $\Pi^*\cap L$ is finite for any strict subset $\Pi\subsetneq \Xi$.  Since $L$ is factor-closed, $\Pi^*\cap L$ is factor-closed.  Since $\Pi^*\cap L$ is factor-closed and finite, $|w|$ is finite for each $w\in\Pi^*\cap L$.
    Therefore, the value
    \begin{equation}
        \ell:=1+\max\{|w|: w\in \Pi^*\cap L,\, \Pi\subsetneq\Xi\} \label{l-def}
    \end{equation}
    is finite.  Now, consider any factor $f$ of any word in $M$ that does not contain every letter of $\Xi$, i.e., $f\in\Pi^*$ for some  $\Pi\subsetneq\Xi$.  Since $\Pi^*\cap L$ is factor closed, $f\in \Pi^*\cap L$ so, by \cref{l-def}, $|f| < \ell$.  Contrapositively, any factor of length at least $\ell$ of any word in $M$ contains every element of $\Xi$, so $M$ is $\ell$-periodic.
\end{proof}

\section{Proof of \cref{main_vertex}}
\label{vertex_colourings}

For each $n\in\N$, let $G_n$ be the $2\times n$ square grid with top row $a_0,\ldots,a_{n-1}$ and bottom row $b_0,\ldots,b_{n-1}$ (see \cref{g_n}).  For the sake of contradiction, suppose that there exists an integer $c$ such that $\afcn(G_n)\le c$ for each $n\in\N$.  Observe that there is a bijection between vertex $c$-colourings of $G_n$ and words of length $2n$ over the alphabet $\Sigma_0:=\{1,\ldots,c\}$ where a word $w:=w_0,\ldots,w_{2n-1}$ corresponds to the colouring $\phi_w:V(G_n)\to\{1,\ldots,c\}$ defined by
$\phi_w(a_i):=w_{2i}$ and $\phi_w(b_i):=w_{2i+1}$, for each $i\in\{0,\ldots,n-1\}$.

\begin{figure}
    \centering{
        \includegraphics{figs/g_n}
    }
    \caption{The graph $G_n$}
    \label{g_n}
\end{figure}

% Let $L_0\subseteq \Sigma_0^*$ consist of all words $w\in \Sigma_0^*$ whose length is a multiple of $8$ and such that $\phi_w$ is an anagram-free colouring of $G_{|w|/2}$. Furthermore, the assumption that $\afcn(G_n)\le c$ for each $n\in\N$ implies that $L_0$ is infinite.

For reasons that will become apparent, it is useful to break $G_n$ up into blocks consisting of $4$ columns each.  Let $\Sigma_1:=\Sigma_0^8$ and define the language $L_1$ to consist of exactly those words $w_1,\ldots,w_r\in \Sigma_1^*$ for which $\phi_{w_1\cdots w_r}$ is an anagram-free colouring of $G_{4r}$.  From the assumption that $\afcn(G_n)\le c$ for all $n\in\N$, it follows that $L_1$ is infinite.  From the definition of anagram-free colouring, it follows that $L_1$ is factor-closed.

By \cref{periodicity}, there exists an infinite periodic factor-closed language $L_2\subseteq L_1$.  Let $x$ be any letter in $\Sigma_{L_2}$.  Note that the word $xx$ is not in $L_2$ since the corresponding colouring $\phi_{xx}$ of $G_8$ has, for example, the anagramish path $a_0,\ldots,a_7$.  Since $L_2$ is factor-closed, no word in $L_2$ contains $xx$ as a factor, i.e., $xx\not\in F(L_2)$.

Let $\Sigma_3\subset F(L_2)$ contain only those factors of $L_2$ that do not contain $x$.  Thus, any word in $L_2$ is of the form $w_0xw_1xw_2\cdots xw_rxw_{r+1}$ where each of $w_1,\ldots,w_{r}$ is in $\Sigma_3$ and each of $w_0,w_{r+1}$ is in $\Sigma_3\cup\{\varepsilon\}$.  Since $L_2$ is periodic, each word in $\Sigma_3$ has bounded length, so $\Sigma_3$ is finite, i.e., $\Sigma_3$ is an alphabet.  Let $L_3$ be the language consisting of all words $w_1,\ldots,w_r\in\Sigma_3^*$ such that $xw_1xw_2\cdots xw_r\in L_2$.  Since $L_2$ is infinite and each word in $L_2$ has the form described above, $L_3$ is infinite.  Since $L_2$ is factor closed, $L_3$ is also factor-closed. Therefore, by \cref{periodicity}, there exists an infinite periodic factor-closed language $L\subseteq L_3$ over the alphabet $\Sigma:=\Sigma_{L_3}$.

See \cref{bigexample} for what follows. Each word $w:=w_1,\ldots,w_r\in L$ defines a word $w':=xw_1xw_2\cdots xw_{r}\in L_1$.  In graph colouring terms, $w$ defines an integer $n_w:=4r+\tfrac{1}{2}\sum_{i=1}^r|w_i|$ and an anagram-free colouring $\phi_w:=\phi_{w'}$ of $G_{n_w}$. The word $w'$ partitions the columns of $G_{n_w}$ into \emph{blocks}, each of which corresponds to a letter $a$ of $w'$ and whose length $|a|/2$ is a multiple of $4$.  There are two types of blocks:
\begin{compactitem}
    \item There are \emph{boring} blocks $Q_0,\ldots,Q_{r-1}$, each containing $4$ columns of $G_{n_w}$, and each corresponding to an occurrence of $x$ in $w'$.
    \item There are \emph{colourful blocks} $H_1,\ldots,H_r$, where $H_i$ is the subgraph of $G_{n_w}$ corresponding to $w_i$ and contains $4|w_i|$ columns of $G_{n_w}$.
\end{compactitem}

\begin{figure}
    \centering{\includegraphics{figs/bigexample-1}}
    \caption{A word $w:=dabacbdabc$, its corresponding word $w':=xdxaxbxaxcxbxdxaxbxc$, and its corresponding coloured graph $G_{n_w}$.}
    \label{bigexample}
\end{figure}

Since $L$ is periodic, there exists an integer $\ell$ such that every word in $L$ is $\ell$-periodic.  Since $L$ is infinite, it contains arbitrarily long words.  These two facts, along with \cref{near_anagram_fourier} immediately imply the following lemma:

\begin{lem}\label{near_anagram_graph}
    If $\afcn(G_n)\le c$ for each $n\in\N$ then, for each $r_0\in\N$ and $\epsilon>0$, there exists $w\in L$ with even length $|w|\ge 2r_0$ and $\tau(w)\le\epsilon r$.
\end{lem}

% \begin{proof}
%     By \cref{block_colouring}, for each $k\in\N$ there exists an $\ell$-periodic word $w'\in\Sigma^k$ such that $\phi_{s'}$ is an anagram-free vertex colouring of $G_{n_{s'}}$. Applying \cref{near_anagram_fourier} to $w'$ proves the existence of the desired word $w$.
% \end{proof}

\cref{near_anagram_graph} shows the existence of a colouring $\phi_w$ of $G_n$ (for arbitrarily large $n$) that is defined by a word $w$ that is $\epsilon$-close to being anagramish.  This implies that the path $P_0:=a_1,\ldots,a_{n}$ is $4\epsilon$-close to being anagramish under $\phi_w$.  The last step, done in the next lemma, is to show that there is enough flexibility to modify $P_0$ to obtain a path is anagramish under $\phi_w$.

\begin{lem}\label{anagramish_path}
    There exists $\epsilon>0$ and $r_0\in\N$ such that, for any $w\in L$ with even length $|w|\ge 2r_0$ and $\tau(w)\le\epsilon r$, the graph $G_{n_w}$ contains a path that is anagramish under $\phi_w$.
\end{lem}

\begin{proof}
    Let $w:=w_1,\ldots,w_{2r}$ so that the colouring $\phi_w$ is defined by the word $xw_1xw_2\cdots xw_r$ (and $r\ge r_0$).
    For each $a\in\Sigma$, define sets $A_a\subseteq\{i\in \{1,\ldots,r-1\}: w_i=a\}$ and $B_a\subseteq\{i\in\{r+1,\ldots,2r-1\}:w_i=a\}$ as follows:
    \begin{compactenum}
        \item If $\delta_a(w)=0$ then $A_a=B_a=\emptyset$.
        \item If $\delta_a(w)>0$ then $|A_a|=2\delta_a(w)=2|B_a|$.
        \item If $\delta_a(w)<0$ then $|A_a|=\delta_a(w)=\tfrac{1}{2}|B_a|$.
    \end{compactenum}
    Let $A:=\bigcup_{a\in\Sigma} A_a$ and $B:=\bigcup_{a\in\Sigma} B_a$.
    The sets $A_a$ and $B_a$ are chosen so that $A\cup B$ satisfies the following \emph{global independence constraint}:  There is no pair $i,j\in A\cup B$ such that $i-j=1$.  To see that this is possible, first observe that, because $r$ is not in $A$ or $B$ we need only concern ourselves with pairs where both $i,j\in\{1,\ldots,r-1\}$ or pairs where both $i,j\in\{r+1,\ldots,2r-1\}$.  Thus, we can choose the elements of $A_a$, for each $a\in\Sigma$ and then independently choose the elements of $B_a$, for each $a\in\Sigma$.

    We show how to choose the elements of $A_a$ for each $a\in\Sigma$.  The same method works for choosing the elements in $B_a$. Observe that, because $w\in L$, $w$ is $\ell$-periodic, so $|\{i\in \{1,\ldots,r-1\}:w_i=a\}|\ge \lfloor(r-1)/\ell\rfloor$ for each $a\in\Sigma$.  This allows us to greedily choose the elements in $A_a$ for each $a\in\Sigma$. At each step we simply avoid choosing $i$ if $i-1$ or $i+1$ have already been chosen in some previous step.  At any step in the process, at most $\epsilon r$ elements have already been chosen in previous steps and each of these eliminates at most $2$ options.  Therefore, there will always be an element available to choose, provided that $2\epsilon r < \lfloor(r-1)/\ell\rfloor$.  In particular, for any $r\ge r_0\ge \ell$, $\epsilon < 1/(4\ell)$ works.

    To ease notation, let $G:=G_{n_w}$.  We now construct the anagramish path $P$ in a piecewise fashion.  Refer to \cref{path_construction}.  For each $i\in\{1,\ldots,2r\}$, let $H_i$ be the subgraph of $G$ induced by the vertices corresponding to $w_i$. (The subgraph $H_1,\ldots,H_{2r}$ are referred to above as colourful blocks.)
    \begin{figure}
        \centering{
            \includegraphics{figs/bigexample-1} \\
            \includegraphics{figs/bigexample-2} \\
            \includegraphics{figs/bigexample-3} \\
            \includegraphics{figs/bigexample-4}
        }
        \caption{Constructing the anagramish path $P$:
            (1) pairs of top and bottom paths are matched with zig-zag paths;
            (2) all remaining colourful blocks receive top paths;
            (3) all boring blocks receive top, updown, or downup paths.
        }
        \label{path_construction}
    \end{figure}
    \begin{compactenum}
        \item For each $a\in\Sigma$ such that $\delta_a(w)>0$, group the elements of $A_a$ into pairs.  For each pair $(i,j)$, $P$ contains the path through the top row of $H_i$ and the path through the bottom row of $H_j$.  For each element $i\in B_a$, $P$ contains the zig-zag path with both endpoints in the top row of $H_i$ and that contains every vertex of $H_i$.  (Note that the zig-zag path begins at the top and ends at the bottom row because $H_i$ contains an even number of columns of $G$.)

        \begin{figure}
            \centering{
                \begin{tabular}{ccc}
                    \includegraphics{figs/paths-1} &
                    \includegraphics{figs/paths-2} &
                    \includegraphics{figs/paths-3}
                \end{tabular}
            }
            \caption{Subpaths of $P$ through colourful blocks: A pair of top and bottom paths contribute the same amount as a single zig-zag path.}
            \label{paths}
        \end{figure}

        % \todo[inline]{Handle this comment: • fig. 4. Maybe do not use the same color for the bottom and the top, it took me a while to notice that the shapes were different and that this meant that the colors were different (or you could simply add a sentence to clarify this).
        % PM: I handled this by making the symbols huge}

        \item For each $a\in\Sigma$ such that $\delta_a(w)<0$ we proceed symmetrically to the previous case, but reversing the roles of $A_a$ and $B_a$.  Specifically, we group the elements of $B_a$ into pairs.  For each pair $(i,j)$, $P$ contains the path through the top row of $H_i$ and the path through the bottom row of $H_j$.  For each element $i\in A_a$, $P$ contains the zig-zag path with both endpoints in the top row of $H_i$ and that contains every vertex of $H_i$.

        \item For each $i\in\{1,\ldots,2r\}\setminus\bigcup_{a\in\Sigma}(I_a\cup B_a)$, $P$ contains the top row of $H_i$.
    \end{compactenum}
    The rules above define the intersection, $P_i$, of $P$ with each colourful block $H_i$ of $G_{n_s}$.  If $P_i$ is the path through the bottom (top) row of $H_i$ then we call $H_i$ a \emph{bottom (top) block}.  If $P_i$ is the zig-zag path that contains every vertex of $H_i$ then we call $H_i$ a \emph{zig-zag block}.  Note that $\sum_{a\in\Sigma} \delta_a = 0$ and this implies that the number of bottom blocks among $H_1,\ldots,H_{r-1}$ is the same as the number of bottom blocks among $H_{r+1},\ldots,H_{2r}$.  Indeed, this number is exactly $\beta:=\tfrac{1}{2}\sum_{a\in\Sigma} |\delta_a(w)|=\tfrac{1}{2}\tau(w)$.

    We now define how $P$ behaves for the boring blocks, denoted by  $Q_0,\ldots,Q_{2r-1}$. The first boring block $Q_0$ comes immediately before $H_1$. Each boring block $Q_j$, for $j\in\{1,\ldots,2r-1\}$ comes immediately after $H_j$ and immediately before $H_{j+1}$.  In almost every case, $P$ uses the path through the top row of $Q_j$.  The only exceptions are when $H_j$ or $H_{j+1}$ are bottom blocks. Note that, because of the global independence constraint, these two cases are mutually exclusive. See \cref{updownup}.

    \begin{figure}
        \centering{
            \begin{tabular}{ccc}
                \includegraphics{figs/crossings-1} &
                \includegraphics{figs/crossings-2} &
                \includegraphics{figs/crossings-3}
            \end{tabular}
            \caption{The paths taken by $P$ through boring blocks: A top path, a downup path, and an updown path.}
            \label{updownup}
        }
    \end{figure}

    \begin{compactenum}
        \item When $H_j$ is a bottom block, $P$ uses a path that begins at the bottom row of $Q_j$ but moves immediately to the top row of $Q_j$ and uses the entire path along the top row. We call this a \emph{downup} path.
        \item When $H_{j+1}$ is a bottom block, $P$ uses a path that begins at the top row of $Q_j$ and moves immediately to to the bottom row of $Q_j$ and uses the entire path along the bottom row.  We call this a \emph{updown} path.
    \end{compactenum}
    This completely defines the path $P:=v_1,\ldots,v_{2m}$. All that remains is to argue that the word $\rho:=\phi_w(v_1),\ldots,\phi_w(v_{2m})$ is anagramish.

    Observe that the number of downup paths and the number of updown paths that appear in $Q_0,\ldots,Q_{r-1}$ is exactly the same as the number of bottom blocks among $H_1,\ldots,H_{r-1}$ which is exactly $\beta$.  Similarly, the number of updown paths and downup paths that appear in $Q_{r+1},\ldots,Q_{2r-1}$ is exactly $\beta$.  Now every path that is neither downup nor updown uses the top row.  This implies that the sequence of colours contributed to $\rho$ by  the intersection of $P$ with $Q_0,\ldots,Q_{r-1}$ is a permutation of the sequence of colours contributed to $\rho$ by the intersection of $P$ with $Q_{r+1},\ldots,Q_{2r-1}$.  These two sequences cancel perfectly.

    Next, by construction, each pair of top and bottom blocks in $H_1,\ldots,H_{r-1}$ contributes exactly the same amount as a single matching zig-zag block in $H_{r+1},\ldots,H_{2r-1}$.  Specifically, if $x,y\in A_a$, $z\in B_a$, $H_x$ is a top block, $H_y$ is a bottom block and $H_z$ is a zig-zag block, then the contributions of $P_x$ and $P_y$ to $\rho$ cancels out the contribution of $P_z$. After doing this cancellation exhaustively, all that remains are top blocks, which also cancel each other perfectly.  This completes the proof.
\end{proof}

To be explicit, we finish the proof of \cref{main_vertex} by pointing out the contradiction between \cref{near_anagram_graph} and \cref{anagramish_path}:

\begin{proof}[Proof of \cref{main_vertex}]
    Assume for the sake of contradiction that there exists some $c\in\N$ such that $\afcn(G_n)\le c$ for each $n\in\N$.  With this assumption, \cref{near_anagram_graph} shows that for every $\epsilon>0$ and $r_0\in\N$ there exists a word $w\in L$ with $|w|\ge 2r_0$, $\tau(w)\le\epsilon r$, and such that $\phi_w$ is an anagram-free colouring of $G_{n_w}$.

    \cref{anagramish_path} shows that there are fixed values $\epsilon >0$ and $r_0\in \N$ such that, for any word $w\in L$ with $|w|\ge r_0$ and $\tau(w)\le\epsilon r$, the graph $G_{n_w}$ contain a path that is anagramish under $\phi_w$.  This certainly contradicts the fact that $\phi_w$ is an anagram-free colouring of $G_{n_w}$. We therefore conclude that, for every $c\in\N$ there exists an $n\in N$ such that $\afcn(G_n)> c$.
\end{proof}

% \todo[inline]{It would be nice to say something about the anagram-free chromatic index.  I'm pretty sure that it's constant for the $2\times n$ grid, but the result is not worth the effort.  More interesting would be a proof that its super-constant for the $2\times n$ grid with one diagonal in each square.  Almost everything we did above works except that we get hung up on a parity problem.  We can't do $2$ versus $1$ cancellation in $w$ but we can do $6$ versus $4$ cancellation. Everything works out except when the number of occurrences of some character is odd. In that case we get stuck with one uncancelled copy.}
%

\section{Proof of \cref{near_anagram_fourier}}
\label{near_anagram_proof}

% \todo[inline]{I suspect this lemma (and probably a better quantitative version) has a proof using Fourier approximation, but I don't know enough about Fourier approximation and my half-assed attempt to learn about it wasn't sufficient.}

All that remains is to prove \cref{near_anagram_fourier}, which we do now.

\begin{proof}[Proof of \cref{near_anagram_fourier}]
    First observe that, for any $\ell$-periodic word $w$, $|\Sigma_w|\le\ell$.
    Define an even-length word $t$ to be \emph{$a$-unbalanced} if $\tau_a(t)>\epsilon|t|/\ell$ and \emph{$a$-balanced} otherwise.  If $t$ is $a$-balanced for each $a\in\Sigma_t$ then $t$ is \emph{balanced}. Observe that, if $t$ is balanced then $\tau(t)\le |\Sigma_t|\epsilon|t|/\ell\le \epsilon|t|$. A word is \emph{everywhere unbalanced} if it contains no balanced factor of length $r\ge r_0$. Our goal therefore is to show that there is an upper bound $n:=n(\ell,\epsilon,r_0)$ on the length of any $\ell$-periodic everywhere unbalanced word.

    Let $h$ be a positive integer (that determines $n$ and whose value will be discussed later), and let $n:=r_02^{h}$. Let $w$ be an $\ell$-periodic everywhere unbalanced word of length $n$ over the alphabet $\Sigma$.  Assume, without loss of generality, that $r_0$ is a multiple of $\ell$.

    Consider the complete binary tree $T$ of height $h$ whose leaves, in order, are length-$r_0$ words whose concatenation is $w$ and for which each internal node is the factor obtained by concatenating the node's left and right child.
    % See \cref{binary_tree}.
    Note that for each $v\in V(T)$ and each $a\in\Sigma$, the fact that $w$ is $\ell$-periodic and $r_0$ is multiple of $\ell$ implies that $|v|_{a}\ge |v|/\ell$.

    % \begin{figure}
    %     \centering{
    %         \includegraphics[width=\textwidth]{figs/binary-tree-1}
    %     }
    %     \caption{Building a binary tree $T$ over the word $w:=4 3 5 3 2\ldots5 4 1 4$.}
    %     \label{binary_tree}
    % \end{figure}

  % For each node $v$ of $T$, let $h(v)$ denote the height of $v$'s subtree
  % and $w(v)=2^{h(v)}$ denote the length of the word $v$. For each
  % $i\in\Sigma$, let $w_i(v):=n_i(v)/s(v)$.  Note that $0\le w_i(v)\le
  % 1/2$ and that $\sum_{i\in\Sigma} w_i(v)=1$.  Furthermore, if $v$
  % has two children $x$ and $y$, then $w_i(v) = (w_i(x)+w_i(y))/2$.
  % The assumption that $w$ has no balanced factor implies $v$ is
  % $i$-unbalanced, for some $i\in\Sigma$.  Assign each internal node
  % $v$ of $T$ the \emph{label} $\ell(v):=\min\{i\in\Sigma: \mbox{$v$
  % is $i$-unbalanced}\}$.

  % \begin{figure}
  %   \centering{
  %      \includegraphics[width=\textwidth]{figs/binary-tree-3}
  %   }
  %   \todo[inline]{This is an old figure and needs updating.}
  %   \caption{The tree $T$ with the vector $(|v|_{1},\ldots,|v|_{5})$ for each
  %    internal node $v$.  The unbalanced index that defines $\ell(v)$ is highlited for each node.}
  %   \label{binary_tree_2}
  % \end{figure}

  For each $a\in\Sigma$, let $S_a:=\{v\in V(T): \text{$v$ is $a$-unbalanced}\}$. Since $w$ is everywhere unbalanced, $\bigcup_{a\in\Sigma} S_a=V(T)$. Therefore,
  \[
     (h+1)n = \sum_{v\in V(T)}|v|\le \sum_{a\in\Sigma} \sum_{v\in S_a} |v|
     \enspace .
  \]
  Therefore, there exists some $\alpha\in\Sigma$ such that $\sum_{v\in S_{\alpha}}|v|\ge (h+1)n/|\Sigma| \ge (h+1)n/\ell$.   At this point we are primarily concerned with appearances of $\alpha$, so let $X:=S_{\alpha}$, and, for each node $v\in V(T)$, let $W(v):=|v|_\alpha$.

  For each non-leaf node $v$ of $T$, let $R(v)$ denote a child of $v$ such that $W(R(v))\le \tfrac12\cdot W(v)$.  (It is helpful to think of $T$ as being ordered so that each right child $y$ with sibling $x$ has $W(y)\le W(x)$.)  For a non-leaf node $v\in X$ the fact that $v$ is $\alpha$-unbalanced implies that
  \[  W(R(v)) \le \tfrac{1}{2}\cdot W(v) - \tfrac{\epsilon}{2\ell}\cdot |v|
      \le (\tfrac12-\tfrac{\epsilon}{2\ell})W(v) \enspace .
  \]

  From this point on we use the following shorthands. For any $S\subseteq V(T)$, $L(S):=\sum_{v\in S}|v|$,
  % \todo{$L$ and $R$ are bad notation choices: $L$ for length, $R$ for right child}
  $W(S):=\sum_{v\in S}W(v)$, and $R(S)=\{R(v):v\in S\}$.  Summarizing, we have a complete binary tree $T$ of height $h$ and
  $X\subseteq V(T)$ with the following properties:
  \begin{compactenum}
    % \item For each $v\in V(T)$, $w(v) \le 1/2$.
    \item For each $v\in V(T)$,  $W(v)\ge |v|/\ell$.
    % \item For each internal node $v\in V(T)$ with children $x$ and $y$,
       % $w(v) = (w(x)+w(y))/2$.
    \item $L(X) \ge (h+1)n/\ell$.
    \item For each non-leaf node $v\in X$,
      $W(R(v)) \le (\tfrac{1}{2}-\tfrac{\epsilon}{2\ell})W(v)$.
  \end{compactenum}
  For each $i\in\{0,\ldots,h\}$, let $X_i\subseteq X$ denote the
  set of nodes $v\in X$ for which the path from the root of $T$ to $v$
  contains exactly $i$ nodes in $X$, excluding $v$.  See \cref{bigtree}. Observe that, since each node in $X_i$ has an ancestor in $X_{i-1}$,
  \[  n \ge L(X_0) \ge L(X_1) \ge \cdots\ge L(X_{h}) \enspace . \]

  We will show that there exists an integer $t:=t(\epsilon,\ell,r_0)$ such that, for each $i\in\{0,\ldots,h-t\}$,
  \begin{equation}
     L(X_{i+t}) \le (1-(1/2)^{t+1}) L(X_i) \enspace . \label{t}
  \end{equation}
  In this way,
  \begin{align*}
     (h+1)n/\ell
        \le L(X) & = \sum_{i=0}^{h} L(X_i) \\
           &\le \sum_{i=0}^{h} L(X_{t\floor{i/t}})
             & \text{(since $i\ge t\floor{i/t}$, so $L(X_i)\le L(X_{i\floor{i/t}}$))} \\
           & = t\cdot\sum_{i=0}^{h/t} L(X_{it})
             & \text{(for $h$ a multiple of $t$)} \\
           &\le t\cdot\sum_{i=0}^{\infty} (1-(1/2)^{t+1})^i L(X_0)
           & \text{(by \cref{t})}\\
           &\le tn\cdot \sum_{i=0}^{\infty} (1-(1/2)^{t+1})^i
           & \text{(since $|X_0|\le n$)} \\
           & = tn2^{t+1} \\
  \end{align*}
  which is a contradiction for sufficiently large $h$; in particular, for $h > \ell t2^{t+1}-1$.

  \begin{figure}
    \centering{
       \includegraphics[width=\textwidth]{figs/bigtree}
    }
    \caption{The partitioning of $X$ into $X_0,\ldots,X_h$. Shaded
    nodes are in $X$ and all nodes in $X_i$ are shaded with the same
    colour.   Starting with $A_0=X_0$, the elements of $A_0,\ldots,A_h$
    are highlighted (in orange).  The elements of $R(A_0),\ldots,R(A_h)$ are also highlighted (in pink).}
    \label{bigtree}
  \end{figure}

  It remains to establish \cref{t}, which we do now.  Fix some $i\in\{0,\ldots,h-1\}$, define $A_0:= X_i$ and, for each $j\ge 1$, define $A_j$ to be the subset of $X_{i+j}$ that are descendants of some node in $R(A_{j-1})$.  See \cref{bigtree}.  To upper bound $L(X_{i+t})$ observe that $X_{i+t}$ can be split into two sets $A_0'$ and $B$ defined as follows:  The nodes $A_0'$ do not have an ancestor in $R(A_0)$ and therefore $L(A_0')\le (1/2)L(A_0)$. The nodes in $B$ do have an ancestor in $R(A_0)$ and therefore have an ancestor in $A_1$.  Iterating this argument, we obtain
 \begin{align*}
      L(X_{i+t})
         &\le (1/2)L(A_0) + (1/2)L(A_1) + \cdots + (1/2)L(A_{t-1}) + L(A_t) \\
         &\le (1/2)L(A_0) + (1/4)L(A_0) + \cdots + (1/2)^t L(A_{0}) + L(A_t) \\
         &  = (1-(1/2)^t)L(A_0) + L(A_t) = (1-(1/2)^t)L(X_i) + L(A_t)  \enspace .
   \end{align*}
  So all that remains to establish \ref{t} is to prove that
  $L(A_t)\le (1/2)^{t+1}L(X_i)$.  To do this, observe that, for each $j\in\{1,\ldots,t\}$,
  \begin{equation}
      W(A_j)
        \le W(R(A_{j-1}))
        \le (\tfrac12-\tfrac{\epsilon}{2\ell})\cdot W(A_{j-1})
  \end{equation}
  which implies
  \[
       W(A_t)
       \le (\tfrac12-\tfrac{\epsilon}{2\ell})^t W(A_0) \le (\tfrac12-\tfrac{\epsilon}{2\ell})^t\cdot L(A_0)
       =  (\tfrac12-\tfrac{\epsilon}{2\ell})^t\cdot L(X_i)
  \]
  Since $w$ is $\ell$-periodic,
  \[
        L(A_t)\le \ell\cdot W(A_t) \le \ell\cdot(\tfrac12-\tfrac{\epsilon}{2\ell})^t\cdot L(X_i) \le L(X_i)/2^{t+1}
  \]
  for $t := \ceil{\log(2\ell)/\log((1/(1-\tfrac{\epsilon}{\ell}))}$.
\end{proof}

\section{Reflections}
\label{reflections}

Although an explicit upper bound on $n:=n(\epsilon,\ell,r_0)$ could be extracted from the proof of \cref{near_anagram_fourier} it would likely be far from tight.  We suspect that there is a Fourier analytic proof that would give better quantitative bounds.  We have not pursued this, because we have no idea how to explicitly upper bound $\ell$, for reasons discussed in the next paragraph.

\cref{periodicity} and its proof give absolutely no clues to help find a concrete bound on $\ell$ or to find a minimal set $\Xi$. Indeed, for some choices of $L$, doing so can be a difficult problem.  Consider the example where $|\Sigma|=5$ and $L$ is the language of all anagram free words in $\Sigma^*$. It is easy to see that this predicate $L$ is factor-closed and the result of \citet{pleasants:non-repetitive}, published in 1970, shows that this $L$ is infinite.  The question of whether $|\Xi|=4$ or $|\Xi|=5$ is then the question of determining whether there exist arbitrarily long anagram-free words on an alphabet of size $4$.  This was the open problem posed by \citet{erdos:some} in 1961 and again by \citet{brown:is} in 1971 and not resolved until 1992 when \citet{keranen:abelian,keranen:powerful} showed that the answer, in this case, is that $|\Xi|=4$.  However, if this were not the case, then determining $\ell$ would be the question of determining the length of the longest anagram-free word over an alphabet of size $4$.

Our proof uses \cref{periodicity} twice and each uses a language $L$ that is considerably more complicated than the language of anagram-free words. It seems unlikely that we will obtain concrete upper bounds on $\ell$ as a function of $c$ except, possibly, through the use of computer search. The resulting value $\ell$ is used in the application of \cref{near_anagram_fourier} and also within the proof of \cref{anagramish_path}.

\section*{Acknowledgement}

We would like to thank an anonymous referee for pointing out existing results and terminology from the field of combinatorics on words that has helped streamline the presentation in \cref{near_anagram_statement,vertex_colourings}.

\bibliographystyle{plainurlnat}
\bibliography{af2t}

\end{document}

### 2021-09-24 dynpro.py

@@ -1,0 +1,64 @@
+import sys
+import itertools
+import numpy
+
+
+
+def e(k, i):
+    return tuple([int(j==i) for j in range(k)])
+
+def rank_zero_colourings(k):
+    return [{e(k,i)} for i in range(k)]
+
+def increase_rank(colourings):
+    s = []
+    i = 0
+    for c1, c2 in itertools.combinations(colourings, 2):
+        if i % 10000 == 0:
+            m = len(colourings)*(len(colourings)-1)//2
+            print("{}/{} ({:.1f}%)\r".format(i, m, 100*i/m), end='')
+        i += 1
+        good_pair = True
+        for p1, p2 in itertools.product(c1, c2):
+            if not pair_has_center(p1, p2):
+                good_pair = False
+                break
+        if good_pair:
+            if not s:
+                print("Able to increase rank")
+            # Found a good pair of colourings, combine them
+            # print("Good pair: {}, {}".format(c1, c2))
+            s.append(combine_colourings(c1, c2))
+            s.append(combine_colourings(c2, c1))
+            # print(len(s))
+    print()
+    return s
+
+def combine_colourings(c1, c2):
+    root = [x for x in c1 if sum(x)==1][0]  # todo, inefficient
+    return c1.union({combine_pair(root, p) for p in c2})
+
+
+
+def pair_has_center(s1, s2):
+    for i in range(len(s1)):
+        if s1[i]+s2[i] == 1: return True
+    return False
+
+def combine_pair(p1, p2):
+    return tuple([min(2, p1[i]+p2[i]) for i in range(len(p1))])
+
+if __name__ == "__main__":
+    k = int(sys.argv[1])
+    colourings = rank_zero_colourings(k)
+    rank = 0
+    while colourings:
+        print("rank {} tree has {} colourings using {} colours".format(rank, len(colourings), k))
+        print("Here's one:")
+        print(colourings[0])
+        # print(colourings)
+        colourings = increase_rank(colourings)
+        rank += 1
+    print("rank {} tree has {} colourings using {} colours".format(rank, len(colourings), k))
+
+    print(colourings)

### 2021-09-24 dynpro2.py

@@ -1,0 +1,117 @@
+import sys
+import itertools
+import numpy
+
+
+
+def e(k, i):
+    return tuple([int(j==i) for j in range(k)])
+
+def rank_zero_colourings(k):
+    return [{e(k,i)} for i in range(k)]
+
+def increase_rank(colourings):
+    s = []
+    i = 0
+    for c1, c2 in itertools.combinations(colourings, 2):
+        if i % 10000 == 0:
+            m = len(colourings)*(len(colourings)-1)//2
+            print("{}/{} ({:.1f}%)\r".format(i, m, 100*i/m), end='')
+        i += 1
+        good_pair = True
+        for p1, p2 in itertools.product(c1, c2):
+            if not pair_has_center(p1, p2):
+                good_pair = False
+                break
+        if good_pair:
+            if not s:
+                print("Able to increase rank")
+            s.append(combine_colourings(c1, c2))
+            s.append(combine_colourings(c2, c1))
+    print()
+    return eliminate_duplicates(s)
+
+
+def eliminate_duplicates(colourings):
+    if not colourings: return colourings
+    a = { str(sorted(colourings[i])) : i for i in range(len(colourings)) }
+    return [colourings[a[c]] for c in a]
+
+def combine_colourings(c1, c2):
+    root = [x for x in c1 if sum(x)==1][0]  # todo, inefficient
+    return c1.union({combine_pair(root, p) for p in c2})
+
+def pair_has_center(s1, s2):
+    for i in range(len(s1)):
+        if s1[i]+s2[i] == 1: return True
+    return False
+
+def combine_pair(p1, p2):
+    return tuple([min(2, p1[i]+p2[i]) for i in range(len(p1))])
+
+def filter_pareto_optimal(colourings):
+    optimal = list()
+    for i1 in range(len(colourings)):
+        c1 = colourings[i1]
+        keep = True
+        # if i % 1000 == 0:
+        m = len(colourings)
+        discards = i1-len(optimal)
+        print("{}/{} ({:.1f}%) [-{}]\r".format(i1, m, 100*i1/m, discards), end='')
+        for i2 in itertools.chain(range(i1+1, len(colourings)), optimal):
+            c2 = colourings[i2]
+            assert(c1 != c2)
+            if c2.issubset(c1):
+                keep = False
+                break
+        if keep:
+            optimal.append(i1)
+    print()
+    return [colourings[i] for i in optimal]
+
+    # optimal = list()
+    # pareto_optimal = True
+    # for i in range(len(colourings)):
[…]

### 2021-09-27 lin-vs-cen.tex
Getting somewhere

@@
 \setlength{\linenumbersep}{2em}
-\linenumbers
+% \linenumbers
 % \rightlinenumbers
-\linenumbers
+% \linenumbers
 \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
@@
 
-\DeclareMathOperator{\ddiv}{div}
-\DeclareMathOperator{\hist}{p}
+% \DeclareMathOperator{\ddiv}{div}
+% \DeclareMathOperator{\hist}{p}
+\DeclareMathOperator{\dist}{dist}
 
@@
 
+\section{Preliminaries}
+
+For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.
+For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.\todo{Surely this next lemma is well-known?}
+
+\begin{lem}\label{hall_vees}
+  Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.
+Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
+\end{lem}
+
+\begin{proof}
+  Let $\Lambda$ be a subgraph of $G$ in which each vertex of $A$ has degree at most two, each vertex of $B$ has degree at most one, and that maximizes $\deg_\Lambda(A)$.
+If $\deg_{\Lambda}(A) = 2|\Lambda|$ then there is nothing to prove.
+Assume therefore, for the sake of contradiction,  that there exists $v_0\in A$ with $\deg_\Lambda(v_0) < 2$.
+
+  We say that a path $v_0,\ldots,v_m$ in $G$ is \emph{$\Lambda$-aternating} if $v_iv_{i+1}\in E(\Lambda)$ for all odd $i\in\{0,\ldots,m-1\}$ and $v_iv_{i+1}\not\in E(\Lambda)$ for all even $i\in\{0,\ldots,m\}$.
+Observe that any such path has even length so that it ends at a vertex $v_m\in A$.
+Otherwise, replacing the edges $\lfloor m/2\rfloor$ edges $\{v_1v_2,v_3v_4,\ldots,v_{m-2}v_{m-1}\}\in E(\Lambda)$ with the $\lceil m/2\rceil$ edges $\{v_0v_1,v_2v_3,\ldots,v_{m-1}v_m\}$ does not change $\deg_\Lambda(v)$ for any $v\in A_0\setminus\{v_0\}$ but does increase  $\deg_\Lambda(v_0)$, contradicting the assumption that $\Lambda$ maximizes $\deg_\Lambda(A)$.
+
+  Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
+  Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.
+Let $a$ be the number of edges of $\Lambda$ having one endpoint $A_0$ and one endpoint in $B_0$, so that $a=\deg_\Lambda(B_0)=|B_0|$.
+Let $b$ be the number of edges of $\Lambda$ having one endpoint in $A_0$ and one endpoint in $B\setminus B_0$, so that $\deg_\Lambda(A_0)=a+b = |B_0|+b$.
+Since $\deg_{\Lambda}(v) \le 2$ for each $v\in A_0$ and $\deg_\Lambda(v_0)<2$, $\deg_\Lambda(A_0)<2|A_0|$.
+This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.
+\end{proof}
+
 \section{The Linear Colouring Number of Pseudogrids}
@@
 
-The $k\times k$ grid graph $G_k$ is the graph with vertex set $V(G_k):=\{1,\ldots,k\}^2$ and that contain an edge with endpoints $(v,w)$ and $x,y$ if and only $|v-x|+|w-y|=1$.
+The $k\times k$ grid graph $G_k$ is the graph with vertex set $V(G_k):=\{1,\ldots,k\}^2$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only $|v-x|+|w-y|=1$.
 A graph $H$ is a $k\times k$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $G$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_k$.
@@
 \begin{proof}
-  Let $\varphi:V(H)\to\{1,\ldots,c\}$ be an arbitrarily $c$-colouring of some $k\times k$ pseudogrid $H$ for some $c\le \epsilon k$, where $\epsilon$ is a small positive constant.
+  Let $\varphi:V(H)\to\{1,\ldots,C\}$ be an arbitrarily $C$-colouring of some $k\times k$ pseudogrid $H$ for some $C\le \epsilon k$, where $\epsilon$ is a small positive constant.
 We will show that $H$ contains a path $P$ which has no center with respect to $c$.
@@
 
-  For each colour $\alpha\in\{1,\ldots,c\}$, define the \emph{multiplicity} of $\alpha$ as the number of vertices and edges of $G_k$ that include $\alpha$ in their colour set.
-Fix some large integer constant $A$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity at most $A$.
+  For each colour $\alpha\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $\alpha$ as the number of vertices and edges of $G_k$ that include $\alpha$ in their colour set.
+Fix some large integer constant $\alpha$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity at most $\alpha$ and \emph{frequent} otherwise.
 
-  If there exists some infrequent colour $\alpha\in\{1,\ldots,c\}$, we identify every row and column of $G_k$ that contains an edge or vertex with $\alpha$ in its colour set.
-We then \emph{remove} each of these (at most $A$) rows and columns from $H$ by removing all edges of $H$ in $P_e$ for each $e$ in the relevant row (column) of $G_k$ and then discarding any isolated vertices.
-After doing this for each row and column containing $\alpha$, the graph $H$ may be disconnected.
-We discard any isolated vertices and any path $P_{v}$ that is disconnected from the rest of $H$.
-
-  The resulting graph $\tilde{H}$ contains a $k'\times k'$ pseudogrid for some $k' \le k-A$ and contains no vertex of colour $\alpha$.
-We then set $H\gets \tilde{H}$, $k\gets k'$ and repeat this process until eventually obtain a new $k'\times k'$ pseudogrid $\tilde{H}$ for some $k'\ge k-cA\ge (1-\epsilon A)k \ge k/2$, for $\epsilon \le 1/2A$.
-From this point onward that that there is no infrequent colour, i.e., every colour $\alpha\in\{1,\ldots,c\}$ has multiplicity at least $A$.
-
-  Next we run a two stage process that, for each $\alpha\in\{1,\ldots,c\}$, will identify a pair of vertices $v_\alpha,w_\alpha$ in $H$ whose colour is $\alpha$.
-These pairs are carefully chosen so that we can construct a simple path $P$ in $H$ that contains $\bigcup_{\alpha\in\{1,\ldots,c\}} \{v_\alpha,w_\alpha\}$.
-
-  \begin{compactenum}[Stage 1]
-    \item We begin this stage by defining each edge and vertex of $G_k$ as being initially \emph{unmarked}.
-Fix some constant $B>0$.
-We iterate over each $\alpha\in\{1,\ldots,c\}$ in turn.
-If $G_k$ contain a pair $(x_\alpha,y_\alpha)$ where each of $x_\alpha$ and $y_\alpha$ is an unmarked vertex or unmarked edge of $G_k$ with $\alpha$ in its colour set and the distance in $G_k$ between $x_\alpha$ and $y_\alpha$ is at least $B$ then Stage~1 \emph{succeeds} for $\alpha$.
-In this case, we take $v_\alpha$ to be any vertex in $P_{x_\alpha}$ of colour $\alpha$ and $w_\alpha$ to be any vertex in $P_{y_\alpha}$ of colour $\alpha$.
-We then \emph{mark} every vertex and edge of $G_k$ whose distance from $x_\alpha$ or $y_\alpha$ is less than $B$.
-
-    If $G_k$ contains no such pair $(x_\alpha,y_\alpha)$, then we leave $v_\alpha$ and $w_\alpha$ undefined and say that Stage~1 \emph{failed} for colour $\alpha$.
-
-    \item Let $S\subset\{1,\ldots,c\}$ be the set of colours $\alpha$ for which Stage~1 succeeded and let $\overline{S}:=\{1,\ldots,c\}\setminus S$ be the set of colours for which Stage~1 failed.
-We construct a bipartite graph $K$ with parts $L(K):=\overline{S}$ and $R(K):=\bigcup_{\alpha\in S}\{x_\alpha,y_\alpha\}$.
-For each marked vertex or edge $z$ of $G_k$, let $s(z)$ be the vertex in $R(k)$ that is closest to $z$ in $G_k$.\todo{Gotta be really precise here to avoid degree 3 issue.}  The edge set of $K$ contains an edge $\alpha\beta$ precisely when there is a marked edge or vertex $z$ of $G_k$ such that $\alpha\in\phi(z)$ and $s(z)=\beta$.
-
-    \begin{clm}
-      $K$ contains two matchings $M_1$ and $M_2$ such that each vertex in $L(K)$ appears in each of $M_1$ and $M_2$ and each vertex in $R(k)$ appears in at most one of $M_1$ or $M_2$.
+  \begin{clm}\label{all_frequent}
+    $H$ contains a $k'\times k'$ pseudogrid in which every colour is frequent, for some $k'\ge (1-\epsilon \alpha)k$.
   \end{clm}
@@
   \begin{proof}
-      Use Hall's Mathcing Theorem/Lemma.
+    If there exists some infrequent colour $c\in\{1,\ldots,C\}$, we identify every row and column of $G_k$ that contains an edge or vertex with $c$ in its colour set.
+We then \emph{remove} each of these (at most $A$) rows and columns from $H$ by removing all edges of $H$ in $P_e$ for each $e$ in the relevant row (column) of $G_k$ and then discarding any isolated vertices.
+After doing this for each row and column containing $c$, the graph $H$ may be disconnected.
+We discard any isolated vertices and any path $P_{v}$ that is disconnected from the rest of $H$.
+
+    The remaining graph contains a $k'\times k'$ pseudogrid for some $k' \ge k-A$ and contains no vertex of colour $c$.
+We then repeat this (at most $c$ times) until the resulting graph has no infrequent colour.
+This leaves a graph that contains a $k'\times k'$ pseudogrid for some $k'\ge k-C\alpha\ge(1-\epsilon \alpha)k$ for which every colour is frequent.
   \end{proof}
 
-    Finally, for each $\alpha\in\overline{S}$, we define $x_\alpha$ to be the the unique element of $R(K)$ such that $M_1$ includes the edge $\alphax_\alpha$ and we define $y_\alpha$ to be the unique element of $R(K)$ such that $M_2$ includes the edge $\alpha y_\alpha$.
-We define $v_\alpha$ and $w_\alpha$ as before.
+  From this point onward, we assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$.
+Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair of vertices $v_c,w_c$ in $H$ whose colour is $c$.
+These pairs are carefully chosen so that we will be able to construct a simple path $P$ in $H$ that contains $\bigcup_{c\in\{1,\ldots,C\}} \{v_c,w_c\}$.
 
-    Now, let $R:=\bigcup_{\alpha\in\{1,\ldots,c\}} \{x_\alpha,y_\alpha\}$.
+  \begin{compactenum}[{Stage} 1.]
+    \item We begin this stage by defining each vertex and edge of $G_k$ as being initially \emph{unmarked}.
+Fix some integer constant $\beta>0$.
+We iterate over each $c\in\{1,\ldots,C\}$ in turn.
+If $G_k$ contain a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is an unmarked vertex or unmarked edge of $G_k$ with $c$ in its colour set and the distance in $G_k$ between $x_c$ and $y_c$ is at least $B$ then Stage~1 \emph{succeeds} for $c$.
+In this case, we take $v_c$ to be any vertex in $P_{x_c}$ of colour $c$ and $w_c$ to be any vertex in $P_{y_c}$ of colour $c$.
+We then \emph{mark} every vertex and edge $z$ of $G_k$ such that $\dist_{G_k}(z, \{x,y\})\le \beta$.
+
+    If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $v_c$ and $w_c$ undefined and say that Stage~1 \emph{fails} for colour $c$.
+Note that Stage~1 can only fail for $c$ if $G_k$ contains few than $B^2$ unmarked vertices/edges with $c$ in their colour sets.
+
+    \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeded and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 failed.
+We construct a bipartite graph $I$ with vertex parts $A:=\overline{S}$ and $B:=\bigcup_{c\in S}\{x_c,y_c\}$.
+For each marked vertex or edge $p$ of $G_k$, let $s(p)$ be a vertex $q\in B$ that minimizes $\dist_{G_k}(p,q)$.\todo{Gotta be really precise here to avoid degree 3 issue.}  The edge set of $I$ contains an edge $xy$ precisely when there is a marked edge or vertex $z$ of $G_k$ such that $x\in\phi(z)$ and $s(z)=y$.
+
+    First observe that, for any node in $y\in B$, the number of $x\in A$ such that $\dist_{G_k}(x,y) \le \beta$ is at most $8\beta^2$.
+Therefore, the number of marked vertices/edges $x$ of $G_k$ with $s(x)=y$ is at most $8\beta^2$.
+Since each colour $c\in A$ is frequent, this implies that $\deg_I(c)\ge \alpha/8\beta^2$.
+Similarly, for each $y\in B$,  $\deg_{I}(y) \le 8\beta^12$.
+This implies that, for each $A_0\subseteq A$, $|N_I(A_0)|\ge \alpha/64\beta^4$.
+In particular, for any $\alpha$ and $\beta$ with $\alpha > 128\beta^4$,  $|N_I(A_0)|\ge 2|A_0|$.
+Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
+Therefore, each $c\in A=\overline{S}$ has two $\Lambda$-neighbours $p$ and $q$ in $B$.
+The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $s(x_c)=p$ and $c\in \phi(x_c)$.
+Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $s(y_c)=q$ and $c\in\phi(y_c)$.
+This defines the pairs $\{x_c,y_c\}$ for each element in $\overline{S}$.
+  \end{compactenum}
+
+  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of $2c$ vertices and edges of $G_k$ where each $c\in \phi(x_c)\cap \phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
+Ultimately, we want to use these to find a path in $H$ that contains at least two vertices of each colour.
+We start by first establishing some properties of $R\subseteq V(G_k)$.
+
+  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$.
+
+  \begin{clm}\label{far_from_home}
+    For each $z\in R_1$ and $z'\in R_2$ with $s(z')\neq z$, $\dist_{G_k}(z,z')\ge \beta /2$.
+  \end{clm}
+
+  \begin{proof}
+    Since $s(z')\neq z$, there exists some other $z''\in R_1\setminus\{z\}$ with $s(z')=s(z'')$.
+Since $z,z''\in R_1$ each of $z$ and $z''$ were selected in Stage~1.
+The rules for selecting vertices in Stage~1 guarantee that $\dist_{G_k}(z,z'')\ge \beta$.
+By the triangle inequality $\dist(z, z')+\dist(z',z'') \ge \dist(z,z'')\ge \beta$.
+Since $s(z')=z''$, $\dist_{G_k}(z', z'') \le \dist_{G_k}(z, z')$, so $\dist_{G_k}(z,z') \ge \dist(z',z'')/2 \ge \beta/2$.
+  \end{proof}
 
@@
 
+  \begin{proof}
+    To prove this, we have to define $s(x)$ a little more carefully.
+  \end{proof}
+
   \begin{clm}
-        Each $B/2\times B_/2$ subgrid of $G_k$ contains at most four elements of $R$.
+    Any $5$-element subset of $R$ contains a pair $x,y$ with $\dist_{G_k}(x,y)\ge\beta/2$.
   \end{clm}
+
+  \begin{proof}
+    Let $R'$ be some $5$-element subset of $R$.
+$R'$ contains two or more elements of $R_1$ then the distance between these two elements is at most $\beta>\beta/2$.
+Therefore, we may assume that $R'$ contains at most one element in $R_1$.
+If $R'$ contains exactly one element $x$ in $R_1$ then, by \cref{far_from_home}, there is at most one other element in $R'$ whose distance to $x$ is at most $\beta/2$, and the three remaining elements have greater distance.
+Therefore, $R'$ contains no elements in $R_1$.
+Therefore, $S:=\{s(x):x\in R'\}$ is a five-element set whose pairwise distances are all at most $\beta$.
+But this implies that two elements in $R'$ have distance less than $\beta$.
+This is a contradiction because all points in $R'$ are in $R_1$ and all pairs of points in $R_1$ have distance at least $\beta$.\todo{Something is off by a factor of 2.}
+  \end{proof}
+
+  \begin{cor}
+      In any $\beta/2\times \beta/2$ subgrid $G_k$ contains at most $4$ elements of $R$.
+  \end{cor}
 
@@
     \end{lem}
-
 \end{proof}
@@
 \end{document}
-
-Two words $s$ and $t$ are \emph{anagrams} of each other if $s$ is a permutation of $t$.
-A single word $w:=w_1,\ldots,w_{2r}$ is \emph{anagramish} if its first half $w_1,\ldots,w_r$ and its second half $w_{r+1},\ldots,w_{2r}$ are anagrams of each other.
-A word is \emph{anagram-free} if it does not contain an anagramish factor.
-Anagramish words are also known as \emph{abelian squares} and anagram-free words are also known as \emph{abelian square free} words \cite{keranen:abelian,keranen:powerful}, \emph{strongly nonrepetitive} words \cite{pleasants:non-repetitive}, or \emph{strongly asymmetric} sequences \cite{evdokimov:strongly,evdokimov:strongly2}.
-
-In 1961, \citet{erdos:some} asked if there exist arbitrarily long words over an alphabet of size $4$ that are anagram-free.\footnote{This was an incredibly prescient question since it is not at all obvious that there exist arbitrarily long anagram-free words over \emph{any} finite alphabet.
-The only justification for choosing the constant $4$ is that a short case analysis rules out the possibility of length-$8$ anagram-free words over an alphabet of size $3$.}  In 1968 \citet{evdokimov:strongly,evdokimov:strongly2} showed the existence of arbitrarily long anagram-free words over an alphabet of size $25$ and in 1971 \citet{pleasants:non-repetitive} showed an alphabet of size $5$ is sufficient.
-Erd\H{o}s's question was not fully resolved until 1992, when \citet{keranen:abelian} answered it in the affirmative.
-
-A path $v_1,\ldots,v_{2r}$ in a graph $G$ is \emph{anagramish} under a vertex $c$-colouring $\phi:V(G)\to\{1,\ldots,c\}$ if $\phi(v_1),\ldots,\phi(v_{2r})$ is an anagramish word.
-The colouring $\phi$ is an \emph{anagram-free colouring} of $G$ if no path in $G$ is anagramish under $\phi$.
-The minimum integer $c$ for which $G$ has an anagram-free vertex $c$-colouring is called the \emph{anagram-free chromatic number} of $G$ and is denoted by $\afcn(G)$.
-
-For a graph family $\mathcal{G}$, $\afcn(\mathcal{G}):=\max\{\afcn(G):G\in\mathcal{G}\}$ or $\afcn(\mathcal{G}):=\infty$ if the maximum is undefined.
-The results on anagram-free words discussed in the preceding paragraph can be interpreted in terms of $\afcn(\mathcal{P})$ where $\mathcal{P}$ is the family of all paths.
-Indeed, Ker\"anen's result shows that $\afcn(\mathcal{P})\le 4$.
-Slightly more complicated than paths are trees.
-\citet{wilson.wood:anagram-free} showed that $\afcn(\mathcal{T})=\infty$ for the family $\mathcal{T}$ of trees and \citet{kamcev.luczak.ea:anagram-free} showed that $\afcn(\mathcal{T}_2)=\infty$ even for the family $\mathcal{T}_2$ of binary trees.
-
-One positive result in this context is that of
-\citet{wilson.wood:anagram-free}, who showed that every tree $T$ of pathwidth $p$ has $\afcn(T)\le 4p+1$.
-Since trees are graphs of treewidth $1$ it is natural to ask if this result can be extended to show that every graph $G$ of treewidth $t$ and pathwidth $p$ has $\afcn(G)\le f(t,p)$ for some function $f:\N^2\to\N$.
-\citet{carmi.dujmovic.ea:anagram-free} showed that such a generalization is not possible for any $t\ge 3$ by giving examples of $n$-vertex graphs of pathwidth $3$ (and treewidth $3$) with $\afcn(G)\in\Omega(\log n)$.
-The obvious remaining gap left by these two works is graphs of treewidth $2$.
-Our main result is to show that $G_n$, the $2\times n$ square grid has $\afcn(G_n)\in\omega_n(1)$.
-Since $G_n$ has pathwidth 2, we have:
-
-\begin{thm}\label{main_vertex}
-    For every $c\in\N$, there exists a graph of pathwidth $2$ that has no anagram-free vertex $c$-colouring.
-\end{thm}
-
-\citet[Section 7.1]{wilson:anagram-free} conjectured that $\afcn(G_n)\in\omega_n(1)$, so this work confirms this conjecture.
-Prior to the current work, it was not even known if the family of $n\times n$ square grids had anagram-free colourings using a bounded number of colours.
-
-In a larger context, this lower bound gives more evidence that, except for a few special cases (paths \cite{evdokimov:strongly,pleasants:non-repetitive,keranen:abelian}, trees of bounded pathwidth \cite{wilson.wood:anagram-free}, and highly subdivided graphs \cite{wilson.wood:anagram-free2}), the qualitative behaviour of anagram-free chromatic number is not much different than that of treedepth/centered colouring \cite{nesetril.ossona:tree-depth}.
-Very roughly: For most graph classes, every graph in the class has an anagram-free colouring using a bounded number of colors precisely when every graph in the class has a colouring using a bounded number of colours in which every path contains a colour that appears only once in the path.
-
-% \todo[inline]{Statement of vertex colouring result (hopefully)}
-%
-% An \emph{edge $c$-colouring} $\varphi:E(G)\to\{1,\ldots,c\}$ is \emph{anagram-free} if $G$ has no path $v_0,\ldots,v_{2r}$ such that $\varphi(v_0v_1),\ldots,\varphi(v_{2r-1}v_{2r})$ is anagramish.
-We also give an edge colouring version of \cref{main_vertex}.
-%
-% \begin{thm}\label{main_edge}
-%     For every $c\in\N$, there exists a graph of $G$ of pathwidth $2$ that has no anagram-free edge $c$-colouring.
-% \end{thm}
-
-% \todo[inline]{Relation to Existing Work}
-
-The remainder of this paper is organized as follows: \Cref{near_anagram_statement} gives some definitions and states a key lemma which shows, under a certain periodicity condition, that every sufficiently long word contains a factor that is $\epsilon$-close to being anagramish.
-% In \cref{edge_colourings} we prove \cref{main_edge}.
-In \cref{vertex_colourings} we use this key lemma to prove \cref{main_vertex}.
-In \cref{near_anagram_proof} we prove the key lemma.
-\cref{reflections} concludes with some final remarks about the (non-)constructiveness of our proof technique.
-
-\section{Periodicity in words}
-\label{near_anagram_statement}
-
-An \emph{alphabet} $\Sigma$ is a finite non-empty set and each element of $\Sigma$ is called a \emph{letter}.
-A \emph{word} over $\Sigma$ is a (possibly empty) sequence $w:=w_1,\ldots,w_n$ with $w_i\in\Sigma$ for each $i\in\{1,\ldots,n\}$.
-The \emph{length} $|w|$ of $w$ is the length, $n$, of the sequence.
-A word of length at least $1$ is \emph{non-empty}.
-The unique word of length $0$ is denoted by $\varepsilon$.
-For a word $w:=w_1,\ldots,w_n$, we let $\Sigma_w:=\{w_i:i\in\{1,\ldots,n\}\}$.
-For each $1 \le i \le j\le n+1$, $w_i,w_{i+1},\ldots,w_{j-1}$ is called a \emph{factor} of $w$.\footnote{Note that this definition includes the empty factor $\varepsilon$ obtained when $i=j$.}  The set of all factors of $w$ is denoted by $F(w)$.
-
-A \emph{language} $L$ over $\Sigma$ is a set of words over $\Sigma$.
-A language $L$ is \emph{infinite} if $|L|=\infty$.
-We let $\Sigma_L:=\bigcup_{w\in L}\Sigma_w$ and $F(L):=\bigcup_{w\in L}F(w)$.
-The language $L$ is \emph{factor-closed} if $F(w)\subseteq L$ for each $w\in L$.
-For any alphabet $\Sigma$ and any $k\in\N$, $\Sigma^k$ (the $k$-fold cartesian product of $\Sigma$ with itself) is the language consisting of all length-$k$ words over $\Sigma$.
-The \emph{Kleene closure} $\Sigma^*:=\bigcup_{k=0}^\infty \Sigma^k$ is the language consisting of all words over $\Sigma$.
-
- % (Note the convention that $w[i\mathbin{:}j]$ does not include $w_j$, so $w[i\mathbin{:}j]$ has length $j-i$.)\todo{Is this notation still needed after revisions?}
-
-Let $w:=w_1,\ldots,w_n$ be a word over an alphabet $\Sigma$ and, for each $a\in \Sigma$, define $|w|_a:=|\{i\in\{1,\ldots,n\}:w_i=a\}|$.
-The \emph{Parikh vector} of $w$ is the integer-valued $|\Sigma|$-vector $\hist(w):=(|w|_{a}:a\in\Sigma)$ that is indexed by elements of $\Sigma$.
-Observe that a word $w_1,\ldots,w_{2r}$ is anagramish if and only if $\hist(w_1,\ldots,w_r)=\hist(w_{r+1},\ldots,w_{2r})$ or, equivalently, $\hist(w_1,\ldots,w_r)-\hist(w_{r+1},\ldots,w_{2r})=\boldsymbol{0}$.
-% Let $\|\cdot\|_1$ denote the L1 norm of a vector.\footnote{The L1 norm of a real-valued $d$-vector $\boldsymbol{v}:=(v_1,\ldots,v_d)$, $\|\boldsymbol{v}\|_1=\sum_{i=1}^d|v_i|$.}
-For each $a\in\Sigma$, let $\delta_a(w):=|w_1,\ldots,w_r|_{a}-|w_{r+1},\ldots,w_{2r}|_{a}$ and let $\tau_a(w):=|\delta_a|$.
-Then $\tau(w):=\sum_{a\in\Sigma}\tau_a(w)$ is a useful measure of how far a word is from being anagramish and that $\tau(w)=0$ if and only if $w$ is anagramish.
-
-A word $w:=w_1,\ldots,w_n$ is \emph{$\ell$-periodic} over an alphabet $\Sigma\supseteq\Sigma_w$ if each length-$\ell$ factor of $w$ contains every letter in $\Sigma$.
-A language $L$ is \emph{$\ell$-periodic}  if each of its words is $\ell$-periodic over $\Sigma_L$.
-A language $L$ is \emph{periodic} if it is $\ell$-periodic for some finite $\ell$.
-We make use of the following key lemma, which states that every sufficiently long $\ell$-periodic word contains arbitrarily long factors that are $\epsilon$-close to being anagramish.
-
-\begin{lem}\label{near_anagram_fourier}
-    For each $r_0,\ell\in\N$ and each $\epsilon>0$, there exists a positive integer $n$ such that each $\ell$-periodic word $w_1,\ldots,w_n$ contains a factor $w:=w_{i+1},\ldots,w_{i+2r}$ of length $2r \ge 2r_0$ such that $\tau(w)\le \epsilon r$.
-\end{lem}
-
-The proof of \cref{near_anagram_fourier} is deferred to \cref{near_anagram_proof}.
-We now give some intuition as to how it is used.
-The process of checking if a word is anagramish can be viewed as finding common letters in the first and second halves and crossing them both out.
-If this results in a complete cancellation of all letters, then the word is an anagram.
-\cref{near_anagram_fourier} tells us that we can always find a long factor $w$ where, after exhaustive cancellation, only an $\epsilon$-fraction of the original letters remain.
-% Informally, the factor $w$ is $\epsilon$-close to being anagramish.
-
-\cref{near_anagram_fourier} says that, if up to $\epsilon r$ letters in each half of $w$ were each allowed to cancel two letters of the same letters in the other half of $w$, then it would be possible to complete the cancellation process.
-To achieve this type of one-versus-two cancellation in our setting, we decompose our coloured pathwidth-$2$ graph into pieces of constant size.
-The vertices in each piece can be covered with one path or partitioned into two paths (see \cref{paths}).
-In this way an occurrence $H_z$ of a particular coloured piece in one half can be matched with two like-coloured pieces $H_x$ and $H_y$ in the other half.
-We construct a single path $P$ that contains all vertices in $H_z$ and only half the vertices in each of $H_x$ and $H_y$.
-In this way, the colours of vertices $P\cap H_z$ can cancel the colours of the vertices in $P\cap(H_x\cup H_y)$.
-
-Since \cref{near_anagram_fourier} requires that the word $w$ be $\ell$-periodic, the following lemma will be helpful in obtaining words that can be used with \cref{near_anagram_fourier}.
-(Although \cref{periodicity} follows immediately from considerably stronger results in combinatorics on words---for example \citet[Proposition~1.5.12]{lothaire:algebraic}---we provide a proof here for the sake of completeness.)
-
-\begin{lem}\label{periodicity}
-    Let $L$ be an infinite factor-closed language over an alphabet $\Sigma$.  Then there exists an infinite periodic factor-closed language $M\subseteq L$.
-\end{lem}
-
-\begin{proof}
-    Let $\Xi\subseteq\Sigma$ be any minimal subset of $\Sigma$ with the property that $M:=\Xi^* \cap L$ is infinite.
-Observe that, since $L$ is factor closed, $M$ is also factor-closed.
-Therefore $M$ is infinite and factor-closed.
-All that remains is to show that $M$ is $\ell$-periodic for some integer $\ell$.
-
-    Since $\Xi$ is minimal, $\Pi^*\cap L$ is finite for any strict subset $\Pi\subsetneq \Xi$.
-Since $L$ is factor-closed, $\Pi^*\cap L$ is factor-closed.
-Since $\Pi^*\cap L$ is factor-closed and finite, $|w|$ is finite for each $w\in\Pi^*\cap L$.
-    Therefore, the value
-    \begin{equation}
-        \ell:=1+\max\{|w|: w\in \Pi^*\cap L,\, \Pi\subsetneq\Xi\} \label{l-def}
-    \end{equation}
-    is finite.
-Now, consider any factor $f$ of any word in $M$ that does not contain every letter of $\Xi$, i.e., $f\in\Pi^*$ for some  $\Pi\subsetneq\Xi$.
-Since $\Pi^*\cap L$ is factor closed, $f\in \Pi^*\cap L$ so, by \cref{l-def}, $|f| < \ell$.
-Contrapositively, any factor of length at least $\ell$ of any word in $M$ contains every element of $\Xi$, so $M$ is $\ell$-periodic.
-\end{proof}
-
-\section{Proof of \cref{main_vertex}}
-\label{vertex_colourings}
-
-For each $n\in\N$, let $G_n$ be the $2\times n$ square grid with top row $a_0,\ldots,a_{n-1}$ and bottom row $b_0,\ldots,b_{n-1}$ (see \cref{g_n}).
-For the sake of contradiction, suppose that there exists an integer $c$ such that $\afcn(G_n)\le c$ for each $n\in\N$.
-Observe that there is a bijection between vertex $c$-colourings of $G_n$ and words of length $2n$ over the alphabet $\Sigma_0:=\{1,\ldots,c\}$ where a word $w:=w_0,\ldots,w_{2n-1}$ corresponds to the colouring $\phi_w:V(G_n)\to\{1,\ldots,c\}$ defined by
-$\phi_w(a_i):=w_{2i}$ and $\phi_w(b_i):=w_{2i+1}$, for each $i\in\{0,\ldots,n-1\}$.
-
-\begin{figure}
-    \centering{
-        \includegraphics{figs/g_n}
-    }
-    \caption{The graph $G_n$}
-    \label{g_n}
-\end{figure}
-
-% Let $L_0\subseteq \Sigma_0^*$ consist of all words $w\in \Sigma_0^*$ whose length is a multiple of $8$ and such that $\phi_w$ is an anagram-free colouring of $G_{|w|/2}$.
-Furthermore, the assumption that $\afcn(G_n)\le c$ for each $n\in\N$ implies that $L_0$ is infinite.
-
-For reasons that will become apparent, it is useful to break $G_n$ up into blocks consisting of $4$ columns each.
-Let $\Sigma_1:=\Sigma_0^8$ and define the language $L_1$ to consist of exactly those words $w_1,\ldots,w_r\in \Sigma_1^*$ for which $\phi_{w_1\cdots w_r}$ is an anagram-free colouring of $G_{4r}$.
-From the assumption that $\afcn(G_n)\le c$ for all $n\in\N$, it follows that $L_1$ is infinite.
-From the definition of anagram-free colouring, it follows that $L_1$ is factor-closed.
-
-By \cref{periodicity}, there exists an infinite periodic factor-closed language $L_2\subseteq L_1$.
-Let $x$ be any letter in $\Sigma_{L_2}$.
-Note that the word $xx$ is not in $L_2$ since the corresponding colouring $\phi_{xx}$ of $G_8$ has, for example, the anagramish path $a_0,\ldots,a_7$.
-Since $L_2$ is factor-closed, no word in $L_2$ contains $xx$ as a factor, i.e., $xx\not\in F(L_2)$.
-
-Let $\Sigma_3\subset F(L_2)$ contain only those factors of $L_2$ that do not contain $x$.
-Thus, any word in $L_2$ is of the form $w_0xw_1xw_2\cdots xw_rxw_{r+1}$ where each of $w_1,\ldots,w_{r}$ is in $\Sigma_3$ and each of $w_0,w_{r+1}$ is in $\Sigma_3\cup\{\varepsilon\}$.
-Since $L_2$ is periodic, each word in $\Sigma_3$ has bounded length, so $\Sigma_3$ is finite, i.e., $\Sigma_3$ is an alphabet.
-Let $L_3$ be the language consisting of all words $w_1,\ldots,w_r\in\Sigma_3^*$ such that $xw_1xw_2\cdots xw_r\in L_2$.
-Since $L_2$ is infinite and each word in $L_2$ has the form described above, $L_3$ is infinite.
-Since $L_2$ is factor closed, $L_3$ is also factor-closed.
-Therefore, by \cref{periodicity}, there exists an infinite periodic factor-closed language $L\subseteq L_3$ over the alphabet $\Sigma:=\Sigma_{L_3}$.
-
-See \cref{bigexample} for what follows.
-Each word $w:=w_1,\ldots,w_r\in L$ defines a word $w':=xw_1xw_2\cdots xw_{r}\in L_1$.
-In graph colouring terms, $w$ defines an integer $n_w:=4r+\tfrac{1}{2}\sum_{i=1}^r|w_i|$ and an anagram-free colouring $\phi_w:=\phi_{w'}$ of $G_{n_w}$.
-The word $w'$ partitions the columns of $G_{n_w}$ into \emph{blocks}, each of which corresponds to a letter $a$ of $w'$ and whose length $|a|/2$ is a multiple of $4$.
-There are two types of blocks:
-\begin{compactitem}
-    \item There are \emph{boring} blocks $Q_0,\ldots,Q_{r-1}$, each containing $4$ columns of $G_{n_w}$, and each corresponding to an occurrence of $x$ in $w'$.
-    \item There are \emph{colourful blocks} $H_1,\ldots,H_r$, where $H_i$ is the subgraph of $G_{n_w}$ corresponding to $w_i$ and contains $4|w_i|$ columns of $G_{n_w}$.
-\end{compactitem}
-
-\begin{figure}
-    \centering{\includegraphics{figs/bigexample-1}}
-    \caption{A word $w:=dabacbdabc$, its corresponding word $w':=xdxaxbxaxcxbxdxaxbxc$, and its corresponding coloured graph $G_{n_w}$.}
-    \label{bigexample}
-\end{figure}
-
-Since $L$ is periodic, there exists an integer $\ell$ such that every word in $L$ is $\ell$-periodic.
-Since $L$ is infinite, it contains arbitrarily long words.
-These two facts, along with \cref{near_anagram_fourier} immediately imply the following lemma:
-
-\begin{lem}\label{near_anagram_graph}
-    If $\afcn(G_n)\le c$ for each $n\in\N$ then, for each $r_0\in\N$ and $\epsilon>0$, there exists $w\in L$ with even length $|w|\ge 2r_0$ and $\tau(w)\le\epsilon r$.
-\end{lem}
-
-% \begin{proof}
-%     By \cref{block_colouring}, for each $k\in\N$ there exists an $\ell$-periodic word $w'\in\Sigma^k$ such that $\phi_{s'}$ is an anagram-free vertex colouring of $G_{n_{s'}}$.
-Applying \cref{near_anagram_fourier} to $w'$ proves the existence of the desired word $w$.
-% \end{proof}
-
-\cref{near_anagram_graph} shows the existence of a colouring $\phi_w$ of $G_n$ (for arbitrarily large $n$) that is defined by a word $w$ that is $\epsilon$-close to being anagramish.
-This implies that the path $P_0:=a_1,\ldots,a_{n}$ is $4\epsilon$-close to being anagramish under $\phi_w$.
-The last step, done in the next lemma, is to show that there is enough flexibility to modify $P_0$ to obtain a path is anagramish under $\phi_w$.
-
-\begin{lem}\label{anagramish_path}
-    There exists $\epsilon>0$ and $r_0\in\N$ such that, for any $w\in L$ with even length $|w|\ge 2r_0$ and $\tau(w)\le\epsilon r$, the graph $G_{n_w}$ contains a path that is anagramish under $\phi_w$.
-\end{lem}
-
-\begin{proof}
-    Let $w:=w_1,\ldots,w_{2r}$ so that the colouring $\phi_w$ is defined by the word $xw_1xw_2\cdots xw_r$ (and $r\ge r_0$).
-    For each $a\in\Sigma$, define sets $A_a\subseteq\{i\in \{1,\ldots,r-1\}: w_i=a\}$ and $B_a\subseteq\{i\in\{r+1,\ldots,2r-1\}:w_i=a\}$ as follows:
-    \begin{compactenum}
-        \item If $\delta_a(w)=0$ then $A_a=B_a=\emptyset$.
-        \item If $\delta_a(w)>0$ then $|A_a|=2\delta_a(w)=2|B_a|$.
-        \item If $\delta_a(w)<0$ then $|A_a|=\delta_a(w)=\tfrac{1}{2}|B_a|$.
-    \end{compactenum}
-    Let $A:=\bigcup_{a\in\Sigma} A_a$ and $B:=\bigcup_{a\in\Sigma} B_a$.
-    The sets $A_a$ and $B_a$ are chosen so that $A\cup B$ satisfies the following \emph{global independence constraint}:  There is no pair $i,j\in A\cup B$ such that $i-j=1$.
-To see that this is possible, first observe that, because $r$ is not in $A$ or $B$ we need only concern ourselves with pairs where both $i,j\in\{1,\ldots,r-1\}$ or pairs where both $i,j\in\{r+1,\ldots,2r-1\}$.
-Thus, we can choose the elements of $A_a$, for each $a\in\Sigma$ and then independently choose the elements of $B_a$, for each $a\in\Sigma$.
-
-    We show how to choose the elements of $A_a$ for each $a\in\Sigma$.
-The same method works for choosing the elements in $B_a$.
-Observe that, because $w\in L$, $w$ is $\ell$-periodic, so $|\{i\in \{1,\ldots,r-1\}:w_i=a\}|\ge \lfloor(r-1)/\ell\rfloor$ for each $a\in\Sigma$.
-This allows us to greedily choose the elements in $A_a$ for each $a\in\Sigma$.
-At each step we simply avoid choosing $i$ if $i-1$ or $i+1$ have already been chosen in some previous step.
-At any step in the process, at most $\epsilon r$ elements have already been chosen in previous steps and each of these eliminates at most $2$ options.
-Therefore, there will always be an element available to choose, provided that $2\epsilon r < \lfloor(r-1)/\ell\rfloor$.
-In particular, for any $r\ge r_0\ge \ell$, $\epsilon < 1/(4\ell)$ works.
-
-    To ease notation, let $G:=G_{n_w}$.
-We now construct the anagramish path $P$ in a piecewise fashion.
-Refer to \cref{path_construction}.
-For each $i\in\{1,\ldots,2r\}$, let $H_i$ be the subgraph of $G$ induced by the vertices corresponding to $w_i$.
-(The subgraph $H_1,\ldots,H_{2r}$ are referred to above as colourful blocks.)
-    \begin{figure}
-        \centering{
-            \includegraphics{figs/bigexample-1} \\
-            \includegraphics{figs/bigexample-2} \\
-            \includegraphics{figs/bigexample-3} \\
-            \includegraphics{figs/bigexample-4}
-        }
-        \caption{Constructing the anagramish path $P$:
-            (1) pairs of top and bottom paths are matched with zig-zag paths;
-            (2) all remaining colourful blocks receive top paths;
-            (3) all boring blocks receive top, updown, or downup paths.
-        }
-        \label{path_construction}
-    \end{figure}
-    \begin{compactenum}
-        \item For each $a\in\Sigma$ such that $\delta_a(w)>0$, group the elements of $A_a$ into pairs.
-For each pair $(i,j)$, $P$ contains the path through the top row of $H_i$ and the path through the bottom row of $H_j$.
-For each element $i\in B_a$, $P$ contains the zig-zag path with both endpoints in the top row of $H_i$ and that contains every vertex of $H_i$.
-(Note that the zig-zag path begins at the top and ends at the bottom row because $H_i$ contains an even number of columns of $G$.)
-
-        \begin{figure}
-            \centering{
-                \begin{tabular}{ccc}
-                    \includegraphics{figs/paths-1} &
-                    \includegraphics{figs/paths-2} &
-                    \includegraphics{figs/paths-3}
-                \end{tabular}
-            }
-            \caption{Subpaths of $P$ through colourful blocks: A pair of top and bottom paths contribute the same amount as a single zig-zag path.}
-            \label{paths}
-        \end{figure}
-
-        % \todo[inline]{Handle this comment: • fig.
-4.
-Maybe do not use the same color for the bottom and the top, it took me a while to notice that the shapes were different and that this meant that the colors were different (or you could simply add a sentence to clarify this).
-        % PM: I handled this by making the symbols huge}
-
-        \item For each $a\in\Sigma$ such that $\delta_a(w)<0$ we proceed symmetrically to the previous case, but reversing the roles of $A_a$ and $B_a$.
-Specifically, we group the elements of $B_a$ into pairs.
-For each pair $(i,j)$, $P$ contains the path through the top row of $H_i$ and the path through the bottom row of $H_j$.
-For each element $i\in A_a$, $P$ contains the zig-zag path with both endpoints in the top row of $H_i$ and that contains every vertex of $H_i$.
-
-        \item For each $i\in\{1,\ldots,2r\}\setminus\bigcup_{a\in\Sigma}(I_a\cup B_a)$, $P$ contains the top row of $H_i$.
-    \end{compactenum}
-    The rules above define the intersection, $P_i$, of $P$ with each colourful block $H_i$ of $G_{n_s}$.
-If $P_i$ is the path through the bottom (top) row of $H_i$ then we call $H_i$ a \emph{bottom (top) block}.
-If $P_i$ is the zig-zag path that contains every vertex of $H_i$ then we call $H_i$ a \emph{zig-zag block}.
-Note that $\sum_{a\in\Sigma} \delta_a = 0$ and this implies that the number of bottom blocks among $H_1,\ldots,H_{r-1}$ is the same as the number of bottom blocks among $H_{r+1},\ldots,H_{2r}$.
-Indeed, this number is exactly $\beta:=\tfrac{1}{2}\sum_{a\in\Sigma} |\delta_a(w)|=\tfrac{1}{2}\tau(w)$.
-
-    We now define how $P$ behaves for the boring blocks, denoted by  $Q_0,\ldots,Q_{2r-1}$.
-The first boring block $Q_0$ comes immediately before $H_1$.
-Each boring block $Q_j$, for $j\in\{1,\ldots,2r-1\}$ comes immediately after $H_j$ and immediately before $H_{j+1}$.
-In almost every case, $P$ uses the path through the top row of $Q_j$.
-The only exceptions are when $H_j$ or $H_{j+1}$ are bottom blocks.
-Note that, because of the global independence constraint, these two cases are mutually exclusive.
-See \cref{updownup}.
-
-    \begin{figure}
-        \centering{
-            \begin{tabular}{ccc}
-                \includegraphics{figs/crossings-1} &
-                \includegraphics{figs/crossings-2} &
-                \includegraphics{figs/crossings-3}
-            \end{tabular}
-            \caption{The paths taken by $P$ through boring blocks: A top path, a downup path, and an updown path.}
-            \label{updownup}
-        }
-    \end{figure}
-
-    \begin{compactenum}
-        \item When $H_j$ is a bottom block, $P$ uses a path that begins at the bottom row of $Q_j$ but moves immediately to the top row of $Q_j$ and uses the entire path along the top row.
-We call this a \emph{downup} path.
-        \item When $H_{j+1}$ is a bottom block, $P$ uses a path that begins at the top row of $Q_j$ and moves immediately to to the bottom row of $Q_j$ and uses the entire path along the bottom row.
-We call this a \emph{updown} path.
-    \end{compactenum}
-    This completely defines the path $P:=v_1,\ldots,v_{2m}$. All that remains is to argue that the word $\rho:=\phi_w(v_1),\ldots,\phi_w(v_{2m})$ is anagramish.
-
-    Observe that the number of downup paths and the number of updown paths that appear in $Q_0,\ldots,Q_{r-1}$ is exactly the same as the number of bottom blocks among $H_1,\ldots,H_{r-1}$ which is exactly $\beta$.
-Similarly, the number of updown paths and downup paths that appear in $Q_{r+1},\ldots,Q_{2r-1}$ is exactly $\beta$.
-Now every path that is neither downup nor updown uses the top row.
-This implies that the sequence of colours contributed to $\rho$ by  the intersection of $P$ with $Q_0,\ldots,Q_{r-1}$ is a permutation of the sequence of colours contributed to $\rho$ by the intersection of $P$ with $Q_{r+1},\ldots,Q_{2r-1}$.
-These two sequences cancel perfectly.
-
-    Next, by construction, each pair of top and bottom blocks in $H_1,\ldots,H_{r-1}$ contributes exactly the same amount as a single matching zig-zag block in $H_{r+1},\ldots,H_{2r-1}$.
-Specifically, if $x,y\in A_a$, $z\in B_a$, $H_x$ is a top block, $H_y$ is a bottom block and $H_z$ is a zig-zag block, then the contributions of $P_x$ and $P_y$ to $\rho$ cancels out the contribution of $P_z$.
-After doing this cancellation exhaustively, all that remains are top blocks, which also cancel each other perfectly.
-This completes the proof.
-\end{proof}
-
-To be explicit, we finish the proof of \cref{main_vertex} by pointing out the contradiction between \cref{near_anagram_graph} and \cref{anagramish_path}:
-
-\begin{proof}[Proof of \cref{main_vertex}]
-    Assume for the sake of contradiction that there exists some $c\in\N$ such that $\afcn(G_n)\le c$ for each $n\in\N$.
-With this assumption, \cref{near_anagram_graph} shows that for every $\epsilon>0$ and $r_0\in\N$ there exists a word $w\in L$ with $|w|\ge 2r_0$, $\tau(w)\le\epsilon r$, and such that $\phi_w$ is an anagram-free colouring of $G_{n_w}$.
-
-    \cref{anagramish_path} shows that there are fixed values $\epsilon >0$ and $r_0\in \N$ such that, for any word $w\in L$ with $|w|\ge r_0$ and $\tau(w)\le\epsilon r$, the graph $G_{n_w}$ contain a path that is anagramish under $\phi_w$.
-This certainly contradicts the fact that $\phi_w$ is an anagram-free colouring of $G_{n_w}$.
-We therefore conclude that, for every $c\in\N$ there exists an $n\in N$ such that $\afcn(G_n)> c$.
-\end{proof}
-
-% \todo[inline]{It would be nice to say something about the anagram-free chromatic index.
-I'm pretty sure that it's constant for the $2\times n$ grid, but the result is not worth the effort.
-More interesting would be a proof that its super-constant for the $2\times n$ grid with one diagonal in each square.
-Almost everything we did above works except that we get hung up on a parity problem.
-We can't do $2$ versus $1$ cancellation in $w$ but we can do $6$ versus $4$ cancellation.
-Everything works out except when the number of occurrences of some character is odd.
-In that case we get stuck with one uncancelled copy.}
-%
-
-\section{Proof of \cref{near_anagram_fourier}}
-\label{near_anagram_proof}
-
-% \todo[inline]{I suspect this lemma (and probably a better quantitative version) has a proof using Fourier approximation, but I don't know enough about Fourier approximation and my half-assed attempt to learn about it wasn't sufficient.}
-
-All that remains is to prove \cref{near_anagram_fourier}, which we do now.
-
-\begin{proof}[Proof of \cref{near_anagram_fourier}]
-    First observe that, for any $\ell$-periodic word $w$, $|\Sigma_w|\le\ell$.
-    Define an even-length word $t$ to be \emph{$a$-unbalanced} if $\tau_a(t)>\epsilon|t|/\ell$ and \emph{$a$-balanced} otherwise.
-If $t$ is $a$-balanced for each $a\in\Sigma_t$ then $t$ is \emph{balanced}.
-Observe that, if $t$ is balanced then $\tau(t)\le |\Sigma_t|\epsilon|t|/\ell\le \epsilon|t|$.
-A word is \emph{everywhere unbalanced} if it contains no balanced factor of length $r\ge r_0$.
-Our goal therefore is to show that there is an upper bound $n:=n(\ell,\epsilon,r_0)$ on the length of any $\ell$-periodic everywhere unbalanced word.
-
-    Let $h$ be a positive integer (that determines $n$ and whose value will be discussed later), and let $n:=r_02^{h}$.
-Let $w$ be an $\ell$-periodic everywhere unbalanced word of length $n$ over the alphabet $\Sigma$.
-Assume, without loss of generality, that $r_0$ is a multiple of $\ell$.
-
-    Consider the complete binary tree $T$ of height $h$ whose leaves, in order, are length-$r_0$ words whose concatenation is $w$ and for which each internal node is the factor obtained by concatenating the node's left and right child.
-    % See \cref{binary_tree}.
-    Note that for each $v\in V(T)$ and each $a\in\Sigma$, the fact that $w$ is $\ell$-periodic and $r_0$ is multiple of $\ell$ implies that $|v|_{a}\ge |v|/\ell$.
-
-    % \begin{figure}
-    %     \centering{
-    %         \includegraphics[width=\textwidth]{figs/binary-tree-1}
-    %     }
-    %     \caption{Building a binary tree $T$ over the word $w:=4 3 5 3 2\ldots5 4 1 4$.}
-    %     \label{binary_tree}
-    % \end{figure}
-
-  % For each node $v$ of $T$, let $h(v)$ denote the height of $v$'s subtree
-  % and $w(v)=2^{h(v)}$ denote the length of the word $v$. For each
-  % $i\in\Sigma$, let $w_i(v):=n_i(v)/s(v)$.  Note that $0\le w_i(v)\le
-  % 1/2$ and that $\sum_{i\in\Sigma} w_i(v)=1$.  Furthermore, if $v$
-  % has two children $x$ and $y$, then $w_i(v) = (w_i(x)+w_i(y))/2$.
-  % The assumption that $w$ has no balanced factor implies $v$ is
-  % $i$-unbalanced, for some $i\in\Sigma$.  Assign each internal node
-  % $v$ of $T$ the \emph{label} $\ell(v):=\min\{i\in\Sigma: \mbox{$v$
-  % is $i$-unbalanced}\}$.
-
-  % \begin{figure}
-  %   \centering{
-  %      \includegraphics[width=\textwidth]{figs/binary-tree-3}
-  %   }
-  %   \todo[inline]{This is an old figure and needs updating.}
-  %   \caption{The tree $T$ with the vector $(|v|_{1},\ldots,|v|_{5})$ for each
-  %    internal node $v$.  The unbalanced index that defines $\ell(v)$ is highlited for each node.}
-  %   \label{binary_tree_2}
-  % \end{figure}
-
-  For each $a\in\Sigma$, let $S_a:=\{v\in V(T): \text{$v$ is $a$-unbalanced}\}$. Since $w$ is everywhere unbalanced, $\bigcup_{a\in\Sigma} S_a=V(T)$. Therefore,
-  \[
-     (h+1)n = \sum_{v\in V(T)}|v|\le \sum_{a\in\Sigma} \sum_{v\in S_a} |v|
-     \enspace .
-  \]
-  Therefore, there exists some $\alpha\in\Sigma$ such that $\sum_{v\in S_{\alpha}}|v|\ge (h+1)n/|\Sigma| \ge (h+1)n/\ell$.
-At this point we are primarily concerned with appearances of $\alpha$, so let $X:=S_{\alpha}$, and, for each node $v\in V(T)$, let $W(v):=|v|_\alpha$.
-
-  For each non-leaf node $v$ of $T$, let $R(v)$ denote a child of $v$ such that $W(R(v))\le \tfrac12\cdot W(v)$.
-(It is helpful to think of $T$ as being ordered so that each right child $y$ with sibling $x$ has $W(y)\le W(x)$.)  For a non-leaf node $v\in X$ the fact that $v$ is $\alpha$-unbalanced implies that
-  \[  W(R(v)) \le \tfrac{1}{2}\cdot W(v) - \tfrac{\epsilon}{2\ell}\cdot |v|
-      \le (\tfrac12-\tfrac{\epsilon}{2\ell})W(v) \enspace .
-  \]
-
-  From this point on we use the following shorthands. For any $S\subseteq V(T)$, $L(S):=\sum_{v\in S}|v|$,
-  % \todo{$L$ and $R$ are bad notation choices: $L$ for length, $R$ for right child}
-  $W(S):=\sum_{v\in S}W(v)$, and $R(S)=\{R(v):v\in S\}$.  Summarizing, we have a complete binary tree $T$ of height $h$ and
-  $X\subseteq V(T)$ with the following properties:
-  \begin{compactenum}
-    % \item For each $v\in V(T)$, $w(v) \le 1/2$.
-    \item For each $v\in V(T)$,  $W(v)\ge |v|/\ell$.
-    % \item For each internal node $v\in V(T)$ with children $x$ and $y$,
-       % $w(v) = (w(x)+w(y))/2$.
-    \item $L(X) \ge (h+1)n/\ell$.
-    \item For each non-leaf node $v\in X$,
-      $W(R(v)) \le (\tfrac{1}{2}-\tfrac{\epsilon}{2\ell})W(v)$.
-  \end{compactenum}
-  For each $i\in\{0,\ldots,h\}$, let $X_i\subseteq X$ denote the
-  set of nodes $v\in X$ for which the path from the root of $T$ to $v$
-  contains exactly $i$ nodes in $X$, excluding $v$.  See \cref{bigtree}. Observe that, since each node in $X_i$ has an ancestor in $X_{i-1}$,
-  \[  n \ge L(X_0) \ge L(X_1) \ge \cdots\ge L(X_{h}) \enspace . \]
-
-  We will show that there exists an integer $t:=t(\epsilon,\ell,r_0)$ such that, for each $i\in\{0,\ldots,h-t\}$,
-  \begin{equation}
-     L(X_{i+t}) \le (1-(1/2)^{t+1}) L(X_i) \enspace . \label{t}
-  \end{equation}
-  In this way,
-  \begin{align*}
-     (h+1)n/\ell
-        \le L(X) & = \sum_{i=0}^{h} L(X_i) \\
-           &\le \sum_{i=0}^{h} L(X_{t\floor{i/t}})
-             & \text{(since $i\ge t\floor{i/t}$, so $L(X_i)\le L(X_{i\floor{i/t}}$))} \\
-           & = t\cdot\sum_{i=0}^{h/t} L(X_{it})
-             & \text{(for $h$ a multiple of $t$)} \\
-           &\le t\cdot\sum_{i=0}^{\infty} (1-(1/2)^{t+1})^i L(X_0)
-           & \text{(by \cref{t})}\\
-           &\le tn\cdot \sum_{i=0}^{\infty} (1-(1/2)^{t+1})^i
-           & \text{(since $|X_0|\le n$)} \\
-           & = tn2^{t+1} \\
-  \end{align*}
-  which is a contradiction for sufficiently large $h$; in particular, for $h > \ell t2^{t+1}-1$.
-
-  \begin{figure}
-    \centering{
-       \includegraphics[width=\textwidth]{figs/bigtree}
-    }
-    \caption{The partitioning of $X$ into $X_0,\ldots,X_h$. Shaded
-    nodes are in $X$ and all nodes in $X_i$ are shaded with the same
-    colour.   Starting with $A_0=X_0$, the elements of $A_0,\ldots,A_h$
-    are highlighted (in orange).  The elements of $R(A_0),\ldots,R(A_h)$ are also highlighted (in pink).}
-    \label{bigtree}
-  \end{figure}
-
-  It remains to establish \cref{t}, which we do now.
-Fix some $i\in\{0,\ldots,h-1\}$, define $A_0:= X_i$ and, for each $j\ge 1$, define $A_j$ to be the subset of $X_{i+j}$ that are descendants of some node in $R(A_{j-1})$.
-See \cref{bigtree}.
-To upper bound $L(X_{i+t})$ observe that $X_{i+t}$ can be split into two sets $A_0'$ and $B$ defined as follows:  The nodes $A_0'$ do not have an ancestor in $R(A_0)$ and therefore $L(A_0')\le (1/2)L(A_0)$.
-The nodes in $B$ do have an ancestor in $R(A_0)$ and therefore have an ancestor in $A_1$.
-Iterating this argument, we obtain
- \begin{align*}
-      L(X_{i+t})
-         &\le (1/2)L(A_0) + (1/2)L(A_1) + \cdots + (1/2)L(A_{t-1}) + L(A_t) \\
-         &\le (1/2)L(A_0) + (1/4)L(A_0) + \cdots + (1/2)^t L(A_{0}) + L(A_t) \\
-         &  = (1-(1/2)^t)L(A_0) + L(A_t) = (1-(1/2)^t)L(X_i) + L(A_t)  \enspace .
-   \end{align*}
-  So all that remains to establish \ref{t} is to prove that
-  $L(A_t)\le (1/2)^{t+1}L(X_i)$.  To do this, observe that, for each $j\in\{1,\ldots,t\}$,
-  \begin{equation}
-      W(A_j)
-        \le W(R(A_{j-1}))
-        \le (\tfrac12-\tfrac{\epsilon}{2\ell})\cdot W(A_{j-1})
-  \end{equation}
-  which implies
-  \[
-       W(A_t)
-       \le (\tfrac12-\tfrac{\epsilon}{2\ell})^t W(A_0) \le (\tfrac12-\tfrac{\epsilon}{2\ell})^t\cdot L(A_0)
-       =  (\tfrac12-\tfrac{\epsilon}{2\ell})^t\cdot L(X_i)
-  \]
-  Since $w$ is $\ell$-periodic,
-  \[
-        L(A_t)\le \ell\cdot W(A_t) \le \ell\cdot(\tfrac12-\tfrac{\epsilon}{2\ell})^t\cdot L(X_i) \le L(X_i)/2^{t+1}
-  \]
-  for $t := \ceil{\log(2\ell)/\log((1/(1-\tfrac{\epsilon}{\ell}))}$.
-\end{proof}
-
-\section{Reflections}
-\label{reflections}
-
-Although an explicit upper bound on $n:=n(\epsilon,\ell,r_0)$ could be extracted from the proof of \cref{near_anagram_fourier} it would likely be far from tight.
-We suspect that there is a Fourier analytic proof that would give better quantitative bounds.
-We have not pursued this, because we have no idea how to explicitly upper bound $\ell$, for reasons discussed in the next paragraph.
-
-\cref{periodicity} and its proof give absolutely no clues to help find a concrete bound on $\ell$ or to find a minimal set $\Xi$.
-Indeed, for some choices of $L$, doing so can be a difficult problem.
-Consider the example where $|\Sigma|=5$ and $L$ is the language of all anagram free words in $\Sigma^*$.
-It is easy to see that this predicate $L$ is factor-closed and the result of \citet{pleasants:non-repetitive}, published in 1970, shows that this $L$ is infinite.
-The question of whether $|\Xi|=4$ or $|\Xi|=5$ is then the question of determining whether there exist arbitrarily long anagram-free words on an alphabet of size $4$.
-This was the open problem posed by \citet{erdos:some} in 1961 and again by \citet{brown:is} in 1971 and not resolved until 1992 when \citet{keranen:abelian,keranen:powerful} showed that the answer, in this case, is that $|\Xi|=4$.
-However, if this were not the case, then determining $\ell$ would be the question of determining the length of the longest anagram-free word over an alphabet of size $4$.
-
-Our proof uses \cref{periodicity} twice and each uses a language $L$ that is considerably more complicated than the language of anagram-free words.
-It seems unlikely that we will obtain concrete upper bounds on $\ell$ as a function of $c$ except, possibly, through the use of computer search.
-The resulting value $\ell$ is used in the application of \cref{near_anagram_fourier} and also within the proof of \cref{anagramish_path}.
-
-\section*{Acknowledgement}
-
-We would like to thank an anonymous referee for pointing out existing results and terminology from the field of combinatorics on words that has helped streamline the presentation in \cref{near_anagram_statement,vertex_colourings}.
-
-\bibliographystyle{plainurlnat}
-\bibliography{af2t}
-
-\end{document}

### 2021-09-28 lin-vs-cen.tex
Working

@@
 
-\title{\MakeUppercase{Linear versus Centered Colouring Numbers}\thanks{This research was partly funded by NSERC.}}
+\title{\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}}
 \author{Prosenjit Bose, Vida Dujmović, Mehrnoosh Javarsineh, and Pat Morin}
@@
 \begin{abstract}
-    These are some notes on the relationships betwee linear colouring numbers and centered colouring numbers.
+    These are some notes on the relationships betwee linear colouring numbers and centred colouring numbers.
 \end{abstract}
@@
 A \emph{$k$-colouring} of $G$ is any function $\varphi:V(G)\to S$ where $S$ is a set of size $k$.
-A vertex $v$ of $G$ is a \emph{center} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.
-The colouring $\varphi$ is \emph{centered} if every connected sugraph $H\subseteq G$ has a center.
-The colouring $\varphi$ is \emph{linear} if every simple path $P\subseteq G$ has a center.
+A vertex $v$ of $G$ is a \emph{centre} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.
+A colouring $\varphi$ of $G$ is \emph{centred} if every connected subgraph of $G$ has a centre with respect to $\varphi$.
+A colouring $\varphi$ of $G$ is \emph{linear} if every simple path in $G$ has a centre with respect to $\varphi$.
 
-The \emph{centered chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centered $c$-colouring.
+The \emph{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.
 The \emph{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.
-Since every path in $G$ is a connected subgraph of $G$, every centered colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.
+Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.
 
@@
 
-The $k\times k$ grid graph $G_k$ is the graph with vertex set $V(G_k):=\{1,\ldots,k\}^2$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only $|v-x|+|w-y|=1$.
-A graph $H$ is a $k\times k$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $G$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_k$.
-Since $G_k$ has maximum degree $4$ any graph $G$ that contains a $G_k$-minor contains a subgraph $H$ that is a $k\times k$ pseudogrid.
+For positive integers $a$ and $b$, the $a\times b$ \emph{grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only $|v-x|+|w-y|=1$.
+For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.
+
+A graph $H$ is an $a\times b$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $H$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_{a\times b}$.
+Since $G_{a\times b}$ has maximum degree $4$ any graph $G$ that contains a $G_{a\times b}$-minor contains a subgraph $H$ that is a $a\times b$ pseudogrid.\footnote{This comes from the fact that any tree $T$ with at most $4$ leaves contains at most two vertices of degree greater than $2$.
+Contracting the path between these two vertices (if they exist) produces a tree with exactly one vertex of degree greater than $2$.}
 
@@
 \begin{proof}
-  Let $\varphi:V(H)\to\{1,\ldots,C\}$ be an arbitrarily $C$-colouring of some $k\times k$ pseudogrid $H$ for some $C\le \epsilon k$, where $\epsilon$ is a small positive constant.
-We will show that $H$ contains a path $P$ which has no center with respect to $c$.
-First we introduce some notations:
+  This proof uses two large integer constants $\alpha$ and $\beta$ and a small positive real constant $\epsilon$.
+The relationships between these constants are that $\beta = \Theta(\alpha^{1/4})$ and $\epsilon=\Theta(1/\alpha)$.
+Let $\varphi:V(H)\to\{1,\ldots,C\}$ be an arbitrarily $C$-colouring of some $k\times k$ pseudogrid $H$ for some $C\le \epsilon k$.
+We will show that $H$ contains a path $P$ which has no centre with respect to $\varphi$.
 
+  First we introduce some notations for mapping edges and vertices of the grid $G_k$ to their corresponding paths in the pseudogrid $H$.
 For each vertex $v$ of $G_k$ there is a path $P_v$ in $\mathcal{P}$ (possibly consisting of one vertex) such that the vertex $v$ appears in $H'$ as a result of contracting $P_v$ in $H$.
 In this way, each vertex $v$ of $G_k$ is associated with a non-empty \emph{colour set} $\phi(v):=\{\varphi(w):w\in V(P_v)\}$.
-Similarly, for each edge $e$ of $G_k$ there is a path $P_e$ in $H'$ that comes from subdividing $e$.
-Thus, each edge $e$ of $G_k$ is associated with (possibly empty) \emph{colour set} $\phi(e):=\{\varphi(w):w\in P^-_e\}$, where $P^-_e$ denotes the (possibly empty) set of internal vertices on the path $P_e$.
+Similarly, for each edge $e$ of $G_k$ there is a path $P_e$ in $H'$ that comes from subdividing $e$ (possibly zero times).
+Thus, each edge $e$ of $G_k$ is associated with a (possibly empty) \emph{colour set} $\phi(e):=\{\varphi(w):w\in P^-_e\}$, where $P^-_e$ denotes the (possibly empty) set of internal vertices on the path $P_e$.
 
-  For each colour $\alpha\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $\alpha$ as the number of vertices and edges of $G_k$ that include $\alpha$ in their colour set.
-Fix some large integer constant $\alpha$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity at most $\alpha$ and \emph{frequent} otherwise.
+  For each colour $c\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $c$ as the number of vertices and edges of $G_k$ that include $c$ in their colour set.
+Fix some large integer constant $\alpha$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity less than $\alpha$ and \emph{frequent} otherwise.
 
   \begin{clm}\label{all_frequent}
-    $H$ contains a $k'\times k'$ pseudogrid in which every colour is frequent, for some $k'\ge (1-\epsilon \alpha)k$.
+    $H$ contains a $k'\times k'$ pseudogrid $H'$ in which every colour in $\{\varphi(v):v\in V(H')\}$ is frequent, for some $k'\ge (1-\epsilon \alpha)k$.
   \end{clm}
@@
   \begin{proof}
-    If there exists some infrequent colour $c\in\{1,\ldots,C\}$, we identify every row and column of $G_k$ that contains an edge or vertex with $c$ in its colour set.
-We then \emph{remove} each of these (at most $A$) rows and columns from $H$ by removing all edges of $H$ in $P_e$ for each $e$ in the relevant row (column) of $G_k$ and then discarding any isolated vertices.
-After doing this for each row and column containing $c$, the graph $H$ may be disconnected.
-We discard any isolated vertices and any path $P_{v}$ that is disconnected from the rest of $H$.
+    This proof uses the notion of \emph{removing} a row or column of a pseudogrid.
+Removing a row is defined as follows (removing a column is similar): For each edge $e$ of $G_k$ in the relevant row, remove all edges of $P_e$ from $H$.
+Next, remove any isolated vertices from $H$ and finally, repeatedly remove degree-$1$ vertices from $H$ until none remain.
+The result is pseudogrid with one less row.
 
-    The remaining graph contains a $k'\times k'$ pseudogrid for some $k' \ge k-A$ and contains no vertex of colour $c$.
-We then repeat this (at most $c$ times) until the resulting graph has no infrequent colour.
-This leaves a graph that contains a $k'\times k'$ pseudogrid for some $k'\ge k-C\alpha\ge(1-\epsilon \alpha)k$ for which every colour is frequent.
+    While there exists some infrequent colour $c\in\{1,\ldots,C\}$, find a row or column of $G_k$ that contains an edge or vertex with $c$ in its colour set and remove that row or column.
+Doing this exhaustively results in the removal of at most $(\alpha-1) C \le \epsilon\alpha k$ rows and at most $\epsilon\alpha k$ columns of $H$.
+This leaves a pseudogrid $H'$ that contains a $k'\times k'$ pseudogrid for some $k'\ge (1-\epsilon \alpha)k$ for which every colour that appears colours at least one vertex of $H'$ is frequent.
   \end{proof}
 
-  From this point onward, we assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$.
-Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair of vertices $v_c,w_c$ in $H$ whose colour is $c$.
-These pairs are carefully chosen so that we will be able to construct a simple path $P$ in $H$ that contains $\bigcup_{c\in\{1,\ldots,C\}} \{v_c,w_c\}$.
+  Because of \cref{all_frequent}, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$.
+Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is an edge or vertex of $G_k$ that contains $c$ in its colour set.
+These pairs are carefully chosen so that we will be able to construct a simple path $P_k$ in $G_k$ that contains $\bigcup_{c\in\{1,\ldots,C\}} \{x_c,y_c\}$ and has some additional nice properties.
+Using these additional properties, we will then show how to convert $P_k$ into a simple path $P$ in $H$ that contains at least two vertices of each colour.
 
-  \begin{compactenum}[{Stage} 1.]
-    \item We begin this stage by defining each vertex and edge of $G_k$ as being initially \emph{unmarked}.
+  \begin{compactenum}[{Stage} 1:]
+    \item We begin this stage by defining each vertex and edge of $G_k$ as being \emph{unmarked}.
 Fix some integer constant $\beta>0$.
 We iterate over each $c\in\{1,\ldots,C\}$ in turn.
-If $G_k$ contain a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is an unmarked vertex or unmarked edge of $G_k$ with $c$ in its colour set and the distance in $G_k$ between $x_c$ and $y_c$ is at least $B$ then Stage~1 \emph{succeeds} for $c$.
-In this case, we take $v_c$ to be any vertex in $P_{x_c}$ of colour $c$ and $w_c$ to be any vertex in $P_{y_c}$ of colour $c$.
-We then \emph{mark} every vertex and edge $z$ of $G_k$ such that $\dist_{G_k}(z, \{x,y\})\le \beta$.
+If $G_k$ contain a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is an unmarked vertex or unmarked edge of $G_k$ with $c$ in its colour set and the distance in $G_k$ between $x_c$ and $y_c$ is at least $B$ then Stage~1 \emph{succeeds} for $c$.\todo{Define distance, including between sets}  We then \emph{mark} each vertex and edge $z$ of $G_k$ such that $\dist_{G_k}(z, \{x,y\})\le \beta$.
 
-    If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $v_c$ and $w_c$ undefined and say that Stage~1 \emph{fails} for colour $c$.
-Note that Stage~1 can only fail for $c$ if $G_k$ contains few than $B^2$ unmarked vertices/edges with $c$ in their colour sets.
+    If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $x_c$ and $y_c$ undefined for now and say that Stage~1 \emph{fails} for colour $c$.
+Note that Stage~1 can only fail for $c$ if $G_k$ contains fewer than $3(\beta/2+1)^2$\todo{check precisely} unmarked vertices/edges with $c$ in their colour sets.\footnote{$G_{a\times a}$ contains fewer than $3a^2$ vertices and edges and contains pairs of vertices/edges at distance $2(a-1)$.}
 
-    \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeded and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 failed.
+    \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeds and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 fails.
 We construct a bipartite graph $I$ with vertex parts $A:=\overline{S}$ and $B:=\bigcup_{c\in S}\{x_c,y_c\}$.
@@
 Since each colour $c\in A$ is frequent, this implies that $\deg_I(c)\ge \alpha/8\beta^2$.
-Similarly, for each $y\in B$,  $\deg_{I}(y) \le 8\beta^12$.
+Similarly, for each $y\in B$,  $\deg_{I}(y) \le 8\beta^2$.
 This implies that, for each $A_0\subseteq A$, $|N_I(A_0)|\ge \alpha/64\beta^4$.
@@
 Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $s(y_c)=q$ and $c\in\phi(y_c)$.
-This defines the pairs $\{x_c,y_c\}$ for each element in $\overline{S}$.
+In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
   \end{compactenum}
 
-  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of $2c$ vertices and edges of $G_k$ where each $c\in \phi(x_c)\cap \phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
+  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \phi(x_c)\cap \phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
 Ultimately, we want to use these to find a path in $H$ that contains at least two vertices of each colour.
@@
 
-  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$.
+  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.
 
@@
 Therefore, $R'$ contains no elements in $R_1$.
-Therefore, $S:=\{s(x):x\in R'\}$ is a five-element set whose pairwise distances are all at most $\beta$.
-But this implies that two elements in $R'$ have distance less than $\beta$.
-This is a contradiction because all points in $R'$ are in $R_1$ and all pairs of points in $R_1$ have distance at least $\beta$.\todo{Something is off by a factor of 2.}
+Therefore, $R'':=\{s(x):x\in R'\}$ is a five-element set whose pairwise distances are all at most $5\beta/2$ and whose minimum distance is at least $\beta$ (Contradiction?)\todo{Finish}.
   \end{proof}

### 2021-09-30 lin-vs-cen.tex
More progress

@@
 
-For positive integers $a$ and $b$, the $a\times b$ \emph{grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only $|v-x|+|w-y|=1$.
+For positive integers $a$ and $b$, the $a\times b$ \emph{grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
 For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.
+An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.
+The \emph{vertex boundary} of $G_{a\times b}$ is $\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.
+The \emph{edge boundary} of $G_{a,b}$ consists of all edges incident to at least one vertex in the vertex boundary of $G_{a,b}$.
+The \emph{boundary} of $G_{a\times b}$ is the union of the edge and vertex boundaries.
+
+\begin{lem}
+  Let $I,L$ be disjoint subsets of $V(G_{10})$, let $Q$ be a subset of $E(G_{10})$ and let $X:=I\cup L\cup Q$ have the following properties:
+  \begin{compactenum}[(Pr1)]\setcounter{enumi}{0}
+    \item $|X|\le 4$.
+    \item $Q$ does not contain the edges of a $4$-cycle.
+    \item No vertex of $G_k$ is incident to three or more edges in $Q$.
+    \item No vertex in $I\cup L$ is incident to two or more edges in $Q$.
+    \item $X$ contains no boundary vertices or edges.
+  \end{compactenum}
+  Then for any pair of boundary vertices $s$ and $t$, $G_{10}$ contains a path $P$ that 
+  \begin{compactenum}
+    \item begins at $s$ and ends at $t$
+    \item contains all elements of $X$;
+    \item goes straight through each vertex in $I$; and
+    \item turns at each vertex in $L$.\todo{Define turns and goes straight}
+    % \item contains no boundary vertices or edges except $s$ and $t$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  I have no idea how to prove this.  Computer search?.  
+\end{proof}
+
+A \emph{grid square} $S_{x,y,\ell}$ of side-length $\ell$ is a point set of the form $\{x,\ldots,x+\ell-1\}\times\{y,\ldots,y+\ell-1\}$ for some $x,y\in\Z^2$.
+The \emph{interior} of $S_{x,y,\ell}$ is the grid square $S_{x+1,y+1,\ell-2}$
+
+\begin{lem}\label{four_points}
+  Let $P$ be a set of at most $4$ points in $\Z^2$.
+Then there exists a set $S$ of disjoint grid squares, each of side-length $10$ such that each point of $P$ is contained in the interior of some square in $S$.
+\end{lem}
+
+\begin{proof}
+  Let $P:=\{p_1,\ldots,p_t\}$ where $p_i:=(x_i,y_i)$.
+Let $X:=\bigcup_{i=1}^t \{x_i\bmod 10,(x_{i+1})\bmod 10\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 10,(y_i+1)\bmod 10\}$.
+Then $|X|\le 2t\le 8$ and $|Y|\le 2t\le 8$.
+In particular, there exists some $x\in\{0,\ldots,9\}\setminus X$ and some $y\in\{0,\ldots,9\}\setminus Y$.
+We use squares of the form $\{x+10r,\ldots,x+10r-1\}\times\{y+10r\ldots,y+10r-1\}$ for each $r\in\Z$.
+\end{proof}
+
+\begin{lem}
+  Let $P$ be a set of points in $\Z^2$ such that no square of side-length $80$ contains more than 4 points of $P$.
+Then there exists a set $S$ of disjoint grid squares, each of side length $10$ and such that each point of $P$ is in the interior of some square in $S$.
+\end{lem}
+
+\begin{proof}
+  Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|p-q\|_\infty \le 20$.
+Observe that each component of $H$ contains at most $4$ points of $P$.
+For each component $C$ of $H$, apply \cref{four_points} to $V(C)$.
+This produces a set of disjoint grid squares of side length $10$ that each contain some point of $V(C)$ in their interior.
+Any two squares $s_1$ and $s_2$ obtained from different components of $H$ are disjoint because $s_1$ contains a point $p_1\in P$ and $s_2$ contains a point $p_2\in P$ such that $\|p_1-p_2\|\infty > 20$, but the squares $s_1$ and $s_2$ each have side length $10$.\todo{Could be tidier here.}
+\end{proof}
+
+\todo[inline]{Adjust the numbers in the point covering lemmas.}
+
+\begin{lem}
+  Let $k$ be a positive integer, let $I,L$ be disjoint subsets of $V(G_{k})$, let $Q$ be a subset of $E(G_{k})$ and let $X:=I\cup L\cup Q$ have the following properties:
+  \begin{compactenum}[(Pr1)]\setcounter{enumi}{0}
+    \item Each $80\times 80$ subgrid of $G_k$ contains at most $4$ elements of $X$.
+    \item $Q$ does not contain all the edges of a $4$-cycle.
+    % \item $X$ contains no boundary vertices or edges.
+    \item No vertex of $G_k$ is incident to three or more edges in $Q$.
+    \item No vertex in $I\cup L$ is incident to two or more edges in $Q$.
+  \end{compactenum}
+  Then $G_k$ contains a path $P$ that
+  \begin{compactenum}
+    \item contains all elements of $X$;
+    \item goes straight through each vertex in $I$; and
+    \item turns at each vertex in $L$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  Start with the snakelike path that visits every $10$th row of $G_k$.
+Cover the elements of $X$ with disjoint squares of side length $10$.
+Each such square is intersected by exactly one row of the snakelike path.
+Use \cref{ten_by_ten} to locally deform the path to pick up the objects in any square that it intersects.
+\end{proof}
 
@@
 \begin{proof}
-  This proof uses two large integer constants $\alpha$ and $\beta$ and a small positive real constant $\epsilon$.
+  This proof uses two large integer constants $\alpha$ and $\beta$ and two small positive real constants $\epsilon$ and $\delta$.
 The relationships between these constants are that $\beta = \Theta(\alpha^{1/4})$ and $\epsilon=\Theta(1/\alpha)$.
@@
     If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $x_c$ and $y_c$ undefined for now and say that Stage~1 \emph{fails} for colour $c$.
-Note that Stage~1 can only fail for $c$ if $G_k$ contains fewer than $3(\beta/2+1)^2$\todo{check precisely} unmarked vertices/edges with $c$ in their colour sets.\footnote{$G_{a\times a}$ contains fewer than $3a^2$ vertices and edges and contains pairs of vertices/edges at distance $2(a-1)$.}
+Note that Stage~1 can only fail for $c$ if $G_k$ contains fewer than $3(\beta/2+1)^2$\todo{check precisely} unmarked vertices/edges with $c$ in their colour sets.\footnote{$G_{a\times a}$ contains fewer than $3a^2$ vertices and edges and contains pairs of vertices/edges at distance $2(a-1)$.}  In particular, for $\alpha > 3(\beta/2+1)^2$, Stage~1 succeeds for $c=1$.
 
     \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeds and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 fails.
-We construct a bipartite graph $I$ with vertex parts $A:=\overline{S}$ and $B:=\bigcup_{c\in S}\{x_c,y_c\}$.
-For each marked vertex or edge $p$ of $G_k$, let $s(p)$ be a vertex $q\in B$ that minimizes $\dist_{G_k}(p,q)$.\todo{Gotta be really precise here to avoid degree 3 issue.}  The edge set of $I$ contains an edge $xy$ precisely when there is a marked edge or vertex $z$ of $G_k$ such that $x\in\phi(z)$ and $s(z)=y$.
+Let $R_1:=\bigcup_{c\in S}\{x_c,y_c\}$ be the set of vertices and edges chosen in $R_1$.
 
-    First observe that, for any node in $y\in B$, the number of $x\in A$ such that $\dist_{G_k}(x,y) \le \beta$ is at most $8\beta^2$.
-Therefore, the number of marked vertices/edges $x$ of $G_k$ with $s(x)=y$ is at most $8\beta^2$.
-Since each colour $c\in A$ is frequent, this implies that $\deg_I(c)\ge \alpha/8\beta^2$.
-Similarly, for each $y\in B$,  $\deg_{I}(y) \le 8\beta^2$.
-This implies that, for each $A_0\subseteq A$, $|N_I(A_0)|\ge \alpha/64\beta^4$.
-In particular, for any $\alpha$ and $\beta$ with $\alpha > 128\beta^4$,  $|N_I(A_0)|\ge 2|A_0|$.
+    As mentioned above, for sufficiently large $\alpha$ (relative to $\beta$), the set $S$ is non-empty and therefore $R_1$ is non-empty.
+In the next paragraph, we will define a function $\nu$ that maps each marked vertex or edge of $G_k$ to some element of $R_1$.
+We use to this to construct a bipartite graph $I$ with vertex parts $A:=\overline{S}$ and $R_1$.
+The edge set of $I$ contains an edge $cx$ precisely when there is a marked edge or vertex $z$ of $G_k$ such that $c\in\phi(z)$ and $\nu(z)=x$.
+
+    We now define the mapping $\nu$.
+Fix some total order on $R_1$ by assigning each element of $R_1$ a distinct rank in $\{1,\ldots,|R_1|\}$.
+For each marked element $z\in V(G_k)\cup E(G_k)$, let $\gamma(z):=\min\{\dist_{G_k}(z,x):x\in R_1\}$, let $\Gamma(z):=\{x\in R_1:\dist_{G_k}(z,x) \le (1+\delta)\gamma(z)$, and let $\nu(z)$ be the element in $\Gamma(z)$ of minimum rank.
+In words, $\gamma(z)$ is the distance to $z$'s nearest neighbour in $R_1$, $\Gamma(z)$ is the set of all approximate nearest neighbours of $z$ in $R_1$ and $\nu(z)$ is the approximate nearest-neighbour of $z$ with smallest rank.
+
+    Note that for any marked element $z$, there exists $x\in R_1$ such that $\dist_{G_k}(z,x)\le\beta$.
+Therefore $\gamma(z)\le\beta$.
+Therefore, for any $x\in R_1$, the set $\nu{^-1}(x)$ only contains marked elements $z$ for which $\dist_{G_k}(x,z)\le (1+\delta)\beta$  Thus, $|\nu^{-1}(x)|\le \zeta\beta^2$ for some $\zeta$ depending only on $\delta$.
+This has two implications:
+    \begin{compactenum}
+      \item For each $x\in R_1$,  $\deg_{I}(x) \le \zeta\beta^2$.
+      \item Since each colour $c\in A$ is frequent, $\deg_I(c)\ge \alpha/\zeta beta^2$.  
+    \end{compactenum}
+    These two conditions imply that, for each $A_0\subseteq A$, $|N_I(A_0)|\ge \alpha/\zeta^2\beta^4$.
+In particular, for any $\alpha$ and $\beta$ with $\alpha > 2\zeta^2\beta^4$,  $|N_I(A_0)|\ge 2|A_0|$.
 Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
@@
   Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.
+The following claim follows immediately from the algorithm used to select pairs in Stage~1.
+  \begin{clm}\label{r1_far}
+      For each distinct pair $z,z'\in R_1$, $\dist_{G_k}(z,z')\ge \beta$.
+  \end{clm}
+
+  The next claim states that any element $z'\in R_2$ is quite far from every element of $R_1$ except possibly its nearest neighbour in $R_1$.
 
   \begin{clm}\label{far_from_home}
-    For each $z\in R_1$ and $z'\in R_2$ with $s(z')\neq z$, $\dist_{G_k}(z,z')\ge \beta /2$.
+    For each $x\in R_1$ and $y\in R_2$ with $\nu(y)\neq x$, $\dist_{G_k}(x,y)\ge \beta /(2+\delta)$.
   \end{clm}
@@
   \begin{proof}
-    Since $s(z')\neq z$, there exists some other $z''\in R_1\setminus\{z\}$ with $s(z')=s(z'')$.
-Since $z,z''\in R_1$ each of $z$ and $z''$ were selected in Stage~1.
-The rules for selecting vertices in Stage~1 guarantee that $\dist_{G_k}(z,z'')\ge \beta$.
-By the triangle inequality $\dist(z, z')+\dist(z',z'') \ge \dist(z,z'')\ge \beta$.
-Since $s(z')=z''$, $\dist_{G_k}(z', z'') \le \dist_{G_k}(z, z')$, so $\dist_{G_k}(z,z') \ge \dist(z',z'')/2 \ge \beta/2$.
+    Since $\nu(y)\neq x$, there exists some other $z\in R_1\setminus\{y\}$ with $\nu(y)=\nu(z)$.
+Since $x,z\in R_1$, \cref{r1_far} implies that $\dist_{G_k}(x,z)\ge \beta$.
+Since $\nu(y)=z$, $\dist_{G_k}(y, z) \le (1+\delta)\gamma(y)\le (1+\delta)\dist_{G_k}(y, x)$.
+By the triangle inequality
+    \[ \beta \le \dist_{G_k}(x,z) 
+        \le \dist_{G_k}(x,y)+\dist_{G_k}(y,z)
+        % \le \dist_{G_k}(x,y) + (1+\delta)\dist_{G_k}(x,y) 
+        \le (2+\delta)\dist_{G_k}(x,y) \enspace . \qedhere
+    \]
+  \end{proof}
+
+  \begin{clm}
+    Any subset of $R$ containing containing $4$ elements contains a pair $(x,y)$ such that $\dist_{G_k}(x,y)\ge \psi\beta$, for some $\psi>0$ that depends only on $\delta$.
+  \end{clm}
+
+  \begin{proof}
+    Let $R':=\{x_1,x_2,x_3,x_4\}$ be a $4$-element subset of $R$.
+If $|R'\cap R_1|\ge 2$ because say $x_1,x_2\in R_1$ then $\dist_{G_k}(x_1,x_2)\ge \beta$, by \cref{r1_far}.
+If $|R'\cap R_1=1$ because say $x_1\in R_1$ then $\nu(x_i)= x_1$ for at most one $i\in\{2,3,4\}$.
+For each of the two $j\in\{2,3,4\}\setminus\{i\}$, $\dist_{G_k}(x_1,x_j)\ge \beta/(2+\delta)$, by \cref{far_from_home}.
+
+    The only possibility that remains is that $R'\subseteq R_2$.
+Let the elements of $R'$ be labelled so that the rank of $\nu(x_i)$ is less than the rank of $\nu(x_j)$ for each $i<j$.
+Suppose that $\dist_{G_k}(x_i,x_j)\le \psi\beta$ for each $i,j\in\{1,2,3\}$, otherwise there is nothing to prove.
+
+    First observe that, for any $i\in\{1,2,3,4\}$,  $\dist_{G_k}(x_i,\nu(x_i)) \ge (\tfrac{1}{2+\delta}-\psi)\beta$ since, otherwise 
+    \[  \dist_{G_k}(x_{i+1},\nu(x_i))\le \dist_{G_k}(x_{i+1},x_i) + 
+    \dist_{G_k}(x_i,\nu(x_i)) < \beta/(2+\delta) \enspace ,
+    \]
+    for any $j\in\{1,2,3,4\}\setminus\{i\}$, 
+    contradicting \cref{far_from_home}.\todo{Make this a separate lemma?}
+
   \end{proof}
@@
 
+  \begin{clm}
+    There is no triple $(v,e_1,e_2)\in R$ where $v$ is a vertex of $G_k$ and $e_1$ and $e_2$ are each edges of $G_k$ incident on $v$.
+  \end{clm}
+
+  \todo[inline]{The following claim is not true.
+It's easy to come up with an example in which a small square contains $8$ different elements.
+We're probably better off adjusting the boundaries of Voronoi cells so that we enforce a minimum distance (say at least $\beta/10$) between Voronoi vertices.}
   \begin{clm}

### 2021-09-30 lin-vs-cen.tex
really getting somewhere now

@@
 \usepackage{todonotes}
+\usepackage{etoolbox}
 
@@
 
-\newcommand{\coloured}[2]{{\color{#1}{#2}}}
-\newenvironment{colourblock}[1]{\color{#1}}{}
-
-\newcommand{\condref}[1]{(C\ref{#1})}
-
 % Taken from
@@
 
-\DeclareMathOperator{\diam}{diam}
 \DeclareMathOperator{\tw}{tw}
-\DeclareMathOperator{\stw}{stw}
-\DeclareMathOperator{\ltw}{ltw}
-\DeclareMathOperator{\pw}{pw}
-\DeclareMathOperator{\lpw}{lpw}
-\DeclareMathOperator{\lhptw}{lhp-tw}
-\DeclareMathOperator{\lhppw}{lhp-pw}
+\DeclareMathOperator{\chicen}{\chi_{\mathrm{cen}}}
+\DeclareMathOperator{\chilin}{\chi_{\mathrm{lin}}}
+\DeclareMathOperator{\dist}{dist}
 
-\DeclareMathOperator{\x}{x}
-\DeclareMathOperator{\depth}{d}
-\DeclareMathOperator{\sh}{cbt}
-\DeclareMathOperator{\cbt}{cbt}
-\DeclareMathOperator{\sgn}{sgn}
-\DeclareMathOperator{\dc}{dc}
-
-\DeclareMathOperator{\afci}{\overline{\chi}_\pi}
-\DeclareMathOperator{\afcn}{\dot{\chi}_\pi}
-
-\newcommand{\ellt}{{\lfloor\ell/2\rfloor}}
+\newrobustcmd{\onesub}{\mathord{\includegraphics{figs/one-sub}}}
+\newrobustcmd{\leftup}{\mathord{\includegraphics{figs/left-up}}}
 
@@
 
-% \DeclareMathOperator{\ddiv}{div}
-% \DeclareMathOperator{\hist}{p}
-\DeclareMathOperator{\dist}{dist}
-
-\newcommand{\colored}[2]{{\color{#1}#2}}
-
-\usepackage{tabularx}
-
-\DeclareMathOperator{\ci}{\overline{\pi}}
-
-\DeclareMathOperator{\chicen}{\chi_{\mathrm{cen}}}
-\DeclareMathOperator{\chilin}{\chi_{\mathrm{lin}}}
-
 \begin{document}
 
-% \begin{titlepage}
 \maketitle
@@
 \end{abstract}
-% \end{titlepage}
-
-% \pagenumbering{roman}
-% \tableofcontents
-%
-% \newpage
-% \pagenumbering{arabic}
 
@@
 
-\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun-obrien-general}
+\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general}
   For any graph $G$, $\chicen(G)\le \chilin^{190}(G)\cdot\log^{O(1)}(\chilin(G))$.
@@
 
-Their proof of \cref{kun-obrien-general} has three steps:
+Their proof of \cref{kun_obrien_general} has three steps:
 \begin{enumerate}
@@
 
-These notes are an attempt to improve the bound in \cref{kun-obrien-general} in the general case as well as for some special classes of graphs.
+These notes are an attempt to improve the bound in \cref{kun_obrien_general} in the general case as well as for some special classes of graphs.
 
 As a first step, we attempt to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.
-This, by itself, reduces the exponent in \cref{kun.obrien.general} from $190$ to $95$.
+This, by itself, reduces the exponent in \cref{kun_obrien_general} from $190$ to $95$.
 Next, we observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.
@@
 
-A \emph{subdivision} $G'$ of a graph $G$ is any graph that can be obtained from $G$ by repeatedly replacing some edge edge $uw$ of $G$ with a path $uvw$, where $v\not\in V(G)$ is a newly introduced degree-$2$ vertex.
-Alternatively, $G'$ is obtained from $G$ by replacing some of the edges of $G$ with paths whose internal vertices have degree $2$.
+A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.
+Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.
+A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.
+For a graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.
 
-For positive integers $a$ and $b$, the $a\times b$ \emph{grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
-For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.
-An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.
+\subsection{Subdivided Grids}
+
+For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
+For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
+We define the \emph{boundary} of $G_{a\times b}$ as $\partial G_{a\times b}:=\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.
+We say that a path $P:=v_0,v_1,\ldots,v_\ell$ in $G_{a\times b}$ \emph{goes straight} at a vertex $v_i$, $i\in\{1,\ldots,k-1\}$ if $v_{i-1}$, $v_i$, and $v_{i+1}$ are contained in a single row or column.
+Otherwise, we say that $P$ \emph{turns} at $v_i$.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{cc}
+      \includegraphics{figs/grid-1} &
+      \includegraphics{figs/grid-2}
+    \end{tabular}
+  \end{center}
+  \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}
+\end{figure}
+
+% An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.
 The \emph{vertex boundary} of $G_{a\times b}$ is $\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.
@@
 
-\begin{lem}
-  Let $I,L$ be disjoint subsets of $V(G_{10})$, let $Q$ be a subset of $E(G_{10})$ and let $X:=I\cup L\cup Q$ have the following properties:
-  \begin{compactenum}[(Pr1)]\setcounter{enumi}{0}
-    \item $|X|\le 4$.
-    \item $Q$ does not contain the edges of a $4$-cycle.
-    \item No vertex of $G_k$ is incident to three or more edges in $Q$.
-    \item No vertex in $I\cup L$ is incident to two or more edges in $Q$.
-    \item $X$ contains no boundary vertices or edges.
+For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.
+We will work extensively with the $1$-subdivision $G^{\onesub}_k$ of $G_k$.
+In this case, we define $\partial G^{\onesub}_k:=\{v\in V(G^{\onesub}_k):\dist_{G^{\onesub}_k}(v,\partial G_k)\le 1\}$.
+That is, the boundary of $G^{\onesub}_k$ includes the vertex boundary of $G_k$ as well as any degree-$2$ vertex of $G^{\onesub}_k$ that is adjacent to at least one boundary vertex of $G_k$.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{figs/grid-3}
+  \end{center}
+  \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
+\end{figure}
+
+\begin{lem}\label{seven_by_seven}
+  Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_k$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
+  \begin{compactenum}[(i)]\setcounter{enumi}{0}
+    \item Each vertex is $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
+    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
+    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
   \end{compactenum}
   Then for any pair of boundary vertices $s$ and $t$, $G_{10}$ contains a path $P$ that 
-  \begin{compactenum}
-    \item begins at $s$ and ends at $t$
-    \item contains all elements of $X$;
+  \begin{compactenum}[(a)]
+    \item begins at $s$ and ends at $t$;
+    \item contains every vertex in $I\cup L\cup Q$;
     \item goes straight through each vertex in $I$; and
@@
 
-A \emph{grid square} $S_{x,y,\ell}$ of side-length $\ell$ is a point set of the form $\{x,\ldots,x+\ell-1\}\times\{y,\ldots,y+\ell-1\}$ for some $x,y\in\Z^2$.
-The \emph{interior} of $S_{x,y,\ell}$ is the grid square $S_{x+1,y+1,\ell-2}$
+It will be helpful to take some natural definitions for $G_k$ and extend them to $G^{\onesub}_k$.
+To do this, we associate each vertex $v$ of $G^{\onesub}_k$ with a vertex $v^{\leftup}$ of $G_k$.
+For each vertex $v\in V(G_k)$, we define $v^{\leftup}:=v$.
+For each subdivision vertex $s$ in $G^{\onesub}_k$ with neighbours $v$ and $w$, we define $s^{\leftup}$ to be the lexicographically smaller of $v$ or $w$.
+For two vertices $v_1,v_2\in V(G^{\onesub}_k)$, with $v^{\leftup}_1:=(x_1,y_1)$ and $v^{\leftup}_2:=(x_2,y_2)$ we define the $L_\infty$-distance between $v_1$ and $v_2$ as $\|v_1v_2\|_\infty = \max\{|x_1-x_2|, |y_1-y_2|\}$.
+For any $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $r$ centered at $v$ is $B_{r}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le r\}$.\todo{Make a figure that illustrates this.}
 
 \begin{lem}\label{four_points}
-  Let $P$ be a set of at most $4$ points in $\Z^2$.
-Then there exists a set $S$ of disjoint grid squares, each of side-length $10$ such that each point of $P$ is contained in the interior of some square in $S$.
+  Let $P$ be a set of at most $3$ vertices of $G^{\onesub}_k$.
+Then there exists a set $S$ of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.\todo{Define the interior of a ball.}
 \end{lem}
@@
 \begin{proof}
-  Let $P:=\{p_1,\ldots,p_t\}$ where $p_i:=(x_i,y_i)$.
+  Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.
 Let $X:=\bigcup_{i=1}^t \{x_i\bmod 10,(x_{i+1})\bmod 10\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 10,(y_i+1)\bmod 10\}$.
-Then $|X|\le 2t\le 8$ and $|Y|\le 2t\le 8$.
-In particular, there exists some $x\in\{0,\ldots,9\}\setminus X$ and some $y\in\{0,\ldots,9\}\setminus Y$.
-We use squares of the form $\{x+10r,\ldots,x+10r-1\}\times\{y+10r\ldots,y+10r-1\}$ for each $r\in\Z$.
+Then $|X|,|Y|\le 2t\le 6$.
+In particular, there exists some $x\in\{0,\ldots,6\}\setminus X$ and some $y\in\{0,\ldots,6\}\setminus Y$.
+For each $i,j\in\Z$, let $c_{i,j}:=(7i+x+3,7j+y+3)$ and observe that the ball $B_{3}(c_{i,j})$ does not have any point of $P$ on its boundary.
+Furthermore, the set of balls $S:=\{B_{3}(c_{i,j}):i,j\in\Z\}$ is a partition of $V(G^{\onesub}_{k})$.
+Therefore $S$ is a set of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.
 \end{proof}
@@
 \begin{lem}
-  Let $P$ be a set of points in $\Z^2$ such that no square of side-length $80$ contains more than 4 points of $P$.
-Then there exists a set $S$ of disjoint grid squares, each of side length $10$ and such that each point of $P$ is in the interior of some square in $S$.
+  Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $9$ contains more than $3$ vertices in $P$.
+Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
 \end{lem}
@@
 \begin{proof}
-  Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|p-q\|_\infty \le 20$.
-Observe that each component of $H$ contains at most $4$ points of $P$.
-For each component $C$ of $H$, apply \cref{four_points} to $V(C)$.
-This produces a set of disjoint grid squares of side length $10$ that each contain some point of $V(C)$ in their interior.
-Any two squares $s_1$ and $s_2$ obtained from different components of $H$ are disjoint because $s_1$ contains a point $p_1\in P$ and $s_2$ contains a point $p_2\in P$ such that $\|p_1-p_2\|\infty > 20$, but the squares $s_1$ and $s_2$ each have side length $10$.\todo{Could be tidier here.}
+  Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$.
+Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $9$.
+Therefore each component of $H$ contains at most $3$ vertices.
+
+  For each component $C$ of $H$, apply \cref{four_points} to $V(C)$ and keep only those balls in $S$ that contain at least one vertex of $V(C)$.
+This produces a set $S_C$ of disjoint balls, each of radius $3$, that each contain some vertex in $V(C)$ in their interior.
+Any two balls $B_{C_1}$ and $B_{C_2}$ obtained from different components of $H$ are disjoint because the first ball contains a vertex $p_1\in P$ and the second ball contains a vertex $p_2\in P$ such that $\|p_1p_2\|\infty > 12$, but the balls $s_1$ and $s_2$ each have radius $3$.
 \end{proof}
 
-\todo[inline]{Adjust the numbers in the point covering lemmas.}
-
-\begin{lem}
-  Let $k$ be a positive integer, let $I,L$ be disjoint subsets of $V(G_{k})$, let $Q$ be a subset of $E(G_{k})$ and let $X:=I\cup L\cup Q$ have the following properties:
-  \begin{compactenum}[(Pr1)]\setcounter{enumi}{0}
-    \item Each $80\times 80$ subgrid of $G_k$ contains at most $4$ elements of $X$.
-    \item $Q$ does not contain all the edges of a $4$-cycle.
-    % \item $X$ contains no boundary vertices or edges.
-    \item No vertex of $G_k$ is incident to three or more edges in $Q$.
-    \item No vertex in $I\cup L$ is incident to two or more edges in $Q$.
+\begin{lem}\label{one_path_enchilada}
+  Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
+  \begin{compactenum}[(1)]\setcounter{enumi}{-1}
+    \item Any ball of radius 9 contains at most $3$ elements of $I\cup L\cup Q$.
+    \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
+    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
+    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
   \end{compactenum}
-  Then $G_k$ contains a path $P$ that
-  \begin{compactenum}
-    \item contains all elements of $X$;
+  Then $G^{\onesub}_k$ contains a path $P$ that
+  \begin{compactenum}[(a)]
+    \item contains all the vertices in $I\cup L\cup Q$;
     \item goes straight through each vertex in $I$; and
@@
 \begin{proof}
-  Start with the snakelike path that visits every $10$th row of $G_k$.
-Cover the elements of $X$ with disjoint squares of side length $10$.
+  Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$.
+Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.
 Each such square is intersected by exactly one row of the snakelike path.
-Use \cref{ten_by_ten} to locally deform the path to pick up the objects in any square that it intersects.
+Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in any square that it intersects.
 \end{proof}
+
+\subsection{Approximate Voronoi Diagrams in $G^{\onesub}_k$}
+
+Here we explain the ranking trick that allows us to make approximate Voronoi diagrams.
+If the minimum-distance between any two sites in these diagrams is $\beta$, then any four points in four different cells of the diagram contains a pair of points at distance $\Omega(\beta)$.
+In particular, there exists a constant $a>0$ such that any ball of radius $a\beta$ contains points from at most $3$ grid cells.
+We will then use this to find vertex sets $P\subseteq V(G^{\onesub}_k)$ that can be used with \cref{one_path_enchilada}.
+
+\subsection{Pseudogrids}
 

### 2021-10-01 lin-vs-cen.tex
Pffffff

@@
 \usepackage{todonotes}
+
+% etoolbox allows for robust commands that don't need \protect, e.g.
+% \newrobustcmd{\onesub}{\mathord{\includegraphics{figs/one-sub}}}
+% \subsection{Approximate Voronoi Diagrams in $G^{\onesub}_k$}
 \usepackage{etoolbox}
@@
 
+% Document-specific commands and math operators
 \DeclareMathOperator{\tw}{tw}
@@
 \DeclareMathOperator{\dist}{dist}
+\DeclareMathOperator{\vor}{Vor}
 
@@
 
+\todo[inline]{This lemma is not true.
+Pick a center $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.
+Then $x$ is incident on all four Voronoi regions.
+We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
+\begin{lem}
+  For any $r\in\N$ there exists $\beta\in\N$ such that, for any $k\in\N$ and any $P\subseteq G^{\onesub}_k$ with $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$ there exists a function $\vor_P:V(G^{\onesub}_k)\to P$ such that
+  \begin{compactenum}[(i)]
+    \item for any vertex $v$ of $G^{\onesub}_k$, $|\{\vor_P(v)\}\cup\{\vor_P(w):w\in N_{G^{\onesub}_k}(v)\}|\le 2$.
+    \item for any ball $B$ of radius $r$, $|\{\vor_P(v):v\in B\}|\le 3$;
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  % The idea behind this proof is to use the Voronoi diagram of $P$ to define the function $\vor$.
+The problem is that the Voronoi diagram can have vertices of high degree and pairs of vertices that are close together, both of these violate the second requirement.
+Instead, we construct an approximate Voronoi diagram in which each point is assigned to one of
+  We first define a function $\vor':V(G^{\onesub}_k)\to P$ that only satisifies the second condition and then show how local modifications can be used to obtain a function $\vor$ that satisifies both conditions.
+
+  Arbitrarily order the points of $P$ as $p_1,\ldots,p_{|P|}$.
+For each vertex $v$ of $:V(G^{\onesub}_k)$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+ar\}$ for some $a>1$ discussed shortly, and let $\vor_P'(v):=\min\{i:p_i\in S(v)\}$.
+In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $ar$), and $\vor_P'(v)$ is the approximate nearest neighbour of $v$ of smallest index.
+
+  We now argue that for sufficiently large $\beta$, depending only on $r$, $\vor'$ satisfies the second condition.
+Let $v_1,\ldots,v_4\in V(G^{\onesub}_k)$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor_P'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.
+Let $P':=\{p_1,\ldots,p_4\}$ and observe that $\vor'_{P'}(v_j)=\vor'_P(v_j)$ for each $j\in\{1,\ldots,4\}$.
+Therefore, it suffices to consider the case where $|P|=4$, so that $i_j=j$ for each $j\in\{1,\ldots,4\}$.
+
+  Claim: Some ball $B_2(x_1)$ of radius $2$ contains three vertices $y_1,y_2,y_3$ such that $\vor'(x_i)=p_i$ for each $i\in\{1,2,3\}$.
+Claim: Some ball $B_2(x_2)$ of radius $2$ contains three vertices $z_1,z_2,z_4$ such that $\vor'(z_i)=p_i$ for each $i\in\{1,2,4\}$.
+(Prove this.)
+
+  We will show that there exists $j_1,j_2\in\{1,\ldots,4\}$ such that $\|v_{j_1}v_{j_2}\|_\infty > 2r$, and therefore no ball of radius $r$ contains $v_1,\ldots,v_4$.
+
+  Let $B'$ be any ball of radius $2$ that contains
+
+\end{proof}
+
 \subsection{Pseudogrids}

### 2021-10-03 lin-vs-cen.tex
Working proof

--- 2021-09-30 lin-vs-cen.tex
@@
 
+% \todo[inline]{This lemma is not true.
+Pick a center $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.
+Then $x$ is incident on all four Voronoi regions.
+We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
+\begin{lem}
+  For any $r\in\N$ there exists $\beta\in\N$ such that, for any $k\in\N$ and any $P\subseteq G^{\onesub}_k$ with $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$ the following statement is true.
+Let $Q:=\bigcup_{p\in P} B_\beta(p)$.
+There exists a function $\vor_P:Q\to P$ such that
+  \begin{compactenum}[(i)]
+    \item for any $v\in Q$, $|\{\vor_P(v)\}\cup\{\vor_P(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
+    \item for any ball $B$ of radius $r$, $|\{\vor_P(v):v\in B\cap Q\}|\le 3$;
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  % The idea behind this proof is to use the Voronoi diagram of $P$ to define the function $\vor$.
+The problem is that the Voronoi diagram can have vertices of high degree and pairs of vertices that are close together, both of these violate the second requirement.
+Instead, we construct an approximate Voronoi diagram in which each point is assigned to one of
+  We first define a function $\vor':V(G^{\onesub}_k)\to P$ that only satisifies the second condition and then show how local modifications can be used to obtain a function $\vor$ that satisifies both conditions.
+
+  Arbitrarily order the points of $P$ as $p_1,\ldots,p_{|P|}$.
+For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+ar\}$ for some $a>1$ discussed shortly, and let $\vor_P'(v):=\min\{i:p_i\in S(v)\}$.
+In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $ar$), and $\vor_P'(v)$ is the approximate nearest neighbour of $v$ of smallest index.
+
+  We now argue that for sufficiently large $\beta$, depending only on $r$, $\vor'$ satisfies the second condition.
+Let $v_1,\ldots,v_4\in V(G^{\onesub}_k)$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor_P'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.
+Suppose that $v_1,\ldots,v_4$ are contained in some ball $B_{x/2}(c)$ of radius $x$
+  % , so that $\max\{\|v_iv_j\|_\infty:i,j\in\{1,\ldots,4\}\}\le x$.  Then
+  \[ \|p_{i_1}c\|_\infty \le \|p_{i_1}v_1\|+x \le \beta + ar + x \]
+  Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
+  \[ \|p_{i_2}v_2\|_\infty
+     < \|p_{i_1}v_2\|_\infty - ar
+     \le \|p_{i_1}c\|_\infty + x - ar
+     \le \beta + 2x
+  \]
+  so $\|p_{i_2}c\|_\infty < \beta + 3x$.
+  Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
+  \[
+     \|p_{i_2}v_3\|_\infty
+     < \|p_{i_2}v_3\|_\infty - ar
+     \le \|p_{i_2}c\|_\infty + x - ar
+     < \beta + 4x - ar
+  \]
+  so $\p_{i_2}c\|_\infty \le \beta + 5x - ar$.  Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
+  \[
+    \|p_{i_4}v_4\|_\infty
+    < \|p_{i_3}v_4\|_\infty - ar
+    \le \|p_{i_3}c\|_\infty +x -ar
+    < \beta + 5x - 2ar \enspace .
+  \]
+  Setting $ar := 3\beta/4$ and $5x=\beta/2$ we find that $\|p_{i,4}v_4\|_\infty < 0$, clearly a contradiction.
+We conclude that, for $ar=3\beta/4$, any ball of radius $\beta/10$ contains $v_1,\ldots,v_4$.
+Thus, $\vor'$ satisfies the second condition of the lemma.\todo{Optimize and cleanup.
+Finish proving (i)}
+\end{proof}
+
 \subsection{Pseudogrids}

### 2021-10-04 lin-vs-cen.tex
Put in explicity constants

--- 2021-09-30 lin-vs-cen.tex
@@
 
+% \todo[inline]{This lemma is not true.
+Pick a center $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.
+Then $x$ is incident on all four Voronoi regions.
+We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
+\begin{lem}
+  For any $\beta, k\in\N$ and any $P\subseteq V(G^{\onesub}_k)$ with $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$ the following statement is true:  Let $Q:=\bigcup_{p\in P} B_\beta(p)$.
+There exists a function $\vor_P:Q\to P$ such that
+  \begin{compactenum}[(i)]
+    \item for any $v\in Q$, $|\{\vor_P(v)\}\cup\{\vor_P(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
+    \item for any ball $B$ of radius $\beta/12$, $|\{\vor_P(v):v\in B\cap Q\}|\le 3$;
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  % The idea behind this proof is to use the Voronoi diagram of $P$ to define the function $\vor$.
+The problem is that the Voronoi diagram can have vertices of high degree and pairs of vertices that are close together, both of these violate the second requirement.
+Instead, we construct an approximate Voronoi diagram in which each point is assigned to one of
+  We first define a function $\vor':Q\to P$ that satisfies the second condition and then show how local modifications can be used to obtain a function $\vor$ that satisfies both conditions.
+
+  Arbitrarily order the vertices of $P$ as $p_1,\ldots,p_{|P|}$.
+For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+3\beta/4\}$, and let $\vor'_P(v):=\min\{i:p_i\in S(v)\}$.
+In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $\beta/12$), and $\vor_P'(v)$ is (the index of) the approximate nearest neighbour of $v$ of smallest index.
+
+  We now argue that $\vor'$ satisfies the second condition.
+Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor_P'(v_j)=i_j$ for each $j\in\{1,\ldots,4\}$.
+Suppose, for the sake of contradiction, that $i_1,\ldots,i_4$ are all distinct and that $v_1,\ldots,v_4$ are contained in some ball $B_{\beta/12}(c)$.
+By the triangle inequality $\|v_iv_j\|_\infty \le \|v_ic\|_\infty+\|cv_j\|_\infty \le \beta/6$ for each  $i,j\in\{1,\ldots,4\}$.
+  Since $v_1\in Q$, $d(v)\le\beta$ so
+  \[ \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \]
+  Since $\vor'(v_2)=i_2\neq i_1$,
+  \[ \|p_{i_2}v_2\|_\infty
+     < \|p_{i_1}v_2\|_\infty - 3\beta/4
+     \le \|p_{i_1}v_1\|_\infty + \beta/6 - 3\beta/4
+     \le \beta + \beta/5 \enspace .
+  \]
+  Since $\vor'(v_3)=i_3\neq i_2$,
+  \[
+     \|p_{i_2}v_3\|_\infty
+     < \|p_{i_2}v_3\|_\infty - 3\beta/4
+     \le \|p_{i_2}v_2\|_\infty + \beta/6 - 3\beta/4
+     < \beta + 2\beta/6 - 3\beta/4 \enspace .
+  \]
+  Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
+  \[
+    \|p_{i_4}v_4\|_\infty
+    < \|p_{i_3}v_4\|_\infty - 3\beta/4
+    \le \|p_{i_3}v_3\|_\infty + \beta/5 - 3\beta/4
+    < \beta + 3\beta/6 - 3\beta/2 = 0 \enspace ,
+  \]
+  which is clearly a contradiction.
+This establishes that $\vor'$ satisfies Condition~(ii).\todo{Optimize and cleanup.
+Finish proving (i)}\todo{Why not use $\beta$ (or $100\beta$) instead of $3\beta/4$?}
+\end{proof}
+
 \subsection{Pseudogrids}

### 2021-10-05 lin-vs-cen.tex
Stuck!

@@
 \usepackage{cancel}
-\usepackage{enumitem}
+%\usepackage{enumitem}
 
@@
 A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.
-For a graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.
+For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.
+Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.
 
@@
 
+Refer to \cref{grid_figure}.
 For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
@@
 We say that a path $P:=v_0,v_1,\ldots,v_\ell$ in $G_{a\times b}$ \emph{goes straight} at a vertex $v_i$, $i\in\{1,\ldots,k-1\}$ if $v_{i-1}$, $v_i$, and $v_{i+1}$ are contained in a single row or column.
-Otherwise, we say that $P$ \emph{turns} at $v_i$.
+Otherwise, we say that $P$ \emph{bends} at $v_i$.
 
@@
   \begin{center}
-    \begin{tabular}{cc}
+    \begin{tabular}{c@{\hspace{2cm}}c}
       \includegraphics{figs/grid-1} &
@@
   \end{center}
-  \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}
+  \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
+  \label{grid_figure}
 \end{figure}
@@
 We will work extensively with the $1$-subdivision $G^{\onesub}_k$ of $G_k$.
-In this case, we define $\partial G^{\onesub}_k:=\{v\in V(G^{\onesub}_k):\dist_{G^{\onesub}_k}(v,\partial G_k)\le 1\}$.
-That is, the boundary of $G^{\onesub}_k$ includes the vertex boundary of $G_k$ as well as any degree-$2$ vertex of $G^{\onesub}_k$ that is adjacent to at least one boundary vertex of $G_k$.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{figs/grid-3}
-  \end{center}
-  \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
-\end{figure}
+In this case, we define $\partial G^{\onesub}_k:= \partial G_k\cup N_{G^{\onesub}_k}[\partial G_k]$.\todo{Define open and closed neighbourhods $N_G(\cdot)$ and $N_G[\cdot]$.}
 
 \begin{lem}\label{seven_by_seven}
-  Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_k$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
+  Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_7$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
   \begin{compactenum}[(i)]\setcounter{enumi}{0}
@@
     \item goes straight through each vertex in $I$; and
-    \item turns at each vertex in $L$.\todo{Define turns and goes straight}
+    \item bends at each vertex in $L$.
     % \item contains no boundary vertices or edges except $s$ and $t$.
@@
 For each subdivision vertex $s$ in $G^{\onesub}_k$ with neighbours $v$ and $w$, we define $s^{\leftup}$ to be the lexicographically smaller of $v$ or $w$.
+In other words, $s^{\leftup}$ is the vertex of $G_k$ that is directly below or the left of $s$.
+
 For two vertices $v_1,v_2\in V(G^{\onesub}_k)$, with $v^{\leftup}_1:=(x_1,y_1)$ and $v^{\leftup}_2:=(x_2,y_2)$ we define the $L_\infty$-distance between $v_1$ and $v_2$ as $\|v_1v_2\|_\infty = \max\{|x_1-x_2|, |y_1-y_2|\}$.
-For any $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $r$ centered at $v$ is $B_{r}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le r\}$.\todo{Make a figure that illustrates this.}
+Refer to \cref{ball_figure}.
+For each $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $\beta$ centered at $v$ is $B_{\beta}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le \beta\}$.
+Because it comes up in several places, it is worth noting that $|B_{\beta}(v)|\le 4(\beta+1)^2$, and this bound is tight for each $v\in\{\beta+1,\ldots,\beta-r-1\}\times\{\beta+1,\ldots,b-\beta-1\}$.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{figs/grid-3}
+  \end{center}
+  \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
+  \label{ball_figure}
+\end{figure}
 
@@
 
-Here we explain the ranking trick that allows us to make approximate Voronoi diagrams.
-If the minimum-distance between any two sites in these diagrams is $\beta$, then any four points in four different cells of the diagram contains a pair of points at distance $\Omega(\beta)$.
-In particular, there exists a constant $a>0$ such that any ball of radius $a\beta$ contains points from at most $3$ grid cells.
-We will then use this to find vertex sets $P\subseteq V(G^{\onesub}_k)$ that can be used with \cref{one_path_enchilada}.
-
 % \todo[inline]{This lemma is not true.
@@
 We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
-\begin{lem}
-  For any $\beta, k\in\N$ and any $P\subseteq V(G^{\onesub}_k)$ with $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$ the following statement is true:  Let $Q:=\bigcup_{p\in P} B_\beta(p)$.
-There exists a function $\vor_P:Q\to P$ such that
+\begin{lem}\label{voronoi_mapping}
+  Let $\beta, k\in\N$, let $P\subseteq V(G^{\onesub}_k)$ be such that $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$, and define  $Q:=\bigcup_{p\in P} B_\beta(p)$.
+Then there exists a function $\vor:Q\to P$ such that
   \begin{compactenum}[(i)]
-    \item for any $v\in Q$, $|\{\vor_P(v)\}\cup\{\vor_P(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
-    \item for any ball $B$ of radius $\beta/12$, $|\{\vor_P(v):v\in B\cap Q\}|\le 3$;
+    \item \label{real_close} For each $p\in P$ and each $v\in B_{\beta/8-1}(p)$, $\vor(v)=p$.
+    \item \label{never_far} For each $v\in Q$, $\|v\vor(p)\|_\infty \le 7\beta/4$.
+    \item \label{three_in_ball} For any ball $B$ of radius $\beta/12$, $|\{\vor(v):v\in B\cap Q\}|\le 3$. 
+    \item \label{twocolour_neighbourhood} For any $v\in Q$, $|\{\vor(v)\}\cup\{\vor(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$. 
   \end{compactenum}
@@
 \begin{proof}
-  % The idea behind this proof is to use the Voronoi diagram of $P$ to define the function $\vor$.
-The problem is that the Voronoi diagram can have vertices of high degree and pairs of vertices that are close together, both of these violate the second requirement.
-Instead, we construct an approximate Voronoi diagram in which each point is assigned to one of
-  We first define a function $\vor':Q\to P$ that satisfies the second condition and then show how local modifications can be used to obtain a function $\vor$ that satisfies both conditions.
+  We first define a function $\vor':Q\to P$ that satisfies (\ref{real_close})--(\ref{three_in_ball}) and then show how local modifications can be used to obtain a function $\vor$ that also satisfies (\ref{twocolour_neighbourhood}).
 
   Arbitrarily order the vertices of $P$ as $p_1,\ldots,p_{|P|}$.
-For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+3\beta/4\}$, and let $\vor'_P(v):=\min\{i:p_i\in S(v)\}$.
-In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $\beta/12$), and $\vor_P'(v)$ is (the index of) the approximate nearest neighbour of $v$ of smallest index.
+For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+3\beta/4\}$, let $i(v):=\min\{i:p_i\in S(v)\}$, and let $\vor'(v)=p_{i(v)}$.
+In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $\beta/12$), and $\vor'(v)$ is the approximate nearest neighbour of $v$ of smallest index.
 
-  We now argue that $\vor'$ satisfies the second condition.
-Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor_P'(v_j)=i_j$ for each $j\in\{1,\ldots,4\}$.
+  To see that $\vor'$ satisfies (\ref{real_close}), observe that, for each distinct $p,q\in P$ and $v\in B_{\beta/8-1}(p)$, the triangle inequality implies that $\|vq\|_\infty \ge \|pq\|_\infty-\|vp\|_\infty \ge \beta-\beta/8+1 > 7\beta/8$.
+On the other hand $d(v)\le \beta/8$, so $d(v)+3\beta/4\le 7\beta/8$.
+Therefore, $q\not\in S(v)$ for any $q\in P\setminus\{p\}$, so $\vor'(v)=p$.
+
+  That $\vor'$ satsifies (\ref{never_far}) follows immediately from the fact that each $v\in Q$, so $d(v)\le \beta$, and therefore $\|pv\|_\infty\le 7\beta/4$ for each $p\in S(v)$.
+
+  We now argue that $\vor'$ satisfies (\ref{three_in_ball}).
+Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.
 Suppose, for the sake of contradiction, that $i_1,\ldots,i_4$ are all distinct and that $v_1,\ldots,v_4$ are contained in some ball $B_{\beta/12}(c)$.
@@
   Since $v_1\in Q$, $d(v)\le\beta$ so
-  \[ \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \]
-  Since $\vor'(v_2)=i_2\neq i_1$,
-  \[ \|p_{i_2}v_2\|_\infty
+  \[ 
+    \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
+  \]
+  Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
+  \[ 
+    \|p_{i_2}v_2\|_\infty
      < \|p_{i_1}v_2\|_\infty - 3\beta/4
@@
   \]
-  Since $\vor'(v_3)=i_3\neq i_2$,
+  Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
   \[
@@
   which is clearly a contradiction.
-This establishes that $\vor'$ satisfies Condition~(ii).\todo{Optimize and cleanup.
-Finish proving (i)}\todo{Why not use $\beta$ (or $100\beta$) instead of $3\beta/4$?}
+This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{Optimize and cleanup.
+Finish proving (iii)}\todo{Why not use $\beta$ (or $100\beta$) instead of $3\beta/4$?}
 \end{proof}
@@
 A graph $H$ is an $a\times b$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $H$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_{a\times b}$.
-Since $G_{a\times b}$ has maximum degree $4$ any graph $G$ that contains a $G_{a\times b}$-minor contains a subgraph $H$ that is a $a\times b$ pseudogrid.\footnote{This comes from the fact that any tree $T$ with at most $4$ leaves contains at most two vertices of degree greater than $2$.
-Contracting the path between these two vertices (if they exist) produces a tree with exactly one vertex of degree greater than $2$.}
+The following lemma connecting grid minors and pseudogrids is stated implicitly by \citet{kun.obrien.ea:polynomial}.
+We give a proof here because some of the details of the connection are required in order to prove our results.
+
+\begin{lem}\label{pseudogrid_minor}
+  If a graph $G$ contains a $G_{a\times b}$ grid minor then $G$ contain a subgraph $H$ that is an $a\times b$ pseudogrid.
+\end{lem}
+
+\begin{proof}
+  (This proof only uses the fact that $G_{a\times b})$ has no vertex of degree greater than $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
+  \begin{inparaenum}[(i)]
+    \item $B_x$ is non-empty and $H[B_x]$ is connected for each $x\in V(G_{a\times b})$;
+    \item $H/\mathcal{B}$ is isomorphic to $G_{a\times b}$ using the isomorphism given by $B_x\mapsto x$.
+  \end{inparaenum}
+  Let $(H,\mathcal{B})$ be chosen so that $H$ the minimum possible number of edges, over all choices $H$ and $\mathcal{B}$ that satisfy (i) and (ii).
+
+  An edge of $vw$ of $H$ is a \emph{border edge} if $v\in B_x$ and $w\in B_y$ for distinct $x,y\in V(G_{a\times b})$.
+For each $x\in V(G_{a\times b})$, let $\partial B_x$ denote the set of border edges incident on one vertex of $B_x$.
+Since $H$ is edge-minimal, $|\partial B_x|=\deg_{G_{a\times b}}(x)\le 4$, for each $x\in V(G_{a\times b})$.
+For each $x\in V(G_{a\times b})$, let $T_x\subseteq H$ be a minimal tree contained in $H[B_x]\cup \partial B_x$ that contains all the border edges in $\partial B_x$.
+We define the partition $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$ as follows:
+\begin{compactenum}
+  \item If $T_x$ has a node $v$ of degree $4$, then $P_x:=\{v\}$.  
+  \item If $T_x$ has two nodes $v$ and $w$ of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
+  \item If $T_x$ has exactly one node $v$ of degree $3$ then $P_{x}:=\{v\}$
+  \item Otherwise $T_x$ is a path and $P_x$ contains exactly one node in $V(T_x)$.
+\end{compactenum}
+It is straightforward to verify that, with this assignment $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$, the quotient graph $H/\mathcal{P}$ is a subdivision of $G_{a\times b}$, so that $H$ is indeed an $a\times b$ pseudogrid.
+\end{proof}
+
+The second case (ii) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.
+This case is especially important in our proof because it corresponds to a case in which some of the vertices of $P_x$ are shared by the $i$th column and the $j$th row of the pseudogrid.
+However, a path in the pseudogrid that uses vertices of $B_x$ may or may not contain all the vertices of $P_x$ depending on whether or not the corresponding path in $G_{a\times b}$ bends at $x$.
+This motivates the following definitions (see \cref{straight_or_bent}):
+\begin{compactenum}
+  \item We say that the $x$ is \emph{straight} if the path from $B_{(i-1,j)}$ to $B_{i+1,j}$ in $T_x$ contains all the vertices of $P_x$.
+  \item Otherwise, we say that $x$ is \emph{bent}.
+\end{compactenum}
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{c@{\hspace{1cm}}c}
+      \includegraphics{figs/bx-paths-1.pdf} & \includegraphics{figs/bx-paths-3} \\
+      (a) & (b)
+    \end{tabular}
+  \end{center}
+  \caption{The vertex $x:=(i,j)$ is (a)~straight if the path from $B_{(i,j-1)}$ to $B_{(i+1,j)}$ uses all vertices of $P_x$ or (b)~bent otherwise.}
+  \label{straight_or_bent}
+\end{figure}
+Note that, if $x$ is straight, then the path from $B_{i,j-1}$ to $B_{i,j+1}$ in $T_x$ also contains all the vertices of $P_x$.
+Furthermore, if $x$ is bent, then, for each $\overline\i,\overline\j\in\{-1,1\}$, the path from $B_{i,j+\overline\j}$ to $B_{i+\overline\i, j}$ in $T_x$ contains all the vertices of $P_x$.
 
@@
 
-  First we introduce some notations for mapping edges and vertices of the grid $G_k$ to their corresponding paths in the pseudogrid $H$.
+  First we introduce some notations for vertices of the subdivided grid $G^{\onesub}_k$ to their corresponding paths in the pseudogrid $H$.
 For each vertex $v$ of $G_k$ there is a path $P_v$ in $\mathcal{P}$ (possibly consisting of one vertex) such that the vertex $v$ appears in $H'$ as a result of contracting $P_v$ in $H$.
-In this way, each vertex $v$ of $G_k$ is associated with a non-empty \emph{colour set} $\phi(v):=\{\varphi(w):w\in V(P_v)\}$.
-Similarly, for each edge $e$ of $G_k$ there is a path $P_e$ in $H'$ that comes from subdividing $e$ (possibly zero times).
-Thus, each edge $e$ of $G_k$ is associated with a (possibly empty) \emph{colour set} $\phi(e):=\{\varphi(w):w\in P^-_e\}$, where $P^-_e$ denotes the (possibly empty) set of internal vertices on the path $P_e$.
+In this way, each vertex $v$ of $G_k$ is associated with a non-empty \emph{colour set} $\Phi(v):=\{\varphi(w):w\in V(P_v)\}$.
+Similarly, for each subdivision vertex $s$ of $G^{\onesub}_k$ that comes from subdividing an edge $e$ of $G_k$, there is a path $P_s$ in $H'$ that comes from subdividing $e$ (possibly zero times).
+Thus, each subdivision vertex $s$ of $G^{\onesub}_k$ is associated with a (possibly empty) \emph{colour set} $\Phi(s):=\{\varphi(w):w\in P^-_e\}$, where $P^-_e$ denotes the (possibly empty) set of internal vertices on the path $P_e$.
 
-  For each colour $c\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $c$ as the number of vertices and edges of $G_k$ that include $c$ in their colour set.
+  For each colour $c\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $c$ as the number of vertices $G^{\onesub}_k$ that include $c$ in their colour set.
 Fix some large integer constant $\alpha$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity less than $\alpha$ and \emph{frequent} otherwise.
@@
   \begin{clm}\label{all_frequent}
-    $H$ contains a $k'\times k'$ pseudogrid $H'$ in which every colour in $\{\varphi(v):v\in V(H')\}$ is frequent, for some $k'\ge (1-\epsilon \alpha)k$.
+    There exists an integer $k'\ge (1-\epsilon)k$ such that 
+    $H$ contains a $k'\times k'$ pseudogrid $H'$ in which every colour in $\{\varphi(v):v\in V(H')\}$ is frequent.
   \end{clm}
@@
     This proof uses the notion of \emph{removing} a row or column of a pseudogrid.
-Removing a row is defined as follows (removing a column is similar): For each edge $e$ of $G_k$ in the relevant row, remove all edges of $P_e$ from $H$.
+Removing a row is defined as follows (removing a column is similar): For each vertex $v$ of $G^{\onesub}_k$ in the relevant row, remove all edges of $P_v$ from $H$.
 Next, remove any isolated vertices from $H$ and finally, repeatedly remove degree-$1$ vertices from $H$ until none remain.
@@
 
-    While there exists some infrequent colour $c\in\{1,\ldots,C\}$, find a row or column of $G_k$ that contains an edge or vertex with $c$ in its colour set and remove that row or column.
+    While there exists some infrequent colour $c\in\{1,\ldots,C\}$, find a row or column of $G^{\onesub}_k$ that contains a vertex with $c$ in its colour set and remove that row or column.
 Doing this exhaustively results in the removal of at most $(\alpha-1) C \le \epsilon\alpha k$ rows and at most $\epsilon\alpha k$ columns of $H$.
-This leaves a pseudogrid $H'$ that contains a $k'\times k'$ pseudogrid for some $k'\ge (1-\epsilon \alpha)k$ for which every colour that appears colours at least one vertex of $H'$ is frequent.
+This leaves a pseudogrid $H'$ that contains a $k'\times k'$ pseudogrid for some $k'\ge (1-\epsilon \alpha)k$ for which every colour that appears in at least one vertex of $H'$ is frequent.
   \end{proof}
@@
   Because of \cref{all_frequent}, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$.
-Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is an edge or vertex of $G_k$ that contains $c$ in its colour set.
-These pairs are carefully chosen so that we will be able to construct a simple path $P_k$ in $G_k$ that contains $\bigcup_{c\in\{1,\ldots,C\}} \{x_c,y_c\}$ and has some additional nice properties.
-Using these additional properties, we will then show how to convert $P_k$ into a simple path $P$ in $H$ that contains at least two vertices of each colour.
 
-  \begin{compactenum}[{Stage} 1:]
-    \item We begin this stage by defining each vertex and edge of $G_k$ as being \emph{unmarked}.
-Fix some integer constant $\beta>0$.
-We iterate over each $c\in\{1,\ldots,C\}$ in turn.
-If $G_k$ contain a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is an unmarked vertex or unmarked edge of $G_k$ with $c$ in its colour set and the distance in $G_k$ between $x_c$ and $y_c$ is at least $B$ then Stage~1 \emph{succeeds} for $c$.\todo{Define distance, including between sets}  We then \emph{mark} each vertex and edge $z$ of $G_k$ such that $\dist_{G_k}(z, \{x,y\})\le \beta$.
+  % Colour sets turn out to be problematic, so we would like to assign a single colour $c\in\Phi(v)$ to each $v\in G^{\onesub}_k$.
+To do this, consider the bipartite graph $X$ with parts $A:=\{1,\ldots,C\}$ and $B:=V(G^{\onesub}_k)$ and which contains the edge $cv$ if and only if $c\in\Phi(v)$.
+Since each colour in $A$ is frequent, $\deg_x(c)\ge a$ for each $c\in C$.
+S
+
+  Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is vertex of $G^{\onesub}_k$ that contains $c$ in its colour set.
+These pairs are carefully chosen so that we will be able to construct a simple path $P_k$ in $G^{\onesub}_k$ that contains every vertex in $X:=\bigcup_{c\in\{1,\ldots,C\}} \{x_c,y_c\}$.
+Furthermore, this path will bend at every bent vertex in $X$ and go straight at every straight vertex in $X$.
+This will ensure that the corresponding path in $H$ contains at least two vertices of each colour.
+
+  \begin{enumerate}[{Stage} 1:]
+    \item We begin this stage by defining each vertex $G^{\onesub}_k$ as being \emph{unmarked}.
+Next we consider each $c\in\{1,\ldots,C\}$ in turn.
+If $G^{\onesub}_k$ contains a pair $(x_c,y_c)$ where
+    \begin{inparaenum}[(i)]
+      \item each of $x_c$ and $y_c$ is unmarked;
+      \item $\|x_cy_c\|_\infty \ge \beta$; and
+      \item $c\in \phi(x_c)$ and $c\in \phi(y_c)$;
+    \end{inparaenum}
+    then we say that Stage~1 \emph{succeeeds} for colour $c$ and we mark each vertex in $B_{\beta}(x_c)$ and $B_\beta(y_c)$.
 
     If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $x_c$ and $y_c$ undefined for now and say that Stage~1 \emph{fails} for colour $c$.
-Note that Stage~1 can only fail for $c$ if $G_k$ contains fewer than $3(\beta/2+1)^2$\todo{check precisely} unmarked vertices/edges with $c$ in their colour sets.\footnote{$G_{a\times a}$ contains fewer than $3a^2$ vertices and edges and contains pairs of vertices/edges at distance $2(a-1)$.}  In particular, for $\alpha > 3(\beta/2+1)^2$, Stage~1 succeeds for $c=1$.
+Note that Stage~1 can only fail for $c$ if $G^{\onesub}_k$ contains fewer than $(2\beta+1)^2$ unmarked vertices with $c$ in their colour sets.
+In particular, for $\alpha > (2\beta+1)^2$, Stage~1 succeeds for $c=1$.
 
@@
     As mentioned above, for sufficiently large $\alpha$ (relative to $\beta$), the set $S$ is non-empty and therefore $R_1$ is non-empty.
-In the next paragraph, we will define a function $\nu$ that maps each marked vertex or edge of $G_k$ to some element of $R_1$.
-We use to this to construct a bipartite graph $I$ with vertex parts $A:=\overline{S}$ and $R_1$.
-The edge set of $I$ contains an edge $cx$ precisely when there is a marked edge or vertex $z$ of $G_k$ such that $c\in\phi(z)$ and $\nu(z)=x$.
+Let $P:=R_1$ and let $Q:=\bigcup_{p\in P} B_\beta(p)$ and apply \cref{voronoi_mapping} to obtain the mapping $\vor:Q\to P$.
+We use to this to construct a bipartite graph $I$ with vertex parts $\overline{S}$ and $R_1$.
+The edge set of $I$ contains an edge $cx$ precisely when there is a marked vertex $z$ of $G^{\onesub}_k$ such that $c\in\phi(z)$ and $\vor(z)=x$.
 
-    We now define the mapping $\nu$.
-Fix some total order on $R_1$ by assigning each element of $R_1$ a distinct rank in $\{1,\ldots,|R_1|\}$.
-For each marked element $z\in V(G_k)\cup E(G_k)$, let $\gamma(z):=\min\{\dist_{G_k}(z,x):x\in R_1\}$, let $\Gamma(z):=\{x\in R_1:\dist_{G_k}(z,x) \le (1+\delta)\gamma(z)$, and let $\nu(z)$ be the element in $\Gamma(z)$ of minimum rank.
-In words, $\gamma(z)$ is the distance to $z$'s nearest neighbour in $R_1$, $\Gamma(z)$ is the set of all approximate nearest neighbours of $z$ in $R_1$ and $\nu(z)$ is the approximate nearest-neighbour of $z$ with smallest rank.
-
-    Note that for any marked element $z$, there exists $x\in R_1$ such that $\dist_{G_k}(z,x)\le\beta$.
-Therefore $\gamma(z)\le\beta$.
-Therefore, for any $x\in R_1$, the set $\nu{^-1}(x)$ only contains marked elements $z$ for which $\dist_{G_k}(x,z)\le (1+\delta)\beta$  Thus, $|\nu^{-1}(x)|\le \zeta\beta^2$ for some $\zeta$ depending only on $\delta$.
+    First observe that, for each $p\in R_1$, \cref{voronoi_mapping}(\ref{never_far}) implies that $|\vor^{-1}(p)|\le|B_{7\beta/4}(p)|\le (7\beta/2+1)^2$.  
+    \todo[inline]{Fuck!
+We run into problems here with large colour sets.
+The next line is not true, for example.
+We get a $C$ in the denominator because the only degree bound we have on $R_1$ is $C(7\beta/2+1)^2$.}
+    This implies that, for any $A_0\subseteq\overline{S}$,
+    \[
+        |N_I(A_0)| \ge \frac{a|A_0|}{(7\beta/2+1)^2} \enspace .    
+    \]  
     This has two implications:
     \begin{compactenum}
-      \item For each $x\in R_1$,  $\deg_{I}(x) \le \zeta\beta^2$.
-      \item Since each colour $c\in A$ is frequent, $\deg_I(c)\ge \alpha/\zeta beta^2$.
+      \item For each $p\in R_1$,  $\deg_{I}(x) \le C(7\beta/2+1)^2$.
+      \item Since each colour $c\in A$ is frequent, $\deg_I(c)\ge \alpha/(7\beta/2+1)^2$.
     \end{compactenum}
@@
 In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
-  \end{compactenum}
+  \end{enumerate}
 

### 2021-10-05 lin-vs-cen.tex
Almost finished (modulo boundary issue) under <=r assumption

@@
 
-For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.
-For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.\todo{Surely this next lemma is well-known?}
+For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.  For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.
 
+For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.
+For any $S\subseteq V(G)$, $N_G(s):=\bigcup_{v\in S}N_G(v)$ and $N_G[s]:=\bigcup_{v\in S}N_G[v]$.
+The subgraph $G[S]$ of $G$ \emph{induced} by $S$ is the graph with vertex set $V(G[S]):=S\cap V(G)$ and edge set $E(G[S]):=\{vw\in E(G):\{v,w\}\subseteq S\}$.
+
+For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.
+When $\mathcal{P}$ consists of disjoint subsets of $V(G)$ that do not necessarily cover $V(G)$ then we use the convention that $G/\mathcal{P}:=G/\mathcal{P'}$ where $\mathcal{P}'$ is the completion of $\mathcal{P}$ obtained by introducing each missing element of $V(G)$ as singleton set.
+More precisely $\mathcal{P'}:=\mathcal{P}\cup\{\{v\}:v\in V(G)\setminus(\cup P)\}$.
+
+A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.
+Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.
+A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.
+For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.
+Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.
+A $\le\!\!\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has at most $\ell$ internal vertices.
+
+\todo[inline]{Surely this next lemma is well-known?}
+We make use of the following generalization of Hall's Marriage Theorem \cite{hall:on}:
 \begin{lem}\label{hall_vees}
@@
 Since $\deg_{\Lambda}(v) \le 2$ for each $v\in A_0$ and $\deg_\Lambda(v_0)<2$, $\deg_\Lambda(A_0)<2|A_0|$.
-This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.
+This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.greater than
 \end{proof}
 
+\begin{rem}
+  The constant $2$ that appears twice in \cref{hall_vees} can be replaced with any positive integer $c$.
+\end{rem}
+
 \section{The Linear Colouring Number of Pseudogrids}
-
-A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.
-Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.
-A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.
-For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.
-Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.
 
@@
   Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.
-Let $X:=\bigcup_{i=1}^t \{x_i\bmod 10,(x_{i+1})\bmod 10\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 10,(y_i+1)\bmod 10\}$.
+Let $X:=\bigcup_{i=1}^t \{x_i\bmod 7,(x_{i+1})\bmod 7\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 7,(y_i+1)\bmod 7\}$.
 Then $|X|,|Y|\le 2t\le 6$.
@@
 \begin{lem}
-  Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $9$ contains more than $3$ vertices in $P$.
+  Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $18$ contains more than $3$ vertices in $P$.
 Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
@@
   Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$.
-Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $9$.
+Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $18$.
 Therefore each component of $H$ contains at most $3$ vertices.
@@
   \begin{compactenum}[(1)]\setcounter{enumi}{-1}
-    \item Any ball of radius 9 contains at most $3$ elements of $I\cup L\cup Q$.
+    \item Any ball of radius 18 contains at most $3$ elements of $I\cup L\cup Q$.
     \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
@@
 Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.
-Each such square is intersected by exactly one row of the snakelike path.
-Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in any square that it intersects.
+Each such ball is intersected by exactly one row of the snakelike path.
+Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in each ball that it intersects.
 \end{proof}
@@
   \]
-  which is clearly a contradiction.
-This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{Optimize and cleanup.
-Finish proving (iii)}\todo{Why not use $\beta$ (or $100\beta$) instead of $3\beta/4$?}
+  which is clearly a contradiction.  This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{TODO: Optimize and cleanup. Finish proving (iii)}
 \end{proof}
@@
 \begin{proof}
-  (This proof only uses the fact that $G_{a\times b})$ has no vertex of degree greater than $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
+  (This proof only uses the fact that $G_{a\times b}$ has maximum degree  $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
   \begin{inparaenum}[(i)]
@@
   \end{inparaenum}
-  Let $(H,\mathcal{B})$ be chosen so that $H$ the minimum possible number of edges, over all choices $H$ and $\mathcal{B}$ that satisfy (i) and (ii).
+  Let $(H,\mathcal{B})$ be chosen so that $|E(H)|$ is minimum, over all choices of $(H,\mathcal{B})$ that satisfy (i) and (ii).
 
@@
 For each $x\in V(G_{a\times b})$, let $\partial B_x$ denote the set of border edges incident on one vertex of $B_x$.
-Since $H$ is edge-minimal, $|\partial B_x|=\deg_{G_{a\times b}}(x)\le 4$, for each $x\in V(G_{a\times b})$.
-For each $x\in V(G_{a\times b})$, let $T_x\subseteq H$ be a minimal tree contained in $H[B_x]\cup \partial B_x$ that contains all the border edges in $\partial B_x$.
-We define the partition $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$ as follows:
+For each $x\in V(G_{a\times b})$, let $T_x:= H[B_x]\cup \partial B_x$.
+Since $H$ is edge-minimal, $|\partial B_x|=\deg_{G_{a\times b}}(x)\le 4$, for each $x\in V(G_{a\times b})$ and $T_x$ is a tree with exactly $\deg_{G_{a\times b}}(x)$ leaves.\todo{Check the preceding claim.}  We define $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$ as follows:
 \begin{compactenum}
-  \item If $T_x$ has a node $v$ of degree $4$, then $P_x:=\{v\}$.  
-  \item If $T_x$ has two nodes $v$ and $w$ of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
-  \item If $T_x$ has exactly one node $v$ of degree $3$ then $P_{x}:=\{v\}$
-  \item Otherwise $T_x$ is a path and $P_x$ contains exactly one node in $V(T_x)$.
+  \item \label{one_special} If $T_x$ has exactly one node $v$ of degree at least three, then $P_x:=\{v\}$.
+  \item \label{two_degree_three} If $T_x$ has two nodes $v$ and $w$ each of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
+  \item \label{just_a_path} Otherwise $T_x$ is a path $v_0,\ldots,v_\ell$ for some $\ell\ge 2$ and $P_x:=v_1$.
 \end{compactenum}
@@
 
-The second case (ii) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.
+The second case (\ref{two_degree_three}) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.
 This case is especially important in our proof because it corresponds to a case in which some of the vertices of $P_x$ are shared by the $i$th column and the $j$th row of the pseudogrid.
@@
 
+\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird conditions.}
+
 \begin{lem}
-  For any $k\times k$ pseudogrid $H$, $\chilin(G)\in\Omega(k)$.
+  %For any $k\times k$ pseudogrid $H$
+  Let $H$ be a $k\times k$ pseudogrid and $\mathcal{P}:=(P_x:x\in V(G_{k}))$ be such that
+  \begin{inparaenum}[(i)]
+      \item for each $x\in V(G_k)$, $H[P_x]$ is a path of length at most $r$ and
+      \item $H/\mathcal{P}$ is a $\le\!\! r$-subdivision of $G_k$.
+  \end{inparaenum}
+  Then $\chilin(G)\in\Omega(k/r)$.
 \end{lem}
@@
 Let $R_1:=\bigcup_{c\in S}\{x_c,y_c\}$ be the set of vertices and edges chosen in $R_1$.
+By construction, $\min\{\|pq\|_\infty: \{p,q\}\in \binom{R_1}{2}\}\ge \beta$.
 
@@
 
-    First observe that, for each $p\in R_1$, \cref{voronoi_mapping}(\ref{never_far}) implies that $|\vor^{-1}(p)|\le|B_{7\beta/4}(p)|\le (7\beta/2+1)^2$.  
-    \todo[inline]{Fuck!
-We run into problems here with large colour sets.
-The next line is not true, for example.
-We get a $C$ in the denominator because the only degree bound we have on $R_1$ is $C(7\beta/2+1)^2$.}
+    First observe that each vertex $x$ of $G^{\onesub}_k$ corresponds to a path $P_x$ in $H$ that contains at most $r$ vertices.\footnote{For some subdivision vertices of $G^{\onesub}_k$, $P_x$ may be empty.}  This implies that $|\Phi(x)|\le r$ for each $x\in V(G^{\onesub}_k)$.
+Next observe that, for each $p\in R_1$, \cref{voronoi_mapping}(\ref{never_far}) implies that $|\vor^{-1}(p)|\le|B_{7\beta/4}(p)|\le (7\beta/2+1)^2$.
+This implies that $\deg_I(p) \le r(7\beta/2+1)^2$ for each $p\in R_1$.
+
     This implies that, for any $A_0\subseteq\overline{S}$,
     \[
-        |N_I(A_0)| \ge \frac{a|A_0|}{(7\beta/2+1)^2} \enspace .    
+        |N_I(A_0)| \ge \frac{a|A_0|}{r(7\beta/2+1)^2} \ge 2|N_I(A_0)| \enspace .
     \]
-    This has two implications:
-    \begin{compactenum}
-      \item For each $p\in R_1$,  $\deg_{I}(x) \le C(7\beta/2+1)^2$.
-      \item Since each colour $c\in A$ is frequent, $\deg_I(c)\ge \alpha/(7\beta/2+1)^2$.
-    \end{compactenum}
-    These two conditions imply that, for each $A_0\subseteq A$, $|N_I(A_0)|\ge \alpha/\zeta^2\beta^4$.
-In particular, for any $\alpha$ and $\beta$ with $\alpha > 2\zeta^2\beta^4$,  $|N_I(A_0)|\ge 2|A_0|$.
-Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
+    for appropriately chosen $\alpha$ and $\beta$ (with $\beta\in O(\sqrt{\alpha}))$.\todo{Specify precise condition.}  Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
 Therefore, each $c\in A=\overline{S}$ has two $\Lambda$-neighbours $p$ and $q$ in $B$.
-The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $s(x_c)=p$ and $c\in \phi(x_c)$.
-Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $s(y_c)=q$ and $c\in\phi(y_c)$.
+The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $s(x_c)=p$ and $c\in \Phi(x_c)$.
+Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $s(y_c)=q$ and $c\in\Phi(y_c)$.
 In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
@@
 
-  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \phi(x_c)\cap \phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
-Ultimately, we want to use these to find a path in $H$ that contains at least two vertices of each colour.
-We start by first establishing some properties of $R\subseteq V(G_k)$.
+  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \Phi(x_c)\cap \Phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
+  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.
 
-  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.
-The following claim follows immediately from the algorithm used to select pairs in Stage~1.
-  \begin{clm}\label{r1_far}
-      For each distinct pair $z,z'\in R_1$, $\dist_{G_k}(z,z')\ge \beta$.
-  \end{clm}
+  \todo[inline]{There is a technical condition missing here.
+To apply \cref{one_path_enchilada} we need the elements of $R$ to be distance at least $3$ from the boundary of $G^{\onesub}_k$.
+When counting multiplicity, we shouldn't count occurrences of a colour that are close to the boundary.}
 
-  The next claim states that any element $z'\in R_2$ is quite far from every element of $R_1$ except possibly its nearest neighbour in $R_1$.
-
-  \begin{clm}\label{far_from_home}
-    For each $x\in R_1$ and $y\in R_2$ with $\nu(y)\neq x$, $\dist_{G_k}(x,y)\ge \beta /(2+\delta)$.
-  \end{clm}
-
-  \begin{proof}
-    Since $\nu(y)\neq x$, there exists some other $z\in R_1\setminus\{y\}$ with $\nu(y)=\nu(z)$.
-Since $x,z\in R_1$, \cref{r1_far} implies that $\dist_{G_k}(x,z)\ge \beta$.
-Since $\nu(y)=z$, $\dist_{G_k}(y, z) \le (1+\delta)\gamma(y)\le (1+\delta)\dist_{G_k}(y, x)$.
-By the triangle inequality
-    \[ \beta \le \dist_{G_k}(x,z)
-        \le \dist_{G_k}(x,y)+\dist_{G_k}(y,z)
-        % \le \dist_{G_k}(x,y) + (1+\delta)\dist_{G_k}(x,y)
-        \le (2+\delta)\dist_{G_k}(x,y) \enspace . \qedhere
-    \]
-  \end{proof}
-
-  \begin{clm}
-    Any subset of $R$ containing containing $4$ elements contains a pair $(x,y)$ such that $\dist_{G_k}(x,y)\ge \psi\beta$, for some $\psi>0$ that depends only on $\delta$.
-  \end{clm}
-
-  \begin{proof}
-    Let $R':=\{x_1,x_2,x_3,x_4\}$ be a $4$-element subset of $R$.
-If $|R'\cap R_1|\ge 2$ because say $x_1,x_2\in R_1$ then $\dist_{G_k}(x_1,x_2)\ge \beta$, by \cref{r1_far}.
-If $|R'\cap R_1=1$ because say $x_1\in R_1$ then $\nu(x_i)= x_1$ for at most one $i\in\{2,3,4\}$.
-For each of the two $j\in\{2,3,4\}\setminus\{i\}$, $\dist_{G_k}(x_1,x_j)\ge \beta/(2+\delta)$, by \cref{far_from_home}.
-
-    The only possibility that remains is that $R'\subseteq R_2$.
-Let the elements of $R'$ be labelled so that the rank of $\nu(x_i)$ is less than the rank of $\nu(x_j)$ for each $i<j$.
-Suppose that $\dist_{G_k}(x_i,x_j)\le \psi\beta$ for each $i,j\in\{1,2,3\}$, otherwise there is nothing to prove.
-
-    First observe that, for any $i\in\{1,2,3,4\}$,  $\dist_{G_k}(x_i,\nu(x_i)) \ge (\tfrac{1}{2+\delta}-\psi)\beta$ since, otherwise
-    \[  \dist_{G_k}(x_{i+1},\nu(x_i))\le \dist_{G_k}(x_{i+1},x_i) +
-    \dist_{G_k}(x_i,\nu(x_i)) < \beta/(2+\delta) \enspace ,
-    \]
-    for any $j\in\{1,2,3,4\}\setminus\{i\}$,
-    contradicting \cref{far_from_home}.\todo{Make this a separate lemma?}
-
-  \end{proof}
-
-  \begin{clm}
-      There is no vertex of $G_k$ that has three or more incident edges in $R$.
-  \end{clm}
-
-  \begin{proof}
-    To prove this, we have to define $s(x)$ a little more carefully.
-  \end{proof}
-
-  \begin{clm}
-    There is no triple $(v,e_1,e_2)\in R$ where $v$ is a vertex of $G_k$ and $e_1$ and $e_2$ are each edges of $G_k$ incident on $v$.
-  \end{clm}
-
-  \todo[inline]{The following claim is not true.
-It's easy to come up with an example in which a small square contains $8$ different elements.
-We're probably better off adjusting the boundaries of Voronoi cells so that we enforce a minimum distance (say at least $\beta/10$) between Voronoi vertices.}
-  \begin{clm}
-    Any $5$-element subset of $R$ contains a pair $x,y$ with $\dist_{G_k}(x,y)\ge\beta/2$.
-  \end{clm}
-
-  \begin{proof}
-    Let $R'$ be some $5$-element subset of $R$.
-$R'$ contains two or more elements of $R_1$ then the distance between these two elements is at most $\beta>\beta/2$.
-Therefore, we may assume that $R'$ contains at most one element in $R_1$.
-If $R'$ contains exactly one element $x$ in $R_1$ then, by \cref{far_from_home}, there is at most one other element in $R'$ whose distance to $x$ is at most $\beta/2$, and the three remaining elements have greater distance.
-Therefore, $R'$ contains no elements in $R_1$.
-Therefore, $R'':=\{s(x):x\in R'\}$ is a five-element set whose pairwise distances are all at most $5\beta/2$ and whose minimum distance is at least $\beta$ (Contradiction?)\todo{Finish}.
-  \end{proof}
-
-  \begin{cor}
-      In any $\beta/2\times \beta/2$ subgrid $G_k$ contains at most $4$ elements of $R$.
-  \end{cor}
-
-    Now classify each vertex of $G_k$ as either being \emph{straight} or \emph{bent}.
-Define what it means for a path $P$ to go straight or bend at a vertex.
-Now we should be able to finish with this lemma:
-
-    \begin{lem}
-        $G_k$ contain a simple path $P$ that contains every edge and vertex in $R$.  Furthermore, $P$ bends at each bent vertex in $R$ and $P$ goes straight at each straight vertex in $R$.
-    \end{lem}
+  Conditions~(\ref{never_far}) and (\ref{three_in_ball}) of \cref{voronoi_mapping} imply that each any ball of radius $\min\{\beta/8-1, \beta/12\}/2$ contains at most three elements of $R$.
+Therefore, (for sufficiently large $\beta$) \cref{one_path_enchilada}, there exists a path $P$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$.
+These latter two conditions guarantee that $V(P)$ contains $P_x$ for each $x\in R$.
+For each colour $c\in\{1,\ldots,C\}$, there are at least two vertices $x_c,y_c\in R$ such that $P_{x_c}$ and $P_{y_c}$ each contain some vertex of colour $c$.
+Therefore the path $P$ contains two vertices of each colour.
+Therefore $P$ has no center.
+Therefore $\varphi$ is not a linear centered colouring of $H$.
+Since $\varphi$ is an arbitrary $\epsilon k$-colouring of $H$, this implies that $\chilin(H)\ge \epsilon k$.
 \end{proof}

### 2021-10-05 lin-vs-cen.tex
Specified some constants

@@
 
-\begin{lem}
+\begin{lem}\label{pseudogrid_lower_bound}
   %For any $k\times k$ pseudogrid $H$
@@
   This proof uses two large integer constants $\alpha$ and $\beta$ and two small positive real constants $\epsilon$ and $\delta$.
-The relationships between these constants are that $\beta = \Theta(\alpha^{1/4})$ and $\epsilon=\Theta(1/\alpha)$.
+The relationships between these constants are that $\alpha=\Theta(r\beta^2)$  $\epsilon=\Theta(1/\alpha)$.
 Let $\varphi:V(H)\to\{1,\ldots,C\}$ be an arbitrarily $C$-colouring of some $k\times k$ pseudogrid $H$ for some $C\le \epsilon k$.
@@
 
-  \begin{proof}
+  \begin{proof}[Proof of \cref{all_frequent}]
     This proof uses the notion of \emph{removing} a row or column of a pseudogrid.
@@
 This leaves a pseudogrid $H'$ that contains a $k'\times k'$ pseudogrid for some $k'\ge (1-\epsilon \alpha)k$ for which every colour that appears in at least one vertex of $H'$ is frequent.
+This completes the proof of \cref{all_frequent}.
   \end{proof}
 
-  Because of \cref{all_frequent}, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$.
+  Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=O(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.
+To avoid overloading notation, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$ in $H$.
+(The remainder of the proof does not depend on the values of $k$ or $C$.)
 
@@
       \item $\|x_cy_c\|_\infty \ge \beta$; and
-      \item $c\in \phi(x_c)$ and $c\in \phi(y_c)$;
+      \item $c\in \Phi(x_c)$ and $c\in \Phi(y_c)$;
     \end{inparaenum}
@@
 We use to this to construct a bipartite graph $I$ with vertex parts $\overline{S}$ and $R_1$.
-The edge set of $I$ contains an edge $cx$ precisely when there is a marked vertex $z$ of $G^{\onesub}_k$ such that $c\in\phi(z)$ and $\vor(z)=x$.
+The edge set of $I$ contains an edge $cx$ precisely when there is a marked vertex $z$ of $G^{\onesub}_k$ such that $c\in\Phi(z)$ and $\vor(z)=x$.
 
@@
     \[
-        |N_I(A_0)| \ge \frac{a|A_0|}{r(7\beta/2+1)^2} \ge 2|N_I(A_0)| \enspace .
+        |N_I(A_0)| \ge \frac{\alpha|A_0|}{r(7\beta/2+1)^2} \ge 2|A_0| \enspace .
     \]
-    for appropriately chosen $\alpha$ and $\beta$ (with $\beta\in O(\sqrt{\alpha}))$.\todo{Specify precise condition.}  Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
-Therefore, each $c\in A=\overline{S}$ has two $\Lambda$-neighbours $p$ and $q$ in $B$.
-The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $s(x_c)=p$ and $c\in \Phi(x_c)$.
-Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $s(y_c)=q$ and $c\in\Phi(y_c)$.
+    for $\alpha\ge 2r(7\beta/2+1)^2$.
+Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
+Therefore, each $c\in A=\overline{S}$ has two $\Lambda$-neighbours $p,q\in R_1$.
+The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $\vor(x_c)=p$ and $c\in \Phi(x_c)$.
+Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $\vor(y_c)=q$ and $c\in\Phi(y_c)$.
 In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
@@
   Conditions~(\ref{never_far}) and (\ref{three_in_ball}) of \cref{voronoi_mapping} imply that each any ball of radius $\min\{\beta/8-1, \beta/12\}/2$ contains at most three elements of $R$.
-Therefore, (for sufficiently large $\beta$) \cref{one_path_enchilada}, there exists a path $P$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$.
+This condition is satisfied for $\beta= 12\cdot18 = 218$ (so the conditions on $\alpha$ are satisified for any $\alpha = r(7\beta/2+1)^2=2331729r$).
+Therefore, by \cref{one_path_enchilada}, there exists a path $P$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$.
 These latter two conditions guarantee that $V(P)$ contains $P_x$ for each $x\in R$.
 For each colour $c\in\{1,\ldots,C\}$, there are at least two vertices $x_c,y_c\in R$ such that $P_{x_c}$ and $P_{y_c}$ each contain some vertex of colour $c$.
-Therefore the path $P$ contains two vertices of each colour.
-Therefore $P$ has no center.
-Therefore $\varphi$ is not a linear centered colouring of $H$.
-Since $\varphi$ is an arbitrary $\epsilon k$-colouring of $H$, this implies that $\chilin(H)\ge \epsilon k$.
+Therefore the path $P$ contains two vertices of each colour, so $P$ has no center and therefore $\varphi$ is not a linear colouring of $H$.
+Since $\varphi$ is an arbitrary $\epsilon k$-colouring of $H$, this implies that $\chilin(H)\ge \epsilon k = \Theta(k/\alpha)=\Theta(k/r)$.
 \end{proof}

### 2021-10-05 lin-vs-cen.tex
Pfffffft

@@
 
-\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird conditions.}
+\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}
 
@@
 
-  Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=O(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.
+  Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=\Theta(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.
 To avoid overloading notation, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$ in $H$.
@@
 
+To finish the proof, we need some lemma which states that highly subdivided graphs don't have linear colouring numbers that are much smaller than mildly subdivided graphs.
+It would be enough to prove the following.
+
+A \emph{refinement} $G'$ of a graph $G$ is a subdivision of $G$ in which each subdivision vertex is adjacent only to degree-$2$ vertices, i.e., $\deg_{G'}(v)=2$ for each $v\in N_{G'}(s)$ and each $s\in V(G')\setminus V(G)$.
+
+\begin{lem}\label{refinement_bound}
+  For any refinement $G'$ of any graph $G$, $\chilin(G') = \Omega(\chilin(G))$.
+\end{lem}
+
+In general, we know that if $G'$ is an $\ell$-subdivision of $G$ then
+\[  
+    \log\ell \le \chilin(G') \le \chilin(G) + \log\ell \enspace .
+\] 
+The lower bound comes from the fact that a path of length $\ell$ has a linear chromatic number at least $\log\ell$.
+The upper bound comes by introducing $\log\ell$ additional colours that are used to colour the paths in $G'-G$.
+Of course, the upper bound holds even when $G'$ is a $\le\!\!\ell$-subdivision of $G$ and the lower bound holds as long as at least one edge of $G$ is subdivided $\ell$ or more times.
+
+Perhaps \cref{refinement_bound} is true even if we replace ``refinement'' with ``subdivision''.
+As a test case, one could start with the $1$-subdivision $K^{\onesub}_k$ of the $k$-clique $K_k$.
+The following result is certainly true, as can be seen from a treedepth decomposition:
+
+\begin{lem}
+  For any subdivision $G'$ of any graph $G$, $\chicen(G') \ge \chicen(G)$.
+\end{lem}
+
 \bibliographystyle{plainurlnat}

### 2021-10-06 lin-vs-cen.tex
Small fix

@@
   Conditions~(\ref{never_far}) and (\ref{three_in_ball}) of \cref{voronoi_mapping} imply that each any ball of radius $\min\{\beta/8-1, \beta/12\}/2$ contains at most three elements of $R$.
-This condition is satisfied for $\beta= 12\cdot18 = 218$ (so the conditions on $\alpha$ are satisified for any $\alpha = r(7\beta/2+1)^2=2331729r$).
+Setting $\beta= 12\cdot2\cdot18 = 436$ (so the conditions on $\alpha$ are satisified for $\alpha = r(7\beta/2+1)^2=2331729r$) gives a point set $R$ that satisfies the requirements for \cref{one_path_enchilada}.
 Therefore, by \cref{one_path_enchilada}, there exists a path $P$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$.

### 2021-10-06 lin-vs-cen.tex
Fixed some confusion between paths in G_k and paths in H.

@@
 Setting $\beta= 12\cdot2\cdot18 = 436$ (so the conditions on $\alpha$ are satisified for $\alpha = r(7\beta/2+1)^2=2331729r$) gives a point set $R$ that satisfies the requirements for \cref{one_path_enchilada}.
-Therefore, by \cref{one_path_enchilada}, there exists a path $P$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$.
-These latter two conditions guarantee that $V(P)$ contains $P_x$ for each $x\in R$.
+Therefore, by \cref{one_path_enchilada}, there exists a path $P'$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$.
+These latter two conditions guarantee that the path $P$ in $H$ that corresponds to $P'$ contains $P_x$ for each $x\in R$.
 For each colour $c\in\{1,\ldots,C\}$, there are at least two vertices $x_c,y_c\in R$ such that $P_{x_c}$ and $P_{y_c}$ each contain some vertex of colour $c$.

### 2021-10-07 lin-vs-cen.tex
Some easy observations that improve exponent to 39

@@
 \DeclareMathOperator{\tw}{tw}
+\DeclareMathOperator{\td}{td}
 \DeclareMathOperator{\chicen}{\chi_{\mathrm{cen}}}
@@
 The \emph{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.
+We note that an easy argument shows that centered chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$.
 The \emph{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.
@@
 These notes are an attempt to improve the bound in \cref{kun_obrien_general} in the general case as well as for some special classes of graphs.
+First we point out that some new results can already reduce the exponent from $190$ to $39$ and eliminate the polylog factor:
 
-As a first step, we attempt to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.
-This, by itself, reduces the exponent in \cref{kun_obrien_general} from $190$ to $95$.
-Next, we observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.
+A recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).
+If we take $\tw(G)=k^{38}$ and $b=k$ then, for any graph $G$ with $\td(G)\ge k^{39}$, one of the following cases applies:
+\begin{enumerate}
+    \item $\tw(G)\ge k^{38}$: In this case the proof follows Steps~2 and 3 as before to show that $\chilin(G)\in\Omega(k)$.
+    \item $G$ contains a tree $T$ of treedepth $k$.
+The same paper \cite[Lemma~1.6]{czerwinski.nadara.ea:improved} shows that any tree $T$ of treedepth $k$ contains a subcubic subtree of treedepth $\Omega(k)$.
+    A theorem of \citet{kun.obrien.ea:polynomial} shows that, for any subcubic tree $T$, $\chicen(T) \le (\log_2 3)\cdot\chilin(T)$.
+Therefore, in this case $\chilin(G)\ge\chilin(T)\ge \chicen(T)/\log_2 3 = \Omega(k)$.
+\end{enumerate}
+
+This gives the following improvement to \cref{kun_obrien_general}
+\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general2}
+  For any graph $G$, $\chicen(G)\in O(\chilin^{39}(G))$.
+\end{thm}
+
+As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.
+This, by itself, reduces the exponent in \cref{kun_obrien_general2} from $39$ to $20$.
+We observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.
 For example, if we consider only planar graphs, then it is well known that any planar graph $G$ of treewidth $k$ contains an $\Omega(k\times k)$ grid minor.
-This already reduces the exponent further (for planar graphs) to $5$.
+This already reduces the exponent in \cref{kun_obrien_general2} further (for planar graphs) to $3$ ($2$ if we can prove the result for grid minors).
 If one could prove a similar result for the first ``heavy hammer'' then this would show that, for any planar graph $G$, $\chicen(G)\in O(\chilin(G))$.

### 2021-10-07 lin-vs-cen.tex
Change of wording

@@
 A recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).
-If we take $\tw(G)=k^{38}$ and $b=k$ then, for any graph $G$ with $\td(G)\ge k^{39}$, one of the following cases applies:
+Therefore, if $\td(G)\ge k^39$ then at least one of the following cases applies:
 \begin{enumerate}
@@
 As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.
-This, by itself, reduces the exponent in \cref{kun_obrien_general2} from $39$ to $20$.
+This, by itself, would reduce the exponent in \cref{kun_obrien_general2} from $39$ to $20$.
 We observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.
@@
 \begin{proof}
-  I have no idea how to prove this.  Computer search?.
+  Case analysis(?).
 \end{proof}
@@
 
-\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}
+\todo[inline]{We can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}
 

### 2021-10-08 lin-vs-cen.tex
Lil' stuff

@@
 These notes are an attempt to improve the bound in \cref{kun_obrien_general} in the general case as well as for some special classes of graphs.
-First we point out that some new results can already reduce the exponent from $190$ to $39$ and eliminate the polylog factor:
+First we point out that some new results can already reduce the exponent from $190$ to $39$ and eliminate the polylog factor.
+(These improvements are already described by \citet[Theorem~1.9]{czerwinski.nadara.ea:improved}.)
 
-A recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).
-Therefore, if $\td(G)\ge k^39$ then at least one of the following cases applies:
+First, \citet{chuzhoy.tan:towards} improved the exponent in the Excluded Grid Theorem:  There exists a constant $c$ such that any graph $G$ with $\tw(G)\ge k^{18}\log^c k$ has a $k\times k$ grid minor.
+This already improves the bound from $k^{190}$ to $k^{90}$.
+Next, a recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).
+Therefore, if $\td(G)\ge k^{19}$ then at least one of the following cases applies:
 \begin{enumerate}
-    \item $\tw(G)\ge k^{38}$: In this case the proof follows Steps~2 and 3 as before to show that $\chilin(G)\in\Omega(k)$.
+    \item $\tw(G)\ge k^{18}$: In this case the proof follows Steps~2 and 3 as before, but using the grid minor result of \citet{chuzhoy.tan:towards} to show that $\chilin(G)\in\Omega(k)$.
     \item $G$ contains a tree $T$ of treedepth $k$.
@@
 This gives the following improvement to \cref{kun_obrien_general}
-\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general2}
-  For any graph $G$, $\chicen(G)\in O(\chilin^{39}(G))$.
+\begin{thm}[\citet{czerwinski.nadara.ea:improved}]\label{kun_obrien_general2}
+  For any graph $G$, $\chicen(G)\in O(\chilin^{19}(G))$.
 \end{thm}

### 2022-03-31 lin-vs-cen.tex
Finally, an approach that's going to work

@@
 
+\newcommand{\defin}[1]{\emph{\color{brown}#1}}
 \setlength{\parskip}{1ex}
@@
 \title{\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}}
-\author{Prosenjit Bose, Vida Dujmović, Mehrnoosh Javarsineh, and Pat Morin}
+\author{Prosenjit~Bose, Vida~Dujmović, Hussein~Houdrouge, Mehrnoosh~Javarsineh, and Pat~Morin}
+
+\DeclareMathOperator{\VE}{VE}
 
@@
 \begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general}
-  For any graph $G$, $\chicen(G)\le \chilin^{190}(G)\cdot\log^{O(1)}(\chilin(G))$.
+  For any graph $G$, $\chicen(G)\le \chilin^{190}(G)\cdot\log^{O(1)}(\chilin(G))$.\todo{Update with most recent numbers}
 \end{thm}
@@
 
-\subsection{Subdivided Grids}
+Refer to \cref{grid_figure}.
+For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
+For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
 
-Refer to \cref{grid_figure}.
+In this work, we need to work with pseudogrids.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
+\begin{compactenum}
+  \item Replace each edge $vw$ of $G_{a\times b}$ with a path $\overline{P}_{vw}$ whose endpoints are $v$ and $w$.
+(In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
+  \item Replace each degree-$4$ vertex $(x,y)$ of $G_{a\times b}$ with a path $P_v$. If $P_v$ has only one vertex then this operation has no effect.  Otherwise, $P_v$ has two endpoints $p$ and $q$ and
+  \begin{compactenum}[(Q1)]
+    \item $p$ is adjacent to $\{(x-1,y), (x,y-1)\}$ and $q$ is adjacent to $(x+1,y),(x,y+1)$;
+    \item $p$ is adjacent to $\{(x,y+1), (x-1,y)\}$ and $q$ is adjacent to $(x,y-1),(x+1,y)$; or
+    \item $p$ is adjacent to $\{(x-1,y),(x+1,y)\}$ and $q$ is adjacent to $\{(x,y-1),(x,y+1)\}$.\label{q_iii}
+  \end{compactenum}
+\end{compactenum}
+
+Let $G$ be an $a\times b$ pseudogrid.
+For an edge $vw$ of $G_{a\times b}$, we let $P_{vw}:=\overline{P}_{vw}-\{v,w\}$ denote the (possibly empty) subpath containing the internal vertices of $\overline{P}_{vw}$.
+For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.
+In this way, $\mathcal{P}:=\{V(P_v):v\in V(G_{a\times b})\}\cup\{V(P_{vw}):vw\in E(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.
+We call $\mathcal{P}$ the \defin{grid-partition} of $G$ and we denote it by $(P_\mu:\mu\in V(G_{a\times b})\cup E(G_{a\times b}))$.
+This defines a mapping $\delta:V(G)\to V(G_{a\times b})\cup E(G_{a\times b})$ where $u\in V(P_{\delta(u)})$ for each $u\in v(G)$.
+
+% Then each vertex $u$ of $G$ naturally corresponds to an edge or vertex $u'$ of $G_{a\times b}$:  If $u$ is a vertex of $P_{vw}$ then $u'=vw$; if $u$ a vertex of $P_v$ then $u'=v$; and otherwise $u'=u$.
+This latter case occurs precisely when $u'$ is one of the boundary vertices of $G_{a\times b}$.
+
+Each row $R':=v_1,\ldots,v_a$ of $G_{a\times b}$ corresponds naturally to a path $R$ of $G$.
+The path $R$ contains $V(\overline{P}_{v_iv_{i+1}})$ for each $i\in\{1,\ldots,a-1\}$.
+However, for $i\in\{2,\ldots,a-1\}$ $R$ may or may not contain $V(P_{v_i})$.
+In particular, if $P_{v_i}$ was created using (Q\ref{q_iii}) then $R$ does not contain any internal vertices in $P_{v_i}$.
+Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
+This correspondence allows us to talk about the rows and columns of $G$, which we will do next.
+
+As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.
+If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$) then we repeatedly remove degree-$1$ vertices until none remain.
+If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.
+Similarly, if we delete column $C$ of $G$, then the resulting graph is an $(a-1)\times b$ pseudogrid.
+
+\subsection{Pseudogrids with Exclusively Frequent Colours}
+
+Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k\times k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
+For any $\mu\in\VE(G_{k\times k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\phi(v):v\in V(P_\mu)\}$.
+For any colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-2}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)$.
+For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-2}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-2}(\alpha)$.
+
+% We extend $\varphi$ to a partial colouring $\phi:V(G_{k\times k})\cup E(G_{k\times k})\to\{1,\ldots,c,\perp\}$ of the edges and vertices of $G_{k\times k}$ as follows:  For each vertex $v$ of $G_{k\times k}$, define $\phi(v)$ be an arbitrary colour in $\{\varphi(x):x\in V(P_v)\}$.
+For each edge $vw$ of $G_{k\times k}$, there are two possibilities:
+% If $P_{vw}$ is empty then $\phi(vw):=\perp$, otherwise $\phi(vw)$ is an arbitrary colour in $\{\varphi(x):x\in V(P_{vw})\}$.
+
+\begin{lem}
+  Let $0 < \epsilon < 1$, let $k\ge 1/\epsilon$ be an integer, let $G$ be a $k\times k$ pseudogrid and let $\varphi:V(G)\to\{1,\ldots,\lfloor\epsilon k\floor\}$ be a vertex colouring of $G$.
+  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ that has a grid-partition $\mathcal{P}':=(P'_\mu:\mu\in \VE(G_{k'\times k'})$ such that
+  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge |A|/2\epsilon$.
+\end{lem}
+
+\begin{proof}
+  The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.
+If there exists no $A\subseteq\{1,\ldots,\lfloor\epsilon k\rfloor\}$ with $\varphi_{\mathcal{P}}^{-2}(A) < |A|/2\epsilon$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
+Now assume that such a set $A$ exists.
+
+  We will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le |A|/2\epsilon$.
+For each $v:=(i,j)\in V(G_{k\times k})$ with $A\cap\varphi(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
+For each edge $vw\in E(G_{k\times k})$ with $A\cap\phi()\alpha\in\Phi(vw)$, we include the row of $G$ that contains $P_{vw}$ in $R$ or we include the column of $G$ that contains $P_{vw}$ in $C$.
+Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
+Now we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-|A|/2\epsilon$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
+In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.
+
+  \todo[inline]{Now apply induction......}
+\end{proof}
+
+\begin{lem}
+  Let $C>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(G)$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge C\cdot|A|+1$.
+Then there exists a partial colouring $\phi:\VE(G_{k\times k})\to\{1,\ldots,c,\perp\}$ with the following properties:
+  \begin{compactenum}[(i)]
+    \item For each $\mu\in\VE(G_{k\times k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
+    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k\times k}):\phi(\mu)=\alpha\}|\ge C$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  Use Hall's Theorem to find $C$ matchings such that each colour $\alpha$ appears in all $C$ matchings and each element of $\VE(G_{k\times k})$ appears in at most one matching.
+\end{proof}
+
+\subsection{Finding Paths Through Well-Separated Vertex/Edge Pairs}
+
+The \defin{$r$-box} centered at a vertex $v:=(i,j)$ of $G_{a\times b}$ is defined as
+\[  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{a,b}) \enspace . \]
+The \defin{$r$-box} centered at an edge $vw$ of $G_{a\times b}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.
+Observe that, for any $\mu\in V(G_{a\times b})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.
+Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{a\times b}[B_r(\mu)]$.
+We extend these definitions to vertices of an $a\times b$ pseudogrid $G$ so that
+\[
+   \tilde{B}_r(v) := \bigcup_{\mu\in V(G_r(\delta(v)) \cup E(G_r(\delta(v))} V(P_\mu)
+\]
+and $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
+
+\begin{lem}\label{pick_up_two}
+  Let $v$ be any vertex of $G$, let $w$ be any vertex of $\tilde{G}_r(v)$, let $s$ be any vertex in the leftmost column of $\tilde{G}_r(v)$ and let $t$ be any vertex in the rightmost column of $\tilde{G}_r(v)$.
+Then there exists an $st$-path\todo{define $st$-path} in $G_r(v)$ that contains $v$ and $w$.
+\end{lem}
+
+\begin{proof}
+  Let $i_s \le i_v \le i_w \le i_t$ be the columns of $\tilde{G}_r(v)$ that contain $s$, $v$, $w$, and $t$, respectively.
+Note that the assumption that $i_v \le i_w$ is without loss of generality and that $i_s \le i_t - r -1$.\todo{finish up}
+\end{proof}
+
+\begin{lem}
+  Let $r$ be an integer and let $S\subseteq V(G)$ be such that $|\tilde{B}_{r}(v)\cap S|=|\tilde{B}_{2r}\cap S|\le 2$ for each $v\in S$.
+Then $G$ contains a path that contains every vertex in $S$.
+\todo{Set specific values for $r$}
+\end{lem}
+
+\begin{proof}[Proof sketch]
+  Choose $S'\subseteq S$ so that, if $B_r(v)\cap S=\{v,w\}$ then exactly one of $v$ or $w$ is in $S'$.
+Now do some scan-line thing with a distance of $2r+1$ between consecutive rows.
+When this scan encounters a box $B_r(v)$ for some $v\in S'$, use \cref{pick_up_two} to detour through the box to pick up the (at most two) vertices in $B_r(v)\cap S$.
+\end{proof}
+
+\subsection{Finding an Uncentered Path}
+
+\begin{lem}
+  Let $G$ be a $k\times k$ pseudogrid and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a colouring of $G$ such that, for each $\alpha\in\{1,\ldots,C\}$, the set $S_\alpha:=\{\mu\in V(G_{k\times k})\cup E(G_{k\times k}): \alpha\in\Phi(\mu)\}$ has size at least $C$.
+Then there exists $S\subseteq V(G_{k\times k})\cup E(G_{k\times k})$ such that
+  \begin{compactenum}[(i)]
+    \item \label{hits_both} $|S\cap S_\alpha|\ge 2$ for each $\alpha\in\{1,\ldots,c\}$; and
+    \item \label{spread_out} $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof}
+  We construct $S$ using a two round process.  We start with an initially empty set $S_1$ and repeat the following for each $\alpha= 1,\ldots,c$:
+  If there exists distinct $\mu_1,\mu_2\in S_{\alpha}$ such that
+  \begin{compactenum}[(a)]
+    \item $\alpha\in\Phi(\mu_1)\cap\Phi(\mu_2)$;
+    \item $\mu_1\not\in B_r(\mu_2)$; and
+    \item $\{\mu_1,\mu_2\}\cap \bigcup_{\mu\in S} B_{r}(\mu)=\emptyset$
+  \end{compactenum}
+  then include $\mu_1$ and $\mu_2$ in $S_1$ and declare this a \defin{success} for colour $\alpha$.  Otherwise, declare this a failure for colour $\alpha$.
+
+  At the end of this process, the resulting set satisfies (\ref{spread_out}) but is only guaranteed to satisfy (\ref{hits_both}) for the successful colours $\alpha$.
+We now use the second round to create a set $S_2$ to fix this.
+Our strategy is to use \cref{hall_vees} to find two matchings between the failed colours and the boxes in $\{B_r(\mu):\mu\in S_1\}$ that contain those colours.
+Each missing colour is matched in both matching, but each box is used in at most one of the matchings.
+This is always possible, since the number of edges and vertices of $G_r(\mu)$ is $\Theta(r^2)$ while the
+
+  Consider a (for now) random permutation $\alpha_1,\ldots,\alpha_c$ of $\{1,\ldots,c\}$. For each $i\in\{1,\ldots,c\}$: Find $2(r+1)^2$ elements $\mu\in S_1$ such that
+  \begin{compactenum}[(a)]
+    \item
+  \end{compactenum}
+\end{proof}
+
+\newpage
+
+% \hrule
+%
+% \subsection{Subdivided Grids}
+%
+% Refer to \cref{grid_figure}.
 For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
@@
 Otherwise, we say that $P$ \emph{bends} at $v_i$.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c@{\hspace{2cm}}c}
-      \includegraphics{figs/grid-1} &
-      \includegraphics{figs/grid-2}
-    \end{tabular}
-  \end{center}
-  \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
-  \label{grid_figure}
-\end{figure}
-
-% An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.
+%
+% \begin{figure}
+%   \begin{center}
+%     \begin{tabular}{c@{\hspace{2cm}}c}
+%       \includegraphics{figs/grid-1} &
+%       \includegraphics{figs/grid-2}
+%     \end{tabular}
+%   \end{center}
+%   \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
+%   \label{grid_figure}
+% \end{figure}
+%
+% % An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.
 The \emph{vertex boundary} of $G_{a\times b}$ is $\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.
@@
 The \emph{boundary} of $G_{a\times b}$ is the union of the edge and vertex boundaries.
-
-For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.
-We will work extensively with the $1$-subdivision $G^{\onesub}_k$ of $G_k$.
-In this case, we define $\partial G^{\onesub}_k:= \partial G_k\cup N_{G^{\onesub}_k}[\partial G_k]$.\todo{Define open and closed neighbourhods $N_G(\cdot)$ and $N_G[\cdot]$.}
-
-\begin{lem}\label{seven_by_seven}
-  Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_7$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
-  \begin{compactenum}[(i)]\setcounter{enumi}{0}
-    \item Each vertex is $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
-    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
-    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
-  \end{compactenum}
-  Then for any pair of boundary vertices $s$ and $t$, $G_{10}$ contains a path $P$ that
-  \begin{compactenum}[(a)]
-    \item begins at $s$ and ends at $t$;
-    \item contains every vertex in $I\cup L\cup Q$;
-    \item goes straight through each vertex in $I$; and
-    \item bends at each vertex in $L$.
-    % \item contains no boundary vertices or edges except $s$ and $t$.
-  \end{compactenum}
-\end{lem}
-
-\begin{proof}
-  Case analysis(?).
-\end{proof}
-
-It will be helpful to take some natural definitions for $G_k$ and extend them to $G^{\onesub}_k$.
+%
+%
+% % Let $G^{\onesub}_k$ denote the $1$-subdivision of $G_k$.
+%
+% A \emph{simple $k\times k$ pseudogrid} is any graph that can be obtained from $G^{\onesub}_k$ by replacing some degree-$4$ vertices of $G_k$ with small binary trees.
+Specifically, we may replace a degree-$4$ vertex with a tree whose leaves are the vertices in $N_{G^{\onesub}_k}(v)$ and whose two internal nodes $v_a$ and $v_b$ are each adjacent to two leaves and to each other.
+%
+% Let $G_0$ be a simple $k\times k$ pseudogrid and observe that there is a natural mapping $f:E(G)\to E(G_k)\cup V(G_k)$ which is fairly obvious except that any edge $v_av_b\in E(G_0)$ introduced when replacing a degree-$4$ vertex $v$ in $G_k$ has $f(v_av_b):= v$.
+%
+% \begin{lem}\label{one_path}
+%    Let $S:=\bigcup_{i=1}^r S_i$ be a subset of $E(G_0)$ with the following properties:
+%    \begin{compactenum}[(a)]
+%      \item $|S_i|\le 2$ for each $i\in\{1,\ldots,r\}$; and
+%      \item for each distinct $i,j\in\{1,\ldots,r\}$, $\dist_{G_k}(\bigcup_{e\in S_i}f(e), \bigcup_{e\in S_j}f(e))\ge C$.\todo{Choose the value of $C$}
+%    \end{compactenum}
+%    Then $G_0$ contains a path that contains each edge in $S$.
+% \end{lem}
+%
+% \begin{proof}
+%   To do.
+% \end{proof}
+%
+%
+% % For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.  We will work extensively with the $1$-subdivision $G^{\onesub}_k$ of $G_k$.
+% %
+% %
+% % In this case, we define $\partial G^{\onesub}_k:= \partial G_k\cup N_{G^{\onesub}_k}[\partial G_k]$.\todo{Define open and closed neighbourhods $N_G(\cdot)$ and $N_G[\cdot]$.}
+% %
+% % \begin{lem}\label{seven_by_seven}
+% %   Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_7$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
+% %   \begin{compactenum}[(i)]\setcounter{enumi}{0}
+% %     \item Each vertex is $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
+% %     \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
+% %     \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
+% %   \end{compactenum}
+% %   Then for any pair of boundary vertices $s$ and $t$, $G_{10}$ contains a path $P$ that
+% %   \begin{compactenum}[(a)]
+% %     \item begins at $s$ and ends at $t$;
+% %     \item contains every vertex in $I\cup L\cup Q$;
+% %     \item goes straight through each vertex in $I$; and
+% %     \item bends at each vertex in $L$.
+% %     % \item contains no boundary vertices or edges except $s$ and $t$.
+% %   \end{compactenum}
+% % \end{lem}
+% %
+% % \begin{proof}
+% %   Case analysis(?).
+% % \end{proof}
+% %
+% % It will be helpful to take some natural definitions for $G_k$ and extend them to $G^{\onesub}_k$.
 To do this, we associate each vertex $v$ of $G^{\onesub}_k$ with a vertex $v^{\leftup}$ of $G_k$.
@@
 In other words, $s^{\leftup}$ is the vertex of $G_k$ that is directly below or the left of $s$.
-
-For two vertices $v_1,v_2\in V(G^{\onesub}_k)$, with $v^{\leftup}_1:=(x_1,y_1)$ and $v^{\leftup}_2:=(x_2,y_2)$ we define the $L_\infty$-distance between $v_1$ and $v_2$ as $\|v_1v_2\|_\infty = \max\{|x_1-x_2|, |y_1-y_2|\}$.
+% %
+% % For two vertices $v_1,v_2\in V(G^{\onesub}_k)$, with $v^{\leftup}_1:=(x_1,y_1)$ and $v^{\leftup}_2:=(x_2,y_2)$ we define the $L_\infty$-distance between $v_1$ and $v_2$ as $\|v_1v_2\|_\infty = \max\{|x_1-x_2|, |y_1-y_2|\}$.
 Refer to \cref{ball_figure}.
@@
 Because it comes up in several places, it is worth noting that $|B_{\beta}(v)|\le 4(\beta+1)^2$, and this bound is tight for each $v\in\{\beta+1,\ldots,\beta-r-1\}\times\{\beta+1,\ldots,b-\beta-1\}$.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{figs/grid-3}
-  \end{center}
-  \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
-  \label{ball_figure}
-\end{figure}
-
-\begin{lem}\label{four_points}
-  Let $P$ be a set of at most $3$ vertices of $G^{\onesub}_k$.
+% %
+% % \begin{figure}
+% %   \begin{center}
+% %     \includegraphics{figs/grid-3}
+% %   \end{center}
+% %   \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
+% %   \label{ball_figure}
+% % \end{figure}
+% %
+% %
+% % \begin{lem}\label{four_points}
+% %   Let $P$ be a set of at most $3$ vertices of $G^{\onesub}_k$.
 Then there exists a set $S$ of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.\todo{Define the interior of a ball.}
-\end{lem}
-
-\begin{proof}
-  Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.
+% % \end{lem}
+% %
+% % \begin{proof}
+% %   Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.
 Let $X:=\bigcup_{i=1}^t \{x_i\bmod 7,(x_{i+1})\bmod 7\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 7,(y_i+1)\bmod 7\}$.
@@
 Therefore $S$ is a set of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.
-\end{proof}
-
-\begin{lem}
-  Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $18$ contains more than $3$ vertices in $P$.
+% % \end{proof}
+% %
+% % \begin{lem}
+% %   Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $18$ contains more than $3$ vertices in $P$.
 Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
-\end{lem}
-
-\begin{proof}
-  Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$.
+% % \end{lem}
+% %
+% % \begin{proof}
+% %   Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$.
 Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $18$.
 Therefore each component of $H$ contains at most $3$ vertices.
-
-  For each component $C$ of $H$, apply \cref{four_points} to $V(C)$ and keep only those balls in $S$ that contain at least one vertex of $V(C)$.
+% %
+% %   For each component $C$ of $H$, apply \cref{four_points} to $V(C)$ and keep only those balls in $S$ that contain at least one vertex of $V(C)$.
 This produces a set $S_C$ of disjoint balls, each of radius $3$, that each contain some vertex in $V(C)$ in their interior.
 Any two balls $B_{C_1}$ and $B_{C_2}$ obtained from different components of $H$ are disjoint because the first ball contains a vertex $p_1\in P$ and the second ball contains a vertex $p_2\in P$ such that $\|p_1p_2\|\infty > 12$, but the balls $s_1$ and $s_2$ each have radius $3$.
-\end{proof}
-
-\begin{lem}\label{one_path_enchilada}
-  Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
-  \begin{compactenum}[(1)]\setcounter{enumi}{-1}
-    \item Any ball of radius 18 contains at most $3$ elements of $I\cup L\cup Q$.
-    \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
-    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
-    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
-  \end{compactenum}
-  Then $G^{\onesub}_k$ contains a path $P$ that
-  \begin{compactenum}[(a)]
-    \item contains all the vertices in $I\cup L\cup Q$;
-    \item goes straight through each vertex in $I$; and
-    \item turns at each vertex in $L$.
-  \end{compactenum}
-\end{lem}
-
-\begin{proof}
-  Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$.
+% % \end{proof}
+% %
+% %
+% % \begin{lem}\label{one_path_enchilada}
+% %   Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
+% %   \begin{compactenum}[(1)]\setcounter{enumi}{-1}
+% %     \item Any ball of radius 18 contains at most $3$ elements of $I\cup L\cup Q$.
+% %     \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
+% %     \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
+% %     \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
+% %   \end{compactenum}
+% %   Then $G^{\onesub}_k$ contains a path $P$ that
+% %   \begin{compactenum}[(a)]
+% %     \item contains all the vertices in $I\cup L\cup Q$;
+% %     \item goes straight through each vertex in $I$; and
+% %     \item turns at each vertex in $L$.
+% %   \end{compactenum}
+% % \end{lem}
+% %
+% % \begin{proof}
+% %   Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$.
 Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.
@@
 Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in each ball that it intersects.
-\end{proof}
-
-\subsection{Approximate Voronoi Diagrams in $G^{\onesub}_k$}
-
-% \todo[inline]{This lemma is not true.
+% % \end{proof}
+% %
+% %
+% % \subsection{Approximate Voronoi Diagrams in $G^{\onesub}_k$}
+% %
+% %
+% % % \todo[inline]{This lemma is not true.
 Pick a center $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.
@@
 We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
-\begin{lem}\label{voronoi_mapping}
-  Let $\beta, k\in\N$, let $P\subseteq V(G^{\onesub}_k)$ be such that $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$, and define  $Q:=\bigcup_{p\in P} B_\beta(p)$.
+% % \begin{lem}\label{voronoi_mapping}
+% %   Let $\beta, k\in\N$, let $P\subseteq V(G^{\onesub}_k)$ be such that $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$, and define  $Q:=\bigcup_{p\in P} B_\beta(p)$.
 Then there exists a function $\vor:Q\to P$ such that
-  \begin{compactenum}[(i)]
-    \item \label{real_close} For each $p\in P$ and each $v\in B_{\beta/8-1}(p)$, $\vor(v)=p$.
-    \item \label{never_far} For each $v\in Q$, $\|v\vor(p)\|_\infty \le 7\beta/4$.
-    \item \label{three_in_ball} For any ball $B$ of radius $\beta/12$, $|\{\vor(v):v\in B\cap Q\}|\le 3$.
-    \item \label{twocolour_neighbourhood} For any $v\in Q$, $|\{\vor(v)\}\cup\{\vor(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
-  \end{compactenum}
-\end{lem}
-
-\begin{proof}
-  We first define a function $\vor':Q\to P$ that satisfies (\ref{real_close})--(\ref{three_in_ball}) and then show how local modifications can be used to obtain a function $\vor$ that also satisfies (\ref{twocolour_neighbourhood}).
-
-  Arbitrarily order the vertices of $P$ as $p_1,\ldots,p_{|P|}$.
+% %   \begin{compactenum}[(i)]
+% %     \item \label{real_close} For each $p\in P$ and each $v\in B_{\beta/8-1}(p)$, $\vor(v)=p$.
+% %     \item \label{never_far} For each $v\in Q$, $\|v\vor(p)\|_\infty \le 7\beta/4$.
+% %     \item \label{three_in_ball} For any ball $B$ of radius $\beta/12$, $|\{\vor(v):v\in B\cap Q\}|\le 3$.
+% %     \item \label{twocolour_neighbourhood} For any $v\in Q$, $|\{\vor(v)\}\cup\{\vor(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
+% %   \end{compactenum}
+% % \end{lem}
+% %
+% % \begin{proof}
+% %   We first define a function $\vor':Q\to P$ that satisfies (\ref{real_close})--(\ref{three_in_ball}) and then show how local modifications can be used to obtain a function $\vor$ that also satisfies (\ref{twocolour_neighbourhood}).
+% %
+% %   Arbitrarily order the vertices of $P$ as $p_1,\ldots,p_{|P|}$.
 For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+3\beta/4\}$, let $i(v):=\min\{i:p_i\in S(v)\}$, and let $\vor'(v)=p_{i(v)}$.
 In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $\beta/12$), and $\vor'(v)$ is the approximate nearest neighbour of $v$ of smallest index.
-
-  To see that $\vor'$ satisfies (\ref{real_close}), observe that, for each distinct $p,q\in P$ and $v\in B_{\beta/8-1}(p)$, the triangle inequality implies that $\|vq\|_\infty \ge \|pq\|_\infty-\|vp\|_\infty \ge \beta-\beta/8+1 > 7\beta/8$.
+% %
+% %   To see that $\vor'$ satisfies (\ref{real_close}), observe that, for each distinct $p,q\in P$ and $v\in B_{\beta/8-1}(p)$, the triangle inequality implies that $\|vq\|_\infty \ge \|pq\|_\infty-\|vp\|_\infty \ge \beta-\beta/8+1 > 7\beta/8$.
 On the other hand $d(v)\le \beta/8$, so $d(v)+3\beta/4\le 7\beta/8$.
 Therefore, $q\not\in S(v)$ for any $q\in P\setminus\{p\}$, so $\vor'(v)=p$.
-
-  That $\vor'$ satsifies (\ref{never_far}) follows immediately from the fact that each $v\in Q$, so $d(v)\le \beta$, and therefore $\|pv\|_\infty\le 7\beta/4$ for each $p\in S(v)$.
-
-  We now argue that $\vor'$ satisfies (\ref{three_in_ball}).
+% %
+% %   That $\vor'$ satsifies (\ref{never_far}) follows immediately from the fact that each $v\in Q$, so $d(v)\le \beta$, and therefore $\|pv\|_\infty\le 7\beta/4$ for each $p\in S(v)$.
+% %
+% %   We now argue that $\vor'$ satisfies (\ref{three_in_ball}).
 Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.
@@
 By the triangle inequality $\|v_iv_j\|_\infty \le \|v_ic\|_\infty+\|cv_j\|_\infty \le \beta/6$ for each  $i,j\in\{1,\ldots,4\}$.
-  Since $v_1\in Q$, $d(v)\le\beta$ so
-  \[
-    \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
-  \]
-  Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
-  \[
-    \|p_{i_2}v_2\|_\infty
-     < \|p_{i_1}v_2\|_\infty - 3\beta/4
-     \le \|p_{i_1}v_1\|_\infty + \beta/6 - 3\beta/4
-     \le \beta + \beta/5 \enspace .
-  \]
-  Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
-  \[
-     \|p_{i_2}v_3\|_\infty
-     < \|p_{i_2}v_3\|_\infty - 3\beta/4
-     \le \|p_{i_2}v_2\|_\infty + \beta/6 - 3\beta/4
-     < \beta + 2\beta/6 - 3\beta/4 \enspace .
-  \]
-  Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
-  \[
-    \|p_{i_4}v_4\|_\infty
-    < \|p_{i_3}v_4\|_\infty - 3\beta/4
-    \le \|p_{i_3}v_3\|_\infty + \beta/5 - 3\beta/4
-    < \beta + 3\beta/6 - 3\beta/2 = 0 \enspace ,
-  \]
-  which is clearly a contradiction.  This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{TODO: Optimize and cleanup. Finish proving (iii)}
-\end{proof}
-
-\subsection{Pseudogrids}
-
-A graph $H$ is an $a\times b$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $H$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_{a\times b}$.
+% %   Since $v_1\in Q$, $d(v)\le\beta$ so
+% %   \[
+% %     \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
+% %   \]
+% %   Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
+% %   \[
+% %     \|p_{i_2}v_2\|_\infty
+% %      < \|p_{i_1}v_2\|_\infty - 3\beta/4
+% %      \le \|p_{i_1}v_1\|_\infty + \beta/6 - 3\beta/4
+% %      \le \beta + \beta/5 \enspace .
+% %   \]
+% %   Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
+% %   \[
+% %      \|p_{i_2}v_3\|_\infty
+% %      < \|p_{i_2}v_3\|_\infty - 3\beta/4
+% %      \le \|p_{i_2}v_2\|_\infty + \beta/6 - 3\beta/4
+% %      < \beta + 2\beta/6 - 3\beta/4 \enspace .
+% %   \]
+% %   Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
+% %   \[
+% %     \|p_{i_4}v_4\|_\infty
+% %     < \|p_{i_3}v_4\|_\infty - 3\beta/4
+% %     \le \|p_{i_3}v_3\|_\infty + \beta/5 - 3\beta/4
+% %     < \beta + 3\beta/6 - 3\beta/2 = 0 \enspace ,
+% %   \]
+% %   which is clearly a contradiction.  This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{TODO: Optimize and cleanup. Finish proving (iii)}
+% % \end{proof}
+%
+% \subsection{Pseudogrids}
+%
+% A graph $H$ is an $a\times b$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $H$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_{a\times b}$.
 The following lemma connecting grid minors and pseudogrids is stated implicitly by \citet{kun.obrien.ea:polynomial}.
 We give a proof here because some of the details of the connection are required in order to prove our results.
-
-\begin{lem}\label{pseudogrid_minor}
-  If a graph $G$ contains a $G_{a\times b}$ grid minor then $G$ contain a subgraph $H$ that is an $a\times b$ pseudogrid.
-\end{lem}
-
-\begin{proof}
-  (This proof only uses the fact that $G_{a\times b}$ has maximum degree  $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
-  \begin{inparaenum}[(i)]
-    \item $B_x$ is non-empty and $H[B_x]$ is connected for each $x\in V(G_{a\times b})$;
-    \item $H/\mathcal{B}$ is isomorphic to $G_{a\times b}$ using the isomorphism given by $B_x\mapsto x$.
-  \end{inparaenum}
-  Let $(H,\mathcal{B})$ be chosen so that $|E(H)|$ is minimum, over all choices of $(H,\mathcal{B})$ that satisfy (i) and (ii).
-
-  An edge of $vw$ of $H$ is a \emph{border edge} if $v\in B_x$ and $w\in B_y$ for distinct $x,y\in V(G_{a\times b})$.
+%
+%
+% \begin{lem}\label{pseudogrid_minor}
+%   If a graph $G$ contains a $G_{a\times b}$ grid minor then $G$ contain a subgraph $H$ that is an $a\times b$ pseudogrid.
+% \end{lem}
+%
+% \begin{proof}
+%   (This proof only uses the fact that $G_{a\times b}$ has maximum degree  $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
+%   \begin{inparaenum}[(i)]
+%     \item $B_x$ is non-empty and $H[B_x]$ is connected for each $x\in V(G_{a\times b})$;
+%     \item $H/\mathcal{B}$ is isomorphic to $G_{a\times b}$ using the isomorphism given by $B_x\mapsto x$.
+%   \end{inparaenum}
+%   Let $(H,\mathcal{B})$ be chosen so that $|E(H)|$ is minimum, over all choices of $(H,\mathcal{B})$ that satisfy (i) and (ii).
+%
+%   An edge of $vw$ of $H$ is a \emph{border edge} if $v\in B_x$ and $w\in B_y$ for distinct $x,y\in V(G_{a\times b})$.
 For each $x\in V(G_{a\times b})$, let $\partial B_x$ denote the set of border edges incident on one vertex of $B_x$.
@@
 Since $H$ is edge-minimal, $|\partial B_x|=\deg_{G_{a\times b}}(x)\le 4$, for each $x\in V(G_{a\times b})$ and $T_x$ is a tree with exactly $\deg_{G_{a\times b}}(x)$ leaves.\todo{Check the preceding claim.}  We define $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$ as follows:
-\begin{compactenum}
-  \item \label{one_special} If $T_x$ has exactly one node $v$ of degree at least three, then $P_x:=\{v\}$.
-  \item \label{two_degree_three} If $T_x$ has two nodes $v$ and $w$ each of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
-  \item \label{just_a_path} Otherwise $T_x$ is a path $v_0,\ldots,v_\ell$ for some $\ell\ge 2$ and $P_x:=v_1$.
-\end{compactenum}
-It is straightforward to verify that, with this assignment $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$, the quotient graph $H/\mathcal{P}$ is a subdivision of $G_{a\times b}$, so that $H$ is indeed an $a\times b$ pseudogrid.
-\end{proof}
-
-The second case (\ref{two_degree_three}) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.
+% \begin{compactenum}
+%   \item \label{one_special} If $T_x$ has exactly one node $v$ of degree at least three, then $P_x:=\{v\}$.
+%   \item \label{two_degree_three} If $T_x$ has two nodes $v$ and $w$ each of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
+%   \item \label{just_a_path} Otherwise $T_x$ is a path $v_0,\ldots,v_\ell$ for some $\ell\ge 2$ and $P_x:=v_1$.
+% \end{compactenum}
+% It is straightforward to verify that, with this assignment $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$, the quotient graph $H/\mathcal{P}$ is a subdivision of $G_{a\times b}$, so that $H$ is indeed an $a\times b$ pseudogrid.
+% \end{proof}
+%
+% % The second case (\ref{two_degree_three}) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.
 This case is especially important in our proof because it corresponds to a case in which some of the vertices of $P_x$ are shared by the $i$th column and the $j$th row of the pseudogrid.
@@
 This motivates the following definitions (see \cref{straight_or_bent}):
-\begin{compactenum}
-  \item We say that the $x$ is \emph{straight} if the path from $B_{(i-1,j)}$ to $B_{i+1,j}$ in $T_x$ contains all the vertices of $P_x$.
-  \item Otherwise, we say that $x$ is \emph{bent}.
-\end{compactenum}
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{c@{\hspace{1cm}}c}
-      \includegraphics{figs/bx-paths-1.pdf} & \includegraphics{figs/bx-paths-3} \\
-      (a) & (b)
-    \end{tabular}
-  \end{center}
-  \caption{The vertex $x:=(i,j)$ is (a)~straight if the path from $B_{(i,j-1)}$ to $B_{(i+1,j)}$ uses all vertices of $P_x$ or (b)~bent otherwise.}
-  \label{straight_or_bent}
-\end{figure}
-Note that, if $x$ is straight, then the path from $B_{i,j-1}$ to $B_{i,j+1}$ in $T_x$ also contains all the vertices of $P_x$.
+% % \begin{compactenum}
+% %   \item We say that the $x$ is \emph{straight} if the path from $B_{(i-1,j)}$ to $B_{i+1,j}$ in $T_x$ contains all the vertices of $P_x$.
+% %   \item Otherwise, we say that $x$ is \emph{bent}.
+% % \end{compactenum}
+% % \begin{figure}
+% %   \begin{center}
+% %     \begin{tabular}{c@{\hspace{1cm}}c}
+% %       \includegraphics{figs/bx-paths-1.pdf} & \includegraphics{figs/bx-paths-3} \\
+% %       (a) & (b)
+% %     \end{tabular}
+% %   \end{center}
+% %   \caption{The vertex $x:=(i,j)$ is (a)~straight if the path from $B_{(i,j-1)}$ to $B_{(i+1,j)}$ uses all vertices of $P_x$ or (b)~bent otherwise.}
+% %   \label{straight_or_bent}
+% % \end{figure}
+% % Note that, if $x$ is straight, then the path from $B_{i,j-1}$ to $B_{i,j+1}$ in $T_x$ also contains all the vertices of $P_x$.
 Furthermore, if $x$ is bent, then, for each $\overline\i,\overline\j\in\{-1,1\}$, the path from $B_{i,j+\overline\j}$ to $B_{i+\overline\i, j}$ in $T_x$ contains all the vertices of $P_x$.
-
-\todo[inline]{We can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}
-
-\begin{lem}\label{pseudogrid_lower_bound}
-  %For any $k\times k$ pseudogrid $H$
-  Let $H$ be a $k\times k$ pseudogrid and $\mathcal{P}:=(P_x:x\in V(G_{k}))$ be such that
-  \begin{inparaenum}[(i)]
-      \item for each $x\in V(G_k)$, $H[P_x]$ is a path of length at most $r$ and
-      \item $H/\mathcal{P}$ is a $\le\!\! r$-subdivision of $G_k$.
-  \end{inparaenum}
-  Then $\chilin(G)\in\Omega(k/r)$.
-\end{lem}
-
-\begin{proof}
-  This proof uses two large integer constants $\alpha$ and $\beta$ and two small positive real constants $\epsilon$ and $\delta$.
+% %
+% %
+% % \todo[inline]{We can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}
+%
+% \begin{lem}\label{pseudogrid_lower_bound}
+%   For any $k\times k$ pseudogrid $H$ $\chilin(G)\in\Omega(k)$.
+%   % Let $H$ be a $k\times k$ pseudogrid and $\mathcal{P}:=(P_x:x\in V(G_{k}))$ be corresponding partition of $V(H)$ into induced paths.
+%   % \begin{inparaenum}[(i)]
+%   %     \item for each $x\in V(G_k)$, $H[P_x]$ is a path of length at most $r$ and
+%   %     \item $H/\mathcal{P}$ is a $\le\!\! r$-subdivision of $G_k$.
+%   % \end{inparaenum}
+%   % Then
+% \end{lem}
+%
+% \begin{proof}
+%   This proof uses two large integer constants $\alpha$ and $\beta$ and two small positive real constants $\epsilon$ and $\delta$.
 The relationships between these constants are that $\alpha=\Theta(r\beta^2)$  $\epsilon=\Theta(1/\alpha)$.
@@
 We will show that $H$ contains a path $P$ which has no centre with respect to $\varphi$.
-
-  First we introduce some notations for vertices of the subdivided grid $G^{\onesub}_k$ to their corresponding paths in the pseudogrid $H$.
+%
+%   First we introduce some notations for vertices of the subdivided grid $G^{\onesub}_k$ to their corresponding paths in the pseudogrid $H$.
 For each vertex $v$ of $G_k$ there is a path $P_v$ in $\mathcal{P}$ (possibly consisting of one vertex) such that the vertex $v$ appears in $H'$ as a result of contracting $P_v$ in $H$.
@@
 Thus, each subdivision vertex $s$ of $G^{\onesub}_k$ is associated with a (possibly empty) \emph{colour set} $\Phi(s):=\{\varphi(w):w\in P^-_e\}$, where $P^-_e$ denotes the (possibly empty) set of internal vertices on the path $P_e$.
-
-  For each colour $c\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $c$ as the number of vertices $G^{\onesub}_k$ that include $c$ in their colour set.
+%
+%   For each colour $c\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $c$ as the number of vertices $G^{\onesub}_k$ that include $c$ in their colour set.
 Fix some large integer constant $\alpha$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity less than $\alpha$ and \emph{frequent} otherwise.
-
-  \begin{clm}\label{all_frequent}
-    There exists an integer $k'\ge (1-\epsilon)k$ such that
-    $H$ contains a $k'\times k'$ pseudogrid $H'$ in which every colour in $\{\varphi(v):v\in V(H')\}$ is frequent.
-  \end{clm}
-
-  \begin{proof}[Proof of \cref{all_frequent}]
-    This proof uses the notion of \emph{removing} a row or column of a pseudogrid.
+%
+%   \begin{clm}\label{all_frequent}
+%     There exists an integer $k'\ge (1-\epsilon)k$ such that
+%     $H$ contains a $k'\times k'$ pseudogrid $H'$ in which every colour in $\{\varphi(v):v\in V(H')\}$ is frequent.
+%   \end{clm}
+%
+%   \begin{proof}[Proof of \cref{all_frequent}]
+%     This proof uses the notion of \emph{removing} a row or column of a pseudogrid.
 Removing a row is defined as follows (removing a column is similar): For each vertex $v$ of $G^{\onesub}_k$ in the relevant row, remove all edges of $P_v$ from $H$.
@@
 The result is pseudogrid with one less row.
-
-    While there exists some infrequent colour $c\in\{1,\ldots,C\}$, find a row or column of $G^{\onesub}_k$ that contains a vertex with $c$ in its colour set and remove that row or column.
+%
+%     While there exists some infrequent colour $c\in\{1,\ldots,C\}$, find a row or column of $G^{\onesub}_k$ that contains a vertex with $c$ in its colour set and remove that row or column.
 Doing this exhaustively results in the removal of at most $(\alpha-1) C \le \epsilon\alpha k$ rows and at most $\epsilon\alpha k$ columns of $H$.
@@
 This completes the proof of \cref{all_frequent}.
-  \end{proof}
-
-  Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=\Theta(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.
+%   \end{proof}
+%
+%   Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=\Theta(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.
 To avoid overloading notation, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$ in $H$.
 (The remainder of the proof does not depend on the values of $k$ or $C$.)
-
-  % Colour sets turn out to be problematic, so we would like to assign a single colour $c\in\Phi(v)$ to each $v\in G^{\onesub}_k$.
+%
+%   % Colour sets turn out to be problematic, so we would like to assign a single colour $c\in\Phi(v)$ to each $v\in G^{\onesub}_k$.
 To do this, consider the bipartite graph $X$ with parts $A:=\{1,\ldots,C\}$ and $B:=V(G^{\onesub}_k)$ and which contains the edge $cv$ if and only if $c\in\Phi(v)$.
@@
 S
-
-  Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is vertex of $G^{\onesub}_k$ that contains $c$ in its colour set.
+%
+%   Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is vertex of $G^{\onesub}_k$ that contains $c$ in its colour set.
 These pairs are carefully chosen so that we will be able to construct a simple path $P_k$ in $G^{\onesub}_k$ that contains every vertex in $X:=\bigcup_{c\in\{1,\ldots,C\}} \{x_c,y_c\}$.
@@
 This will ensure that the corresponding path in $H$ contains at least two vertices of each colour.
-
-  \begin{enumerate}[{Stage} 1:]
-    \item We begin this stage by defining each vertex $G^{\onesub}_k$ as being \emph{unmarked}.
+%
+%   \begin{enumerate}[{Stage} 1:]
+%     \item We begin this stage by defining each vertex $G^{\onesub}_k$ as being \emph{unmarked}.
 Next we consider each $c\in\{1,\ldots,C\}$ in turn.
 If $G^{\onesub}_k$ contains a pair $(x_c,y_c)$ where
-    \begin{inparaenum}[(i)]
-      \item each of $x_c$ and $y_c$ is unmarked;
-      \item $\|x_cy_c\|_\infty \ge \beta$; and
-      \item $c\in \Phi(x_c)$ and $c\in \Phi(y_c)$;
-    \end{inparaenum}
-    then we say that Stage~1 \emph{succeeeds} for colour $c$ and we mark each vertex in $B_{\beta}(x_c)$ and $B_\beta(y_c)$.
-
-    If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $x_c$ and $y_c$ undefined for now and say that Stage~1 \emph{fails} for colour $c$.
+%     \begin{inparaenum}[(i)]
+%       \item each of $x_c$ and $y_c$ is unmarked;
+%       \item $\|x_cy_c\|_\infty \ge \beta$; and
+%       \item $c\in \Phi(x_c)$ and $c\in \Phi(y_c)$;
+%     \end{inparaenum}
+%     then we say that Stage~1 \emph{succeeeds} for colour $c$ and we mark each vertex in $B_{\beta}(x_c)$ and $B_\beta(y_c)$.
+%
+%     If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $x_c$ and $y_c$ undefined for now and say that Stage~1 \emph{fails} for colour $c$.
 Note that Stage~1 can only fail for $c$ if $G^{\onesub}_k$ contains fewer than $(2\beta+1)^2$ unmarked vertices with $c$ in their colour sets.
 In particular, for $\alpha > (2\beta+1)^2$, Stage~1 succeeds for $c=1$.
-
-    \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeds and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 fails.
+%
+%     \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeds and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 fails.
 Let $R_1:=\bigcup_{c\in S}\{x_c,y_c\}$ be the set of vertices and edges chosen in $R_1$.
 By construction, $\min\{\|pq\|_\infty: \{p,q\}\in \binom{R_1}{2}\}\ge \beta$.
-
-    As mentioned above, for sufficiently large $\alpha$ (relative to $\beta$), the set $S$ is non-empty and therefore $R_1$ is non-empty.
+%
+%     As mentioned above, for sufficiently large $\alpha$ (relative to $\beta$), the set $S$ is non-empty and therefore $R_1$ is non-empty.
 Let $P:=R_1$ and let $Q:=\bigcup_{p\in P} B_\beta(p)$ and apply \cref{voronoi_mapping} to obtain the mapping $\vor:Q\to P$.
@@
 The edge set of $I$ contains an edge $cx$ precisely when there is a marked vertex $z$ of $G^{\onesub}_k$ such that $c\in\Phi(z)$ and $\vor(z)=x$.
-
-    First observe that each vertex $x$ of $G^{\onesub}_k$ corresponds to a path $P_x$ in $H$ that contains at most $r$ vertices.\footnote{For some subdivision vertices of $G^{\onesub}_k$, $P_x$ may be empty.}  This implies that $|\Phi(x)|\le r$ for each $x\in V(G^{\onesub}_k)$.
+%
+%     First observe that each vertex $x$ of $G^{\onesub}_k$ corresponds to a path $P_x$ in $H$ that contains at most $r$ vertices.\footnote{For some subdivision vertices of $G^{\onesub}_k$, $P_x$ may be empty.}  This implies that $|\Phi(x)|\le r$ for each $x\in V(G^{\onesub}_k)$.
 Next observe that, for each $p\in R_1$, \cref{voronoi_mapping}(\ref{never_far}) implies that $|\vor^{-1}(p)|\le|B_{7\beta/4}(p)|\le (7\beta/2+1)^2$.
 This implies that $\deg_I(p) \le r(7\beta/2+1)^2$ for each $p\in R_1$.
-
-    This implies that, for any $A_0\subseteq\overline{S}$,
-    \[
-        |N_I(A_0)| \ge \frac{\alpha|A_0|}{r(7\beta/2+1)^2} \ge 2|A_0| \enspace .
-    \]
-    for $\alpha\ge 2r(7\beta/2+1)^2$.
+%
+%     This implies that, for any $A_0\subseteq\overline{S}$,
+%     \[
+%         |N_I(A_0)| \ge \frac{\alpha|A_0|}{r(7\beta/2+1)^2} \ge 2|A_0| \enspace .
+%     \]
+%     for $\alpha\ge 2r(7\beta/2+1)^2$.
 Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
@@
 In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
-  \end{enumerate}
-
-  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \Phi(x_c)\cap \Phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
-  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.
-
-  \todo[inline]{There is a technical condition missing here.
+%   \end{enumerate}
+%
+%   Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \Phi(x_c)\cap \Phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
+%   Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.
+%
+%   \todo[inline]{There is a technical condition missing here.
 To apply \cref{one_path_enchilada} we need the elements of $R$ to be distance at least $3$ from the boundary of $G^{\onesub}_k$.
 When counting multiplicity, we shouldn't count occurrences of a colour that are close to the boundary.}
-
-  Conditions~(\ref{never_far}) and (\ref{three_in_ball}) of \cref{voronoi_mapping} imply that each any ball of radius $\min\{\beta/8-1, \beta/12\}/2$ contains at most three elements of $R$.
+%
+%   Conditions~(\ref{never_far}) and (\ref{three_in_ball}) of \cref{voronoi_mapping} imply that each any ball of radius $\min\{\beta/8-1, \beta/12\}/2$ contains at most three elements of $R$.
 Setting $\beta= 12\cdot2\cdot18 = 436$ (so the conditions on $\alpha$ are satisified for $\alpha = r(7\beta/2+1)^2=2331729r$) gives a point set $R$ that satisfies the requirements for \cref{one_path_enchilada}.
@@
 Since $\varphi$ is an arbitrary $\epsilon k$-colouring of $H$, this implies that $\chilin(H)\ge \epsilon k = \Theta(k/\alpha)=\Theta(k/r)$.
-\end{proof}
-
-To finish the proof, we need some lemma which states that highly subdivided graphs don't have linear colouring numbers that are much smaller than mildly subdivided graphs.
+% \end{proof}
+%
+% To finish the proof, we need some lemma which states that highly subdivided graphs don't have linear colouring numbers that are much smaller than mildly subdivided graphs.
 It would be enough to prove the following.
-
-A \emph{refinement} $G'$ of a graph $G$ is a subdivision of $G$ in which each subdivision vertex is adjacent only to degree-$2$ vertices, i.e., $\deg_{G'}(v)=2$ for each $v\in N_{G'}(s)$ and each $s\in V(G')\setminus V(G)$.
-
-\begin{lem}\label{refinement_bound}
-  For any refinement $G'$ of any graph $G$, $\chilin(G') = \Omega(\chilin(G))$.
-\end{lem}
-
-In general, we know that if $G'$ is an $\ell$-subdivision of $G$ then
-\[
-    \log\ell \le \chilin(G') \le \chilin(G) + \log\ell \enspace .
-\]
-The lower bound comes from the fact that a path of length $\ell$ has a linear chromatic number at least $\log\ell$.
+%
+% A \emph{refinement} $G'$ of a graph $G$ is a subdivision of $G$ in which each subdivision vertex is adjacent only to degree-$2$ vertices, i.e., $\deg_{G'}(v)=2$ for each $v\in N_{G'}(s)$ and each $s\in V(G')\setminus V(G)$.
+%
+% \begin{lem}\label{refinement_bound}
+%   For any refinement $G'$ of any graph $G$, $\chilin(G') = \Omega(\chilin(G))$.
+% \end{lem}
+%
+% In general, we know that if $G'$ is an $\ell$-subdivision of $G$ then
+% \[
+%     \log\ell \le \chilin(G') \le \chilin(G) + \log\ell \enspace .
+% \]
+% The lower bound comes from the fact that a path of length $\ell$ has a linear chromatic number at least $\log\ell$.
 The upper bound comes by introducing $\log\ell$ additional colours that are used to colour the paths in $G'-G$.
 Of course, the upper bound holds even when $G'$ is a $\le\!\!\ell$-subdivision of $G$ and the lower bound holds as long as at least one edge of $G$ is subdivided $\ell$ or more times.
-
-Perhaps \cref{refinement_bound} is true even if we replace ``refinement'' with ``subdivision''.
+%
+% Perhaps \cref{refinement_bound} is true even if we replace ``refinement'' with ``subdivision''.
 As a test case, one could start with the $1$-subdivision $K^{\onesub}_k$ of the $k$-clique $K_k$.
 The following result is certainly true, as can be seen from a treedepth decomposition:
-
-\begin{lem}
-  For any subdivision $G'$ of any graph $G$, $\chicen(G') \ge \chicen(G)$.
-\end{lem}
+%
+% \begin{lem}
+%   For any subdivision $G'$ of any graph $G$, $\chicen(G') \ge \chicen(G)$.
+% \end{lem}
+%
+%
+%
+%
+%
+%
+%
+%
+%
 

### 2022-04-01 lin-vs-cen.tex
Making progress

@@
 \usepackage[noend]{algorithmic}
+\usepackage{bbm}  % needed for \mathbbm{1}
 
@@
 
+\DeclareMathOperator{\binomial}{binomial}
+
 \newrobustcmd{\onesub}{\mathord{\includegraphics{figs/one-sub}}}
@@
 
-\DeclareMathOperator{\VE}{VE}
+\DeclareMathOperator{\VE}{\mathit{VE}}
 
@@
 \maketitle
+\renewcommand{\E}{\mathbb{E}}
+\renewcommand{\Pr}{\mathbb{P}}
 
@@
 
-For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.  For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.
+% For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.
+For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.
 
-For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.
+% For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.
 For any $S\subseteq V(G)$, $N_G(s):=\bigcup_{v\in S}N_G(v)$ and $N_G[s]:=\bigcup_{v\in S}N_G[v]$.
@@
 
-For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.
+In this paper, all graphs are simple and undirected. For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$.
+
+% For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.
 When $\mathcal{P}$ consists of disjoint subsets of $V(G)$ that do not necessarily cover $V(G)$ then we use the convention that $G/\mathcal{P}:=G/\mathcal{P'}$ where $\mathcal{P}'$ is the completion of $\mathcal{P}$ obtained by introducing each missing element of $V(G)$ as singleton set.
@@
 
-A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.
+% A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.
 Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.
@@
 
-\todo[inline]{Surely this next lemma is well-known?}
-We make use of the following generalization of Hall's Marriage Theorem \cite{hall:on}:
-\begin{lem}\label{hall_vees}
-  Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.
-Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
+We make use of the following (polygamous) generalisation of Hall's Marriage Theorem, which follows from Hall's Marriage Theorem by replacing each vertex $x\in X$ with a $d$-tuple $x_1,\ldots,x_d$ each having the same neighbours as $x$.
+
+% \begin{lem}\label{normal_hall}
+%   Let $G$ be a bipartite graph with parts $X$ and $Y$ such that, for any $A\subseteq X$, $|N_G(A)|\ge |A|+1$.
+Then $G$ contains a subgraph $M$ (a matching) such that $\deg_M(x)=1$ for each $x\in X$ and $\deg_M(y)\le 1$ for each $y\in Y$.
+% \end{lem}
+
+\begin{cor}\label{d_hall}
+  Let $d\ge 1$ be an integer and let $G$ be a bipartite graph with parts $X$ and $Y$ such that, for any $A\subseteq X$, $|N_G(A)|\ge d|A|$.
+Then $G$ contains a subgraph $M$ such that $\deg_M(x)=d$ for each $x\in X$ and $\deg_M(y)\le 1$ for each $y\in Y$.
+\end{cor}
+
+\begin{lem}\label{weighted_lovasz}
+  Let $E:=\{E_1,\ldots,E_n\}$ be a set of events in some probability space $(\Omega,\Pr)$.
+For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq E$ be such that the event $E_i$ is mutually independent of $\Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:E\to(0,1)$ be such that
+  \[
+      \Pr(E_i) \le w(E_i)\cdot\prod_{E_j\in\Gamma_i}(1-w(E_j))  \enspace .
+  \]
+  Then $\Pr(\overline{E_1}\cap\cdots\cap\overline{E_n}) > 0$.
 \end{lem}
 
-\begin{proof}
-  Let $\Lambda$ be a subgraph of $G$ in which each vertex of $A$ has degree at most two, each vertex of $B$ has degree at most one, and that maximizes $\deg_\Lambda(A)$.
+% \begin{proof}
+%   \todo[inline]{Surely this is known}
+% \end{proof}
+%
+%
+% \todo[inline]{Surely this next lemma is well-known?}
+% We make use of the following generalization of Hall's Marriage Theorem \cite{hall:on}:
+% \begin{lem}\label{hall_vees}
+%   Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.
+Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
+% \end{lem}
+%
+% \begin{proof}
+%   Let $\Lambda$ be a subgraph of $G$ in which each vertex of $A$ has degree at most two, each vertex of $B$ has degree at most one, and that maximizes $\deg_\Lambda(A)$.
 If $\deg_{\Lambda}(A) = 2|\Lambda|$ then there is nothing to prove.
 Assume therefore, for the sake of contradiction,  that there exists $v_0\in A$ with $\deg_\Lambda(v_0) < 2$.
-
-  We say that a path $v_0,\ldots,v_m$ in $G$ is \emph{$\Lambda$-aternating} if $v_iv_{i+1}\in E(\Lambda)$ for all odd $i\in\{0,\ldots,m-1\}$ and $v_iv_{i+1}\not\in E(\Lambda)$ for all even $i\in\{0,\ldots,m\}$.
+%
+%   We say that a path $v_0,\ldots,v_m$ in $G$ is \emph{$\Lambda$-aternating} if $v_iv_{i+1}\in E(\Lambda)$ for all odd $i\in\{0,\ldots,m-1\}$ and $v_iv_{i+1}\not\in E(\Lambda)$ for all even $i\in\{0,\ldots,m\}$.
 Observe that any such path has even length so that it ends at a vertex $v_m\in A$.
 Otherwise, replacing the edges $\lfloor m/2\rfloor$ edges $\{v_1v_2,v_3v_4,\ldots,v_{m-2}v_{m-1}\}\in E(\Lambda)$ with the $\lceil m/2\rceil$ edges $\{v_0v_1,v_2v_3,\ldots,v_{m-1}v_m\}$ does not change $\deg_\Lambda(v)$ for any $v\in A_0\setminus\{v_0\}$ but does increase  $\deg_\Lambda(v_0)$, contradicting the assumption that $\Lambda$ maximizes $\deg_\Lambda(A)$.
-
-  Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
-  Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.
+%
+%   Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
+%   Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.
 Let $a$ be the number of edges of $\Lambda$ having one endpoint $A_0$ and one endpoint in $B_0$, so that $a=\deg_\Lambda(B_0)=|B_0|$.
@@
 This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.greater than
-\end{proof}
-
-\begin{rem}
-  The constant $2$ that appears twice in \cref{hall_vees} can be replaced with any positive integer $c$.
-\end{rem}
+% \end{proof}
+%
+% \begin{rem}
+%   The constant $2$ that appears twice in \cref{hall_vees} can be replaced with any positive integer $c$.
+% \end{rem}
 
@@
 
-Refer to \cref{grid_figure}.
+% Refer to \cref{grid_figure}.
 For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
@@
   \begin{compactenum}[(Q1)]
-    \item $p$ is adjacent to $\{(x-1,y), (x,y-1)\}$ and $q$ is adjacent to $(x+1,y),(x,y+1)$;
-    \item $p$ is adjacent to $\{(x,y+1), (x-1,y)\}$ and $q$ is adjacent to $(x,y-1),(x+1,y)$; or
-    \item $p$ is adjacent to $\{(x-1,y),(x+1,y)\}$ and $q$ is adjacent to $\{(x,y-1),(x,y+1)\}$.\label{q_iii}
+    \item \label{q_i} $p$ is adjacent to $\{(x-1,y), (x,y-1)\}$ and $q$ is adjacent to $(x+1,y),(x,y+1)$;
+    \item \label{q_ii} $p$ is adjacent to $\{(x,y+1), (x-1,y)\}$ and $q$ is adjacent to $(x,y-1),(x+1,y)$; or
+    \item \label{q_iii} $p$ is adjacent to $\{(x-1,y),(x+1,y)\}$ and $q$ is adjacent to $\{(x,y-1),(x,y+1)\}$.
   \end{compactenum}
@@
 For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.
-In this way, $\mathcal{P}:=\{V(P_v):v\in V(G_{a\times b})\}\cup\{V(P_{vw}):vw\in E(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.
-We call $\mathcal{P}$ the \defin{grid-partition} of $G$ and we denote it by $(P_\mu:\mu\in V(G_{a\times b})\cup E(G_{a\times b}))$.
-This defines a mapping $\delta:V(G)\to V(G_{a\times b})\cup E(G_{a\times b})$ where $u\in V(P_{\delta(u)})$ for each $u\in v(G)$.
+In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.
+We call $\mathcal{P}$ the \defin{grid-partition} of $G$.
+% and we denote it by $(P_\mu:\mu\in \VE(G_{a\times b}))$.
+% This defines a mapping $\delta:V(G)\to V(G_{a\times b})\cup E(G_{a\times b})$ where $u\in V(P_{\delta(u)})$ for each $u\in v(G)$.
 
@@
 However, for $i\in\{2,\ldots,a-1\}$ $R$ may or may not contain $V(P_{v_i})$.
-In particular, if $P_{v_i}$ was created using (Q\ref{q_iii}) then $R$ does not contain any internal vertices in $P_{v_i}$.
+In particular, if $P_{v_i}$ was created using (Q\ref{q_iii})\todo{Fix this!} then $R$ does not contain any internal vertices in $P_{v_i}$.
 Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
@@
 As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.
-If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$) then we repeatedly remove degree-$1$ vertices until none remain.
+If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_a$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove degree-$1$ vertices until none remain.
 If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.
@@
 
-Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k\times k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
-For any $\mu\in\VE(G_{k\times k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\phi(v):v\in V(P_\mu)\}$.
-For any colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-2}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)$.
-For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-2}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-2}(\alpha)$.
+For convenience, let $G_k:=G_{k\times k}$.
+Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
+For any $\mu\in\VE(G_{k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\varphi(v):v\in V(P_\mu)\}$.
+For any colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-1}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)\}$.
+For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-1}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-1}(\alpha)$.
 
@@
 
-\begin{lem}
-  Let $0 < \epsilon < 1$, let $k\ge 1/\epsilon$ be an integer, let $G$ be a $k\times k$ pseudogrid and let $\varphi:V(G)\to\{1,\ldots,\lfloor\epsilon k\floor\}$ be a vertex colouring of $G$.
-  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ that has a grid-partition $\mathcal{P}':=(P'_\mu:\mu\in \VE(G_{k'\times k'})$ such that
-  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge |A|/2\epsilon$.
+\begin{lem}\label{only_frequent}
+  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/d$ colours.
+  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - d|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
+  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d|A|$.
 \end{lem}
@@
   The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.
-If there exists no $A\subseteq\{1,\ldots,\lfloor\epsilon k\rfloor\}$ with $\varphi_{\mathcal{P}}^{-2}(A) < |A|/2\epsilon$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
-Now assume that such a set $A$ exists.
-
-  We will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le |A|/2\epsilon$.
-For each $v:=(i,j)\in V(G_{k\times k})$ with $A\cap\varphi(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
-For each edge $vw\in E(G_{k\times k})$ with $A\cap\phi()\alpha\in\Phi(vw)$, we include the row of $G$ that contains $P_{vw}$ in $R$ or we include the column of $G$ that contains $P_{vw}$ in $C$.
-Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
-Now we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-|A|/2\epsilon$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
+If there exists no $A\subseteq \varphi(V(G))$ with $\varphi_{\mathcal{P}}^{-1}(A) < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
+Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|$ to eliminate all vertices with colours in $A$, as follows:
+  \begin{compactitem}
+    \item For each $v:=(i,j)\in V(G_{k})$ with $A\cap\varphi(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
+    \item For each edge $vw\in E(G_{k\times k})$ with $A\cap\phi()\alpha\in\Phi(vw)$, we include the row of $G$ that contains $P_{vw}$ in $R$ or we include the column of $G$ that contains $P_{vw}$ in $C$.
+    \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
+  \end{compactitem}
+  Now we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
 In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.
 
-  \todo[inline]{Now apply induction......}
+  Now apply induction on $G_0$ to get a $k'\times k'$ pseudogrid with
+  \[  k'\ge k_0-d|\varphi(V(G_0))|  \ge k-d|A|-d|\varphi(V(G_0))|= k - d|\varphi(V(G))|
+  \]
+  that satisfies the conditions of the lemma.
 \end{proof}
 
-\begin{lem}
-  Let $C>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(G)$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge C\cdot|A|+1$.
-Then there exists a partial colouring $\phi:\VE(G_{k\times k})\to\{1,\ldots,c,\perp\}$ with the following properties:
+\begin{lem}\label{one_colour_per_object}
+  Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d\cdot|A|$.
+Then there exists a (partial) colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
   \begin{compactenum}[(i)]
-    \item For each $\mu\in\VE(G_{k\times k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
-    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k\times k}):\phi(\mu)=\alpha\}|\ge C$.
+    \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
+    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
   \end{compactenum}
@@
 \begin{proof}
-  Use Hall's Theorem to find $C$ matchings such that each colour $\alpha$ appears in all $C$ matchings and each element of $\VE(G_{k\times k})$ appears in at most one matching.
+  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(G_k)$ and edge set
+  \[
+    \{ (\alpha,\mu)\in X\times Y: \alpha\in\varphi(\mu) \}
+  \]
+  By \cref{d_hall}, $H$ contains a subgraph $M$ with $\deg_M(x)=d$ for each $\alpha\in X$ and $\deg_M(y)\le 1$ for each $\mu\in Y$.
+For each edge $\alpha\mu$ in $M$, set $\phi(\mu):=\alpha$.
+This defines $\phi(\mu)$ for any $\mu\in X$ with $\deg_M(\mu)=1$.
+For each $\mu\in X$ with $\deg_M(\mu)=0$ set $\phi(\mu):=\perp$.
 \end{proof}
 
-\subsection{Finding Paths Through Well-Separated Vertex/Edge Pairs}
+\subsection{Finding Paths Through Well-Separated Pairs}
+
+Next we show that, given a sufficiently ``well-separated'' set $S$ of pairs of vertices in $G$, we can always find a path in $G$ that contains every vertex in $S$.
+For this, we need some definitions of boxes in $G$ and in $G_k$.
 
 The \defin{$r$-box} centered at a vertex $v:=(i,j)$ of $G_{a\times b}$ is defined as
-\[  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{a,b}) \enspace . \]
+\[
+  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{a,b}) \enspace .
+\]
 The \defin{$r$-box} centered at an edge $vw$ of $G_{a\times b}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.
@@
 Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{a\times b}[B_r(\mu)]$.
-We extend these definitions to vertices of an $a\times b$ pseudogrid $G$ so that
+We extend these definitions to vertices of an $a\times b$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G)\}$.
+For any $v\in V(G)$, define
 \[
-   \tilde{B}_r(v) := \bigcup_{\mu\in V(G_r(\delta(v)) \cup E(G_r(\delta(v))} V(P_\mu)
+   \tilde{B}_r(v) := \bigcup_{\mu\in \VE(G_r(\mu_v))} V(P_\mu)
 \]
-and $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
+where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.
+For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
 
@@
 
-\begin{lem}
+\begin{lem}\label{pick_up_everything}
   Let $r$ be an integer and let $S\subseteq V(G)$ be such that $|\tilde{B}_{r}(v)\cap S|=|\tilde{B}_{2r}\cap S|\le 2$ for each $v\in S$.
@@
 
+\begin{lem}\label{packing_lemma}
+  Some packing lemma
+\end{lem}
+
 \subsection{Finding an Uncentered Path}
 
+Next we show how, given a colouring like that guaranteed by \cref{only_frequent}, we can find a set of vertices that hit every colour at least twice and that are compatible with \cref{pick_up_everything}.
+
 \begin{lem}
-  Let $G$ be a $k\times k$ pseudogrid and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a colouring of $G$ such that, for each $\alpha\in\{1,\ldots,C\}$, the set $S_\alpha:=\{\mu\in V(G_{k\times k})\cup E(G_{k\times k}): \alpha\in\Phi(\mu)\}$ has size at least $C$.
-Then there exists $S\subseteq V(G_{k\times k})\cup E(G_{k\times k})$ such that
+  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.\todo{Use a specific value of $d$}.
+  Then there exists $S\subseteq V(G)$ such that
   \begin{compactenum}[(i)]
-    \item \label{hits_both} $|S\cap S_\alpha|\ge 2$ for each $\alpha\in\{1,\ldots,c\}$; and
-    \item \label{spread_out} $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
+    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|\ge 2$; and
+    \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
   \end{compactenum}
@@
 \begin{proof}
-  We construct $S$ using a two round process.  We start with an initially empty set $S_1$ and repeat the following for each $\alpha= 1,\ldots,c$:
-  If there exists distinct $\mu_1,\mu_2\in S_{\alpha}$ such that
+  We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi$ of $\VE(G_k)$.
+Observe that it is now sufficient to find $Q\subset\VE(G_k)$ such that $|\{\mu\in Q:\phi(v)=\alpha\}|\ge 2$ and $|B_r(\mu)\cap Q|\le 2$ for each $\alpha\in\varphi(V(G))$ and each $\mu\in Q$.
+Indeed, with such a $Q$ we can obtain $S$ by taking one vertex of colour $\phi(\mu)$ from $P_\mu$ for each $\mu\in Q$.
+
+  We construct $Q$ using a two round process.  We start with an initially empty set $Q_1$ and repeat the following for each $\alpha\in\varphi(V(G))$:
+  If there exists distinct $\mu_1,\mu_2\in\VE(G_k)$ such that
   \begin{compactenum}[(a)]
-    \item $\alpha\in\Phi(\mu_1)\cap\Phi(\mu_2)$;
+    \item $\phi(\mu_1)=\phi(\mu_2)=\alpha$;
     \item $\mu_1\not\in B_r(\mu_2)$; and
-    \item $\{\mu_1,\mu_2\}\cap \bigcup_{\mu\in S} B_{r}(\mu)=\emptyset$
+    \item $\{\mu_1,\mu_2\}\cap \bigcup_{\mu\in S} \tilde{B}_{r}(\mu)=\emptyset$
   \end{compactenum}
-  then include $\mu_1$ and $\mu_2$ in $S_1$ and declare this a \defin{success} for colour $\alpha$.  Otherwise, declare this a failure for colour $\alpha$.
+  then we include $\mu_1$ and $\mu_2$ in $Q_1$ and declare this a \defin{success} for colour $\alpha$.  Otherwise, declare this a \defin{failure} for colour $\alpha$.
 
-  At the end of this process, the resulting set satisfies (\ref{spread_out}) but is only guaranteed to satisfy (\ref{hits_both}) for the successful colours $\alpha$.
-We now use the second round to create a set $S_2$ to fix this.
-Our strategy is to use \cref{hall_vees} to find two matchings between the failed colours and the boxes in $\{B_r(\mu):\mu\in S_1\}$ that contain those colours.
-Each missing colour is matched in both matching, but each box is used in at most one of the matchings.
-This is always possible, since the number of edges and vertices of $G_r(\mu)$ is $\Theta(r^2)$ while the
+  Let $K\subseteq\varphi(V(G))$ be the set of colours for which the first round was successful and let $X:=\varphi(V(G))\setminus K$ be the set of colours for which the first round failed.
+At the end of this process, the set $Q_1$ satisfies (\ref{spread_out}) but is only guaranteed to satisfy (\ref{hits_both}) for the colours in $K$.
+We now use the second round to create a set $Q_2$ to fix this.
+Our strategy is to use \cref{d_hall} to find two matchings between the colours in $\overline{K}$ and the boxes in $\{\tilde{B}_r(\mu):\mu\in Q_1\}$ that contain those colours.
 
-  Consider a (for now) random permutation $\alpha_1,\ldots,\alpha_c$ of $\{1,\ldots,c\}$. For each $i\in\{1,\ldots,c\}$: Find $2(r+1)^2$ elements $\mu\in S_1$ such that
+  More precisely, we make a bipartite graph $H$ with parts $X$ and $Y:=Q_1\subseteq\VE(G_k)$.
+Let $\mu_1,\ldots,\mu_{2|K|}$ be a random permutation of the elements of $Y$.
+We include an edge $\alpha\mu_i$ in $H$ if and only if
+  \[
+    \VE(G_r(\mu_i))\setminus\left(\bigcup_{j=1}^{i-1}\VE(G_{2r}(\mu_j))\right)
+  \]
+  contains some object $\nu$ of colour $\phi(\nu)=\alpha$.
+We now want to use \cref{weighted_lovasz} to show that, with probability greater than zero, $\deg_H(\alpha)\in \Omega(d/r^2)$ for each $\alpha\in X$.
+
+  % To begin, define a supergraph $H^+\supseteq H$ that contains an edge $\alpha\mu$ if there exists at least one $\nu\in\VE(G_r(\mu))\cap X$ with $\phi(\nu)=\alpha$.
+
+  For each $\alpha\in X$, let $Z_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|Z_\alpha| \ge d$.
+Furthermore all but at most one element of $Z_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_r(\mu))$ since, otherwise the first round would have succeeded for colour $\alpha$.
+For each $\mu\in Y$, $|\VE(B_r(\mu))|\in O(r^2)$.
+Therefore, there are $\Omega(|Z_\alpha|/r^2)$ objects $\mu\in Y$ such that $B_r(\mu)\cap Z_\alpha\neq\emptyset$.
+
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\Omega(|Z_\alpha|/r^2),1/?)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 100r^2)$.
+  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap Z_\alpha$ contains at least one element $\nu$. Then, by \cref{packing_lemma}, the set
+  \[
+     X_\nu := \{\tau\in Y\setminus\{\mu\}:\nu\in\VE(G_r(\tau))\}
+  \]
+  has size at most $?$.
+Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation of $Y$, $\nu$ appears before any element of $X_\nu$.
+This happens with probability exactly $1/|X_\nu|\ge 1/?$.
+This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|Z_\alpha|/r^2)$.
+
+  In order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.
+To do this, consider the graph $J$ with vertex set $V(J)=Y$ and which contains an edge $\mu_1\mu_2$ if $B_{2r}(\mu_1)\cap B_{2r}(\mu_2)\neq\emptyset$.
+Now, consider the greedy process of repeatedly choosing a vertex $\mu$ of $J$ such that $B_r(\mu)\cap Z_\alpha$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.
+This process continues until no suitable vertex of $J$ remains.
+Each step in this process eliminates $O(r^2)$ objects $\nu$ with $\varphi(\nu)=\alpha$,\todo{Explain better} so the number of iterations is $r\in\Omega(d/r^2)$.
+
+  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and $\nu_1,\ldots,\nu_r$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$ and $\phi(\nu_i)=\alpha$, for each $i\in\{1,\ldots,t\}$.
+The important observation now is that the sets $X_{\nu_1},\ldots,X_{\nu_t}$ are disjoint.
+Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears before any elements of $X_{\nu_i}$ in our random permutation of $Y$'' then the events $U_1,\ldots,U_k$ are mutually independent.
+Indeed, each $U_i$ depends only on the relative order of $\{\mu_i'\}\cup X_{\nu_i}$ in the random permutation of $Y$.
+For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/?$.
+Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/?)$ random variable.
+Therefore,
+  \begin{align*}
+    \Pr\left(\deg_H(\alpha)< 100r^2\right)
+    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
+    & \le \Pr(\binomial(t,1/?)\le 100r^2) \\
+    & \le (2t)^{100r^2}(1/?)^{r^2}(1-1/?)^{t-100r^2} \\
+    & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
+    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/?)) \\
+    & = \exp(ar^2 - bt)
+  \end{align*}
+  for fixed positive $a,b>0$.
+  \todo[inline]{Work the preceding calculation out}
+  We are now ready for an application of \cref{weighted_lovasz}.
+For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $Z_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.
+The event $X_\alpha$ is mutually independent of $\Gamma_\alpha$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le ?|Z_\alpha|$.
+
+  \begin{align*}
+      \Pr(E_\alpha)
+      & \le \exp(ar^2-bt) \\
+      & \le \exp(ar^2-c|Z_\alpha|/r^2) \\
+      & \le \exp(|Z_\alpha|(ar^2/|Z_\alpha|)-c/r^2) \\
+      & \le \exp(|Z_\alpha|(1-c/r^2) & \text{for $d>r^2/a$} \enspace .
+  \end{align*}
+
+  For each $\alpha\in X$, define $w(E_\alpha)=c/2r^2$ and observe that
+  \begin{align*}
+    p(1-p)^{|\Gamma_\alpha|}
+    & \le  p(1-p)^{?|Z_\alpha|} = \exp(\log p + ?|Z_\alpha|\log(1-p))
+  \end{align*}
+
+  % (Think of $\nu$ as belonging to the first box that covers it.) Now, for each $\alpha\in X$, $\deg_H(\alpha)\ge d-1$ since, otherwise the first round would have been successful for $\alpha$.
+Straightforward counting shows that, for each $\mu\in Y$,  $\deg_H(\mu)\le |\VE(G_r(\mu))|\le 2(6r^2+6r+1)=:C$.
+
+  If $d$ and $r$ are chosen so that $(d-1)/C\ge 2$ then \cref{d_hall} guarantees that we can find a set $Q_2$ so that $Q_1\cup Q_2$ satisfies (\ref{hits_both}) and such that $|B_r(\mu)\cap Q_2|\le 1$ for each $u\in Q_1$, which comes very close to satisfying (\ref{spread_out}).
+Unfortunately, this approach is too naïve: it can produce triples (and even $4$-tuples) of objects in $Q$ are all contained in a small box and therefore do not satisfy (\ref{spread_out}).
+We will overcome this difficulty with an application of the Lovász Local Lemma.
+
+  Consider a (for now) random permutation $\mu_1,\ldots,\mu_{2|K|}$ of $Q_1$.  We remove any edge from $\alpha\mu_i$ from $H$ if
+  \[
+    \alpha\not\in \phi(\VE(G_r(\mu_i)\setminus(\bigcup_{j=1}^{i-1}\VE(G_{2r}(\mu_j))))
+  \]
+  contains no object $\nu$ with $\phi(\nu)=\alpha$.  (Note the extra factor of $2$ in the radii on the right hand side.)
+  The effect of this is that each $\mu_j\in Q_1$ defines a box of radius $2r$ around $\mu_j$ that is forbidden after the $j$th step of the second round.
+We will now use \cref{weighted_lovasz} to prove that, with some positive probability, each vertex
+
+  find $2(r+1)^2$ elements $\mu\in S_1$ such that
   \begin{compactenum}[(a)]

### 2022-04-04 lin-vs-cen.tex
Last minute airplane stuff

@@
 \begin{lem}
+  For every integer $r\ge 1$ there exists $d\in O(r^4)$ such that the following is true.
   Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.\todo{Use a specific value of $d$}.
@@
 We now want to use \cref{weighted_lovasz} to show that, with probability greater than zero, $\deg_H(\alpha)\in \Omega(d/r^2)$ for each $\alpha\in X$.
-
-  % To begin, define a supergraph $H^+\supseteq H$ that contains an edge $\alpha\mu$ if there exists at least one $\nu\in\VE(G_r(\mu))\cap X$ with $\phi(\nu)=\alpha$.
 
@@
 For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $Z_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.
-The event $X_\alpha$ is mutually independent of $\Gamma_\alpha$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le ?|Z_\alpha|$.
+The event $X_\alpha$ is mutually independent of $\{X_\beta:\beta\in X\setminus\{\alpha\}\setminus\Gamma_\alpha\}$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le ?|Z_\alpha|$.
 
@@
     p(1-p)^{|\Gamma_\alpha|}
-    & \le  p(1-p)^{?|Z_\alpha|} = \exp(\log p + ?|Z_\alpha|\log(1-p))
+    & \ge  p(1-p)^{?|Z_\alpha|} = \exp(\log c/2r^2 + ?|Z_\alpha|\log(1-c/2r^2))
   \end{align*}
+  \todo[inline]{Sort out these calculations}
 
-  % (Think of $\nu$ as belonging to the first box that covers it.) Now, for each $\alpha\in X$, $\deg_H(\alpha)\ge d-1$ since, otherwise the first round would have been successful for $\alpha$.
-Straightforward counting shows that, for each $\mu\in Y$,  $\deg_H(\mu)\le |\VE(G_r(\mu))|\le 2(6r^2+6r+1)=:C$.
+  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge 100r^2$ for each $\alpha\in X$.
+On the other hand, $\deg_{H}(y)\le 8r^2$ for each $y\in Y$.
+Therefore, for any $A\subseteq X$, $|N_H(A)|\ge 100|A|/8 \ge 2|A|$.
+Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$.
+Each edge $\alpha\mu$ of $M$ corresponds to some $\nu\in X_\alpha\cap\VE(G_r(\mu))$, which we place in $Q_2$.
 
-  If $d$ and $r$ are chosen so that $(d-1)/C\ge 2$ then \cref{d_hall} guarantees that we can find a set $Q_2$ so that $Q_1\cup Q_2$ satisfies (\ref{hits_both}) and such that $|B_r(\mu)\cap Q_2|\le 1$ for each $u\in Q_1$, which comes very close to satisfying (\ref{spread_out}).
-Unfortunately, this approach is too naïve: it can produce triples (and even $4$-tuples) of objects in $Q$ are all contained in a small box and therefore do not satisfy (\ref{spread_out}).
-We will overcome this difficulty with an application of the Lovász Local Lemma.
-
-  Consider a (for now) random permutation $\mu_1,\ldots,\mu_{2|K|}$ of $Q_1$.  We remove any edge from $\alpha\mu_i$ from $H$ if
-  \[
-    \alpha\not\in \phi(\VE(G_r(\mu_i)\setminus(\bigcup_{j=1}^{i-1}\VE(G_{2r}(\mu_j))))
-  \]
-  contains no object $\nu$ with $\phi(\nu)=\alpha$.  (Note the extra factor of $2$ in the radii on the right hand side.)
-  The effect of this is that each $\mu_j\in Q_1$ defines a box of radius $2r$ around $\mu_j$ that is forbidden after the $j$th step of the second round.
-We will now use \cref{weighted_lovasz} to prove that, with some positive probability, each vertex
-
-  find $2(r+1)^2$ elements $\mu\in S_1$ such that
-  \begin{compactenum}[(a)]
-    \item
-  \end{compactenum}
+  Now it is straightforward to verify that $Q:=Q_1\cup Q_2$ satisifies (\ref{hits_both}).
+Furthermore, for each $\mu\in Q_1$, $B_r(\mu)$ contains no object in $Q_1\setminus\{\mu\}$ and contains at most one element of $Q_2$.
+Similarly, for each $\mu\in Q_2$, $B_r(\mu)$ contains no element of $Q_2\setminus\{\mu\}$ and contains at most one element of $Q_1$.
+Therefore $Q$ satisfies (\ref{spread_out}).
 \end{proof}

### 2022-04-04 lin-vs-cen.tex
Pffft

@@
 
+\section*{Acknowledgement}
+
+The main subject of this paper was posed as an open problem at the 9th Annual Bellairs Workshop on Geometry and Graphs (v2), which was held January 21--28, 2022 and again at the 16th Annual Workshop on Probability and Combinatorics, which was held March 25--April 1, 2022.
+We are grateful to the organizers and participants of both workshops for providing a stimulating research environment.
+
 \bibliographystyle{plainurlnat}

### 2022-04-04 lin-vs-cen.tex
Pfffft

@@
 The \emph{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.
-We note that an easy argument shows that centered chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$.
+We note that an easy argument shows that centred chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$.
 The \emph{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.
@@
 \begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general}
-  For any graph $G$, $\chicen(G)\le \chilin^{190}(G)\cdot\log^{O(1)}(\chilin(G))$.\todo{Update with most recent numbers}
+  For any graph $G$, $\chicen(G)\le \chilin^{19}(G)\cdot\log^{O(1)}(\chilin(G))$.
 \end{thm}
@@
 \begin{enumerate}
-  \item A theorem of \citet{kawarabayashi.rossman:polynomial} shows that, if $\chicen(G)\ge k^{190}\log^{O(1)} k$ then $\chilin(G)\ge k^{38}$ or $G$ has treewidth $\tw(G)\ge k^{38}$.
-In the former case there is nothing left to prove.
-  \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy:improved} shows that, if $\tw(G)\ge k^{38}$, then $G$ contains an $\Omega(k^2\times k^2)$ grid minor.
-  \item A technical lemma \cite[Lemma~5]{kun.obrien.ea:polynomial} shows that, if $G$ contains a $k^2\times k^2$ grid minor, then $\chilin(G)\in\Omega(k)$.
+  \item A theorem of \citet{kawarabayashi.rossman:polynomial} shows that, if $\chicen(G)\ge k^{19}\log^{O(1)} k$ then $G$ contains a subcubic tree of treedepth $\Omega(k)$, or $\tw(G)\in\Omega(k^{18})$.
+In the former case, an asymptotically optimal result of \citet{kun.obrien.ea:polynomial} on subcubic trees completes the proof, so we are left with the case where $\tw(G)\in\Omega(k^{18})$.
+  \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy:improved} shows that, if $\tw(G)\ge k^{18}$, then $G$ contains an $\Omega(k^2\times k^2)$ grid minor.\todo{Update reference}
+  \item A technical lemma \cite[Lemma~5]{kun.obrien.ea:polynomial} shows that, if $G$ contains a $k^2\times k^2$ grid minor, then $\chilin(G)\in\Omega(k)$.\todo{Update reference}
 \end{enumerate}
 
-These notes are an attempt to improve the bound in \cref{kun_obrien_general} in the general case as well as for some special classes of graphs.
+In the current work, we improve the technical lemma in Point~3 to show that a $k\times k$ grid minor is sufficient to guarantee that $\chilin(G)\in\Omega(k)$.
+This improves the exponent in \cref{kun_obrien_general} from $19$ to $10$.
+
+This gives the following improvement to \cref{kun_obrien_general}
+\begin{thm}\label{kun_obrien_general2}
+  For any graph $G$, $\chicen(G)\in O(\chilin^{10}(G))$.
+\end{thm}
+
+In addition to the improvement to \cref{kun_obrien_general}, this results in considerably better bounds for some specific graph classes.
+For example, the Excluded Grid Theorem for planar graphs asserts that any planar graph of treewidth $k$ contains a $\Omega(k)\times \Omega(k)$ grid minor.
+Using this result instead of Point~2, above, implies that for any planar graph $G$, $\chicen(G)\in O(\chilin^2(G))$.
+
+Finally, \citet{kun.obrien.ea:polynomial} boldly conjecture that, for any graph $G$, $\chicen(G)\le 2\chilin(G)$.
+However, before the current work, the only class of graphs for which centred chromatic number was known to be bounded by a linear function of linear chromatic number was bounded degree trees.
+The current result adds further evidence for this conjecture by establishing that, when $G$ is a $k\times k$ grid (or even a pseudogrid), then $\chicen(G)\in\Theta(\chilin(G))$.
+% 
+% These notes are an attempt to improve the bound in \cref{kun_obrien_general} in the general case as well as for some special classes of graphs.
 First we point out that some new results can already reduce the exponent from $190$ to $39$ and eliminate the polylog factor.
 (These improvements are already described by \citet[Theorem~1.9]{czerwinski.nadara.ea:improved}.)
+% 
+% First, \citet{chuzhoy.tan:towards} improved the exponent in the Excluded Grid Theorem:  There exists a constant $c$ such that any graph $G$ with $\tw(G)\ge k^{18}\log^c k$ has a $k\times k$ grid minor.
+This already improves the bound from $k^{190}$ to $k^{90}$.
+% Next, a recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).
+Therefore, if $\td(G)\ge k^{19}$ then at least one of the following cases applies:
+% \begin{enumerate}
+%     \item $\tw(G)\ge k^{18}$: In this case the proof follows Steps~2 and 3 as before, but using the grid minor result of \citet{chuzhoy.tan:towards} to show that $\chilin(G)\in\Omega(k)$.
+%     \item $G$ contains a tree $T$ of treedepth $k$.
+The same paper \cite[Lemma~1.6]{czerwinski.nadara.ea:improved} shows that any tree $T$ of treedepth $k$ contains a subcubic subtree of treedepth $\Omega(k)$.
+%     A theorem of \citet{kun.obrien.ea:polynomial} shows that, for any subcubic tree $T$, $\chicen(T) \le (\log_2 3)\cdot\chilin(T)$.
+Therefore, in this case $\chilin(G)\ge\chilin(T)\ge \chicen(T)/\log_2 3 = \Omega(k)$.
+% \end{enumerate}
 
-First, \citet{chuzhoy.tan:towards} improved the exponent in the Excluded Grid Theorem:  There exists a constant $c$ such that any graph $G$ with $\tw(G)\ge k^{18}\log^c k$ has a $k\times k$ grid minor.
-This already improves the bound from $k^{190}$ to $k^{90}$.
-Next, a recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).
-Therefore, if $\td(G)\ge k^{19}$ then at least one of the following cases applies:
-\begin{enumerate}
-    \item $\tw(G)\ge k^{18}$: In this case the proof follows Steps~2 and 3 as before, but using the grid minor result of \citet{chuzhoy.tan:towards} to show that $\chilin(G)\in\Omega(k)$.
-    \item $G$ contains a tree $T$ of treedepth $k$.
-The same paper \cite[Lemma~1.6]{czerwinski.nadara.ea:improved} shows that any tree $T$ of treedepth $k$ contains a subcubic subtree of treedepth $\Omega(k)$.
-    A theorem of \citet{kun.obrien.ea:polynomial} shows that, for any subcubic tree $T$, $\chicen(T) \le (\log_2 3)\cdot\chilin(T)$.
-Therefore, in this case $\chilin(G)\ge\chilin(T)\ge \chicen(T)/\log_2 3 = \Omega(k)$.
-\end{enumerate}
-
-This gives the following improvement to \cref{kun_obrien_general}
-\begin{thm}[\citet{czerwinski.nadara.ea:improved}]\label{kun_obrien_general2}
-  For any graph $G$, $\chicen(G)\in O(\chilin^{19}(G))$.
-\end{thm}
-
-As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.
+% As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.
 This, by itself, would reduce the exponent in \cref{kun_obrien_general2} from $39$ to $20$.
@@
 If one could prove a similar result for the first ``heavy hammer'' then this would show that, for any planar graph $G$, $\chicen(G)\in O(\chilin(G))$.
+%
 
@@
 
+We will also make use of the following version of the Lovász Local Lemma:
+
 \begin{lem}\label{weighted_lovasz}
   Let $E:=\{E_1,\ldots,E_n\}$ be a set of events in some probability space $(\Omega,\Pr)$.
-For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq E$ be such that the event $E_i$ is mutually independent of $\Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:E\to(0,1)$ be such that
+For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq E$ be such that the event $E_i$ is mutually independent of $E\setminus \Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:E\to(0,1)$ be such that
   \[
@@
 In this work, we need to work with pseudogrids.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
-\begin{compactenum}
+\begin{compactitem}
   \item Replace each edge $vw$ of $G_{a\times b}$ with a path $\overline{P}_{vw}$ whose endpoints are $v$ and $w$.
 (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
-  \item Replace each degree-$4$ vertex $(x,y)$ of $G_{a\times b}$ with a path $P_v$. If $P_v$ has only one vertex then this operation has no effect.  Otherwise, $P_v$ has two endpoints $p$ and $q$ and
+  \item Replace each degree-$4$ vertex $(x,y)$ of $G_{a\times b}$ with a path $P_v$.
+If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each of $(x\pm 1,y\pm 1)$ and this has effect.
+Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $(x\pm 1,y\pm 1)$.
+It is useful to consider three possible cases:
   \begin{compactenum}[(Q1)]
@@
   \end{compactenum}
-\end{compactenum}
+\end{compactitem}
 
@@
 For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.
-In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.
-We call $\mathcal{P}$ the \defin{grid-partition} of $G$.
+In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths and the quotient graph $G/\mathcal{P}$ is isomorphic to $G_{a\times b}$.
+We call $\mathcal{P}$ a \defin{grid-partition} of $G$.
 % and we denote it by $(P_\mu:\mu\in \VE(G_{a\times b}))$.
@@
 However, for $i\in\{2,\ldots,a-1\}$ $R$ may or may not contain $V(P_{v_i})$.
-In particular, if $P_{v_i}$ was created using (Q\ref{q_iii})\todo{Fix this!} then $R$ does not contain any internal vertices in $P_{v_i}$.
+In particular, if $P_{v_i}$ was created using (Q\ref{q_iii}) then $R$ does not contain any internal vertices in $P_{v_i}$.
 Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
@@
 If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.
-Similarly, if we delete column $C$ of $G$, then the resulting graph is an $(a-1)\times b$ pseudogrid.
+Similarly, if we delete column $C$ of $G$, then the resulting graph is a $(a-1)\times b$ pseudogrid.
+
+\subsection{Proof Outline}
+
+If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph.
+Thus, our goal is to show that that $\chilin(G)\in \Omega(k)$ for any $k\times k$ pseudogrid.
+We do this by showing that, any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
+We will do this in several steps (see \cref{outline}):
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{ccc}
+      \includegraphics{figs/outline-2} &
+      \includegraphics{figs/outline-3} &
+      \includegraphics{figs/outline-4} \\
+      $S_1$ & $S_1\cup S_2$ & $P$
+    \end{tabular}
+  \end{center}
+  \caption{Constructing the sets $S_1$ and $S_2$ and the path $P$.}
+  \label{outline}
+\end{figure}
+\begin{compactenum}
+
+  \item We first delete rows and columns from $G$ so that each colour that appears in $G$ appears at least $C$ times, for some large constant $C$.
+From this point on the goal is to construct $P$ so that it contains each colour at least twice.
+
+  \item We do our best to find a set $S_1$ of vertices in $G$ that contains two vertices of each colour and that is `well-separated' in the sense that the corresponding set of vertices/edges in $G_{k\times k}$ have a minimum distance between them.
+In \cref{outline} this minimum distance corresponds to the fact that the box drawn centered at each vertex in $S_1$ contains no other vertices of $S_1$.
+  % This makes it easy to find a single path in $G$ that contains each of these vertices.  
+  Unfortunately, this process will fail for some subset of the colours that appear in $G$.
+
+  \item For these failed colours, we use a variant of Hall's Theorem to match each failed colour $\alpha$ with two vertices in $S_1$ that each contain a vertex of colour $\alpha$ nearby.
+In this matching, no vertex of $S_1$ is matched with more than one failed colour.
+From this we identify a set $S_2$ that contains two vertices of each of the missing colours and such that no vertex of $S_2$ is close to any other vertex of $S_2$.
+
+  \item We construct a path $P$ that contains each vertex in $S_1\cup S_2$.  This is possible (easy, even) because each vertex in $S_1$ is `close to' at most one vertex of $S_2$ and vice-versa.
+\end{compactenum}
+
+The most challenging aspect of this proof is the construction of $S_2$, which requires the use of the Local Lemma to ensure that no vertex chosen to take part in $S_2$ is close to any other vertex in $S_2$.
+The difficulty is illustrated in \cref{outline} by the cluster of three points of $S_1$ in the top right corner.
+These three points of $S_1$ are well-separated, but choosing one point from each of the three boxes to take part in $S_2$ could result in three points of $S_2$ being very close to each other.
+The local lemma is used to show that, in situations like this, we can still obtain two vertices of $S_2$ from each missing colour class while avoiding taking two vertices of $S_2$ from these three boxes.
+
+We now proceed with the various steps in the proof.
 
@@
 Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
-For any $\mu\in\VE(G_{k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\varphi(v):v\in V(P_\mu)\}$.
-For any colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-1}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)\}$.
+The partition $\mathcal{P}$ associates each vertex of $G$ with an edge or vertex $\mu$ of $G_k$, so $\varphi$ associates a colour set with each object in $\VE(G_k)$, as follows.
+For each $\mu\in\VE(G_{k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\varphi(v):v\in V(P_\mu)\}$.
+For each colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-1}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)\}$.
 For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-1}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-1}(\alpha)$.
+(Throughout this section, we will use the idiom $\varphi(V(G))$ to denote the set of all colours used by $\varphi$ to colour the vertices of $G$.)
 
@@
   \begin{compactitem}
-    \item For each $v:=(i,j)\in V(G_{k})$ with $A\cap\varphi(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
-    \item For each edge $vw\in E(G_{k\times k})$ with $A\cap\phi()\alpha\in\Phi(vw)$, we include the row of $G$ that contains $P_{vw}$ in $R$ or we include the column of $G$ that contains $P_{vw}$ in $C$.
+    \item For each $v:=(i,j)\in V(G_{k})$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
+    \item For each horizontal edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
+    \item For each vertical edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.\todo{Define horizontal and vertical edges of $G$.}
     \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
   \end{compactitem}
-  Now we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
+  At this point $|R|=|C|\le d|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
 In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.
@@
   Now apply induction on $G_0$ to get a $k'\times k'$ pseudogrid with
-  \[  k'\ge k_0-d|\varphi(V(G_0))|  \ge k-d|A|-d|\varphi(V(G_0))|= k - d|\varphi(V(G))|
+  \[  k'\ge k_0-d|\varphi(V(G_0))|  \ge k-d|A|-d|\varphi(V(G_0))|\ge k - d|\varphi(V(G))|
   \]
@@
 
+Having each vertex or edge $\mu$ of $G_k$ associated with a set $\varphi_\mathcal{P}(\mu)$ of colours rather than a single colour will be problematic for what we want to to do next.
+The following lemma allows us to choose one representative colour $\phi(\mu)$ from $\varphi_{\mathcal{P}}(\mu)$ for each $\mu\in \VE(G_k)$ while still ensuring that each colour appears frequently.
+
 \begin{lem}\label{one_colour_per_object}
   Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d\cdot|A|$.
-Then there exists a (partial) colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
+Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
   \begin{compactenum}[(i)]
@@
   \[
-    \{ (\alpha,\mu)\in X\times Y: \alpha\in\varphi(\mu) \}
+    E(H) := \{ (\alpha,\mu)\in X\times Y: \alpha\in\varphi_\mathcal{P}(\mu) \}
   \]
@@
 For each edge $\alpha\mu$ in $M$, set $\phi(\mu):=\alpha$.
-This defines $\phi(\mu)$ for any $\mu\in X$ with $\deg_M(\mu)=1$.
-For each $\mu\in X$ with $\deg_M(\mu)=0$ set $\phi(\mu):=\perp$.
+This defines $\phi(\mu)$ for any $\mu\in Y$ with $\deg_M(\mu)=1$.
+For each $\mu\in Y$ with $\deg_M(\mu)=0$ set $\phi(\mu):=\perp$.
 \end{proof}
@@
 
-The \defin{$r$-box} centered at a vertex $v:=(i,j)$ of $G_{a\times b}$ is defined as
+The \defin{$r$-box} centred at a vertex $v:=(i,j)$ of $G_{k}$ is defined as
 \[
-  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{a,b}) \enspace .
+  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{k}) \enspace .
 \]
-The \defin{$r$-box} centered at an edge $vw$ of $G_{a\times b}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.
-Observe that, for any $\mu\in V(G_{a\times b})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.
-Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{a\times b}[B_r(\mu)]$.
-We extend these definitions to vertices of an $a\times b$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G)\}$.
-For any $v\in V(G)$, define
+The \defin{$r$-box} centred at an edge $vw$ of $G_{k}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.  
+% Observe that, for any $\mu\in V(G_{k})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.  
+Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.
+We extend these definitions to vertices of an $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows.
+For any $\mu\in V(G_k)$, define
 \[
-   \tilde{B}_r(v) := \bigcup_{\mu\in \VE(G_r(\mu_v))} V(P_\mu)
+   \tilde{B}_r(\mu) := \bigcup_{\nu\in \VE(G_r(\mu))} V(P_\nu)
 \]
-where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.
-For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
+and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$
+where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never any ambiguity as to which partition is being used.}
+% For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
 
@@
 
+Before finishing with this section we recall a fairly standard looking packing lemma.
+
 \begin{lem}\label{packing_lemma}
-  Some packing lemma
+  Let $Q\subseteq \VE(G_k)$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$, 
+  $|\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}| \le 25$.
 \end{lem}
 
-\subsection{Finding an Uncentered Path}
+\todo[inline]{Fix notation conflict:  $G_r(\mu)$ versus $G_k$?  Maybe go back to $G_{k\times k}$?}
 
-Next we show how, given a colouring like that guaranteed by \cref{only_frequent}, we can find a set of vertices that hit every colour at least twice and that are compatible with \cref{pick_up_everything}.
+\subsection{Finding an Uncentred Path}
+
+Next we show how, given a colouring $\varphi$ like that guaranteed by \cref{only_frequent}, we can find a set of vertices in $G$ that contain two vertices of each colour and that is compatible with \cref{pick_up_everything}.
 
 \begin{lem}
-  For every integer $r\ge 1$ there exists $d\in O(r^4)$ such that the following is true.
-  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.\todo{Use a specific value of $d$}.
+  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
+  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.
   Then there exists $S\subseteq V(G)$ such that
   \begin{compactenum}[(i)]
-    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|\ge 2$; and
+    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|= 2$; and
     \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
@@
   We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi$ of $\VE(G_k)$.
-Observe that it is now sufficient to find $Q\subset\VE(G_k)$ such that $|\{\mu\in Q:\phi(v)=\alpha\}|\ge 2$ and $|B_r(\mu)\cap Q|\le 2$ for each $\alpha\in\varphi(V(G))$ and each $\mu\in Q$.
+Observe that it is now sufficient to find a $2|\varphi(V(G))|$-object subset $Q\subseteq\VE(G_k)$ such that
+  \begin{compactenum}[(i)]
+    \item $|\{\mu\in Q:\phi(v)=\alpha\}|= 2$ and 
+    \item $|\VE(G_r(\mu))\cap Q|\le 2$ for each $\alpha\in\varphi(V(G))$ and each $\mu\in Q$.  
+  \end{compactenum}
   Indeed, with such a $Q$ we can obtain $S$ by taking one vertex of colour $\phi(\mu)$ from $P_\mu$ for each $\mu\in Q$.
 
-  We construct $Q$ using a two round process.  We start with an initially empty set $Q_1$ and repeat the following for each $\alpha\in\varphi(V(G))$:
+  We construct $Q$ in two rounds.  In the first round, we start with an initially empty set $Q_1$ and repeat the following for each $\alpha\in\varphi(V(G))$:
   If there exists distinct $\mu_1,\mu_2\in\VE(G_k)$ such that
   \begin{compactenum}[(a)]
-    \item $\phi(\mu_1)=\phi(\mu_2)=\alpha$;
-    \item $\mu_1\not\in B_r(\mu_2)$; and
-    \item $\{\mu_1,\mu_2\}\cap \bigcup_{\mu\in S} \tilde{B}_{r}(\mu)=\emptyset$
+    \item $\phi(\mu_1)=\phi(\mu_2)=\alpha$; and\label{hits_both_q}
+    \item for each $b\in\{1,2\}$, $\mu_b\not\in \VE(G_r(\mu_{3-b})) \cup \bigcup_{\mu\in Q_1} \VE(G_r(\mu))$.\label{spread_out_q} 
   \end{compactenum}
-  then we include $\mu_1$ and $\mu_2$ in $Q_1$ and declare this a \defin{success} for colour $\alpha$.  Otherwise, declare this a \defin{failure} for colour $\alpha$.
+  then we include $\mu_1$ and $\mu_2$ in $Q_1$ and declare the first round \defin{$\alpha$-successful}.  Otherwise, we declare the first round an \defin{$\alpha$-failure}.
 
-  Let $K\subseteq\varphi(V(G))$ be the set of colours for which the first round was successful and let $X:=\varphi(V(G))\setminus K$ be the set of colours for which the first round failed.
-At the end of this process, the set $Q_1$ satisfies (\ref{spread_out}) but is only guaranteed to satisfy (\ref{hits_both}) for the colours in $K$.
+  Let $K\subseteq\varphi(V(G))$ be the set of colours $\alpha$ for which the first round was $\alpha$-successful.
+At the end of this process, the set $Q_1$ certainly satisfies (\ref{spread_out}) since (\ref{spread_out_q}) implies that $\VE(G_r(\mu))\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$.
+However, $Q_1$ only satisfies (\ref{hits_both}) for the colours in $K$.
 We now use the second round to create a set $Q_2$ to fix this.
-Our strategy is to use \cref{d_hall} to find two matchings between the colours in $\overline{K}$ and the boxes in $\{\tilde{B}_r(\mu):\mu\in Q_1\}$ that contain those colours.
+  % Our strategy is to use \cref{d_hall} to find two matchings between the colours in $X$ and the boxes in $\{\tilde{B}_r(\mu):\mu\in Q_1\}$ that contain those colours.
 
-  More precisely, we make a bipartite graph $H$ with parts $X$ and $Y:=Q_1\subseteq\VE(G_k)$.
-Let $\mu_1,\ldots,\mu_{2|K|}$ be a random permutation of the elements of $Y$.
+  Define a bipartite graph $H$ with parts $X:=\varphi(V(G))\setminus K$ and $Y:=Q_1$.
+Let $\pi:=\mu_1,\ldots,\mu_{2|K|}$ be a random permutation of the elements of $Y$.
 We include an edge $\alpha\mu_i$ in $H$ if and only if
@@
   contains some object $\nu$ of colour $\phi(\nu)=\alpha$.
-We now want to use \cref{weighted_lovasz} to show that, with probability greater than zero, $\deg_H(\alpha)\in \Omega(d/r^2)$ for each $\alpha\in X$.
+Note the asymmetry here: In the first round, each element $\mu_j$ that takes part in $Q_1$ prevents elements in $\VE(G_r(\mu))$ from also being part of $Q_1$.
+In the second round, $\mu_j$ defines a larger box $\VE(G_{2r}(\mu_j))$ that must be avoided when choosing an element from $\VE(G_r(\mu_i))$, for $i>j$.
 
-  For each $\alpha\in X$, let $Z_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|Z_\alpha| \ge d$.
-Furthermore all but at most one element of $Z_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_r(\mu))$ since, otherwise the first round would have succeeded for colour $\alpha$.
-For each $\mu\in Y$, $|\VE(B_r(\mu))|\in O(r^2)$.
-Therefore, there are $\Omega(|Z_\alpha|/r^2)$ objects $\mu\in Y$ such that $B_r(\mu)\cap Z_\alpha\neq\emptyset$.
+  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\in \Omega(d/r^2)$ for each $\alpha\in X$.
+  For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.
+Furthermore all but at most one element of $\Phi_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since, otherwise the first round would have succeeded for colour $\alpha$.
+For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$.
+Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\Omega(|Z_\alpha|/r^2),1/?)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 100r^2)$.
-  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap Z_\alpha$ contains at least one element $\nu$. Then, by \cref{packing_lemma}, the set
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\Omega(|\Phi_\alpha|/r^2),1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 100r^2)$.
+(We make no attempt to optimize the constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$.
+Then, by \cref{packing_lemma}, the set
   \[
-     X_\nu := \{\tau\in Y\setminus\{\mu\}:\nu\in\VE(G_r(\tau))\}
+     Y_\nu := \{\tau\in Y\setminus\{\mu\}:\nu\in\VE(G_r(\tau))\}
   \]
-  has size at most $?$.
-Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation of $Y$, $\nu$ appears before any element of $X_\nu$.
-This happens with probability exactly $1/|X_\nu|\ge 1/?$.
-This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|Z_\alpha|/r^2)$.
+  has size at most $25$.
+Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any element of $Y_\nu$.
+This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
+This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.
 
@@
 To do this, consider the graph $J$ with vertex set $V(J)=Y$ and which contains an edge $\mu_1\mu_2$ if $B_{2r}(\mu_1)\cap B_{2r}(\mu_2)\neq\emptyset$.
-Now, consider the greedy process of repeatedly choosing a vertex $\mu$ of $J$ such that $B_r(\mu)\cap Z_\alpha$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.
+Now, consider the greedy process of repeatedly choosing a vertex $\mu$ of $J$ such that $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.
 This process continues until no suitable vertex of $J$ remains.
@@
 
-  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and $\nu_1,\ldots,\nu_r$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$ and $\phi(\nu_i)=\alpha$, for each $i\in\{1,\ldots,t\}$.
+  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$, for each $i\in\{1,\ldots,t\}$.
 The important observation now is that the sets $X_{\nu_1},\ldots,X_{\nu_t}$ are disjoint.
@@
 Indeed, each $U_i$ depends only on the relative order of $\{\mu_i'\}\cup X_{\nu_i}$ in the random permutation of $Y$.
-For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/?$.
-Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/?)$ random variable.
+For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.
+Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/25)$ random variable.
 Therefore,
@@
     & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
-    & \le \Pr(\binomial(t,1/?)\le 100r^2) \\
-    & \le (2t)^{100r^2}(1/?)^{r^2}(1-1/?)^{t-100r^2} \\
+    & \le \Pr(\binomial(t,1/25)\le 100r^2) \\
+    & \le (2t)^{100r^2}(1/25)^{r^2}(1-1/25)^{t-100r^2} \\
     & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
-    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/?)) \\
+    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
     & = \exp(ar^2 - bt)
   \end{align*}
-  for fixed positive $a,b>0$.
+  for fixed positive $a,b>0$.  Recall that we are working in the regime in which $|\Phi_\alpha|\ge d\ge Cr^2$ and $t\in\Omega(d/r^2)$ so, for sufficiently large $C$,
+  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-\Omega(C|\Phi_\alpha|/r^2))$.
   \todo[inline]{Work the preceding calculation out}
-  We are now ready for an application of \cref{weighted_lovasz}.
-For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $Z_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.
-The event $X_\alpha$ is mutually independent of $\{X_\beta:\beta\in X\setminus\{\alpha\}\setminus\Gamma_\alpha\}$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le ?|Z_\alpha|$.
+  We are now ready for an application of \cref{weighted_lovasz}.\todo{Continue here}  For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $\Phi_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.
+The event $X_\alpha$ is mutually independent of $\{X_\beta:\beta\in X\setminus\{\alpha\}\setminus\Gamma_\alpha\}$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.
 
@@
       & \le \exp(ar^2-bt) \\
-      & \le \exp(ar^2-c|Z_\alpha|/r^2) \\
-      & \le \exp(|Z_\alpha|(ar^2/|Z_\alpha|)-c/r^2) \\
-      & \le \exp(|Z_\alpha|(1-c/r^2) & \text{for $d>r^2/a$} \enspace .
+      & \le \exp(ar^2-c|\Phi_\alpha|/r^2) \\
+      & \le \exp(|\Phi_\alpha|(ar^2/|\Phi_\alpha|)-c/r^2) \\
+      & \le \exp(|\Phi_\alpha|(1-c/r^2) & \text{for $d>r^2/a$} \enspace .
   \end{align*}
@@
     p(1-p)^{|\Gamma_\alpha|}
-    & \ge  p(1-p)^{?|Z_\alpha|} = \exp(\log c/2r^2 + ?|Z_\alpha|\log(1-c/2r^2))
+    & \ge  p(1-p)^{?|\Phi_\alpha|} = \exp(\log c/2r^2 + ?|\Phi_\alpha|\log(1-c/2r^2))
   \end{align*}
@@
 % \begin{figure}
-%   \begin{center}
+%   \begin{centre}
 %     \begin{tabular}{c@{\hspace{2cm}}c}
@@
 %     \end{tabular}
-%   \end{center}
+%   \end{centre}
 %   \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
@@
 Refer to \cref{ball_figure}.
-For each $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $\beta$ centered at $v$ is $B_{\beta}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le \beta\}$.
+For each $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $\beta$ centred at $v$ is $B_{\beta}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le \beta\}$.
 Because it comes up in several places, it is worth noting that $|B_{\beta}(v)|\le 4(\beta+1)^2$, and this bound is tight for each $v\in\{\beta+1,\ldots,\beta-r-1\}\times\{\beta+1,\ldots,b-\beta-1\}$.
@@
 % % \begin{figure}
-% %   \begin{center}
+% %   \begin{centre}
 % %     \includegraphics{figs/grid-3}
-% %   \end{center}
+% %   \end{centre}
 % %   \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
@@
 % % % \todo[inline]{This lemma is not true.
-Pick a center $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.
+Pick a centre $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.
 Then $x$ is incident on all four Voronoi regions.
@@
 % % \begin{figure}
-% %   \begin{center}
+% %   \begin{centre}
 % %     \begin{tabular}{c@{\hspace{1cm}}c}
@@
 % %     \end{tabular}
-% %   \end{center}
+% %   \end{centre}
 % %   \caption{The vertex $x:=(i,j)$ is (a)~straight if the path from $B_{(i,j-1)}$ to $B_{(i+1,j)}$ uses all vertices of $P_x$ or (b)~bent otherwise.}
@@
 For each colour $c\in\{1,\ldots,C\}$, there are at least two vertices $x_c,y_c\in R$ such that $P_{x_c}$ and $P_{y_c}$ each contain some vertex of colour $c$.
-Therefore the path $P$ contains two vertices of each colour, so $P$ has no center and therefore $\varphi$ is not a linear colouring of $H$.
+Therefore the path $P$ contains two vertices of each colour, so $P$ has no centre and therefore $\varphi$ is not a linear colouring of $H$.
 Since $\varphi$ is an arbitrary $\epsilon k$-colouring of $H$, this implies that $\chilin(H)\ge \epsilon k = \Theta(k/\alpha)=\Theta(k/r)$.
@@
 We are grateful to the organizers and participants of both workshops for providing a stimulating research environment.
+Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
 

### 2022-04-04 lin-vs-cen.tex
Pffft

@@
   \end{align*}
-  for fixed positive $a,b>0$.  Recall that we are working in the regime in which $|\Phi_\alpha|\ge d\ge Cr^2$ and $t\in\Omega(d/r^2)$ so, for sufficiently large $C$,
+  for fixed positive $a,b>0$.
+Recall that we are working in the regime in  $t\in\Omega(|\Phi_\alpha|/r^2)$ and $|\Phi_\alpha|\in\Omega(d/r^2)$ so, for any fixed $C>0$, there exists $d\in\Omega(r^4)$ such that
   $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-\Omega(C|\Phi_\alpha|/r^2))$.

### 2022-04-04 lin-vs-cen.tex
Pfffffffffft

--- 2022-04-04 lin-vs-cen.tex
@@
 The current result adds further evidence for this conjecture by establishing that, when $G$ is a $k\times k$ grid (or even a pseudogrid), then $\chicen(G)\in\Theta(\chilin(G))$.
-% 
-% These notes are an attempt to improve the bound in \cref{kun_obrien_general} in the general case as well as for some special classes of graphs.
-First we point out that some new results can already reduce the exponent from $190$ to $39$ and eliminate the polylog factor.
-(These improvements are already described by \citet[Theorem~1.9]{czerwinski.nadara.ea:improved}.)
-% 
-% First, \citet{chuzhoy.tan:towards} improved the exponent in the Excluded Grid Theorem:  There exists a constant $c$ such that any graph $G$ with $\tw(G)\ge k^{18}\log^c k$ has a $k\times k$ grid minor.
-This already improves the bound from $k^{190}$ to $k^{90}$.
-% Next, a recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).
-Therefore, if $\td(G)\ge k^{19}$ then at least one of the following cases applies:
-% \begin{enumerate}
-%     \item $\tw(G)\ge k^{18}$: In this case the proof follows Steps~2 and 3 as before, but using the grid minor result of \citet{chuzhoy.tan:towards} to show that $\chilin(G)\in\Omega(k)$.
-%     \item $G$ contains a tree $T$ of treedepth $k$.
-The same paper \cite[Lemma~1.6]{czerwinski.nadara.ea:improved} shows that any tree $T$ of treedepth $k$ contains a subcubic subtree of treedepth $\Omega(k)$.
-%     A theorem of \citet{kun.obrien.ea:polynomial} shows that, for any subcubic tree $T$, $\chicen(T) \le (\log_2 3)\cdot\chilin(T)$.
-Therefore, in this case $\chilin(G)\ge\chilin(T)\ge \chicen(T)/\log_2 3 = \Omega(k)$.
-% \end{enumerate}
-
-% As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.
-This, by itself, would reduce the exponent in \cref{kun_obrien_general2} from $39$ to $20$.
-We observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.
-For example, if we consider only planar graphs, then it is well known that any planar graph $G$ of treewidth $k$ contains an $\Omega(k\times k)$ grid minor.
-This already reduces the exponent in \cref{kun_obrien_general2} further (for planar graphs) to $3$ ($2$ if we can prove the result for grid minors).
-If one could prove a similar result for the first ``heavy hammer'' then this would show that, for any planar graph $G$, $\chicen(G)\in O(\chilin(G))$.
-%
 
@@
 
-% For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.
-For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.
-
-% For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.
-For any $S\subseteq V(G)$, $N_G(s):=\bigcup_{v\in S}N_G(v)$ and $N_G[s]:=\bigcup_{v\in S}N_G[v]$.
-The subgraph $G[S]$ of $G$ \emph{induced} by $S$ is the graph with vertex set $V(G[S]):=S\cap V(G)$ and edge set $E(G[S]):=\{vw\in E(G):\{v,w\}\subseteq S\}$.
-
 In this paper, all graphs are simple and undirected. For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$.
 
-% For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.
-When $\mathcal{P}$ consists of disjoint subsets of $V(G)$ that do not necessarily cover $V(G)$ then we use the convention that $G/\mathcal{P}:=G/\mathcal{P'}$ where $\mathcal{P}'$ is the completion of $\mathcal{P}$ obtained by introducing each missing element of $V(G)$ as singleton set.
-More precisely $\mathcal{P'}:=\mathcal{P}\cup\{\{v\}:v\in V(G)\setminus(\cup P)\}$.
-
-% A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.
-Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.
-A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.
-For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.
-Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.
-A $\le\!\!\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has at most $\ell$ internal vertices.
-
 We make use of the following (polygamous) generalisation of Hall's Marriage Theorem, which follows from Hall's Marriage Theorem by replacing each vertex $x\in X$ with a $d$-tuple $x_1,\ldots,x_d$ each having the same neighbours as $x$.
-
-% \begin{lem}\label{normal_hall}
-%   Let $G$ be a bipartite graph with parts $X$ and $Y$ such that, for any $A\subseteq X$, $|N_G(A)|\ge |A|+1$.
-Then $G$ contains a subgraph $M$ (a matching) such that $\deg_M(x)=1$ for each $x\in X$ and $\deg_M(y)\le 1$ for each $y\in Y$.
-% \end{lem}
 
@@
 
-We will also make use of the following version of the Lovász Local Lemma:
+We will also make use of the following weighted version of the Lovász Local Lemma:
 
@@
 
-If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph.
-Thus, our goal is to show that that $\chilin(G)\in \Omega(k)$ for any $k\times k$ pseudogrid.
-We do this by showing that, any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
+If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph.  Therefore to prove \cref{kun_obrien_general2}, it suffices to prove:
+
+\begin{lem}\label{pseudogrid_lower_bound}
+  For any $k\times k$ pseudogrid $G$, $\chilin(G)\in\Omega(k)$.
+\end{lem}
+
+We do this by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
 We will do this in several steps (see \cref{outline}):
@@
 
-\begin{lem}
+\begin{lem}\label{doubled_colour_set}
   For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
@@
   \end{align*}
-  for fixed positive $a,b>0$.  Recall that we are working in the regime in which $|\Phi_\alpha|\ge d\ge Cr^2$ and $t\in\Omega(d/r^2)$ so, for sufficiently large $C$,
-  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-\Omega(C|\Phi_\alpha|/r^2))$.
+  for fixed positive $a,b>0$.
+Recall that we are working in the regime in  $t\in\Omega(|\Phi_\alpha|/r^2)$ and $|\Phi_\alpha|\in\Omega(d/r^2)$ so, for any fixed $C>0$, there exists $d\in\Omega(r^4)$ such that
+  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-C|\Phi_\alpha|/r^2)$.
   \todo[inline]{Work the preceding calculation out}
@@
   \begin{align*}
-      \Pr(E_\alpha)
-      & \le \exp(ar^2-bt) \\
-      & \le \exp(ar^2-c|\Phi_\alpha|/r^2) \\
-      & \le \exp(|\Phi_\alpha|(ar^2/|\Phi_\alpha|)-c/r^2) \\
-      & \le \exp(|\Phi_\alpha|(1-c/r^2) & \text{for $d>r^2/a$} \enspace .
+      \Pr(E_\alpha) \le \exp(-|\Phi_\alpha|\cdot C/r^2) \enspace .
   \end{align*}
 
-  For each $\alpha\in X$, define $w(E_\alpha)=c/2r^2$ and observe that
+  For each $\alpha\in X$, define $w(E_\alpha)=C/2r^2$ and observe that
   \begin{align*}
     p(1-p)^{|\Gamma_\alpha|}
-    & \ge  p(1-p)^{?|\Phi_\alpha|} = \exp(\log c/2r^2 + ?|\Phi_\alpha|\log(1-c/2r^2))
+    & \ge  p(1-p)^{25|\Phi_\alpha|} = p\exp(|\Phi_\alpha|\cdot 25\log(1-c/2r^2))
   \end{align*}
@@
 Furthermore, for each $\mu\in Q_1$, $B_r(\mu)$ contains no object in $Q_1\setminus\{\mu\}$ and contains at most one element of $Q_2$.
-Similarly, for each $\mu\in Q_2$, $B_r(\mu)$ contains no element of $Q_2\setminus\{\mu\}$ and contains at most one element of $Q_1$.
+Similarly, for each $\nu\in Q_2$, $B_r(\nu)$ contains no element of $Q_2\setminus\{\nu\}$ and contains at most one element of $Q_1$.
 Therefore $Q$ satisfies (\ref{spread_out}).
@@
 
-\newpage
+We now have all the pieces needed to complete our lower bound on the linear chromatic number of pseudogrids:
 
-% \hrule
-%
-% \subsection{Subdivided Grids}
-%
-% Refer to \cref{grid_figure}.
-For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
-For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
-We define the \emph{boundary} of $G_{a\times b}$ as $\partial G_{a\times b}:=\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.
-We say that a path $P:=v_0,v_1,\ldots,v_\ell$ in $G_{a\times b}$ \emph{goes straight} at a vertex $v_i$, $i\in\{1,\ldots,k-1\}$ if $v_{i-1}$, $v_i$, and $v_{i+1}$ are contained in a single row or column.
-Otherwise, we say that $P$ \emph{bends} at $v_i$.
-%
-% \begin{figure}
-%   \begin{centre}
-%     \begin{tabular}{c@{\hspace{2cm}}c}
-%       \includegraphics{figs/grid-1} &
-%       \includegraphics{figs/grid-2}
-%     \end{tabular}
-%   \end{centre}
-%   \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
-%   \label{grid_figure}
-% \end{figure}
-%
-% % An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.
-The \emph{vertex boundary} of $G_{a\times b}$ is $\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.
-The \emph{edge boundary} of $G_{a,b}$ consists of all edges incident to at least one vertex in the vertex boundary of $G_{a,b}$.
-The \emph{boundary} of $G_{a\times b}$ is the union of the edge and vertex boundaries.
-%
-%
-% % Let $G^{\onesub}_k$ denote the $1$-subdivision of $G_k$.
-%
-% A \emph{simple $k\times k$ pseudogrid} is any graph that can be obtained from $G^{\onesub}_k$ by replacing some degree-$4$ vertices of $G_k$ with small binary trees.
-Specifically, we may replace a degree-$4$ vertex with a tree whose leaves are the vertices in $N_{G^{\onesub}_k}(v)$ and whose two internal nodes $v_a$ and $v_b$ are each adjacent to two leaves and to each other.
-%
-% Let $G_0$ be a simple $k\times k$ pseudogrid and observe that there is a natural mapping $f:E(G)\to E(G_k)\cup V(G_k)$ which is fairly obvious except that any edge $v_av_b\in E(G_0)$ introduced when replacing a degree-$4$ vertex $v$ in $G_k$ has $f(v_av_b):= v$.
-%
-% \begin{lem}\label{one_path}
-%    Let $S:=\bigcup_{i=1}^r S_i$ be a subset of $E(G_0)$ with the following properties:
-%    \begin{compactenum}[(a)]
-%      \item $|S_i|\le 2$ for each $i\in\{1,\ldots,r\}$; and
-%      \item for each distinct $i,j\in\{1,\ldots,r\}$, $\dist_{G_k}(\bigcup_{e\in S_i}f(e), \bigcup_{e\in S_j}f(e))\ge C$.\todo{Choose the value of $C$}
-%    \end{compactenum}
-%    Then $G_0$ contains a path that contains each edge in $S$.
-% \end{lem}
-%
-% \begin{proof}
-%   To do.
-% \end{proof}
-%
-%
-% % For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.  We will work extensively with the $1$-subdivision $G^{\onesub}_k$ of $G_k$.
-% %
-% %
-% % In this case, we define $\partial G^{\onesub}_k:= \partial G_k\cup N_{G^{\onesub}_k}[\partial G_k]$.\todo{Define open and closed neighbourhods $N_G(\cdot)$ and $N_G[\cdot]$.}
-% %
-% % \begin{lem}\label{seven_by_seven}
-% %   Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_7$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
-% %   \begin{compactenum}[(i)]\setcounter{enumi}{0}
-% %     \item Each vertex is $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
-% %     \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
-% %     \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
-% %   \end{compactenum}
-% %   Then for any pair of boundary vertices $s$ and $t$, $G_{10}$ contains a path $P$ that
-% %   \begin{compactenum}[(a)]
-% %     \item begins at $s$ and ends at $t$;
-% %     \item contains every vertex in $I\cup L\cup Q$;
-% %     \item goes straight through each vertex in $I$; and
-% %     \item bends at each vertex in $L$.
-% %     % \item contains no boundary vertices or edges except $s$ and $t$.
-% %   \end{compactenum}
-% % \end{lem}
-% %
-% % \begin{proof}
-% %   Case analysis(?).
-% % \end{proof}
-% %
-% % It will be helpful to take some natural definitions for $G_k$ and extend them to $G^{\onesub}_k$.
-To do this, we associate each vertex $v$ of $G^{\onesub}_k$ with a vertex $v^{\leftup}$ of $G_k$.
-For each vertex $v\in V(G_k)$, we define $v^{\leftup}:=v$.
-For each subdivision vertex $s$ in $G^{\onesub}_k$ with neighbours $v$ and $w$, we define $s^{\leftup}$ to be the lexicographically smaller of $v$ or $w$.
-In other words, $s^{\leftup}$ is the vertex of $G_k$ that is directly below or the left of $s$.
-% %
-% % For two vertices $v_1,v_2\in V(G^{\onesub}_k)$, with $v^{\leftup}_1:=(x_1,y_1)$ and $v^{\leftup}_2:=(x_2,y_2)$ we define the $L_\infty$-distance between $v_1$ and $v_2$ as $\|v_1v_2\|_\infty = \max\{|x_1-x_2|, |y_1-y_2|\}$.
-Refer to \cref{ball_figure}.
-For each $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $\beta$ centred at $v$ is $B_{\beta}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le \beta\}$.
-Because it comes up in several places, it is worth noting that $|B_{\beta}(v)|\le 4(\beta+1)^2$, and this bound is tight for each $v\in\{\beta+1,\ldots,\beta-r-1\}\times\{\beta+1,\ldots,b-\beta-1\}$.
-% %
-% % \begin{figure}
-% %   \begin{centre}
-% %     \includegraphics{figs/grid-3}
-% %   \end{centre}
-% %   \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
-% %   \label{ball_figure}
-% % \end{figure}
-% %
-% %
-% % \begin{lem}\label{four_points}
-% %   Let $P$ be a set of at most $3$ vertices of $G^{\onesub}_k$.
-Then there exists a set $S$ of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.\todo{Define the interior of a ball.}
-% % \end{lem}
-% %
-% % \begin{proof}
-% %   Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.
-Let $X:=\bigcup_{i=1}^t \{x_i\bmod 7,(x_{i+1})\bmod 7\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 7,(y_i+1)\bmod 7\}$.
-Then $|X|,|Y|\le 2t\le 6$.
-In particular, there exists some $x\in\{0,\ldots,6\}\setminus X$ and some $y\in\{0,\ldots,6\}\setminus Y$.
-For each $i,j\in\Z$, let $c_{i,j}:=(7i+x+3,7j+y+3)$ and observe that the ball $B_{3}(c_{i,j})$ does not have any point of $P$ on its boundary.
-Furthermore, the set of balls $S:=\{B_{3}(c_{i,j}):i,j\in\Z\}$ is a partition of $V(G^{\onesub}_{k})$.
-Therefore $S$ is a set of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.
-% % \end{proof}
-% %
-% % \begin{lem}
-% %   Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $18$ contains more than $3$ vertices in $P$.
-Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
-% % \end{lem}
-% %
-% % \begin{proof}
-% %   Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$.
-Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $18$.
-Therefore each component of $H$ contains at most $3$ vertices.
-% %
-% %   For each component $C$ of $H$, apply \cref{four_points} to $V(C)$ and keep only those balls in $S$ that contain at least one vertex of $V(C)$.
-This produces a set $S_C$ of disjoint balls, each of radius $3$, that each contain some vertex in $V(C)$ in their interior.
-Any two balls $B_{C_1}$ and $B_{C_2}$ obtained from different components of $H$ are disjoint because the first ball contains a vertex $p_1\in P$ and the second ball contains a vertex $p_2\in P$ such that $\|p_1p_2\|\infty > 12$, but the balls $s_1$ and $s_2$ each have radius $3$.
-% % \end{proof}
-% %
-% %
-% % \begin{lem}\label{one_path_enchilada}
-% %   Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
-% %   \begin{compactenum}[(1)]\setcounter{enumi}{-1}
-% %     \item Any ball of radius 18 contains at most $3$ elements of $I\cup L\cup Q$.
-% %     \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
-% %     \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
-% %     \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
-% %   \end{compactenum}
-% %   Then $G^{\onesub}_k$ contains a path $P$ that
-% %   \begin{compactenum}[(a)]
-% %     \item contains all the vertices in $I\cup L\cup Q$;
-% %     \item goes straight through each vertex in $I$; and
-% %     \item turns at each vertex in $L$.
-% %   \end{compactenum}
-% % \end{lem}
-% %
-% % \begin{proof}
-% %   Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$.
-Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.
-Each such ball is intersected by exactly one row of the snakelike path.
-Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in each ball that it intersects.
-% % \end{proof}
-% %
-% %
-% % \subsection{Approximate Voronoi Diagrams in $G^{\onesub}_k$}
-% %
-% %
-% % % \todo[inline]{This lemma is not true.
-Pick a centre $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.
-Then $x$ is incident on all four Voronoi regions.
-We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
-% % \begin{lem}\label{voronoi_mapping}
-% %   Let $\beta, k\in\N$, let $P\subseteq V(G^{\onesub}_k)$ be such that $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$, and define  $Q:=\bigcup_{p\in P} B_\beta(p)$.
-Then there exists a function $\vor:Q\to P$ such that
-% %   \begin{compactenum}[(i)]
-% %     \item \label{real_close} For each $p\in P$ and each $v\in B_{\beta/8-1}(p)$, $\vor(v)=p$.
-% %     \item \label{never_far} For each $v\in Q$, $\|v\vor(p)\|_\infty \le 7\beta/4$.
-% %     \item \label{three_in_ball} For any ball $B$ of radius $\beta/12$, $|\{\vor(v):v\in B\cap Q\}|\le 3$.
-% %     \item \label{twocolour_neighbourhood} For any $v\in Q$, $|\{\vor(v)\}\cup\{\vor(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
-% %   \end{compactenum}
-% % \end{lem}
-% %
-% % \begin{proof}
-% %   We first define a function $\vor':Q\to P$ that satisfies (\ref{real_close})--(\ref{three_in_ball}) and then show how local modifications can be used to obtain a function $\vor$ that also satisfies (\ref{twocolour_neighbourhood}).
-% %
-% %   Arbitrarily order the vertices of $P$ as $p_1,\ldots,p_{|P|}$.
-For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+3\beta/4\}$, let $i(v):=\min\{i:p_i\in S(v)\}$, and let $\vor'(v)=p_{i(v)}$.
-In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $\beta/12$), and $\vor'(v)$ is the approximate nearest neighbour of $v$ of smallest index.
-% %
-% %   To see that $\vor'$ satisfies (\ref{real_close}), observe that, for each distinct $p,q\in P$ and $v\in B_{\beta/8-1}(p)$, the triangle inequality implies that $\|vq\|_\infty \ge \|pq\|_\infty-\|vp\|_\infty \ge \beta-\beta/8+1 > 7\beta/8$.
-On the other hand $d(v)\le \beta/8$, so $d(v)+3\beta/4\le 7\beta/8$.
-Therefore, $q\not\in S(v)$ for any $q\in P\setminus\{p\}$, so $\vor'(v)=p$.
-% %
-% %   That $\vor'$ satsifies (\ref{never_far}) follows immediately from the fact that each $v\in Q$, so $d(v)\le \beta$, and therefore $\|pv\|_\infty\le 7\beta/4$ for each $p\in S(v)$.
-% %
-% %   We now argue that $\vor'$ satisfies (\ref{three_in_ball}).
-Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.
-Suppose, for the sake of contradiction, that $i_1,\ldots,i_4$ are all distinct and that $v_1,\ldots,v_4$ are contained in some ball $B_{\beta/12}(c)$.
-By the triangle inequality $\|v_iv_j\|_\infty \le \|v_ic\|_\infty+\|cv_j\|_\infty \le \beta/6$ for each  $i,j\in\{1,\ldots,4\}$.
-% %   Since $v_1\in Q$, $d(v)\le\beta$ so
-% %   \[
-% %     \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
-% %   \]
-% %   Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
-% %   \[
-% %     \|p_{i_2}v_2\|_\infty
-% %      < \|p_{i_1}v_2\|_\infty - 3\beta/4
-% %      \le \|p_{i_1}v_1\|_\infty + \beta/6 - 3\beta/4
-% %      \le \beta + \beta/5 \enspace .
-% %   \]
-% %   Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
-% %   \[
-% %      \|p_{i_2}v_3\|_\infty
-% %      < \|p_{i_2}v_3\|_\infty - 3\beta/4
-% %      \le \|p_{i_2}v_2\|_\infty + \beta/6 - 3\beta/4
-% %      < \beta + 2\beta/6 - 3\beta/4 \enspace .
-% %   \]
-% %   Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
-% %   \[
-% %     \|p_{i_4}v_4\|_\infty
-% %     < \|p_{i_3}v_4\|_\infty - 3\beta/4
-% %     \le \|p_{i_3}v_3\|_\infty + \beta/5 - 3\beta/4
-% %     < \beta + 3\beta/6 - 3\beta/2 = 0 \enspace ,
-% %   \]
-% %   which is clearly a contradiction.  This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{TODO: Optimize and cleanup. Finish proving (iii)}
-% % \end{proof}
-%
-% \subsection{Pseudogrids}
-%
-% A graph $H$ is an $a\times b$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $H$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_{a\times b}$.
-The following lemma connecting grid minors and pseudogrids is stated implicitly by \citet{kun.obrien.ea:polynomial}.
-We give a proof here because some of the details of the connection are required in order to prove our results.
-%
-%
-% \begin{lem}\label{pseudogrid_minor}
-%   If a graph $G$ contains a $G_{a\times b}$ grid minor then $G$ contain a subgraph $H$ that is an $a\times b$ pseudogrid.
-% \end{lem}
-%
-% \begin{proof}
-%   (This proof only uses the fact that $G_{a\times b}$ has maximum degree  $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
-%   \begin{inparaenum}[(i)]
-%     \item $B_x$ is non-empty and $H[B_x]$ is connected for each $x\in V(G_{a\times b})$;
-%     \item $H/\mathcal{B}$ is isomorphic to $G_{a\times b}$ using the isomorphism given by $B_x\mapsto x$.
-%   \end{inparaenum}
-%   Let $(H,\mathcal{B})$ be chosen so that $|E(H)|$ is minimum, over all choices of $(H,\mathcal{B})$ that satisfy (i) and (ii).
-%
-%   An edge of $vw$ of $H$ is a \emph{border edge} if $v\in B_x$ and $w\in B_y$ for distinct $x,y\in V(G_{a\times b})$.
-For each $x\in V(G_{a\times b})$, let $\partial B_x$ denote the set of border edges incident on one vertex of $B_x$.
-For each $x\in V(G_{a\times b})$, let $T_x:= H[B_x]\cup \partial B_x$.
-Since $H$ is edge-minimal, $|\partial B_x|=\deg_{G_{a\times b}}(x)\le 4$, for each $x\in V(G_{a\times b})$ and $T_x$ is a tree with exactly $\deg_{G_{a\times b}}(x)$ leaves.\todo{Check the preceding claim.}  We define $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$ as follows:
-% \begin{compactenum}
-%   \item \label{one_special} If $T_x$ has exactly one node $v$ of degree at least three, then $P_x:=\{v\}$.
-%   \item \label{two_degree_three} If $T_x$ has two nodes $v$ and $w$ each of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
-%   \item \label{just_a_path} Otherwise $T_x$ is a path $v_0,\ldots,v_\ell$ for some $\ell\ge 2$ and $P_x:=v_1$.
-% \end{compactenum}
-% It is straightforward to verify that, with this assignment $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$, the quotient graph $H/\mathcal{P}$ is a subdivision of $G_{a\times b}$, so that $H$ is indeed an $a\times b$ pseudogrid.
-% \end{proof}
-%
-% % The second case (\ref{two_degree_three}) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.
-This case is especially important in our proof because it corresponds to a case in which some of the vertices of $P_x$ are shared by the $i$th column and the $j$th row of the pseudogrid.
-However, a path in the pseudogrid that uses vertices of $B_x$ may or may not contain all the vertices of $P_x$ depending on whether or not the corresponding path in $G_{a\times b}$ bends at $x$.
-This motivates the following definitions (see \cref{straight_or_bent}):
-% % \begin{compactenum}
-% %   \item We say that the $x$ is \emph{straight} if the path from $B_{(i-1,j)}$ to $B_{i+1,j}$ in $T_x$ contains all the vertices of $P_x$.
-% %   \item Otherwise, we say that $x$ is \emph{bent}.
-% % \end{compactenum}
-% % \begin{figure}
-% %   \begin{centre}
-% %     \begin{tabular}{c@{\hspace{1cm}}c}
-% %       \includegraphics{figs/bx-paths-1.pdf} & \includegraphics{figs/bx-paths-3} \\
-% %       (a) & (b)
-% %     \end{tabular}
-% %   \end{centre}
-% %   \caption{The vertex $x:=(i,j)$ is (a)~straight if the path from $B_{(i,j-1)}$ to $B_{(i+1,j)}$ uses all vertices of $P_x$ or (b)~bent otherwise.}
-% %   \label{straight_or_bent}
-% % \end{figure}
-% % Note that, if $x$ is straight, then the path from $B_{i,j-1}$ to $B_{i,j+1}$ in $T_x$ also contains all the vertices of $P_x$.
-Furthermore, if $x$ is bent, then, for each $\overline\i,\overline\j\in\{-1,1\}$, the path from $B_{i,j+\overline\j}$ to $B_{i+\overline\i, j}$ in $T_x$ contains all the vertices of $P_x$.
-% %
-% %
-% % \todo[inline]{We can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}
-%
-% \begin{lem}\label{pseudogrid_lower_bound}
-%   For any $k\times k$ pseudogrid $H$ $\chilin(G)\in\Omega(k)$.
-%   % Let $H$ be a $k\times k$ pseudogrid and $\mathcal{P}:=(P_x:x\in V(G_{k}))$ be corresponding partition of $V(H)$ into induced paths.
-%   % \begin{inparaenum}[(i)]
-%   %     \item for each $x\in V(G_k)$, $H[P_x]$ is a path of length at most $r$ and
-%   %     \item $H/\mathcal{P}$ is a $\le\!\! r$-subdivision of $G_k$.
-%   % \end{inparaenum}
-%   % Then
-% \end{lem}
-%
-% \begin{proof}
-%   This proof uses two large integer constants $\alpha$ and $\beta$ and two small positive real constants $\epsilon$ and $\delta$.
-The relationships between these constants are that $\alpha=\Theta(r\beta^2)$  $\epsilon=\Theta(1/\alpha)$.
-Let $\varphi:V(H)\to\{1,\ldots,C\}$ be an arbitrarily $C$-colouring of some $k\times k$ pseudogrid $H$ for some $C\le \epsilon k$.
-We will show that $H$ contains a path $P$ which has no centre with respect to $\varphi$.
-%
-%   First we introduce some notations for vertices of the subdivided grid $G^{\onesub}_k$ to their corresponding paths in the pseudogrid $H$.
-For each vertex $v$ of $G_k$ there is a path $P_v$ in $\mathcal{P}$ (possibly consisting of one vertex) such that the vertex $v$ appears in $H'$ as a result of contracting $P_v$ in $H$.
-In this way, each vertex $v$ of $G_k$ is associated with a non-empty \emph{colour set} $\Phi(v):=\{\varphi(w):w\in V(P_v)\}$.
-Similarly, for each subdivision vertex $s$ of $G^{\onesub}_k$ that comes from subdividing an edge $e$ of $G_k$, there is a path $P_s$ in $H'$ that comes from subdividing $e$ (possibly zero times).
-Thus, each subdivision vertex $s$ of $G^{\onesub}_k$ is associated with a (possibly empty) \emph{colour set} $\Phi(s):=\{\varphi(w):w\in P^-_e\}$, where $P^-_e$ denotes the (possibly empty) set of internal vertices on the path $P_e$.
-%
-%   For each colour $c\in\{1,\ldots,C\}$, define the \emph{multiplicity} of $c$ as the number of vertices $G^{\onesub}_k$ that include $c$ in their colour set.
-Fix some large integer constant $\alpha$ and say that $\alpha$ is \emph{infrequent} if it has multiplicity less than $\alpha$ and \emph{frequent} otherwise.
-%
-%   \begin{clm}\label{all_frequent}
-%     There exists an integer $k'\ge (1-\epsilon)k$ such that
-%     $H$ contains a $k'\times k'$ pseudogrid $H'$ in which every colour in $\{\varphi(v):v\in V(H')\}$ is frequent.
-%   \end{clm}
-%
-%   \begin{proof}[Proof of \cref{all_frequent}]
-%     This proof uses the notion of \emph{removing} a row or column of a pseudogrid.
-Removing a row is defined as follows (removing a column is similar): For each vertex $v$ of $G^{\onesub}_k$ in the relevant row, remove all edges of $P_v$ from $H$.
-Next, remove any isolated vertices from $H$ and finally, repeatedly remove degree-$1$ vertices from $H$ until none remain.
-The result is pseudogrid with one less row.
-%
-%     While there exists some infrequent colour $c\in\{1,\ldots,C\}$, find a row or column of $G^{\onesub}_k$ that contains a vertex with $c$ in its colour set and remove that row or column.
-Doing this exhaustively results in the removal of at most $(\alpha-1) C \le \epsilon\alpha k$ rows and at most $\epsilon\alpha k$ columns of $H$.
-This leaves a pseudogrid $H'$ that contains a $k'\times k'$ pseudogrid for some $k'\ge (1-\epsilon \alpha)k$ for which every colour that appears in at least one vertex of $H'$ is frequent.
-This completes the proof of \cref{all_frequent}.
-%   \end{proof}
-%
-%   Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=\Theta(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.
-To avoid overloading notation, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$ in $H$.
-(The remainder of the proof does not depend on the values of $k$ or $C$.)
-%
-%   % Colour sets turn out to be problematic, so we would like to assign a single colour $c\in\Phi(v)$ to each $v\in G^{\onesub}_k$.
-To do this, consider the bipartite graph $X$ with parts $A:=\{1,\ldots,C\}$ and $B:=V(G^{\onesub}_k)$ and which contains the edge $cv$ if and only if $c\in\Phi(v)$.
-Since each colour in $A$ is frequent, $\deg_x(c)\ge a$ for each $c\in C$.
-S
-%
-%   Next we run a two stage process that, for each $c\in\{1,\ldots,C\}$, will identify a pair $(x_c,y_c)$ where each of $x_c$ and $y_c$ is vertex of $G^{\onesub}_k$ that contains $c$ in its colour set.
-These pairs are carefully chosen so that we will be able to construct a simple path $P_k$ in $G^{\onesub}_k$ that contains every vertex in $X:=\bigcup_{c\in\{1,\ldots,C\}} \{x_c,y_c\}$.
-Furthermore, this path will bend at every bent vertex in $X$ and go straight at every straight vertex in $X$.
-This will ensure that the corresponding path in $H$ contains at least two vertices of each colour.
-%
-%   \begin{enumerate}[{Stage} 1:]
-%     \item We begin this stage by defining each vertex $G^{\onesub}_k$ as being \emph{unmarked}.
-Next we consider each $c\in\{1,\ldots,C\}$ in turn.
-If $G^{\onesub}_k$ contains a pair $(x_c,y_c)$ where
-%     \begin{inparaenum}[(i)]
-%       \item each of $x_c$ and $y_c$ is unmarked;
-%       \item $\|x_cy_c\|_\infty \ge \beta$; and
-%       \item $c\in \Phi(x_c)$ and $c\in \Phi(y_c)$;
-%     \end{inparaenum}
-%     then we say that Stage~1 \emph{succeeeds} for colour $c$ and we mark each vertex in $B_{\beta}(x_c)$ and $B_\beta(y_c)$.
-%
-%     If $G_k$ contains no such pair $(x_c,y_c)$, then we leave $x_c$ and $y_c$ undefined for now and say that Stage~1 \emph{fails} for colour $c$.
-Note that Stage~1 can only fail for $c$ if $G^{\onesub}_k$ contains fewer than $(2\beta+1)^2$ unmarked vertices with $c$ in their colour sets.
-In particular, for $\alpha > (2\beta+1)^2$, Stage~1 succeeds for $c=1$.
-%
-%     \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeds and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 fails.
-Let $R_1:=\bigcup_{c\in S}\{x_c,y_c\}$ be the set of vertices and edges chosen in $R_1$.
-By construction, $\min\{\|pq\|_\infty: \{p,q\}\in \binom{R_1}{2}\}\ge \beta$.
-%
-%     As mentioned above, for sufficiently large $\alpha$ (relative to $\beta$), the set $S$ is non-empty and therefore $R_1$ is non-empty.
-Let $P:=R_1$ and let $Q:=\bigcup_{p\in P} B_\beta(p)$ and apply \cref{voronoi_mapping} to obtain the mapping $\vor:Q\to P$.
-We use to this to construct a bipartite graph $I$ with vertex parts $\overline{S}$ and $R_1$.
-The edge set of $I$ contains an edge $cx$ precisely when there is a marked vertex $z$ of $G^{\onesub}_k$ such that $c\in\Phi(z)$ and $\vor(z)=x$.
-%
-%     First observe that each vertex $x$ of $G^{\onesub}_k$ corresponds to a path $P_x$ in $H$ that contains at most $r$ vertices.\footnote{For some subdivision vertices of $G^{\onesub}_k$, $P_x$ may be empty.}  This implies that $|\Phi(x)|\le r$ for each $x\in V(G^{\onesub}_k)$.
-Next observe that, for each $p\in R_1$, \cref{voronoi_mapping}(\ref{never_far}) implies that $|\vor^{-1}(p)|\le|B_{7\beta/4}(p)|\le (7\beta/2+1)^2$.
-This implies that $\deg_I(p) \le r(7\beta/2+1)^2$ for each $p\in R_1$.
-%
-%     This implies that, for any $A_0\subseteq\overline{S}$,
-%     \[
-%         |N_I(A_0)| \ge \frac{\alpha|A_0|}{r(7\beta/2+1)^2} \ge 2|A_0| \enspace .
-%     \]
-%     for $\alpha\ge 2r(7\beta/2+1)^2$.
-Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.
-Therefore, each $c\in A=\overline{S}$ has two $\Lambda$-neighbours $p,q\in R_1$.
-The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $\vor(x_c)=p$ and $c\in \Phi(x_c)$.
-Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $\vor(y_c)=q$ and $c\in\Phi(y_c)$.
-In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
-%   \end{enumerate}
-%
-%   Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \Phi(x_c)\cap \Phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
-%   Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.
-%
-%   \todo[inline]{There is a technical condition missing here.
-To apply \cref{one_path_enchilada} we need the elements of $R$ to be distance at least $3$ from the boundary of $G^{\onesub}_k$.
-When counting multiplicity, we shouldn't count occurrences of a colour that are close to the boundary.}
-%
-%   Conditions~(\ref{never_far}) and (\ref{three_in_ball}) of \cref{voronoi_mapping} imply that each any ball of radius $\min\{\beta/8-1, \beta/12\}/2$ contains at most three elements of $R$.
-Setting $\beta= 12\cdot2\cdot18 = 436$ (so the conditions on $\alpha$ are satisified for $\alpha = r(7\beta/2+1)^2=2331729r$) gives a point set $R$ that satisfies the requirements for \cref{one_path_enchilada}.
-Therefore, by \cref{one_path_enchilada}, there exists a path $P'$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$.
-These latter two conditions guarantee that the path $P$ in $H$ that corresponds to $P'$ contains $P_x$ for each $x\in R$.
-For each colour $c\in\{1,\ldots,C\}$, there are at least two vertices $x_c,y_c\in R$ such that $P_{x_c}$ and $P_{y_c}$ each contain some vertex of colour $c$.
-Therefore the path $P$ contains two vertices of each colour, so $P$ has no centre and therefore $\varphi$ is not a linear colouring of $H$.
-Since $\varphi$ is an arbitrary $\epsilon k$-colouring of $H$, this implies that $\chilin(H)\ge \epsilon k = \Theta(k/\alpha)=\Theta(k/r)$.
-% \end{proof}
-%
-% To finish the proof, we need some lemma which states that highly subdivided graphs don't have linear colouring numbers that are much smaller than mildly subdivided graphs.
-It would be enough to prove the following.
-%
-% A \emph{refinement} $G'$ of a graph $G$ is a subdivision of $G$ in which each subdivision vertex is adjacent only to degree-$2$ vertices, i.e., $\deg_{G'}(v)=2$ for each $v\in N_{G'}(s)$ and each $s\in V(G')\setminus V(G)$.
-%
-% \begin{lem}\label{refinement_bound}
-%   For any refinement $G'$ of any graph $G$, $\chilin(G') = \Omega(\chilin(G))$.
-% \end{lem}
-%
-% In general, we know that if $G'$ is an $\ell$-subdivision of $G$ then
-% \[
-%     \log\ell \le \chilin(G') \le \chilin(G) + \log\ell \enspace .
-% \]
-% The lower bound comes from the fact that a path of length $\ell$ has a linear chromatic number at least $\log\ell$.
-The upper bound comes by introducing $\log\ell$ additional colours that are used to colour the paths in $G'-G$.
-Of course, the upper bound holds even when $G'$ is a $\le\!\!\ell$-subdivision of $G$ and the lower bound holds as long as at least one edge of $G$ is subdivided $\ell$ or more times.
-%
-% Perhaps \cref{refinement_bound} is true even if we replace ``refinement'' with ``subdivision''.
-As a test case, one could start with the $1$-subdivision $K^{\onesub}_k$ of the $k$-clique $K_k$.
-The following result is certainly true, as can be seen from a treedepth decomposition:
-%
-% \begin{lem}
-%   For any subdivision $G'$ of any graph $G$, $\chicen(G') \ge \chicen(G)$.
-% \end{lem}
-%
-%
-%
-%
-%
-%
-%
-%
-%
+\begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
+  Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.
+Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon$.
+In particular, some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $B_r(\mu)\cap S = \{\mu\}$ for each $\mu\in S$.
+By By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.
+Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
+\end{proof}
 

### 2022-04-05 lin-vs-cen.tex
Pfffft

@@
 
-In this paper, all graphs are simple and undirected. For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$.
+In this paper, all graphs are simple and undirected.
+For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$ denotes the set of vertices and edges of $G$.
+For a vertex $v\in V(G)$ we use the notations $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[H]:=N_G(H)\cup\{v\}$ to denote the open and close neighbourhoods of $v$ in $G$.
+For a set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$ and $N_G[S]:=\bigcup_{v\in S} N_G[v]$.
 
-We make use of the following (polygamous) generalisation of Hall's Marriage Theorem, which follows from Hall's Marriage Theorem by replacing each vertex $x\in X$ with a $d$-tuple $x_1,\ldots,x_d$ each having the same neighbours as $x$.
+For a  bipartite graph $H$, the two parts of $V(H)$ are denoted by $L(H)$ and $R(H)$ and we use the convention that an edge $vw$ has $v\in L(H)$ and $w\in R(H)$.
+A \defin{matching} $M$ in a bipartite graph $H$ is a subgraph of $H$ in which each vertex has degree at most $M$.
+We say that $M$ \defin{saturates} a set $S\subseteq V(H)$ if $\deg_M(v)=1$ for each $v\in S$.
+We make use of Hall's Famous Marriage Theorem (see, for example \citet[Theorem~2.1.2]{diestel:graph}):
+
+\begin{lem}\label{hall}
+  Let $H$ be a bipartite graph with the property that $|N_H(A)|\ge |A|$ for each $A\subseteq L(H)$.  Then $H$ contains a matching that saturates $L(H)$.
+\end{lem}
+
+More specifically, we make use of the following (polygamous) generalisation of \cref{hall}, which follows from \cref{hall} by replacing each vertex $x\in L(H)$ with $d$ vertices $x_1,\ldots,x_d$ each having the same neighbourhood as $x$.
+More precisely, we apply \cref{hall} on the graph $H':=H-\{x\}\cup\{x_iy:i\in\{1,\ldots,d\},\, y\in N_H(x)\}$.
 
 \begin{cor}\label{d_hall}
-  Let $d\ge 1$ be an integer and let $G$ be a bipartite graph with parts $X$ and $Y$ such that, for any $A\subseteq X$, $|N_G(A)|\ge d|A|$.
-Then $G$ contains a subgraph $M$ such that $\deg_M(x)=d$ for each $x\in X$ and $\deg_M(y)\le 1$ for each $y\in Y$.
+  Let $d\ge 1$ be an integer and let $H$ be a bipartite graph with the property that $|N_H(A)|\ge d|A|$ for each $A\subseteq L(H)$.
+Then $H$ contains a subgraph $M$ such that $\deg_M(v)=d$ for each $v\in L(H)$ and $\deg_M(w)\le 1$ for each $y\in R(H)$.
 \end{cor}
 
-We will also make use of the following weighted version of the Lovász Local Lemma:
+Late in the game, we will make use of the following asymmetric version of the Lovász Local Lemma (see, for example, \citet[Lemma~5.1.1]{alon.spencer:probabilistic}):
 
 \begin{lem}\label{weighted_lovasz}
-  Let $E:=\{E_1,\ldots,E_n\}$ be a set of events in some probability space $(\Omega,\Pr)$.
-For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq E$ be such that the event $E_i$ is mutually independent of $E\setminus \Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:E\to(0,1)$ be such that
+  Let $\mathcal{E}:=\{E_1,\ldots,E_n\}$ be a set of events in some probability space $(\Omega,\Pr)$.
+For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq \mathcal{E}$ be such that the event $E_i$ is mutually independent of $\mathcal{E}\setminus \Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:E\to[0,1)$ be such that
   \[
@@
   \]
-  Then $\Pr(\overline{E_1}\cap\cdots\cap\overline{E_n}) > 0$.
+  Then $\Pr(\overline{E}_1\cap\cdots\cap\overline{E}_n) > 0$.
 \end{lem}
-
-% \begin{proof}
-%   \todo[inline]{Surely this is known}
-% \end{proof}
-%
-%
-% \todo[inline]{Surely this next lemma is well-known?}
-% We make use of the following generalization of Hall's Marriage Theorem \cite{hall:on}:
-% \begin{lem}\label{hall_vees}
-%   Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.
-Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
-% \end{lem}
-%
-% \begin{proof}
-%   Let $\Lambda$ be a subgraph of $G$ in which each vertex of $A$ has degree at most two, each vertex of $B$ has degree at most one, and that maximizes $\deg_\Lambda(A)$.
-If $\deg_{\Lambda}(A) = 2|\Lambda|$ then there is nothing to prove.
-Assume therefore, for the sake of contradiction,  that there exists $v_0\in A$ with $\deg_\Lambda(v_0) < 2$.
-%
-%   We say that a path $v_0,\ldots,v_m$ in $G$ is \emph{$\Lambda$-aternating} if $v_iv_{i+1}\in E(\Lambda)$ for all odd $i\in\{0,\ldots,m-1\}$ and $v_iv_{i+1}\not\in E(\Lambda)$ for all even $i\in\{0,\ldots,m\}$.
-Observe that any such path has even length so that it ends at a vertex $v_m\in A$.
-Otherwise, replacing the edges $\lfloor m/2\rfloor$ edges $\{v_1v_2,v_3v_4,\ldots,v_{m-2}v_{m-1}\}\in E(\Lambda)$ with the $\lceil m/2\rceil$ edges $\{v_0v_1,v_2v_3,\ldots,v_{m-1}v_m\}$ does not change $\deg_\Lambda(v)$ for any $v\in A_0\setminus\{v_0\}$ but does increase  $\deg_\Lambda(v_0)$, contradicting the assumption that $\Lambda$ maximizes $\deg_\Lambda(A)$.
-%
-%   Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
-%   Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.
-Let $a$ be the number of edges of $\Lambda$ having one endpoint $A_0$ and one endpoint in $B_0$, so that $a=\deg_\Lambda(B_0)=|B_0|$.
-Let $b$ be the number of edges of $\Lambda$ having one endpoint in $A_0$ and one endpoint in $B\setminus B_0$, so that $\deg_\Lambda(A_0)=a+b = |B_0|+b$.
-Since $\deg_{\Lambda}(v) \le 2$ for each $v\in A_0$ and $\deg_\Lambda(v_0)<2$, $\deg_\Lambda(A_0)<2|A_0|$.
-This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.greater than
-% \end{proof}
-%
-% \begin{rem}
-%   The constant $2$ that appears twice in \cref{hall_vees} can be replaced with any positive integer $c$.
-% \end{rem}
 
@@
 
-% Refer to \cref{grid_figure}.
 For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
+Such an edge is \defin{vertical} if $v=x$ and \defin{horizontal} if $w=y$.
 For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
 
-In this work, we need to work with pseudogrids.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
+\begin{figure}
+  \begin{center}
+    \includegraphics{figs/pseudogrid-2}
+  \end{center}
+  \caption{A $6\times 4$ pseudogrid.  Black vertices are contained in $P_v$ for some vertex $v$ of $G_{6\times 4}$ and white vertices are contained in $P_{vw}$ for some edge of $G_{6\times 4}$.}
+  \label{pseudogrid_fig}
+\end{figure}
+
+In this work, we need to work with pseudogrids.  Refer to \cref{pseudogrid_fig}.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
 \begin{compactitem}
@@
 We call $\mathcal{P}$ a \defin{grid-partition} of $G$.
-% and we denote it by $(P_\mu:\mu\in \VE(G_{a\times b}))$.
-% This defines a mapping $\delta:V(G)\to V(G_{a\times b})\cup E(G_{a\times b})$ where $u\in V(P_{\delta(u)})$ for each $u\in v(G)$.
-
-% Then each vertex $u$ of $G$ naturally corresponds to an edge or vertex $u'$ of $G_{a\times b}$:  If $u$ is a vertex of $P_{vw}$ then $u'=vw$; if $u$ a vertex of $P_v$ then $u'=v$; and otherwise $u'=u$.
-This latter case occurs precisely when $u'$ is one of the boundary vertices of $G_{a\times b}$.
 
@@
 
-We now proceed with the various steps in the proof.
+In \cref{only_frequent_sec} we explain how to remove rows and columns of $G$ to obtain a sub-pseudogrid in which every colour appears frequently.
+In \cref{path_finding} we explain how, given a well-separated set $S$ of vertices in $G$, to find a path that contains every vertex in $S$.
+In \cref{packing_lemma_sec} we introduce fairly standard-looking packing lemma that is needed in several places later.
+In \cref{well_separated} we explain how find a well-separated set $S$ that contains two vertices of each colour.
+Finally \cref{wrapping_up} shows how to assemble these various pieces to prove \cref{pseudogrid_lower_bound}.
 
 \subsection{Pseudogrids with Exclusively Frequent Colours}
+\label{only_frequent_sec}
 
@@
 (Throughout this section, we will use the idiom $\varphi(V(G))$ to denote the set of all colours used by $\varphi$ to colour the vertices of $G$.)
-
-% We extend $\varphi$ to a partial colouring $\phi:V(G_{k\times k})\cup E(G_{k\times k})\to\{1,\ldots,c,\perp\}$ of the edges and vertices of $G_{k\times k}$ as follows:  For each vertex $v$ of $G_{k\times k}$, define $\phi(v)$ be an arbitrary colour in $\{\varphi(x):x\in V(P_v)\}$.
-For each edge $vw$ of $G_{k\times k}$, there are two possibilities:
-% If $P_{vw}$ is empty then $\phi(vw):=\perp$, otherwise $\phi(vw)$ is an arbitrary colour in $\{\varphi(x):x\in V(P_{vw})\}$.
 
@@
     \item For each horizontal edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
-    \item For each vertical edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.\todo{Define horizontal and vertical edges of $G$.}
+    \item For each vertical edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
     \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
@@
 \subsection{Finding Paths Through Well-Separated Pairs}
+\label{path_finding}
 
-Next we show that, given a sufficiently ``well-separated'' set $S$ of pairs of vertices in $G$, we can always find a path in $G$ that contains every vertex in $S$.
+Next we show that, given a sufficiently `well-separated' set $S$ of pairs of vertices in $G$, we can always find a path in $G$ that contains every vertex in $S$.
 For this, we need some definitions of boxes in $G$ and in $G_k$.
@@
 
-Before finishing with this section we recall a fairly standard looking packing lemma.
+\subsection{A Packing Lemma}
+\label{packing_lemma_sec}
+
+We will make use of the following fairly standard looking packing lemma.
 
@@
 
-\todo[inline]{Fix notation conflict:  $G_r(\mu)$ versus $G_k$?  Maybe go back to $G_{k\times k}$?}
-
-\subsection{Finding an Uncentred Path}
+\subsection{Finding a Well-Separated Set}
+\label{well_separated}
 
@@
 Furthermore all but at most one element of $\Phi_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since, otherwise the first round would have succeeded for colour $\alpha$.
-For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$.
-Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.
+  % For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$. Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.
 
@@
   \[
-     Y_\nu := \{\tau\in Y\setminus\{\mu\}:\nu\in\VE(G_r(\tau))\}
+     Y_\nu := \{\tau\in Y:\nu\in\VE(G_{2r}(\tau))\}
   \]
   has size at most $25$.
-Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any element of $Y_\nu$.
+Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.
 This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
-This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.
+   % This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.
 
   In order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.
-To do this, consider the graph $J$ with vertex set $V(J)=Y$ and which contains an edge $\mu_1\mu_2$ if $B_{2r}(\mu_1)\cap B_{2r}(\mu_2)\neq\emptyset$.
-Now, consider the greedy process of repeatedly choosing a vertex $\mu$ of $J$ such that $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.
-This process continues until no suitable vertex of $J$ remains.
-Each step in this process eliminates $O(r^2)$ objects $\nu$ with $\varphi(\nu)=\alpha$,\todo{Explain better} so the number of iterations is $r\in\Omega(d/r^2)$.
+To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $B_r(\mu)\cap J=\emptyset$ for each $\mu\in Y$.
+Since $|\VE(G_{10r}(\mu))|\in O(r^2)$, each step in this process eliminates $O(r^2)$ objects from $J$ so the number $t$ of iterations is in $\Omega(|\Phi_\alpha|/r^2)\subseteq \Omega(d/r^2)$.
 
-  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$, for each $i\in\{1,\ldots,t\}$.
-The important observation now is that the sets $X_{\nu_1},\ldots,X_{\nu_t}$ are disjoint.
-Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears before any elements of $X_{\nu_i}$ in our random permutation of $Y$'' then the events $U_1,\ldots,U_k$ are mutually independent.
-Indeed, each $U_i$ depends only on the relative order of $\{\mu_i'\}\cup X_{\nu_i}$ in the random permutation of $Y$.
+  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be the corresponding elements of $\Phi_\alpha$.
+The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are disjoint.
+Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_k$ are mutually independent.
+Indeed, each $U_i$ depends only on the relative order of $Y_{\nu_i}$ in the random permutation of $Y$.
 For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.
-Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/25)$ random variable.
+Therefore, $\sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/25)$ random variable.
 Therefore,
@@
 Recall that we are working in the regime in  $t\in\Omega(|\Phi_\alpha|/r^2)$ and $|\Phi_\alpha|\in\Omega(d/r^2)$ so, for any fixed $C>0$, there exists $d\in\Omega(r^4)$ such that
-  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-C|\Phi_\alpha|/r^2)$.
-  \todo[inline]{Work the preceding calculation out}
-  We are now ready for an application of \cref{weighted_lovasz}.\todo{Continue here}  For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $\Phi_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.
-The event $X_\alpha$ is mutually independent of $\{X_\beta:\beta\in X\setminus\{\alpha\}\setminus\Gamma_\alpha\}$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.
+  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-C|\Phi_\alpha|/r^2)$.\todo{This could be done better.} Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  By the preceding discussion, $\Pr(E_\alpha)\le \exp(-C|\Phi_\alpha|/r^2)$.
+
+  We are now ready for an application of \cref{weighted_lovasz}.  Let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$ and, for each $\alpha\in X$, define
+  \[
+     \Gamma_\alpha : \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\neq\emptyset\right\} \enspace .
+  \]
+  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
+
+  For each $\alpha\in X$, let $w(E_\alpha):=1-e^{-C/50r^2}$.  Then
+  \begin{align}
+    w(E_\alpha)\prod_{E_\beta\in\Gamma_\alpha} (1-w(E_\beta))
+    & = (1-e^{-C/50r^2})e^{-C|\Gamma_\alpha|/50r^2} \\
+    & = e^{-C|\Gamma_\alpha|/50r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2} \\
+    & \ge e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2} \enspace . \label{ratioo}
+    % & \ge (1-e^{-C/50r^2})e^{-C|\Phi_\alpha|/2r^2} \\
+    % & \ge (1-e^{-C/50r^2})e^{-C|\Phi_\alpha|/2r^2} \\
+  \end{align}
+  Taking the ratio of \cref{ratioo} and $\Pr(E_\alpha)$ we obtain
+  \begin{align*}
+    \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{\Pr(E_\alpha)}
+    & \ge \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
+    & \ge e^{C|\Phi_\alpha|/2r^2} - \frac{e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
+  \end{align*}
+
+  \todo[inline]{Hmmmmm. Not working.  Try something like $w(E_\alpha)=1-e^{-1/50}$}
+
+  Finally, by \cref{packing_lemma} $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.
 
@@
 
+\subsection{Wrapping Up}
+\label{wrapping_up}
+
 We now have all the pieces needed to complete our lower bound on the linear chromatic number of pseudogrids:

### 2022-04-05 lin-vs-cen.tex
Finally a full application of the LLL

@@
 % Observe that, for any $\mu\in V(G_{k})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.
-Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.
-We extend these definitions to vertices of an $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows.
-For any $\mu\in V(G_k)$, define
+Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.  Straightforward counting shows that, for any $\mu\in\VE(G_k)$,
+\begin{equation}
+   |\VE(G_r(\mu))| \le 3(2r+1)^2 \enspace . \label{rbox_size} % rbox size
+\end{equation}
+We extend these definitions to vertices of a $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in V(G_k)$, define
 \[
@@
 
-  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\in \Omega(d/r^2)$ for each $\alpha\in X$.
+  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 6(2r+1)^2$ for each $\alpha\in X$.
   For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.
@@
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\Omega(|\Phi_\alpha|/r^2),1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 100r^2)$.
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/3(20r+1)^2},1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 6(2r+1)^2)$.
 (We make no attempt to optimize the constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$.
@@
 To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $B_r(\mu)\cap J=\emptyset$ for each $\mu\in Y$.
-Since $|\VE(G_{10r}(\mu))|\in O(r^2)$, each step in this process eliminates $O(r^2)$ objects from $J$ so the number $t$ of iterations is in $\Omega(|\Phi_\alpha|/r^2)\subseteq \Omega(d/r^2)$.
+Since $|\VE(G_{10r}(\mu))|\le 3(20r+1)^2$, each step in this process eliminates at most $3(20r+1)^2$ objects from $J$ so the number $t$ of iterations is at least $\ceil{|\Phi_\alpha|/3(20r+1)^2}$
 
@@
 Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_k$ are mutually independent.
-Indeed, each $U_i$ depends only on the relative order of $Y_{\nu_i}$ in the random permutation of $Y$.
+Indeed, each $U_i$ depends only on the relative order of $Y_{\nu_i}$ in the  permutation $\pi$.
 For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.
-Therefore, $\sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/25)$ random variable.
-Therefore,
-  \begin{align*}
-    \Pr\left(\deg_H(\alpha)< 100r^2\right)
-    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
-    & \le \Pr(\binomial(t,1/25)\le 100r^2) \\
-    & \le (2t)^{100r^2}(1/25)^{r^2}(1-1/25)^{t-100r^2} \\
-    & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
-    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
-    & = \exp(ar^2 - bt)
-  \end{align*}
-  for fixed positive $a,b>0$.
+Therefore, $\sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/25)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
+  \begin{align}
+    \Pr\left(\deg_H(\alpha)< 6(2r+1)^2\right) 
+    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 6(2r+1)^2\right) \notag \\
+    & \le \Pr(\binomial(t,1/25)\le 6(2r+1)^2) \notag \\
+    & \le (2t)^{6(2r+1)^2}(1/25)^{r^2}(1-1/25)^{t-6(2r+1)^2} \notag \\
+    % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
+    & \le \exp(6(2r+1)^2(1+\log(2t))-t/25) \notag \\
+    & \le \exp(6(2r+1)^2(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/75(20r+1)^2) \label{probability}
+    % & \le \exp(-6(2r+1)^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
+    % & = \exp(ar^2 - bt)
+  \end{align}
+  % for fixed positive $a,b>0$.
 Recall that we are working in the regime in  $t\in\Omega(|\Phi_\alpha|/r^2)$ and $|\Phi_\alpha|\in\Omega(d/r^2)$ so, for any fixed $C>0$, there exists $d\in\Omega(r^4)$ such that
-  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-C|\Phi_\alpha|/r^2)$.\todo{This could be done better.} Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  By the preceding discussion, $\Pr(E_\alpha)\le \exp(-C|\Phi_\alpha|/r^2)$.
+  % $\Pr(\deg_H(\alpha)< 6(2r+1)^2)\le \exp(-C|\Phi_\alpha|/r^2)$.\todo{This could be done better.}   By the preceding discussion, $\Pr(E_\alpha)\le \exp(-C|\Phi_\alpha|/r^2)$.
 
-  We are now ready for an application of \cref{weighted_lovasz}.  Let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$ and, for each $\alpha\in X$, define
+  We are now ready for an application of \cref{weighted_lovasz}.
+For each $\alpha\in X$, let Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 6(2r+1)^2$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
+For each $\alpha\in X$, define
   \[
-     \Gamma_\alpha : \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\neq\emptyset\right\} \enspace .
+     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\neq\emptyset\right\} 
   \]
+  and observe that, by \ref{rbox_size}, $|\Gamma_\alpha|\le 3(20r+1)^2|\Phi_\alpha$.
   The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
 
-  For each $\alpha\in X$, let $w(E_\alpha):=1-e^{-C/50r^2}$.  Then
+  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>0$ to be determined shortly.  Then
   \begin{align}
     w(E_\alpha)\prod_{E_\beta\in\Gamma_\alpha} (1-w(E_\beta))
-    & = (1-e^{-C/50r^2})e^{-C|\Gamma_\alpha|/50r^2} \\
-    & = e^{-C|\Gamma_\alpha|/50r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2} \\
-    & \ge e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2} \enspace . \label{ratioo}
-    % & \ge (1-e^{-C/50r^2})e^{-C|\Phi_\alpha|/2r^2} \\
-    % & \ge (1-e^{-C/50r^2})e^{-C|\Phi_\alpha|/2r^2} \\
+    & = \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|-1} \notag \\
+    & \ge \exp\left(-\log\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
+    & \ge \exp\left(-\log\tau - 3(20r+1)^2|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
   \end{align}
-  Taking the ratio of \cref{ratioo} and $\Pr(E_\alpha)$ we obtain
-  \begin{align*}
-    \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{\Pr(E_\alpha)}
-    & \ge \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
-    & \ge e^{C|\Phi_\alpha|/2r^2} - \frac{e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
-  \end{align*}
 
-  \todo[inline]{Hmmmmm. Not working.  Try something like $w(E_\alpha)=1-e^{-1/50}$}
+  Taking $\tau = 450(20r+1)^4+1$, and using \cref{weight_product,probability}, we find that
+  \[
+     \Pr\left(E_\alpha\right) \le w(E_\alpha)\prod_{\beta\in\Gamma_\alpha}(1-w(\Gamma_\alpha))
+  \]
+  provided that 
+  \[   |\Phi_\alpha| \ge 150(20r+1)^2\left(\log\left(450(20r+1)^4+1\right) + 6(2r+1)^2(1+\log(2|\Phi_\alpha|)) \right) \enspace .
+  \]
+  The right hand side of this last equation is of the form $O(r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4)$.
 
-  Finally, by \cref{packing_lemma} $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.
+  % Taking the ratio of \cref{ratioo} and $\Pr(E_\alpha)$ we obtain
+  % \begin{align*}
+  %   \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{\Pr(E_\alpha)}
+  %   & \ge \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
+  %   & \ge e^{C|\Phi_\alpha|/2r^2} - \frac{e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
+  % \end{align*}
+  % 
+  % \todo[inline]{Hmmmmm. Not working.  Try something like $w(E_\alpha)=1-e^{-1/50}$}
+  % 
+  % Finally, by \cref{packing_lemma} $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.
+  % 
+  % \begin{align*}
+  %     \Pr(E_\alpha) \le \exp(-|\Phi_\alpha|\cdot C/r^2) \enspace .
+  % \end{align*}
+  % 
+  % For each $\alpha\in X$, define $w(E_\alpha)=C/2r^2$ and observe that
+  % \begin{align*}
+  %   p(1-p)^{|\Gamma_\alpha|}
+  %   & \ge  p(1-p)^{25|\Phi_\alpha|} = p\exp(|\Phi_\alpha|\cdot 25\log(1-c/2r^2))
+  % \end{align*}
+  % \todo[inline]{Sort out these calculations}
 
-  \begin{align*}
-      \Pr(E_\alpha) \le \exp(-|\Phi_\alpha|\cdot C/r^2) \enspace .
-  \end{align*}
-
-  For each $\alpha\in X$, define $w(E_\alpha)=C/2r^2$ and observe that
-  \begin{align*}
-    p(1-p)^{|\Gamma_\alpha|}
-    & \ge  p(1-p)^{25|\Phi_\alpha|} = p\exp(|\Phi_\alpha|\cdot 25\log(1-c/2r^2))
-  \end{align*}
-  \todo[inline]{Sort out these calculations}
-
-  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge 100r^2$ for each $\alpha\in X$.
-On the other hand, $\deg_{H}(y)\le 8r^2$ for each $y\in Y$.
-Therefore, for any $A\subseteq X$, $|N_H(A)|\ge 100|A|/8 \ge 2|A|$.
+  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge 6(2r+1)^2$ for each $\alpha\in X$.
+On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 \le 3(2r+1)^2$ for each $y\in Y$.
+Therefore, for any $A\subseteq X$, $|N_H(A)|\ge 6(r+1)^2|A|/3(2r+1)^2 = 2|A|$.
 Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$.
@@
 In particular, some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $B_r(\mu)\cap S = \{\mu\}$ for each $\mu\in S$.
-By By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.
+By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.
 Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.

### 2022-04-05 lin-vs-cen.tex
Pfffft

@@
 
-For a  bipartite graph $H$, the two parts of $V(H)$ are denoted by $L(H)$ and $R(H)$ and we use the convention that an edge $vw$ has $v\in L(H)$ and $w\in R(H)$.
+For a  bipartite graph $H$, the two parts of $V(H)$ are denoted by $L(H)$ and $R(H)$ and we use the convention of writing an edge $xy$ so that its first endpoint $x$ is in $L(H)$ and its second endpoint $y$ is in $R(H)$.
 A \defin{matching} $M$ in a bipartite graph $H$ is a subgraph of $H$ in which each vertex has degree at most $M$.
 We say that $M$ \defin{saturates} a set $S\subseteq V(H)$ if $\deg_M(v)=1$ for each $v\in S$.
-We make use of Hall's Famous Marriage Theorem (see, for example \citet[Theorem~2.1.2]{diestel:graph}):
+We make use of (the difficult half of) Hall's Marriage Theorem (see, for example \citet[Theorem~2.1.2]{diestel:graph}):
 
-\begin{lem}\label{hall}
+\begin{thm}[\citet{hall:on}]\label{hall}
   Let $H$ be a bipartite graph with the property that $|N_H(A)|\ge |A|$ for each $A\subseteq L(H)$.  Then $H$ contains a matching that saturates $L(H)$.
-\end{lem}
+\end{thm}
 
-More specifically, we make use of the following (polygamous) generalisation of \cref{hall}, which follows from \cref{hall} by replacing each vertex $x\in L(H)$ with $d$ vertices $x_1,\ldots,x_d$ each having the same neighbourhood as $x$.
+More specifically, we make use of the following (polygamous) version of the Marriage Theorem, which is obtained by replacing each vertex $x\in L(H)$ with $d$ vertices $x_1,\ldots,x_d$ each having the same neighbourhood as $x$.
 More precisely, we apply \cref{hall} on the graph $H':=H-\{x\}\cup\{x_iy:i\in\{1,\ldots,d\},\, y\in N_H(x)\}$.
@@
   Let $d\ge 1$ be an integer and let $H$ be a bipartite graph with the property that $|N_H(A)|\ge d|A|$ for each $A\subseteq L(H)$.
-Then $H$ contains a subgraph $M$ such that $\deg_M(v)=d$ for each $v\in L(H)$ and $\deg_M(w)\le 1$ for each $y\in R(H)$.
+Then $H$ contains a subgraph $M$ such that $\deg_M(v)=d$ for each $x\in L(H)$ and $\deg_M(y)\le 1$ for each $y\in R(H)$.
 \end{cor}
@@
 
+  Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
   Taking $\tau = 450(20r+1)^4+1$, and using \cref{weight_product,probability}, we find that

### 2022-04-05 lin-vs-cen.tex
Pffffffft

@@
 For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$ denotes the set of vertices and edges of $G$.
-For a vertex $v\in V(G)$ we use the notations $N_G(v):=\{w\in V(G):vw\in E(G)\}$ and $N_G[H]:=N_G(H)\cup\{v\}$ to denote the open and close neighbourhoods of $v$ in $G$.
-For a set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$ and $N_G[S]:=\bigcup_{v\in S} N_G[v]$.
+For a vertex $v\in V(G)$ we use the notations $N_G(v):=\{w\in V(G):vw\in E(G)\}$  to denote the open neighbourhood of $v$ in $G$.
+For a set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$.
 
@@
 For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.
-In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths and the quotient graph $G/\mathcal{P}$ is isomorphic to $G_{a\times b}$.
+In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths is isomorphic to a $\le\!
+1$-subdivision of $G_{a\times b}$.
 We call $\mathcal{P}$ a \defin{grid-partition} of $G$.
@@
 
+\todo[inline]{Consider constructing this set so that colours that appear on the boundary don't count.  This just makes the factor of $d$ jump to a $d+2$.}
+
 \begin{lem}\label{only_frequent}
@@
   that satisfies the conditions of the lemma.
+\end{proof}
+
+\begin{lem}
+  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid, let $X\subseteq V(G)$ denote the set of vertices in the first and last row and column of $G$, and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/d$ colours.
+  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - d|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
+  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus X| \ge d|A|$.
+\end{lem}
+
+\begin{proof}
+
 \end{proof}
@@
 \begin{lem}\label{pick_up_two}
-  Let $v$ be any vertex of $G$, let $w$ be any vertex of $\tilde{G}_r(v)$, let $s$ be any vertex in the leftmost column of $\tilde{G}_r(v)$ and let $t$ be any vertex in the rightmost column of $\tilde{G}_r(v)$.
-Then there exists an $st$-path\todo{define $st$-path} in $G_r(v)$ that contains $v$ and $w$.
+  Let $G$ be a $k\times k$ pseudogrid with $k\ge??$,  let $s,v,w,t$ be vertices of $G$ with $s$ in the first column of $G$, neither $v$ nor $w$ in the first or last column of $G$ and $t$ in the last column of $G$.
+Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
 \end{lem}
@@
 \begin{proof}
-  Let $i_s \le i_v \le i_w \le i_t$ be the columns of $\tilde{G}_r(v)$ that contain $s$, $v$, $w$, and $t$, respectively.
-Note that the assumption that $i_v \le i_w$ is without loss of generality and that $i_s \le i_t - r -1$.\todo{finish up}
+  Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_k)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.  
 \end{proof}
@@
 \begin{lem}\label{pick_up_everything}
-  Let $r$ be an integer and let $S\subseteq V(G)$ be such that $|\tilde{B}_{r}(v)\cap S|=|\tilde{B}_{2r}\cap S|\le 2$ for each $v\in S$.
+  Let $r\ge 12$ be divisible by $3$ and let $S\subseteq V(G)$ be such that $|\tilde{B}_{r}(v)\cap S|=|\tilde{B}_{2r}\cap S|\le 2$ for each $v\in S$.
 Then $G$ contains a path that contains every vertex in $S$.
-\todo{Set specific values for $r$}
 \end{lem}
-
-\begin{proof}[Proof sketch]
-  Choose $S'\subseteq S$ so that, if $B_r(v)\cap S=\{v,w\}$ then exactly one of $v$ or $w$ is in $S'$.
-Now do some scan-line thing with a distance of $2r+1$ between consecutive rows.
-When this scan encounters a box $B_r(v)$ for some $v\in S'$, use \cref{pick_up_two} to detour through the box to pick up the (at most two) vertices in $B_r(v)\cap S$.
-\end{proof}
 
@@
 \begin{lem}\label{packing_lemma}
-  Let $Q\subseteq \VE(G_k)$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
+  Let $r\ge 114$ be an even integer and let $Q\subseteq \VE(G_k)$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
   $|\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}| \le 25$.
 \end{lem}
+
+\begin{proof}
+  The packing condition that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $B_{r/2-1}(\mu_1)\cap B_{r/2-1}(\mu_2)=\emptyset$ for each $\mu_1,\mu_2\in Q$.
+
+  For simplicity, assume that $G_k$ is embedded in $G_{k+2r+1}$ with the elements of $Q$ at distance at least $G_{2r}$ from the first and last row and column.
+In particular, for each $\mu_1\in Q$, $|B_{\rho}(\nu)| = 3(2\rho+1)^2$ for any $\rho\in\{0,\ldots,2r\}$.
+
+  Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}$.  For any $\mu_1\in Q'$, 
+  $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.  
+  Since this is true for all $\mu_1\in Q$ we have
+  \[  3(5r+1)^2 \ge |\VE(G_{3r/2}(\mu))| \ge \sum_{\mu_1\in Q} B_{r/2}(\mu_1)
+  = 3|Q'|(r-2)^2 \enspace .
+  \]
+  which implies that $|Q'|\le (5r+1)^2/(r-2)^2$.  For $r > 2$, this bound is decreasing in $r$ and is less than $26$ for any $r> 11\sqrt{26} + 57$.  In particular, $|Q'|< 26$ for any $r\ge 114$.
+\end{proof}
 
@@
   \[
-     Y_\nu := \{\tau\in Y:\nu\in\VE(G_{2r}(\tau))\}
+     Y_\nu := \{\xi\in Y:\nu\in\VE(G_{2r}(\xi))\}
   \]

### 2022-04-06 lin-vs-cen.tex
Boundary work

@@
 
-\newcommand{\defin}[1]{\emph{\color{brown}#1}}
+\definecolor{brightmaroon}{rgb}{0.76, 0.13, 0.28}
+\definecolor{linkblue}{rgb}{0, 0.337, 0.227}
+\newcommand{\defin}[1]{\emph{\color{brightmaroon}#1}}
 \setlength{\parskip}{1ex}
@@
 \begin{abstract}
-    These are some notes on the relationships betwee linear colouring numbers and centred colouring numbers.
+  We prove that linear colouring number of any $k\times k$ pseudogrid is $\Omega(k)$.
+By an argument of Kun et al, this result gives a tighter upper bound on the treedepth of a graph as a function of its linear colouring number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear colouring number.
 \end{abstract}
@@
 Let $G$ be a simple undirected graph.
-A \emph{$k$-colouring} of $G$ is any function $\varphi:V(G)\to S$ where $S$ is a set of size $k$.
-A vertex $v$ of $G$ is a \emph{centre} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.
-A colouring $\varphi$ of $G$ is \emph{centred} if every connected subgraph of $G$ has a centre with respect to $\varphi$.
-A colouring $\varphi$ of $G$ is \emph{linear} if every simple path in $G$ has a centre with respect to $\varphi$.
+A \defin{$k$-colouring} of $G$ is any function $\varphi:V(G)\to S$ where $S$ is a set of size $k$.
+A vertex $v$ of $G$ is a \defin{centre} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.
+A colouring $\varphi$ of $G$ is \defin{centred} if every connected subgraph of $G$ has a centre with respect to $\varphi$.
+A colouring $\varphi$ of $G$ is \defin{linear} if every simple path in $G$ has a centre with respect to $\varphi$.
 
-The \emph{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.
+The \defin{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.
 We note that an easy argument shows that centred chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$.
-The \emph{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.
+The \defin{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.
 Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.
 
-The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet[Theorem~1]{kun.obrien.ea:polynomial} who showed the following result:
+The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet{kun.obrien.ea:polynomial} who showed the following result:
 
@@
 
-For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
+For positive integers $a$ and $b$, the \defin{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
 Such an edge is \defin{vertical} if $v=x$ and \defin{horizontal} if $w=y$.
-For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
+For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
+The \defin{boundary} $\partial G_{a\times b}$ of $G_{a\times b}$ a the set of vertices in the first and last row and column and all edges incident to these vertices.
 
@@
 We call $\mathcal{P}$ a \defin{grid-partition} of $G$.
+The \defin{boundary} of $G$ is defined as $\partial G:=\bigcup_{\mu\in\partial G_{a\times b}} V(P_\mu)$.
 
@@
 
-  \item We first delete rows and columns from $G$ so that each colour that appears in $G$ appears at least $C$ times, for some large constant $C$.
+  \item We first delete rows and columns from $G$ so that each colour that appears in $G$ appears at least $d$ times in the interior of $G$, for some large constant $d$.
 From this point on the goal is to construct $P$ so that it contains each colour at least twice.
@@
 
-\todo[inline]{Consider constructing this set so that colours that appear on the boundary don't count.  This just makes the factor of $d$ jump to a $d+2$.}
-
 \begin{lem}\label{only_frequent}
-  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/d$ colours.
-  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - d|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
-  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d|A|$.
+  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/(d+2)$ colours.
+  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - (d+2)|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
+  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus \partial G_k| \ge d|A|$.
 \end{lem}
@@
   The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.
-If there exists no $A\subseteq \varphi(V(G))$ with $\varphi_{\mathcal{P}}^{-1}(A) < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
-Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|$ to eliminate all vertices with colours in $A$, as follows:
+If there exists no $A\subseteq \varphi(V(G))$ with $|\varphi_{\mathcal{P}}^{-1}(A)\setminus\partial G_k| < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
+Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le (d+2)|A|$ to eliminate all vertices with colours in $A$, as follows:
   \begin{compactitem}
-    \item For each $v:=(i,j)\in V(G_{k})$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
-    \item For each horizontal edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
-    \item For each vertical edge $vw\in E(G_{k\times k})$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
+    \item For each $v:=(i,j)\in V(G_{k})\setminus \partial G_k$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
+    \item For each horizontal edge $vw\in E(G_{k\times k})\setminus \partial G_k$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
+    \item For each vertical edge $vw\in E(G_{k\times k})\setminus\partial G_k$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
+    \item We add the first and last row to $R$ and the first and last column to $C$.
     \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
   \end{compactitem}
-  At this point $|R|=|C|\le d|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
+  At this point $|R|=|C|\le (d+2)|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
 In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.
@@
   Now apply induction on $G_0$ to get a $k'\times k'$ pseudogrid with
-  \[  k'\ge k_0-d|\varphi(V(G_0))|  \ge k-d|A|-d|\varphi(V(G_0))|\ge k - d|\varphi(V(G))|
+  \[
+    k'\ge k_0-(d+2)|\varphi(V(G_0))|
+      \ge k-(d+2)|A|-(d+2)|\varphi(V(G_0))|
+      \ge k - (d+2)|\varphi(V(G))|
   \]
   that satisfies the conditions of the lemma.
-\end{proof}
-
-\begin{lem}
-  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid, let $X\subseteq V(G)$ denote the set of vertices in the first and last row and column of $G$, and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/d$ colours.
-  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - d|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
-  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus X| \ge d|A|$.
-\end{lem}
-
-\begin{proof}
-
 \end{proof}
@@
 \begin{lem}\label{one_colour_per_object}
-  Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)| \ge d\cdot|A|$.
+  Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus \partial G_k| \ge d\cdot|A|$.
 Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
   \begin{compactenum}[(i)]
-    \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
-    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
+    \item For each $\mu\in\partial G_{k}$, $\phi(\mu)=\perp$.
+    \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi_\mathcal{P}(\mu)$.
+    \item For each $\alpha\in\varphi(V(G))$,
+    $|\phi^{-1}(\alpha)|\ge d$.
+    % $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
   \end{compactenum}
@@
 \begin{proof}
-  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(G_k)$ and edge set
+  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(G_k)\setminus\partial G_k$ and edge set
   \[
@@
 \begin{lem}\label{pick_up_two}
-  Let $G$ be a $k\times k$ pseudogrid with $k\ge??$,  let $s,v,w,t$ be vertices of $G$ with $s$ in the first column of $G$, neither $v$ nor $w$ in the first or last column of $G$ and $t$ in the last column of $G$.
+  Let $G$ be a $a\times a$ pseudogrid with $a\ge??$,  let $s,v,w,t$ be vertices of $G$ with $s$ in the first column of $G$, neither $v$ nor $w$ in the first or last column of $G$ and $t$ in the last column of $G$.
 Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
@@
   Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_k)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.
+  todo[inline]{Finish this.}
 \end{proof}
@@
 \begin{lem}\label{pick_up_everything}
-  Let $r\ge 12$ be divisible by $3$ and let $S\subseteq V(G)$ be such that $|\tilde{B}_{r}(v)\cap S|=|\tilde{B}_{2r}\cap S|\le 2$ for each $v\in S$.
+  Let $r\ge ??$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(G)\setminus \partial G$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$\todo{Set constraint on $R$} for each $v\in S$.
 Then $G$ contains a path that contains every vertex in $S$.
 \end{lem}
+
+\begin{proof}[Proof Sketch]
+  We adjust things so that we get a set of disjoint boxes, each of which contains one or two points of $S$ in its interior.
+  We do the scan-line thing depicted in \cref{outline}.
+\end{proof}
 
@@
 
-  Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}$.  For any $\mu_1\in Q'$, 
+  Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu_1))\}$.  For any $\mu_1\in Q'$,
   $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.
   Since this is true for all $\mu_1\in Q$ we have
-  \[  3(5r+1)^2 \ge |\VE(G_{3r/2}(\mu))| \ge \sum_{\mu_1\in Q} B_{r/2}(\mu_1)
+  \[  3(5r+1)^2 \ge |\VE(G_{5r/2}(\mu))| \ge \sum_{\mu_1\in Q'} |\VE(G_{r/2}(\mu_1))|
   = 3|Q'|(r-2)^2 \enspace .
@@
   Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.
-Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon$.
-In particular, some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $B_r(\mu)\cap S = \{\mu\}$ for each $\mu\in S$.
+Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon-2$.
+In particular, for some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
 By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.

### 2022-04-06 lin-vs-cen.tex
Switched to symbolic volume computations

@@
 \usepackage{bbm}  % needed for \mathbbm{1}
+\usepackage{listings}
 
@@
 
+\newcommand{\vol}[1]{\boxplus_{#1}}
+
 % Taken from
@@
 \begin{abstract}
-  We prove that linear colouring number of any $k\times k$ pseudogrid is $\Omega(k)$.
+  We prove that the linear colouring number of any $k\times k$ pseudogrid is $\Omega(k)$.
 By an argument of Kun et al, this result gives a tighter upper bound on the treedepth of a graph as a function of its linear colouring number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear colouring number.
@@
 Let $G$ be a simple undirected graph.
-A \defin{$k$-colouring} of $G$ is any function $\varphi:V(G)\to S$ where $S$ is a set of size $k$.
+A \defin{$k$-colouring} of $G$ is any function $\varphi:V(G)\to C$ where $C$ is a set of size $k$.
 A vertex $v$ of $G$ is a \defin{centre} of $G$ with respect to $\varphi$ if $\varphi(v)\not\in\{\varphi(w):w\in V(G)\setminus\{v\}\}$, i.e., $v$ is the unique vertex of $G$ having colour $\varphi(v)$.
@@
 The \defin{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.
-We note that an easy argument shows that centred chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$.
+An easy argument shows that centred chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$ \cite{nesetril.ossona:sparsity}.
 The \defin{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.
@@
 
-The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet{kun.obrien.ea:polynomial} who showed the following result:
+The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet{kun.obrien.ea:polynomial,czerwinski.nadara.ea:improved} who prove:
 
-\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general}
-  For any graph $G$, $\chicen(G)\le \chilin^{19}(G)\cdot\log^{O(1)}(\chilin(G))$.
+\begin{thm}[\cite{kun.obrien.ea:polynomial,czerwinski.nadara.ea:improved}]\label{kun_obrien_general}
+  For any graph $G$, $\chicen(G)\le (\chilin(G))^{19}\cdot(\log(\chilin(G)))^{O(1)}$.
 \end{thm}
@@
 \begin{enumerate}
-  \item A theorem of \citet{kawarabayashi.rossman:polynomial} shows that, if $\chicen(G)\ge k^{19}\log^{O(1)} k$ then $G$ contains a subcubic tree of treedepth $\Omega(k)$, or $\tw(G)\in\Omega(k^{18})$.
-In the former case, an asymptotically optimal result of \citet{kun.obrien.ea:polynomial} on subcubic trees completes the proof, so we are left with the case where $\tw(G)\in\Omega(k^{18})$.
-  \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy:improved} shows that, if $\tw(G)\ge k^{18}$, then $G$ contains an $\Omega(k^2\times k^2)$ grid minor.\todo{Update reference}
-  \item A technical lemma \cite[Lemma~5]{kun.obrien.ea:polynomial} shows that, if $G$ contains a $k^2\times k^2$ grid minor, then $\chilin(G)\in\Omega(k)$.\todo{Update reference}
+  \item A theorem of \citet{czerwinski.nadara.ea:improved} shows that, if $\chicen(G)\ge k^{19}\log^{q} k$ then $G$ contains a subcubic tree of treedepth $\Omega(k)$, or $\tw(G)\in\Omega(k^{18}\log^q k)$.\footnote{The result of \citet{czerwinski.nadara.ea:improved} is, of course, more general:  If $\chicen(G)\ge a\times b$, then $G$ contains an $\Omega(a)\times\Omega(a)$ grid minor or $G$ contains a subcubic tree of treedepth $\Omega(b)$.
+This is just an application of their result with $a=k^{18}\log^q k$ and $b=k$.}  In the former case, an asymptotically optimal result of \citet{kun.obrien.ea:polynomial} on subcubic trees completes the proof, so we are left with the case where $\tw(G)\in\Omega(k^{18}\log^q k)$.
+  \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy.tan:towards} shows that, if $\tw(G)\in\Omega(k^{18}\log^q k)$ (for a particular fixed positive $q$), then $G$ contains an $\Omega(k^2)\times \Omega(k^2)$ grid minor.
+  \item A technical lemma \cite[Lemma~5]{kun.obrien.ea:polynomial} shows that, if $G$ contains a $\Omega(k^2)\times \Omega(k^2)$ grid minor, then $\chilin(G)\in\Omega(k)$.
 \end{enumerate}
 
-In the current work, we improve the technical lemma in Point~3 to show that a $k\times k$ grid minor is sufficient to guarantee that $\chilin(G)\in\Omega(k)$.
-This improves the exponent in \cref{kun_obrien_general} from $19$ to $10$.
-
-This gives the following improvement to \cref{kun_obrien_general}
+In the current work, we spend considerable effort to improve the technical lemma in Point~3 to show that a $\Omega(k)\times\Omega(k)$ grid minor is sufficient to guarantee that $\chilin(G)\in\Omega(k)$.
+This improves the exponent in \cref{kun_obrien_general} from $19$ to $10$, yielding the following improvement to \cref{kun_obrien_general}:
 \begin{thm}\label{kun_obrien_general2}
@@
 
-In addition to the improvement to \cref{kun_obrien_general}, this results in considerably better bounds for some specific graph classes.
+\todo[inline]{Open problem here: Is there anything special about the relationship between treedepth and grid minors in planar graphs or product structured graphs?
+It's probably worth reading \citet{czerwinski.nadara.ea:improved} to see how their proof works.}
+
+% In addition to the improvement to \cref{kun_obrien_general}, this results in considerably better bounds for some specific graph classes.
 For example, the Excluded Grid Theorem for planar graphs asserts that any planar graph of treewidth $k$ contains a $\Omega(k)\times \Omega(k)$ grid minor.
@@
 
-Finally, \citet{kun.obrien.ea:polynomial} boldly conjecture that, for any graph $G$, $\chicen(G)\le 2\chilin(G)$.
-However, before the current work, the only class of graphs for which centred chromatic number was known to be bounded by a linear function of linear chromatic number was bounded degree trees.
-The current result adds further evidence for this conjecture by establishing that, when $G$ is a $k\times k$ grid (or even a pseudogrid), then $\chicen(G)\in\Theta(\chilin(G))$.
+\citet{kun.obrien.ea:polynomial} boldly conjecture that, for any graph $G$, $\chicen(G)\le 2\chilin(G)$.
+However, before the current work, the only class of graphs for which centred chromatic number was known to be bounded by a linear function of linear chromatic number was bounded degree trees \cite[Theorem~4]{kun.obrien.ea:polynomial}.
+The current result adds further evidence in support of this conjecture by establishing that, when $G$ is a $k\times k$ grid (or even a pseudogrid), then $\chicen(G)\in\Theta(\chilin(G))$.
 
@@
 For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$ denotes the set of vertices and edges of $G$.
-For a vertex $v\in V(G)$ we use the notations $N_G(v):=\{w\in V(G):vw\in E(G)\}$  to denote the open neighbourhood of $v$ in $G$.
-For a set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$.
+For a vertex $v\in V(G)$,  $N_G(v):=\{w\in V(G):vw\in E(G)\}$  denotes the open neighbourhood of $v$ in $G$ and for any set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$.
 
@@
 
-More specifically, we make use of the following (polygamous) version of the Marriage Theorem, which is obtained by replacing each vertex $x\in L(H)$ with $d$ vertices $x_1,\ldots,x_d$ each having the same neighbourhood as $x$.
-More precisely, we apply \cref{hall} on the graph $H':=H-\{x\}\cup\{x_iy:i\in\{1,\ldots,d\},\, y\in N_H(x)\}$.
-
+We make use of the following (polygamous) version of the Marriage Theorem:
 \begin{cor}\label{d_hall}
@@
 
+\cref{d_hall} can be deduced by applying adding $d-1$ twins $x_2,\ldots,x_d$ for each vertex $x\in L(H)$, i.e., by applying \cref{hall} on the graph $H':=H\cup\{x_iy:x\in L(H),\, i\in\{2,\ldots,d\},\, y\in N_H(x)\}$.
+
 Late in the game, we will make use of the following asymmetric version of the Lovász Local Lemma (see, for example, \citet[Lemma~5.1.1]{alon.spencer:probabilistic}):
@@
   Let $\mathcal{E}:=\{E_1,\ldots,E_n\}$ be a set of events in some probability space $(\Omega,\Pr)$.
-For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq \mathcal{E}$ be such that the event $E_i$ is mutually independent of $\mathcal{E}\setminus \Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:E\to[0,1)$ be such that
+For each $i\in\{1,\ldots,n\}$, let $\Gamma_i\subseteq \mathcal{E}$ be such that the event $E_i$ is mutually independent of $\mathcal{E}\setminus \Gamma_i$,\footnote{An event $A$ is mutually independent of a set $\{B_1,\ldots,B_r\}$ of events if, for any disjoint sets $I,J\subseteq\{1,\ldots,r\}$, $\Pr(A\cap\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j=\Pr(A)\Pr(\bigcap_{i\in I} B_i\cap\bigcap_{j\in J} \overline{B}_j)$.} and let $w:\mathcal{E}\to[0,1)$ be such that
   \[
-      \Pr(E_i) \le w(E_i)\cdot\prod_{E_j\in\Gamma_i}(1-w(E_j))  \enspace .
+      \Pr(E_i) \le w(E_i)\cdot\prod_{E_j\in\Gamma_i}(1-w(E_j))  \enspace ,
   \]
+  for each $i\in\{1,\ldots,n\}$.
   Then $\Pr(\overline{E}_1\cap\cdots\cap\overline{E}_n) > 0$.
@@
 
-For positive integers $a$ and $b$, the \defin{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.
-Such an edge is \defin{vertical} if $v=x$ and \defin{horizontal} if $w=y$.
+For positive integers $a$ and $b$, the \defin{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(i_1,j_1)$ and $(i_2,j_2)$ if and only if $|i_1-i_2|+|j_1-j_2|=1$.
+Such an edge is \defin{vertical} if $i_1=i_2$ and \defin{horizontal} if $j_1=j_2$.
 For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
@@
 (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
-  \item Replace each degree-$4$ vertex $(x,y)$ of $G_{a\times b}$ with a path $P_v$.
-If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each of $(x\pm 1,y\pm 1)$ and this has effect.
-Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $(x\pm 1,y\pm 1)$.
+  \item Replace each degree-$4$ vertex $(i,j)$ of $G_{a\times b}$ with a path $P_v$.
+If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each of $(i\pm 1,j\pm 1)$ and this has no effect on the underlying graph.
+Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $(i\pm 1,j\pm 1)$.
 It is useful to consider three possible cases:
   \begin{compactenum}[(Q1)]
-    \item \label{q_i} $p$ is adjacent to $\{(x-1,y), (x,y-1)\}$ and $q$ is adjacent to $(x+1,y),(x,y+1)$;
-    \item \label{q_ii} $p$ is adjacent to $\{(x,y+1), (x-1,y)\}$ and $q$ is adjacent to $(x,y-1),(x+1,y)$; or
-    \item \label{q_iii} $p$ is adjacent to $\{(x-1,y),(x+1,y)\}$ and $q$ is adjacent to $\{(x,y-1),(x,y+1)\}$.
+    \item \label{q_i} $p$ is adjacent to $\{(i-1,j), (i,j-1)\}$ and $q$ is adjacent to $(i+1,j),(i,j+1)$;
+    \item \label{q_ii} $p$ is adjacent to $\{(i,j+1), (i-1,j)\}$ and $q$ is adjacent to $(i,j-1),(i+1,j)$; or
+    \item \label{q_iii} $p$ is adjacent to $\{(i-1,j),(i+1,j)\}$ and $q$ is adjacent to $\{(i,j-1),(i,j+1)\}$.
   \end{compactenum}
@@
 For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.
-In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths is isomorphic to a $\le\!
-1$-subdivision of $G_{a\times b}$.
+In this way, $\mathcal{P}:=\{V(P_\mu):\mu\in \VE(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.
 We call $\mathcal{P}$ a \defin{grid-partition} of $G$.
 The \defin{boundary} of $G$ is defined as $\partial G:=\bigcup_{\mu\in\partial G_{a\times b}} V(P_\mu)$.
+The \defin{interior} of $G$ is $G-\partial G$.
 
@@
 Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
-This correspondence allows us to talk about the rows and columns of $G$, which we will do next.
+This correspondence allows us to talk about the rows and columns of $G$, which we will do immediately.
 
@@
 
-If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph.  Therefore to prove \cref{kun_obrien_general2}, it suffices to prove:
+If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph \cite{kun.obrien.ea:polynomial}.
+Therefore to prove \cref{kun_obrien_general2}, it suffices to prove:
 
@@
 We do this by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
-We will do this in several steps (see \cref{outline}):
+We will prove the existence of $P$ in several steps (see \cref{outline}):
 
@@
 
-  \item We do our best to find a set $S_1$ of vertices in $G$ that contains two vertices of each colour and that is `well-separated' in the sense that the corresponding set of vertices/edges in $G_{k\times k}$ have a minimum distance between them.
+  \item We greedily choose a set $S_1$ of vertices in $G$ that contains two vertices of each colour and that is `well-separated' in the sense that the corresponding set of vertices/edges in $G_{k\times k}$ have a minimum distance between them.
 In \cref{outline} this minimum distance corresponds to the fact that the box drawn centered at each vertex in $S_1$ contains no other vertices of $S_1$.
@@
 
-The most challenging aspect of this proof is the construction of $S_2$, which requires the use of the Local Lemma to ensure that no vertex chosen to take part in $S_2$ is close to any other vertex in $S_2$.
-The difficulty is illustrated in \cref{outline} by the cluster of three points of $S_1$ in the top right corner.
+The most challenging aspect of this proof is the construction of $S_2$, which requires the use of the Local Lemma (\cref{weighted_lovasz}) to ensure that no vertex chosen to take part in $S_2$ is close to any other vertex in $S_2$.
+The difficulty is illustrated in \cref{outline} by the cluster of three points of $S_1$ in the top right corner whose boxes overlap.
 These three points of $S_1$ are well-separated, but choosing one point from each of the three boxes to take part in $S_2$ could result in three points of $S_2$ being very close to each other.
-The local lemma is used to show that, in situations like this, we can still obtain two vertices of $S_2$ from each missing colour class while avoiding taking two vertices of $S_2$ from these three boxes.
+The local lemma is used to show that, in situations like this, we can still obtain two vertices of $S_2$ from each missing colour class while avoiding taking more than one vertex of $S_2$ from these three boxes.
 
@@
 For convenience, let $G_k:=G_{k\times k}$.
-Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
+Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=\{P_\mu:\mu\in\VE(G_{k})\}$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
 The partition $\mathcal{P}$ associates each vertex of $G$ with an edge or vertex $\mu$ of $G_k$, so $\varphi$ associates a colour set with each object in $\VE(G_k)$, as follows.
@@
 If there exists no $A\subseteq \varphi(V(G))$ with $|\varphi_{\mathcal{P}}^{-1}(A)\setminus\partial G_k| < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
-Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le (d+2)|A|$ to eliminate all vertices with colours in $A$, as follows:
+Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|+2 \le (d+2)|A|$ to eliminate all vertices with colours in $A$, as follows:
   \begin{compactitem}
@@
 
-Having each vertex or edge $\mu$ of $G_k$ associated with a set $\varphi_\mathcal{P}(\mu)$ of colours rather than a single colour will be problematic for what we want to to do next.
+Having each vertex or edge $\mu$ of $G_k$ associated with a set $\varphi_\mathcal{P}(\mu)$ of colours rather than a single colour is problematic for what we want to to do next.
 The following lemma allows us to choose one representative colour $\phi(\mu)$ from $\varphi_{\mathcal{P}}(\mu)$ for each $\mu\in \VE(G_k)$ while still ensuring that each colour appears frequently.
@@
   \]
-  By \cref{d_hall}, $H$ contains a subgraph $M$ with $\deg_M(x)=d$ for each $\alpha\in X$ and $\deg_M(y)\le 1$ for each $\mu\in Y$.
+  By \cref{d_hall}, $H$ contains a subgraph $M$ with $\deg_M(\alpha)=d$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$.
 For each edge $\alpha\mu$ in $M$, set $\phi(\mu):=\alpha$.
@@
 % Observe that, for any $\mu\in V(G_{k})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.
-Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.  Straightforward counting shows that, for any $\mu\in\VE(G_k)$,
+For any $\mu\in\VE(G_k)$, the $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.  Straightforward counting shows that, for any $\mu\in\VE(G_k)$,
 \begin{equation}
-   |\VE(G_r(\mu))| \le 3(2r+1)^2 \enspace . \label{rbox_size} % rbox size
+   |\VE(G_r(\mu))| \le |\VE(G_r)| = (2r+1)(6r+1) \enspace . \label{rbox_size} 
 \end{equation}
+For convenience, we define $\vol{r}:=(2r+1)(6r+1)$ and the reader should keep in mind that $\vol{r} \in\Theta(r^2)$.
+
+\begin{lstlisting}[frame=single,language=python]
+  # Sage Code
+  f(r)= (2*r+1)^2 + 4*r*(2*r+1)
+  show("f: ", f)
+  show("f(1): ", f(1))
+  show("f(2): ", f(2))
+  show("f(3): ", f(3))
+  show("f.full_simplify(): ", f.full_simplify())
+  show("f.factor(): ", f.factor())
+  f = f.factor()
+\end{lstlisting}
+
 We extend these definitions to vertices of a $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in V(G_k)$, define
@@
 and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$
-where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never any ambiguity as to which partition is being used.}
+where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never be any ambiguity as to which partition is being used.}
 % For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
@@
   Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_k)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.
-  todo[inline]{Finish this.}
+  \todo[inline]{Finish this.}
 \end{proof}
@@
 \begin{lem}\label{pick_up_everything}
-  Let $r\ge ??$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(G)\setminus \partial G$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$\todo{Set constraint on $R$} for each $v\in S$.
+  Let $r\ge ??$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(G)\setminus \partial G$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$\todo{Set constraint on $r$} for each $v\in S$.
 Then $G$ contains a path that contains every vertex in $S$.
@@
 \begin{lem}\label{packing_lemma}
-  Let $r\ge 114$ be an even integer and let $Q\subseteq \VE(G_k)$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
+  Let $r\ge 2$ be an even integer and let $Q\subseteq \VE(G_k)$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
   $|\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}| \le 25$.
@@
 \begin{proof}
-  The packing condition that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $B_{r/2-1}(\mu_1)\cap B_{r/2-1}(\mu_2)=\emptyset$ for each $\mu_1,\mu_2\in Q$.
+  For simplicity, assume that $G_k$ is embedded in $G_{k+2r+1}$ with the elements of $Q$ at distance at least $G_{2r}$ from the first and last row and column.
+In particular, for each $\mu_1\in Q$, $|B_{\rho}(\nu)| =\vol{\rho}$ for any $\rho\in\{0,\ldots,2r\}$.
 
-  For simplicity, assume that $G_k$ is embedded in $G_{k+2r+1}$ with the elements of $Q$ at distance at least $G_{2r}$ from the first and last row and column.
-In particular, for each $\mu_1\in Q$, $|B_{\rho}(\nu)| = 3(2\rho+1)^2$ for any $\rho\in\{0,\ldots,2r\}$.
-
-  Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu_1))\}$.  For any $\mu_1\in Q'$,
+  The packing condition that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $\VE(G_{r/2-1}(\mu_1))\cap \VE(G_{r/2-1}(\mu_2))=\emptyset$ for each $\mu_1,\mu_2\in Q$.
+Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu_1))\}$ so that our task is to show that $|Q'|\le 25$.
+For any $\mu_1\in Q'$,
   $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.
   Since this is true for all $\mu_1\in Q$ we have
-  \[  3(5r+1)^2 \ge |\VE(G_{5r/2}(\mu))| \ge \sum_{\mu_1\in Q'} |\VE(G_{r/2}(\mu_1))|
-  = 3|Q'|(r-2)^2 \enspace .
+  \[  \vol{5r/2} \ge |\VE(G_{5r/2}(\mu))| \ge \sum_{\mu_1\in Q'} |\VE(G_{r/2}(\mu_1))|
+  = |Q'|\cdot \vol{r/2} \enspace .
   \]
-  which implies that $|Q'|\le (5r+1)^2/(r-2)^2$.  For $r > 2$, this bound is decreasing in $r$ and is less than $26$ for any $r> 11\sqrt{26} + 57$.  In particular, $|Q'|< 26$ for any $r\ge 114$.
+  \begin{lstlisting}[frame=single,language=python]
+    # Sage Code
+    g(r) = f(5*r/2)/f(r/2)
+    show("g: ", g)
+    show("g.full_simplify(): ", g.full_simplify())
+    show("limit(g, r=oo): ", limit(g, r=oo))
+    show(plot(g, r, 1, 200))
+    show(solve(g(r)==25, r))
+  \end{lstlisting}
+  which implies that $|Q'|\le \vol{5r/2}/\vol{r/2}$.  For $r > 2$, this bound is increasing in $r$ and is less than $25$ for any even $r\ge 2$.
 \end{proof}
@@
 \begin{lem}\label{doubled_colour_set}
-  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
-  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.
+  For every integer $r\ge 1$\todo{Really no lower bound on $r$?} there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
+  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\setminus\partial G|\ge d|A|$.
   Then there exists $S\subseteq V(G)$ such that
   \begin{compactenum}[(i)]
-    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|= 2$; and
-    \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
+    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\varphi^{-1}(\alpha)\cap S|= 2$; and
+    \item \label{spread_out} for each $v\in S$, $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
   \end{compactenum}
@@
   We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi$ of $\VE(G_k)$.
-Observe that it is now sufficient to find a $2|\varphi(V(G))|$-object subset $Q\subseteq\VE(G_k)$ such that
+Observe that it is now sufficient to find a $2|\varphi(V(G))|$-element subset $Q\subseteq\VE(G_k)$ such that
   \begin{compactenum}[(i)]
-    \item $|\{\mu\in Q:\phi(v)=\alpha\}|= 2$ and
-    \item $|\VE(G_r(\mu))\cap Q|\le 2$ for each $\alpha\in\varphi(V(G))$ and each $\mu\in Q$.
+    \item $|\phi^{-1}(\alpha)\cap Q|= 2$ for each $\alpha\in\varphi(V(G))$ and
+    \item $|\VE(G_r(\mu))\cap Q|\le 2$ for each $\mu\in Q$.
   \end{compactenum}
@@
     \item $\phi(\mu_1)=\phi(\mu_2)=\alpha$; and\label{hits_both_q}
-    \item for each $b\in\{1,2\}$, $\mu_b\not\in \VE(G_r(\mu_{3-b})) \cup \bigcup_{\mu\in Q_1} \VE(G_r(\mu))$.\label{spread_out_q}
+    \item for each $b\in\{1,2\}$, $\mu_b\not\in \VE(G_r(\mu_{3-b})) \cup \bigcup_{\mu\in Q_1} \VE(G_r(\mu))$,\label{spread_out_q}
   \end{compactenum}
-  then we include $\mu_1$ and $\mu_2$ in $Q_1$ and declare the first round \defin{$\alpha$-successful}.  Otherwise, we declare the first round an \defin{$\alpha$-failure}.
+  then we include $\mu_1$ and $\mu_2$ in $Q_1$ and declare the first round an \defin{$\alpha$-success}.  Otherwise, we declare the first round an \defin{$\alpha$-failure}.
 
-  Let $K\subseteq\varphi(V(G))$ be the set of colours $\alpha$ for which the first round was $\alpha$-successful.
+  Let $\overline{X}\subseteq\varphi(V(G))$ be the set of colours $\alpha$ for which the first round was an $\alpha$-success.
 At the end of this process, the set $Q_1$ certainly satisfies (\ref{spread_out}) since (\ref{spread_out_q}) implies that $\VE(G_r(\mu))\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$.
-However, $Q_1$ only satisfies (\ref{hits_both}) for the colours in $K$.
+However, $Q_1$ only satisfies (\ref{hits_both}) for the colours in $\overline{X}$.
 We now use the second round to create a set $Q_2$ to fix this.
@@
 
-  Define a bipartite graph $H$ with parts $X:=\varphi(V(G))\setminus K$ and $Y:=Q_1$.
-Let $\pi:=\mu_1,\ldots,\mu_{2|K|}$ be a random permutation of the elements of $Y$.
+  Define a bipartite graph $H$ with parts $X:=\varphi(V(G))\setminus \overline{X}$ and $Y:=Q_1$.
+Let $\pi:=\mu_1,\ldots,\mu_{2|K|}$ be a random permutation of $Y$.
 We include an edge $\alpha\mu_i$ in $H$ if and only if
@@
 Note the asymmetry here: In the first round, each element $\mu_j$ that takes part in $Q_1$ prevents elements in $\VE(G_r(\mu))$ from also being part of $Q_1$.
-In the second round, $\mu_j$ defines a larger box $\VE(G_{2r}(\mu_j))$ that must be avoided when choosing an element from $\VE(G_r(\mu_i))$, for $i>j$.
+In the second round, $\mu_j$ defines a larger box $\VE(G_{2r}(\mu_j))$ that must be avoided when choosing a second element from $\VE(G_r(\mu_i))$, for $i>j$.
 
-  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 6(2r+1)^2$ for each $\alpha\in X$.
+  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{r}$ for each $\alpha\in X$.\todo{Constants are going change everywhere now...}
   For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.
-Furthermore all but at most one element of $\Phi_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since, otherwise the first round would have succeeded for colour $\alpha$.
+Furthermore all but at most one element of $\Phi_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since, otherwise the first round would have been an $\alpha$-success.
   % For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$. Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/3(20r+1)^2},1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 6(2r+1)^2)$.
-(We make no attempt to optimize the constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$.
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{10r}},1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{r})$.
+(We make no attempt to optimize constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$.
 Then, by \cref{packing_lemma}, the set
@@
 This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
-   % This already establishes that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.
+This already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.
 
   In order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.
-To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $B_r(\mu)\cap J=\emptyset$ for each $\mu\in Y$.
-Since $|\VE(G_{10r}(\mu))|\le 3(20r+1)^2$, each step in this process eliminates at most $3(20r+1)^2$ objects from $J$ so the number $t$ of iterations is at least $\ceil{|\Phi_\alpha|/3(20r+1)^2}$
+To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $\VE(G_r(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
+Since $|\VE(G_{10r}(\mu))|\le \vol{10r}$, each step in this process eliminates at most $\vol{10r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{10r}}$
 
@@
   \begin{align}
-    \Pr\left(\deg_H(\alpha)< 6(2r+1)^2\right)
-    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 6(2r+1)^2\right) \notag \\
-    & \le \Pr(\binomial(t,1/25)\le 6(2r+1)^2) \notag \\
-    & \le (2t)^{6(2r+1)^2}(1/25)^{r^2}(1-1/25)^{t-6(2r+1)^2} \notag \\
+    \Pr\left(\deg_H(\alpha)< 2\vol{r}\right)
+    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 2\vol{r}\right) \notag \\
+    & \le \Pr(\binomial(t,1/25)\le 2\vol{r}) \notag \\
+    & \le (2t)^{2\vol{r}}(1/25)^{r^2}(1-1/25)^{t-2\vol{r}} \notag \\
     % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
-    & \le \exp(6(2r+1)^2(1+\log(2t))-t/25) \notag \\
-    & \le \exp(6(2r+1)^2(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/75(20r+1)^2) \label{probability}
+    & \le \exp({2\vol{r}}(1+\log(2t))-t/25) \notag \\
+    & \le \exp({2\vol{r}}(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/25\vol{10r}) \label{probability}
     % & \le \exp(-6(2r+1)^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
@@
   We are now ready for an application of \cref{weighted_lovasz}.
-For each $\alpha\in X$, let Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 6(2r+1)^2$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
+For each $\alpha\in X$, let Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{r}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
 For each $\alpha\in X$, define
@@
   \]
-  and observe that, by \ref{rbox_size}, $|\Gamma_\alpha|\le 3(20r+1)^2|\Phi_\alpha$.
+  and observe that $|\Gamma_\alpha|\le \vol{10r}\cdot|\Phi_\alpha|$.
   The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
@@
     & \ge \exp\left(-\log\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
-    & \ge \exp\left(-\log\tau - 3(20r+1)^2|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
+    & \ge \exp\left(-\log\tau - \vol{10r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
   \end{align}
@@
   Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
-  Taking $\tau = 450(20r+1)^4+1$, and using \cref{weight_product,probability}, we find that
+  Taking $\tau = {50\vol{10r}}+1$, and using \cref{weight_product,probability}, we find that
   \[
@@
   provided that
-  \[   |\Phi_\alpha| \ge 150(20r+1)^2\left(\log\left(450(20r+1)^4+1\right) + 6(2r+1)^2(1+\log(2|\Phi_\alpha|)) \right) \enspace .
+  \[   |\Phi_\alpha| \ge {50\vol{10r}}\left(\log({50\vol{10r}}+1) + {2\vol{r}}(1+\log(2|\Phi_\alpha|))\right) \enspace .
   \]
-  The right hand side of this last equation is of the form $O(r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4)$.
+  The right hand side of this last equation is of the form $O(r^4+r^2\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4)$.
 
@@
 
-  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge 6(2r+1)^2$ for each $\alpha\in X$.
-On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 \le 3(2r+1)^2$ for each $y\in Y$.
-Therefore, for any $A\subseteq X$, $|N_H(A)|\ge 6(r+1)^2|A|/3(2r+1)^2 = 2|A|$.
+  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
+On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 \le {\vol{r}}$ for each $y\in Y$.
+Therefore, for any $A\subseteq X$, $|N_H(A)|\ge {2\vol{r}}|A|/{\vol{r}} = 2|A|$.
 Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$.

### 2022-04-06 lin-vs-cen.tex
Pffft

@@
 \begin{lem}\label{doubled_colour_set}
-  For every integer $r\ge 1$\todo{Really no lower bound on $r$?} there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
+  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
   Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\setminus\partial G|\ge d|A|$.
@@
 
-  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{r}$ for each $\alpha\in X$.\todo{Constants are going change everywhere now...}
+  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{r}$ for each $\alpha\in X$.
   For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.

### 2022-04-07 lin-vs-cen.tex
More proving.....

@@
 \newcommand{\vol}[1]{\boxplus_{#1}}
+
+\DeclareMathOperator{\interior}{Int}
 
@@
 For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
-The \defin{boundary} $\partial G_{a\times b}$ of $G_{a\times b}$ a the set of vertices in the first and last row and column and all edges incident to these vertices.
+For any integer $0\le r\le\min\{a,b\}/2$, the \defin{r-interior} of $G_{a\times b}$ defined as $\interior_r(G_{a\times b}):=G_{a\times b}[\{1+r,\ldots,a-r\}\times\{1+r,\ldots,b-r\}]$.
+For $r\ge\min\{a,b\}/2$, $\interior_r(G_{a\times b})$ is the empty graph.
 
@@
 We call $\mathcal{P}$ a \defin{grid-partition} of $G$.
-The \defin{boundary} of $G$ is defined as $\partial G:=\bigcup_{\mu\in\partial G_{a\times b}} V(P_\mu)$.
-The \defin{interior} of $G$ is $G-\partial G$.
+The \defin{$r$-interior} of $G$ is $\interior_r(G):=\bigcup_{\mu\in\VE(\interior_r(G_{a\times b}))} V(P_\mu)$.
+
+ % The \defin{boundary} of $G$ is defined as $\partial G:=\bigcup_{\mu\in\partial G_{a\times b}} V(P_\mu)$.  The \defin{interior} of $G$ is $G-\partial G$.
 
@@
 \begin{lem}\label{only_frequent}
-  Let $d\ge 1$ and $k\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/(d+2)$ colours.
-  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - (d+2)|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
-  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus \partial G_k| \ge d|A|$.
+  Let $d,k,r\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/(d+2r)$ colours.
+  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - (d+2r)|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
+  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\cap \interior_r(G')| \ge d|A|$.
 \end{lem}
@@
   The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.
-If there exists no $A\subseteq \varphi(V(G))$ with $|\varphi_{\mathcal{P}}^{-1}(A)\setminus\partial G_k| < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
-Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|+2 \le (d+2)|A|$ to eliminate all vertices with colours in $A$, as follows:
+If there exists no $A\subseteq \varphi(V(G))$ with $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)| < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
+Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|+2r \le (d+2r)|A|$ to eliminate all vertices with colours in $A$, as follows:
   \begin{compactitem}
-    \item For each $v:=(i,j)\in V(G_{k})\setminus \partial G_k$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
-    \item For each horizontal edge $vw\in E(G_{k\times k})\setminus \partial G_k$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
-    \item For each vertical edge $vw\in E(G_{k\times k})\setminus\partial G_k$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
-    \item We add the first and last row to $R$ and the first and last column to $C$.
+    \item For each $v:=(i,j)\in V(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
+    \item For each horizontal edge $vw\in E(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
+    \item For each vertical edge $vw\in E(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
+    \item We add the first and last $r$ rows to $R$ and the first and last $r$ columns to $C$.
     \item Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.
   \end{compactitem}
-  At this point $|R|=|C|\le (d+2)|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
+  At this point $|R|=|C|\le (d+2r)|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
 In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.
@@
   \[
-    k'\ge k_0-(d+2)|\varphi(V(G_0))|
-      \ge k-(d+2)|A|-(d+2)|\varphi(V(G_0))|
-      \ge k - (d+2)|\varphi(V(G))|
+    k'\ge k_0-(d+2r)|\varphi(V(G_0))|
+      \ge k-(d+2r)|A|-(d+2r)|\varphi(V(G_0))|
+      \ge k - (d+2r)|\varphi(V(G))|
   \]
@@
 \begin{lem}\label{one_colour_per_object}
-  Let $d>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\setminus \partial G_k| \ge d\cdot|A|$.
+  Let $d,r>1$ be integers, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\cap\interior_r(G_k)| \ge d\cdot|A|$.
 Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
   \begin{compactenum}[(i)]
-    \item For each $\mu\in\partial G_{k}$, $\phi(\mu)=\perp$.
+    \item For each $\mu\not\in\interior_r(G_{k})$, $\phi(\mu)=\perp$.
     \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi_\mathcal{P}(\mu)$.
@@
 \begin{proof}
-  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(G_k)\setminus\partial G_k$ and edge set
+  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(\interior_r(G_k))$ and edge set
   \[
@@
 \end{equation}
-For convenience, we define $\vol{r}:=(2r+1)(6r+1)$ and the reader should keep in mind that $\vol{r} \in\Theta(r^2)$.
+For convenience, we define $\vol{r}:=(2r+1)(6r+1)$ and the important thing to  keep in mind is that $\vol{r} \in\Theta(r^2)$.
 
@@
 
+\begin{lem}\label{make_disjoint}
+  Let $r\ge 10$ be an integer divisible by $10$, let $G$ be a $k\times k$ pseudogrid and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$ for each $v\in S$.
+  Then there exists a set $X\subseteq V(G)$ such that
+  \begin{compactenum}[(i)]
+    \item \label{covers_s} $S\subseteq\bigcup_{x\in X}\tilde{B}_{r/10-1}(x)$;
+    \item \label{two_per_box} $|\tilde{B}_{r/10-1}(x)\cap S|\le 2$ for each $x\in X$; and
+    \item \label{disjoint_boxes} $\tilde{B}_{r/10}(x)\cap \tilde{B}_{r/10}(y)=\emptyset$ for each distinct $x,y\in X$.
+  \end{compactenum}
+\end{lem}
+
+\begin{proof} 
+  For any integer $q$, say that two elements $v,w\in S$ are a \defin{$q$-pair} if $v\in B_{q}(w)$ or $w\in B_q(v)$.
+First observe that any $v\in S$ takes part in at most one $(r-1)$-pair since, otherwise, $|\tilde{B}_{r}(v)\cap S|\ge 3$.
+
+  Let $S_1$ be the subset of $S$ containing only those vertices that do not take part in any $(r-1)$-pair.  Let $X_1:= S_1$.  Of course $S_1\subseteq \bigcup_{x\in X_1}\tilde{B}_{r/10-1}(x)$.
+
+  Let $S_2$ be the subset of $S\setminus S_1$ containing those vertices that take part in a $(r/5-2)$-pair.
+We now define a set $X_2$ to handle vertices in $S_2$.
+For each $(r/5-2)$-pair $v,w\in S_2$, there exists a $x\in V(\interior_r(G))$ such that $v,w\in \tilde{B}_{r/10-1}(x)$, and we include $x$ in $X_2$.
+
+  All that remains are the vertices in $S_3:=S\setminus (S_1\cup S_2)$ which take part in some $(r-1)$-pair but do not take part in any $(r/5-2)$ pair.
+For each $(r-1)$-pair $v,w\in S_3$, there exists two vertices $x,y\in V(G)$ such that $v\in \tilde{B}_{r/10-1}(x)$, $w\in\tilde{B}_{r/10-1}(y)$ and $B_{r/10}(x)\cap B_{r/10}(y)=\emptyset$.\todo{Give details or a figure.}, and we include these in $X_3$.
+
+  Let $X:=X_1\cup X_2\cup X_3$.
+By construction, $X$ satisfies (\ref{covers_s}), so it only remains to show that $X$ satisfies (\ref{two_per_box}) and (\ref{disjoint_boxes}).
+By construction, for any $x\in X$, $B_{r/10}(x)$ contains at least one element of $S$.
+That $X$ satisfies (\ref{two_per_box}) follows easily from (\ref{covers_s}) and (\ref{disjoint_boxes}) since our construction associates each $x\in X$ with at most two elements of $S$ that are contained in $B_{r/3-1}(x)$.
+By (\ref{disjoint_boxes}) $B_{r/3-1}(x)$ does not intersect $B_{r/3-1}(y)$ for any $y\in X\setminus\{x\}$ so $B_{r/3-1}(x)$ does contain any elements of $S$ other than the at most two elements associated with it.
+Thus, we need only argue that $X$ satisfies (\ref{disjoint_boxes}):
+  \begin{compactenum}
+    \item For each $x\in X_1$, $B_{r-1}(x)\cap S=\{x\}$.  Therefore, for any $v\in S\setminus\{x\}$, any box $B_{r/10}(v')$ that contains $v$ does not intersect $B_{r/10}(x)$.
+
+    \item For each $x\in X_2$, $B_{r/10}(x)\cap S=\{v,w\}$ for some $(r/5-2)$-pair $v,w\in S_2$.
+Furthermore, with an appropriate choice of $x$, $B_{7r/10}(x)\subseteq B_{r-1}(v)\cup B_{r-1}(w)$.
+Since $(B_{r-1}(v)\cup B_{r-1}(w))\cap S=\{v,w\}$ this again implies that, for any $u\in S\setminus\{v,w\}$, any box $B_{r/10}(z)$ that contains $u$ does not intersect $B_{r/10}(x)$.
+
+    \item For each $(r-1)$-pair $v,w\in S_3$, there is a corresponding pair of vertices $x,y\in X_3$.
+By choice, $B_{r/10}(x)$ and $B_{r/10}(y)$ do not intersect each other.
+$B_{r/10}(x)$ contains $v$ and $B_{r-1}(v)\cap S=\{v,w\}$.
+Again, for any $u\in S\setminus\{v,w\}$, any box $B_{r/10}(z)$ that contains $u$ does not intersect $B_{r/10}(x)$.
+The same argument, applied to $w$ and $y$ shows that no such box intersects $B_{r/10}(y)$.
+\qedhere
+  \end{compactenum} 
+\end{proof}
+
 \begin{lem}\label{pick_up_everything}
-  Let $r\ge ??$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(G)\setminus \partial G$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$\todo{Set constraint on $r$} for each $v\in S$.
+  Let $r\ge 10$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$\todo{Set constraint on $r$} for each $v\in S$.
 Then $G$ contains a path that contains every vertex in $S$.
@@
 
-\begin{proof}[Proof Sketch]
-  We adjust things so that we get a set of disjoint boxes, each of which contains one or two points of $S$ in its interior.
-  We do the scan-line thing depicted in \cref{outline}.
+\begin{proof}
+  Apply \cref{make_disjoint} with $r:=10\floor{r/10}$ to obtain the set $X$.
+Now, consider a `snake-like' path $P_0$ that contains every row $i$ with $i\equiv 1\pmod 2r/3+1$ and, aside from those rows contains only vertices of columns $1$ and $k$.
+For each $x\in X$, this path intersects exactly one row of $B_{r/3}(x)$.
+For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{r/3}(x)$ with a path that is contained in $B_{r/3}(x)$ and contains the (at most two) vertices of $B_{r/3}(x)\cap S$.
+After doing this for each $x\in X$ we obtain a path $P$ that contains every vertex in $S$.
 \end{proof}
@@
 \begin{lem}\label{packing_lemma}
-  Let $r\ge 2$ be an even integer and let $Q\subseteq \VE(G_k)$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
+  Let $r\ge 2$ be an even integer and let $Q\subseteq \VE(\interior_r(G_k))$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
   $|\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}| \le 25$.
@@
 \begin{proof}
-  For simplicity, assume that $G_k$ is embedded in $G_{k+2r+1}$ with the elements of $Q$ at distance at least $G_{2r}$ from the first and last row and column.
-In particular, for each $\mu_1\in Q$, $|B_{\rho}(\nu)| =\vol{\rho}$ for any $\rho\in\{0,\ldots,2r\}$.
+  For simplicity, assume that $G_k$ is embedded in $G_{k+2r}$ with the elements of $Q$ all contained in $\VE(\interior_{2r}(G_{k+2r})$.
+In particular, for each $\mu_1\in Q$, this implies that $|B_{\rho}(\nu)| =\vol{\rho}$ for any $\rho\in\{0,\ldots,2r\}$.
 
@@
   For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
-  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\setminus\partial G|\ge d|A|$.
+  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G)|\ge d|A|$.
   Then there exists $S\subseteq V(G)$ such that

### 2022-04-08 lin-vs-cen.tex
Pfffft

@@
 For convenience, we define $\vol{r}:=(2r+1)(6r+1)$ and the important thing to  keep in mind is that $\vol{r} \in\Theta(r^2)$.
-
-\begin{lstlisting}[frame=single,language=python]
-  # Sage Code
-  f(r)= (2*r+1)^2 + 4*r*(2*r+1)
-  show("f: ", f)
-  show("f(1): ", f(1))
-  show("f(2): ", f(2))
-  show("f(3): ", f(3))
-  show("f.full_simplify(): ", f.full_simplify())
-  show("f.factor(): ", f.factor())
-  f = f.factor()
-\end{lstlisting}
+% 
+% \begin{lstlisting}[frame=single,language=python]
+%   # Sage Code
+%   f(r)= (2*r+1)^2 + 4*r*(2*r+1)
+%   show("f: ", f)
+%   show("f(1): ", f(1))
+%   show("f(2): ", f(2))
+%   show("f(3): ", f(3))
+%   show("f.full_simplify(): ", f.full_simplify())
+%   show("f.factor(): ", f.factor())
+%   f = f.factor()
+% \end{lstlisting}
 
@@
 \begin{lem}\label{make_disjoint}
-  Let $r\ge 10$ be an integer divisible by $10$, let $G$ be a $k\times k$ pseudogrid and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$ for each $v\in S$.
+  Let $r,p$ be positive integers with $r-1\ge 4p+4$, let $G$ be a $k\times k$ pseudogrid and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$ for each $v\in S$.
   Then there exists a set $X\subseteq V(G)$ such that
   \begin{compactenum}[(i)]
-    \item \label{covers_s} $S\subseteq\bigcup_{x\in X}\tilde{B}_{r/10-1}(x)$;
-    \item \label{two_per_box} $|\tilde{B}_{r/10-1}(x)\cap S|\le 2$ for each $x\in X$; and
-    \item \label{disjoint_boxes} $\tilde{B}_{r/10}(x)\cap \tilde{B}_{r/10}(y)=\emptyset$ for each distinct $x,y\in X$.
+    \item \label{covers_s} $S\subseteq\bigcup_{x\in X}\tilde{B}_{p}(x)$;
+    \item \label{two_per_box} $|\tilde{B}_{p}(x)\cap S|\le 2$ for each $x\in X$; and
+    \item \label{disjoint_boxes} $\tilde{B}_{p+1}(x)\cap \tilde{B}_{p+1}(y)=\emptyset$ for each distinct $x,y\in X$.
   \end{compactenum}
@@
 
-  Let $S_1$ be the subset of $S$ containing only those vertices that do not take part in any $(r-1)$-pair.  Let $X_1:= S_1$.  Of course $S_1\subseteq \bigcup_{x\in X_1}\tilde{B}_{r/10-1}(x)$.
+  Refer to \cref{make_disjoint_fig}.
+We will define the set $X$ so that it satisfies (\ref{covers_s}) and for each $x\in X$ we will choose $1$ or $2$ elements of $B_{p}(x)\cap S$ and say that $x$ \defin{covers} those elements.
+  Let $S_1$ be the subset of $S$ containing only those elements that do not take part in any $(r-1)$-pair.  Let $X_1:= S_1$ and we say that each $x\in X_1$ covers itself.
 
-  Let $S_2$ be the subset of $S\setminus S_1$ containing those vertices that take part in a $(r/5-2)$-pair.
-We now define a set $X_2$ to handle vertices in $S_2$.
-For each $(r/5-2)$-pair $v,w\in S_2$, there exists a $x\in V(\interior_r(G))$ such that $v,w\in \tilde{B}_{r/10-1}(x)$, and we include $x$ in $X_2$.
+  \begin{figure}
+    \begin{center}
+      \begin{tabular}{ccc}
+        \includegraphics[width=.3\textwidth]{figs/make_disjoint-1} &
+        \includegraphics[width=.3\textwidth]{figs/make_disjoint-2} &
+        \includegraphics[width=.3\textwidth]{figs/make_disjoint-3} \\
+        $X_1$ & $X_2$ & $X_3$
+      \end{tabular}
+    \end{center}
+    \caption{The proof of \cref{make_disjoint}.
+Disks are points in $S$ and crosses are points the resulting elements of $X$.
+The large yellow boxes are $B_{r-1}(v)$ for some $v\in S$ and the small mauve and green boxes are $B_{p+1}(x)$ and $B_{p}(x)$, respectively, for some $x\in X$.}
+    \label{make_disjoint_fig}
+  \end{figure}
+  Let $S_2$ be the subset of $S\setminus S_1$ containing those vertices that take part in a $2p$-pair.
+For each $2p$-pair $v,w\in S_2$, there exists a $x\in V(\interior_r(G))$ such that $v,w\in \tilde{B}_{p}(x)$.
+We include $x$ in $X_2$ and say that $x$ covers $v$ and $w$.
 
-  All that remains are the vertices in $S_3:=S\setminus (S_1\cup S_2)$ which take part in some $(r-1)$-pair but do not take part in any $(r/5-2)$ pair.
-For each $(r-1)$-pair $v,w\in S_3$, there exists two vertices $x,y\in V(G)$ such that $v\in \tilde{B}_{r/10-1}(x)$, $w\in\tilde{B}_{r/10-1}(y)$ and $B_{r/10}(x)\cap B_{r/10}(y)=\emptyset$.\todo{Give details or a figure.}, and we include these in $X_3$.
+  All that remains are the vertices in $S_3:=S\setminus (S_1\cup S_2)$ which take part in some $(r-1)$-pair but do not take part in any $2p$ pair.
+For each $(r-1)$-pair $v,w\in S_3$, there exists two vertices $x,y\in V(G)$ such that $v\in \tilde{B}_{p}(x)$, $w\in\tilde{B}_{p}(y)$ and $B_{p+1}(x)\cap B_{p+1}(y)=\emptyset$.
+We include $x$ and $y$ in $X_3$ and say that $x$ covers $v$ and $y$ covers $w$.
 
@@
 By construction, $X$ satisfies (\ref{covers_s}), so it only remains to show that $X$ satisfies (\ref{two_per_box}) and (\ref{disjoint_boxes}).
-By construction, for any $x\in X$, $B_{r/10}(x)$ contains at least one element of $S$.
-That $X$ satisfies (\ref{two_per_box}) follows easily from (\ref{covers_s}) and (\ref{disjoint_boxes}) since our construction associates each $x\in X$ with at most two elements of $S$ that are contained in $B_{r/3-1}(x)$.
-By (\ref{disjoint_boxes}) $B_{r/3-1}(x)$ does not intersect $B_{r/3-1}(y)$ for any $y\in X\setminus\{x\}$ so $B_{r/3-1}(x)$ does contain any elements of $S$ other than the at most two elements associated with it.
-Thus, we need only argue that $X$ satisfies (\ref{disjoint_boxes}):
-  \begin{compactenum}
-    \item For each $x\in X_1$, $B_{r-1}(x)\cap S=\{x\}$.  Therefore, for any $v\in S\setminus\{x\}$, any box $B_{r/10}(v')$ that contains $v$ does not intersect $B_{r/10}(x)$.
+That $X$ satisfies (\ref{two_per_box}) follows easily from (\ref{covers_s}) and the fact that each $x\in X$ covers at most two elements of $S$.
 
-    \item For each $x\in X_2$, $B_{r/10}(x)\cap S=\{v,w\}$ for some $(r/5-2)$-pair $v,w\in S_2$.
-Furthermore, with an appropriate choice of $x$, $B_{7r/10}(x)\subseteq B_{r-1}(v)\cup B_{r-1}(w)$.
-Since $(B_{r-1}(v)\cup B_{r-1}(w))\cap S=\{v,w\}$ this again implies that, for any $u\in S\setminus\{v,w\}$, any box $B_{r/10}(z)$ that contains $u$ does not intersect $B_{r/10}(x)$.
-
-    \item For each $(r-1)$-pair $v,w\in S_3$, there is a corresponding pair of vertices $x,y\in X_3$.
-By choice, $B_{r/10}(x)$ and $B_{r/10}(y)$ do not intersect each other.
-$B_{r/10}(x)$ contains $v$ and $B_{r-1}(v)\cap S=\{v,w\}$.
-Again, for any $u\in S\setminus\{v,w\}$, any box $B_{r/10}(z)$ that contains $u$ does not intersect $B_{r/10}(x)$.
-The same argument, applied to $w$ and $y$ shows that no such box intersects $B_{r/10}(y)$.
-\qedhere
-  \end{compactenum} 
+  To see that $X$ satisifies (\ref{disjoint_boxes}), observe that if $x\in X$ covers $v\in S$ then $B_{p+1}(x)\subseteq B_{2p+2}(v)$.
+Therefore, if $x\in X$ covers $v\in S$ and $y\in X$ covers $w\in S$ and $B_{p+1}(x)\cap B_{p+1}(y)\neq\emptyset$ then $w\in B_{4p+4}(v)$ or $v\in B_{4p+4}(w)$.
+Since $r-1 \ge 4p+4$, this implies that $v=w$ or $v$ and $w$ are an $(r-1)$-pair.
+In the later case, $x=y$ and there is nothing to prove.
+Otherwise $v,w\in S_2$ or $v,w\in S_3$.
+If $v,w\in S_2$ then, again $x=y$ and there is nothing to prove.
+If $v,w\in S_3$ then $x$ and $y$ are specifically chosen so that $B_{p+1}(x)$ and $B_{p+1}(y)$ are disjoint.
 \end{proof}
@@
 \begin{lem}\label{pick_up_everything}
-  Let $r\ge 10$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$\todo{Set constraint on $r$} for each $v\in S$.
+  Let $r\ge 9$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$ for each $v\in S$.
 Then $G$ contains a path that contains every vertex in $S$.
@@
 \begin{proof}
-  Apply \cref{make_disjoint} with $r:=10\floor{r/10}$ to obtain the set $X$.
-Now, consider a `snake-like' path $P_0$ that contains every row $i$ with $i\equiv 1\pmod 2r/3+1$ and, aside from those rows contains only vertices of columns $1$ and $k$.
-For each $x\in X$, this path intersects exactly one row of $B_{r/3}(x)$.
-For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{r/3}(x)$ with a path that is contained in $B_{r/3}(x)$ and contains the (at most two) vertices of $B_{r/3}(x)\cap S$.
+  Apply \cref{make_disjoint} with $p=\floor{(r-5)/4}$ to obtain the set $X$.
+Now, consider a `snake-like' path $P_0$ that contains every row $i$ with $i\equiv 1\pmod 2p+1$ and, aside from those rows contains only vertices of columns $1$ and $k$.
+For each $x\in X$, this path intersects exactly one row of $B_{p+1}(x)$.
+For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{p+1}(x)$ with a path that is contained in $B_{p+1}(x)$ and contains the (at most two) vertices of $B_{p+1}(x)\cap S$.
 After doing this for each $x\in X$ we obtain a path $P$ that contains every vertex in $S$.
@@
   \]
-  \begin{lstlisting}[frame=single,language=python]
-    # Sage Code
-    g(r) = f(5*r/2)/f(r/2)
-    show("g: ", g)
-    show("g.full_simplify(): ", g.full_simplify())
-    show("limit(g, r=oo): ", limit(g, r=oo))
-    show(plot(g, r, 1, 200))
-    show(solve(g(r)==25, r))
-  \end{lstlisting}
+%   /* \begin{lstlisting}[frame=single,language=python]
+%     # Sage Code
+%     g(r) = f(5*r/2)/f(r/2)
+%     show("g: ", g)
+%     show("g.full_simplify(): ", g.full_simplify())
+%     show("limit(g, r=oo): ", limit(g, r=oo))
+%     show(plot(g, r, 1, 200))
+%     show(solve(g(r)==25, r))
+%   \end{lstlisting}
   which implies that $|Q'|\le \vol{5r/2}/\vol{r/2}$.  For $r > 2$, this bound is increasing in $r$ and is less than $25$ for any even $r\ge 2$.

### 2022-04-08 lin-vs-cen.tex
Pffft

@@
   We prove that the linear colouring number of any $k\times k$ pseudogrid is $\Omega(k)$.
-By an argument of Kun et al, this result gives a tighter upper bound on the treedepth of a graph as a function of its linear colouring number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear colouring number.
+By an argument of Kun et al (Algorithmica, 2021), this result gives a tighter upper bound on the treedepth of a graph as a function of its linear colouring number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear colouring number.
 \end{abstract}
@@
 
-The question of upper bounding $\chicen(G)$ by some function of $\chilin(G)$ was considered by \citet{kun.obrien.ea:polynomial,czerwinski.nadara.ea:improved} who prove:
+Motivated by finding efficiently-computable approximations of treedepth in bounded expansion classes, linear colourings were introduced by \citet{kun.obrien.ea:polynomial} who were able to establish a polynomial upper bound on $\chicen(G)$ as a function of $\chilin(G)$.
+Combined with recent improvements by \citet{czerwinski.nadara.ea:improved} on the relationship between treedepth and treewidth, they obtain the following result:
 
@@
 
-Their proof of \cref{kun_obrien_general} has three steps:
+Their proof of \cref{kun_obrien_general} has three main ingredients:
 \begin{enumerate}
@@
 \begin{thm}\label{kun_obrien_general2}
-  For any graph $G$, $\chicen(G)\in O(\chilin^{10}(G))$.
+  For any graph $G$, $\chicen(G)\in (\chilin(G))^{10}\log(\chilin(G))^{O(1)}$.
 \end{thm}
 
-\todo[inline]{Open problem here: Is there anything special about the relationship between treedepth and grid minors in planar graphs or product structured graphs?
+% \todo[inline]{Open problem here: Is there anything special about the relationship between treedepth and grid minors in planar graphs or product structured graphs?
 It's probably worth reading \citet{czerwinski.nadara.ea:improved} to see how their proof works.}
@@
 For a vertex $v\in V(G)$,  $N_G(v):=\{w\in V(G):vw\in E(G)\}$  denotes the open neighbourhood of $v$ in $G$ and for any set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$.
+We use $\deg_G(v):=|N_G(v)|$ to denote the degree of the vertex $v$ in the graph $G$.
 
 For a  bipartite graph $H$, the two parts of $V(H)$ are denoted by $L(H)$ and $R(H)$ and we use the convention of writing an edge $xy$ so that its first endpoint $x$ is in $L(H)$ and its second endpoint $y$ is in $R(H)$.
-A \defin{matching} $M$ in a bipartite graph $H$ is a subgraph of $H$ in which each vertex has degree at most $M$.
+A \defin{matching} $M$ in a bipartite graph $H$ is a subgraph of $H$ in which each vertex has degree at most $1$.
 We say that $M$ \defin{saturates} a set $S\subseteq V(H)$ if $\deg_M(v)=1$ for each $v\in S$.
@@
 
-\cref{d_hall} can be deduced by applying adding $d-1$ twins $x_2,\ldots,x_d$ for each vertex $x\in L(H)$, i.e., by applying \cref{hall} on the graph $H':=H\cup\{x_iy:x\in L(H),\, i\in\{2,\ldots,d\},\, y\in N_H(x)\}$.
+\cref{d_hall} can be deduced by adding $d-1$ twins $x_2,\ldots,x_d$ for each vertex $x\in L(H)$, i.e., by applying \cref{hall} on the graph $H':=H\cup\{x_iy:x\in L(H),\, i\in\{2,\ldots,d\},\, y\in N_H(x)\}$.
 
@@
 Such an edge is \defin{vertical} if $i_1=i_2$ and \defin{horizontal} if $j_1=j_2$.
-For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.
+For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the \defin{$j$th row} is the vertex set $\{(1,j),\ldots,(a,j)\}$.
 For any integer $0\le r\le\min\{a,b\}/2$, the \defin{r-interior} of $G_{a\times b}$ defined as $\interior_r(G_{a\times b}):=G_{a\times b}[\{1+r,\ldots,a-r\}\times\{1+r,\ldots,b-r\}]$.
@@
 
-In this work, we need to work with pseudogrids.  Refer to \cref{pseudogrid_fig}.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
+Refer to \cref{pseudogrid_fig}.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
 \begin{compactitem}
@@
 (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
-  \item Replace each degree-$4$ vertex $(i,j)$ of $G_{a\times b}$ with a path $P_v$.
+  \item Replace each degree-$4$ vertex $(i,j)$ of $G_{a\times b}$ with a non-empty path $P_v$.
 If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each of $(i\pm 1,j\pm 1)$ and this has no effect on the underlying graph.
@@
 In particular, if $P_{v_i}$ was created using (Q\ref{q_iii}) then $R$ does not contain any internal vertices in $P_{v_i}$.
-Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
+Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path $C$  in $G$ that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
 This correspondence allows us to talk about the rows and columns of $G$, which we will do immediately.
@@
 As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.
-If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_a$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove degree-$1$ vertices until none remain.
+If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_r$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove degree-$1$ vertices until none remain.
 If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.
@@
 These three points of $S_1$ are well-separated, but choosing one point from each of the three boxes to take part in $S_2$ could result in three points of $S_2$ being very close to each other.
-The local lemma is used to show that, in situations like this, we can still obtain two vertices of $S_2$ from each missing colour class while avoiding taking more than one vertex of $S_2$ from these three boxes.
+The local lemma is used to show that, in situations like this, we can still obtain two vertices from each missing colour class to put into $S_2$ without having to take more than one vertex from these three boxes.
 
@@
   \begin{compactitem}
-    \item For each $v:=(i,j)\in V(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
+    \item For each vertex $v:=(i,j)\in V(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
     \item For each horizontal edge $vw\in E(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
@@
   \end{compactitem}
-  At this point $|R|=|C|\le (d+2r)|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-d|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
+  At this point $|R|=|C|\le (d+2r)|A|$ and we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-(d+2r)|A|$ and such that $\varphi(V(G_0))\cap A=\emptyset$.
 In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.
@@
 \begin{lem}\label{one_colour_per_object}
-  Let $d,r>1$ be integers, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\cap\interior_r(G_k)| \ge d\cdot|A|$.
+  Let $d,r>1$ be integers, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)| \ge d\cdot|A|$.
 Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
@@
 For this, we need some definitions of boxes in $G$ and in $G_k$.
+We base our definition of `well-separated' on the concept of boxes, which we now define.
 
@@
 
-We extend these definitions to vertices of a $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in V(G_k)$, define
+We extend these definitions to vertices of a $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in\VE(G_k)$, define
 \[
@@
 \begin{lem}\label{pick_up_two}
-  Let $G$ be a $a\times a$ pseudogrid with $a\ge??$,  let $s,v,w,t$ be vertices of $G$ with $s$ in the first column of $G$, neither $v$ nor $w$ in the first or last column of $G$ and $t$ in the last column of $G$.
+  Let $G$ be a $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in\interior_1(G)$, and $t$ in column $a$ of $G$.
 Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
@@
 
-\begin{proof}
-  Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_k)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.
-  \todo[inline]{Finish this.}
-\end{proof}
+The proof of \cref{pick_up_two} is a tedious case analysis that appears in \cref{pick_up_two_proof}.
 
@@
 By construction, $X$ satisfies (\ref{covers_s}), so it only remains to show that $X$ satisfies (\ref{two_per_box}) and (\ref{disjoint_boxes}).
-That $X$ satisfies (\ref{two_per_box}) follows easily from (\ref{covers_s}) and the fact that each $x\in X$ covers at most two elements of $S$.
+That $X$ satisfies (\ref{two_per_box}) follows easily from (\ref{disjoint_boxes}), the fact that each element of $S$ is covered by some $x\in X$, and the fact that each $x\in X$ covers at most two elements of $S$.
 
@@
 Since $r-1 \ge 4p+4$, this implies that $v=w$ or $v$ and $w$ are an $(r-1)$-pair.
-In the later case, $x=y$ and there is nothing to prove.
-Otherwise $v,w\in S_2$ or $v,w\in S_3$.
+In the former case, $x=y$ and there is nothing to prove.
+In the latter case, $v,w\in S_2$ or $v,w\in S_3$.
 If $v,w\in S_2$ then, again $x=y$ and there is nothing to prove.
@@
   Apply \cref{make_disjoint} with $p=\floor{(r-5)/4}$ to obtain the set $X$.
-Now, consider a `snake-like' path $P_0$ that contains every row $i$ with $i\equiv 1\pmod 2p+1$ and, aside from those rows contains only vertices of columns $1$ and $k$.
+Now, consider a `snake-like' path $P_0$ that contains every row $i$ with $i\equiv 1\pmod{2p+1}$ and, aside from those rows contains only vertices of columns $1$ and $k$.
 For each $x\in X$, this path intersects exactly one row of $B_{p+1}(x)$.
@@
 \end{proof}
+
+\todo[inline]{A bunch of $B$'s should be $\tilde{B}$'s}
 
@@
 In particular, for each $\mu_1\in Q$, this implies that $|B_{\rho}(\nu)| =\vol{\rho}$ for any $\rho\in\{0,\ldots,2r\}$.
+(This assumption is not critical, but eliminates the need to discuss boundary conditions that have no effect on the final result.)
 
@@
   For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
-  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G)|\ge d|A|$.
-  Then there exists $S\subseteq V(G)$ such that
+  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)|\ge d|A|$.
+  Then there exists $S\subseteq V(\interior_r(G))$ such that
   \begin{compactenum}[(i)]
@@
   We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi$ of $\VE(G_k)$.
-Observe that it is now sufficient to find a $2|\varphi(V(G))|$-element subset $Q\subseteq\VE(G_k)$ such that
+Observe that it is now sufficient to find a $2|\varphi(V(G))|$-element subset $Q\subseteq\VE(\interior_r(G_k))$ such that
   \begin{compactenum}[(i)]
@@
 
-  Define a bipartite graph $H$ with parts $X:=\varphi(V(G))\setminus \overline{X}$ and $Y:=Q_1$.
+  Define a bipartite graph $H$ with parts $L(H):=X:=\varphi(V(G))\setminus \overline{X}$ and $R(H):=Y:=Q_1$.
 Let $\pi:=\mu_1,\ldots,\mu_{2|K|}$ be a random permutation of $Y$.
@@
 Note the asymmetry here: In the first round, each element $\mu_j$ that takes part in $Q_1$ prevents elements in $\VE(G_r(\mu))$ from also being part of $Q_1$.
-In the second round, $\mu_j$ defines a larger box $\VE(G_{2r}(\mu_j))$ that must be avoided when choosing a second element from $\VE(G_r(\mu_i))$, for $i>j$.
+In the second round, $\mu_j$ defines a larger box $\VE(G_{2r}(\mu_j))$ that must be avoided when choosing a second element from $\VE(G_r(\mu_i))$, for each $i>j$.
 
@@
   For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.
-Furthermore all but at most one element of $\Phi_\alpha$ is contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since, otherwise the first round would have been an $\alpha$-success.
+Furthermore all but at most $\vol{r}$ elements of $\Phi_\alpha$ are contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.\todo{Explain more or redefine $\Phi_\alpha$ as $\Phi_\alpha\cap (\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu)))$)}
   % For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$. Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.
@@
   Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
-On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 \le {\vol{r}}$ for each $y\in Y$.
+On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 < {\vol{r}}$ for each $y\in Y$.
 Therefore, for any $A\subseteq X$, $|N_H(A)|\ge {2\vol{r}}|A|/{\vol{r}} = 2|A|$.
@@
   Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.
-Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon-2$.
+Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)\cap\interior_r(G_{k'})|\ge d|A|$, for any $d\le 1/2\epsilon-2$.
 In particular, for some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
@@
 
+\appendix
+\section{Proof of \cref{pick_up_two}}
+\label{pick_up_two_proof}
+
+\begin{proof}[Proof of \cref{pick_up_two}]
+  Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_a)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.
+Define $\mu_s$ and $\mu_t$ similarly, with respect to $s$ and $t$.
+Very roughly, this lemma says that the grid $G_a$ contains a path $P_a$ with $\mu_v,\mu_w\in\VE(P_a)$ whose first edge/vertex is $\mu_s$, and whose last edge/vertex is $\mu_t$.
+
+  Although this is a simplification of the problem, we first describe how to solve it.
+For this simpler problem, we may assume that each of $\mu_s$, $\mu_v$, $\mu_w$, and $\mu_t$ is an edge of $G_a$ since we can replace any vertex of $G_a$ with one of its incident edges.
+Refer to \cref{generic_path}.
+Beginning at $\mu_s$, $P_a$ can traverse the boundary of $G_a$ until reaching the first column $i$ that contains an endpoint of $\mu_v$ or $\mu_w$, then vertically in this column to collect $\mu_v$ (say).
+What happens next depends on whether or not $\mu_w$ also intersects column $i$.
+  \begin{figure}[htbp]
+    \begin{center}
+      \begin{tabular}{cc}
+        \includegraphics{figs/generic_path-1} &
+        \includegraphics{figs/generic_path-2} 
+      \end{tabular}
+    \end{center}
+    \caption{A simplification of \cref{pick_up_two}}
+    \label{generic_path}
+  \end{figure}
+  \begin{compactitem}
+    \item If $\mu_w$ also intersects column $i$, then $P_a$ can immediately return to column $i$ if necessary (if $\mu_v$ was a horizontal edge) and collect $\mu_w$.
+The only care that needs to be taken in this case occurs when $\mu_v$ and $\mu_w$ are a pair of horizontal and vertical edges that share an endpoint.
+When this happens, the initial traversal on the boundary should be done (clockwise or counterclockwise) so that the vertical edge appears first in $P_a$.
+
+    \item If $\mu_w$ does not intersect column $i$, then after collecting $\mu_v$, $P_a$ can proceed horizontally to the first column $j$ that intersects $\mu_w$ and then vertically in column $j$ to collect $\mu_w$.
+  \end{compactitem}
+  In either case, after collecting $\mu_w$, $P_a$ can then proceed vertically to return to the boundary and then traverse the boundary to finish at $\mu_t$.
+
+  The rough description given above works perfectly when $\mu_v$ and $\mu_w$ are each edges of $G_a$.
+However, some mild complications arise when one or both of $\mu_v$ or $\mu_w$ are vertices of $G_a$.
+This is due to the fact that a path $P_a$ in $G_a$ that contains a vertex $\nu$ may not correspond to a path $P$ in $G$ that contains $P_\nu$.
+Indeed, this depends on whether $P_a$ `turns' at $\nu$ and whether $P_\nu$ was created using (Q\ref{q_i}), (Q\ref{q_ii}), or (Q\ref{q_iii}).
+We say that a $\mu_v$ or $\mu_w$ is \defin{straight} if it was created using (Q\ref{q_i}) or (Q\ref{q_ii}) and $\mu_v$ or $\mu_w$ is \defin{bent} if it was created using (Q\ref{q_iii}).
+Without loss of generality, assume that $\mu_v$ is a vertex in column $i$ of $G_a$ and that $\mu_w$ does not intersect columns $2,\ldots,i-1$.
+Our strategy is to replace $\mu_v$ (and possibly also $\mu_w$) with a pair of edges incident on $\mu_v$ in such a way that the algorithm described above is able to construct a path $P_a$ that contains the resulting collection of $3$ or $4$ edges.
+What follows is a (boring) case analysis (see \cref{cases}):
+
+  \begin{figure}
+    \begin{center}
+      \includegraphics{figs/dumdum}
+    \end{center}
+    \caption{Cases in the proof of \cref{pick_up_two}}
+    \label{cases}
+  \end{figure}
+
+  \begin{compactenum}
+    \item If $\mu_w$ is a vertical edge incident on $\mu_v$ then there are two possibilities:
+    \begin{compactitem}
+      \item If $\mu_v$ is straight then we replace $\mu_v$ with the other vertical edge incident on $\mu_v$.
+      \item If $\mu_v$ is bent then we replace $\mu_v$ with the horizontal edge joining $\mu_v$ to a vertex in column $i+1$.
+    \end{compactitem}
+    \item If  $\mu_w$ is a horizontal edge incident to $\mu_v$ then there are two possibilities:
+    \begin{compactitem}
+      \item If $\mu_v$ is straight then we replace it with the other horizontal edge incident to $\mu_v$.
+      \item If $\mu_v$ is bent then we replace it with one of the vertical edges incident to it.
+    \end{compactitem}
+    \item If $\mu_w$ intersects column $i$ but is not an edge incident to $\mu_v$ then we may assume, without loss of generality, that $\mu_w$ is above $\mu_v$.  There are two cases to consider:
+    \begin{compactitem} 
+      \item If $\mu_v$ is straight, then we replace $\mu_v$ with the two vertical edges incident on $\mu_v$.
+      \item If $\mu_v$ is bent, then we replace $\mu_v$ with the horizontal edge joining $\mu_v$ to a vertex in column $i-1$ and the vertical edge incident to $\mu_v$ and above $\mu_v$.
+    \end{compactitem}
+    If $\mu_w$ is an edge of $G_a$ then there is nothing else to. If $\mu_w$ is a vertex of $G_a$ then there are two cases to consider:
+    \begin{compactitem} 
+      \item If $\mu_w$ is straight, then we replace $\mu_w$ with the two vertical edges incident on $\mu_w$.
+      \item If $\mu_w$ is bent, then we replace $\mu_w$ with the horizontal edge joining $\mu_v$ to a vertex in column $i+1$ and the vertical edge incident to $\mu_w$ and below $\mu_w$.
+    \end{compactitem}
+    \item If $\mu_w$ does not intersect column $i$ there are two possibilities:
+    \begin{compactitem} 
+      \item If $\mu_v$ is straight, then we replace $\mu_v$ with the two vertical edges incident on $\mu_v$.
+      \item If $\mu_v$ is bent, then we replace $\mu_v$ with the horizontal edge joining $\mu_v$ to a vertex in column $i-1$ and a vertical edge in column $i$.
+    \end{compactitem}
+    If $\mu_w$ is an edge of $G_a$ then there is nothing further to do. If $\mu_w$ is a vertex of $G_a$ in column $j>i$ then there are again two possibilities:
+    \begin{compactitem} 
+      \item If $\mu_w$ is straight, then we replace $\mu_w$ with the two vertical edges incident on $\mu_w$.
+      \item If $\mu_w$ is bent, then we replace $\mu_w$ with the horizontal edge joining $\mu_v$ to a vertex in column $j+1$ and a vertical edge in column $j$.
+    \end{compactitem}
+  \end{compactenum}
+  Now, exactly the same strategy used above can be used to construct a path $P_a$ that contains the (up to four) required edges of $G_a$ and the corresponding path $P$ in $G$ satisfies the requirements of the lemma.
+\end{proof}
+
 \end{document}

### 2022-04-13 lin-vs-cen.tex
Pfffft

@@
 where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never be any ambiguity as to which partition is being used.}
-% For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
+For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.\todo{Use $B$, $G$, $\tilde{B}$, and $\tilde{G}$ consistently!}
 

### 2022-04-13 lin-vs-cen.tex
Small changes

@@
 
-\section*{Acknowledgement}
-
-The main subject of this paper was posed as an open problem at the 9th Annual Bellairs Workshop on Geometry and Graphs (v2), which was held January 21--28, 2022 and again at the 16th Annual Workshop on Probability and Combinatorics, which was held March 25--April 1, 2022.
+% \section*{Acknowledgement}
+%
+% The main subject of this paper was posed as an open problem at the 9th Annual Bellairs Workshop on Geometry and Graphs (v2), which was held January 21--28, 2022 and again at the 16th Annual Workshop on Probability and Combinatorics, which was held March 25--April 1, 2022.
 We are grateful to the organizers and participants of both workshops for providing a stimulating research environment.
-Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
+% Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
 

### 2022-04-26 lin-vs-cen.tex
Readded the ack
Fixed a lie

--- 2022-04-13 lin-vs-cen.tex
@@
 Motivated by finding efficiently-computable approximations of treedepth in bounded expansion classes, linear colourings were introduced by \citet{kun.obrien.ea:polynomial} who were able to establish a polynomial upper bound on $\chicen(G)$ as a function of $\chilin(G)$.
-Combined with recent improvements by \citet{czerwinski.nadara.ea:improved} on the relationship between treedepth and treewidth, they obtain the following result:
+Their arguments, combined with recent improvements by \citet{czerwinski.nadara.ea:improved} on the relationship between treedepth and treewidth, they obtain the following result:
 
@@
 \begin{enumerate}
-  \item A theorem of \citet{czerwinski.nadara.ea:improved} shows that, if $\chicen(G)\ge k^{19}\log^{q} k$ then $G$ contains a subcubic tree of treedepth $\Omega(k)$, or $\tw(G)\in\Omega(k^{18}\log^q k)$.\footnote{The result of \citet{czerwinski.nadara.ea:improved} is, of course, more general:  If $\chicen(G)\ge a\times b$, then $G$ contains an $\Omega(a)\times\Omega(a)$ grid minor or $G$ contains a subcubic tree of treedepth $\Omega(b)$.
+  \item A theorem of \citet{czerwinski.nadara.ea:improved} shows that, if $\chicen(G)\ge k^{19}\log^{q} k$ then $G$ contains a subcubic tree of treedepth $\Omega(k)$, or $\tw(G)\in\Omega(k^{18}\log^q k)$.\footnote{The result of \citet{czerwinski.nadara.ea:improved} is, of course, more general:  If $\chicen(G)\ge a\times b$, then $\tw(G)\in\Omega(a)$ or $G$ contains a subcubic tree of treedepth $\Omega(b)$.
 This is just an application of their result with $a=k^{18}\log^q k$ and $b=k$.}  In the former case, an asymptotically optimal result of \citet{kun.obrien.ea:polynomial} on subcubic trees completes the proof, so we are left with the case where $\tw(G)\in\Omega(k^{18}\log^q k)$.

### 2022-05-05 lin-vs-cen.tex
Some random changes?

@@
 
-\title{\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}}
+\title{\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}}
 \author{Prosenjit~Bose, Vida~Dujmović, Hussein~Houdrouge, Mehrnoosh~Javarsineh, and Pat~Morin}
@@
 \begin{abstract}
-  We prove that the linear colouring number of any $k\times k$ pseudogrid is $\Omega(k)$.
-By an argument of Kun et al (Algorithmica, 2021), this result gives a tighter upper bound on the treedepth of a graph as a function of its linear colouring number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear colouring number.
+  We prove that the linear chromatic number of any $k\times k$ pseudogrid is $\Omega(k)$.
+By an argument of Kun et al (Algorithmica, 2021), this result gives a tighter upper bound on the treedepth of a graph as a function of its linear chromatic number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear chromatic number.
 \end{abstract}
@@
 Motivated by finding efficiently-computable approximations of treedepth in bounded expansion classes, linear colourings were introduced by \citet{kun.obrien.ea:polynomial} who were able to establish a polynomial upper bound on $\chicen(G)$ as a function of $\chilin(G)$.
-Their arguments, combined with recent improvements by \citet{czerwinski.nadara.ea:improved} on the relationship between treedepth and treewidth, they obtain the following result:
+By combining their arguments with recent improvements by \citet{czerwinski.nadara.ea:improved} on the relationship between treedepth and treewidth, the following theorem is established:
 
@@
 
-Their proof of \cref{kun_obrien_general} has three main ingredients:
+\noindent The proof of \cref{kun_obrien_general} has three main ingredients:
 \begin{enumerate}
@@
 For a graph $G$, $V(G)$ denotes the vertex set of $G$, $E(G)$ denotes the edge set of $G$ and $\VE(G):=V(G)\cup E(G)$ denotes the set of vertices and edges of $G$.
+We will usually refer to an arbitrary element/edge/vertex in $\VE(G)$ as an \defin{object}.
 For a vertex $v\in V(G)$,  $N_G(v):=\{w\in V(G):vw\in E(G)\}$  denotes the open neighbourhood of $v$ in $G$ and for any set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$.
@@
 
-We make use of the following (polygamous) version of the Marriage Theorem:
+We make use of the following (polygamous) consequence of Hall's Marriage Theorem:
 \begin{cor}\label{d_hall}
@@
 
-\section{The Linear Colouring Number of Pseudogrids}
+\section{The Linear Chromatic Number of Pseudogrids}
 
@@
 As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.
-If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_r$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove degree-$1$ vertices until none remain.
+If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_r$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove vertices of degree at most $1$ until none remain.
 If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.
@@
 If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph \cite{kun.obrien.ea:polynomial}.
-Therefore to prove \cref{kun_obrien_general2}, it suffices to prove:
+Therefore to prove \cref{kun_obrien_general2}, it suffices to establish:
 
@@
 
-We do this by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
+We will prove \cref{pseudogrid_lower_bound} by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
 We will prove the existence of $P$ in several steps (see \cref{outline}):
@@
   \item We first delete rows and columns from $G$ so that each colour that appears in $G$ appears at least $d$ times in the interior of $G$, for some large constant $d$.
-From this point on the goal is to construct $P$ so that it contains each colour at least twice.
+From this point on the goal is to construct $P$ so that it contains each remaining colour at least twice.
 
@@
 
-  \item For these failed colours, we use a variant of Hall's Theorem to match each failed colour $\alpha$ with two vertices in $S_1$ that each contain a vertex of colour $\alpha$ nearby.
-In this matching, no vertex of $S_1$ is matched with more than one failed colour.
-From this we identify a set $S_2$ that contains two vertices of each of the missing colours and such that no vertex of $S_2$ is close to any other vertex of $S_2$.
+  \item For these failed colours, we use \cref{d_hall} (the Polygamous Marriage Theorem) to find a $d$-to-$1$ matching $M$ that associates each failed colour $\alpha$ with $d$ vertices in $S_1$ that each contain a vertex of colour $\alpha$ nearby.
+In $M$, no vertex of $S_1$ is matched with more than one failed colour.
 
-  \item We construct a path $P$ that contains each vertex in $S_1\cup S_2$.  This is possible (easy, even) because each vertex in $S_1$ is `close to' at most one vertex of $S_2$ and vice-versa.
+  \item Using $M$ and \cref{weighted_lovasz} (the Lovász Local Lemma) we identify a set $S_2$ that contains two vertices of each of the missing colours and such that no vertex of $S_2$ is close to any other vertex of $S_2$.
+
+  \item We construct a path $P$ that contains each vertex in $S_1\cup S_2$.
+This is possible because each vertex in $S_1$ is `close to' at most one vertex of $S_2$ and vice-versa.
+Aside from these pairs, no pair of vertices is close to each other.
 \end{compactenum}
@@
 The difficulty is illustrated in \cref{outline} by the cluster of three points of $S_1$ in the top right corner whose boxes overlap.
-These three points of $S_1$ are well-separated, but choosing one point from each of the three boxes to take part in $S_2$ could result in three points of $S_2$ being very close to each other.
+These vertices of $S_1$ are well-separated, but choosing one point from each of the three boxes to take part in $S_2$ could result in three vertices of $S_2$ being very close to each other.
+In particular, these three points could be vertices of $P_{\mu_1}$, $P_{\mu_2}$ and $P_{\mu_2}$ where $\mu_1$, $\mu_2$, and $\mu_3$ are objects in $\VE(G_{k\times k})$ that all contain a common vertex, making it impossible to find a path that contains all three (see \cref{bad_examples}).
 The local lemma is used to show that, in situations like this, we can still obtain two vertices from each missing colour class to put into $S_2$ without having to take more than one vertex from these three boxes.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{ccc}
+      \includegraphics{figs/bad-example-1} &
+      \includegraphics{figs/bad-example-2} &
+      \includegraphics{figs/bad-example-3} 
+    \end{tabular}
+  \end{center}
+  \caption{Some examples of $3$- and $4$-element vertex subsets that cannot all be contained in a single path (that must also contain other vertices not pictured).}
+\end{figure}
 
@@
 In \cref{path_finding} we explain how, given a well-separated set $S$ of vertices in $G$, to find a path that contains every vertex in $S$.
-In \cref{packing_lemma_sec} we introduce fairly standard-looking packing lemma that is needed in several places later.
+In \cref{packing_lemma_sec} we introduce a fairly standard-looking packing lemma that is needed in several places later.
 In \cref{well_separated} we explain how find a well-separated set $S$ that contains two vertices of each colour.
@@
 
+The following lemma gives conditions that allow us to delete rows and columns from $G$ to obtain a sub-pseudogrid in which every colour occurs frequently in the its interior.
+
 \begin{lem}\label{only_frequent}
@@
   Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - (d+2r)|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
-  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\cap \interior_r(G')| \ge d|A|$.
+  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\cap \interior_r(G_{k'})| \ge d|A|$.
 \end{lem}
@@
 \begin{lem}\label{one_colour_per_object}
-  Let $d,r>1$ be integers, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)| \ge d\cdot|A|$.
+  Let $d,r>1$ be integers, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)| \ge d|A|$.
 Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
@@
 Next we show that, given a sufficiently `well-separated' set $S$ of pairs of vertices in $G$, we can always find a path in $G$ that contains every vertex in $S$. 
-For this, we need some definitions of boxes in $G$ and in $G_k$.
+% For this, we need some definitions of boxes in $G$ and in $G_k$.  
 We base our definition of `well-separated' on the concept of boxes, which we now define.
@@
 \]
-The \defin{$r$-box} centred at an edge $vw$ of $G_{k}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.
+The \defin{$r$-box} centred at an edge $vw$ of $G_{k}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.\footnote{Technically the notations for $B_r(v)$ and $B_r(\mu)$ should include the value of $k$, but we omit this since there will never be any ambiguity.}
 % Observe that, for any $\mu\in V(G_{k})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.
 For any $\mu\in\VE(G_k)$, the $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{k}[B_r(\mu)]$.  Straightforward counting shows that, for any $\mu\in\VE(G_k)$,
+\[
+   |B_r(\mu)| \le |V(G_{2(r+1)\times (2r+1)})| = 2(r+1)(2r+1) = 4r^2+6r+2
+\]
+and
 \begin{equation}
-   |\VE(G_r(\mu))| \le |\VE(G_r)| = (2r+1)(6r+1) \enspace . \label{rbox_size}
+   |\VE(G_r(\mu))| \le |\VE(G_{2(r+1)\times (2r+1)})| = 12r^2+14r+3 \enspace . \label{rbox_size}
 \end{equation}
-For convenience, we define $\vol{r}:=(2r+1)(6r+1)$ and the important thing to  keep in mind is that $\vol{r} \in\Theta(r^2)$.
+For convenience, we define $\vol{r}:=12r^2+14r+3$ and the important thing to  keep in mind is that $\vol{r} \in\Theta(r^2)$.
 %
@@
 %   # Sage Code
-%   f(r)= (2*r+1)^2 + 4*r*(2*r+1)
-%   show("f: ", f)
-%   show("f(1): ", f(1))
-%   show("f(2): ", f(2))
-%   show("f(3): ", f(3))
-%   show("f.full_simplify(): ", f.full_simplify())
-%   show("f.factor(): ", f.factor())
-%   f = f.factor()
+%   f(r)=2*(r+1)*(2*r+1 + 2*r) + (2*r+1)^2
+%   show(f)
+%   show(f.factor())
 % \end{lstlisting}
@@
 \]
-and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$
-where $\mu_v$ denotes the unique element in $\VE(G_k)$ such that $v\in P_{\mu_v}$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never be any ambiguity as to which partition is being used.}
+and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$ where $P_{\mu_v}$ is the unique part in $\mathcal{P}$ that contains $v$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never be any ambiguity as to which partition is being used.}
 For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.\todo{Use $B$, $G$, $\tilde{B}$, and $\tilde{G}$ consistently!}
+
+The following lemma, whose proof is a (somewhat tedious) case analysis that appears in \cref{pick_up_two_proof}, is the main tool we use to build a path that contains a set of vertices that can be paired off in such a way that each pair is far from all other vertices.
 
@@
 
-The proof of \cref{pick_up_two} is a tedious case analysis that appears in \cref{pick_up_two_proof}.
+The next lemma shows how to take a well separated collection of pairs and cover them with disjoint boxes that can be used with \cref{pick_up_two}.
 
@@
   Refer to \cref{make_disjoint_fig}.
-We will define the set $X$ so that it satisfies (\ref{covers_s}) and for each $x\in X$ we will choose $1$ or $2$ elements of $B_{p}(x)\cap S$ and say that $x$ \defin{covers} those elements.
-  Let $S_1$ be the subset of $S$ containing only those elements that do not take part in any $(r-1)$-pair.  Let $X_1:= S_1$ and we say that each $x\in X_1$ covers itself.
+We will define the set $X$ so that it satisfies (\ref{covers_s}) and for each $x\in X$ we will choose $1$ or $2$ elements of $\tilde{B}_{p}(x)\cap S$ and say that $x$ \defin{covers} those elements.
+Let $S_1$ be the subset of $S$ containing only those elements that do not take part in any $(r-1)$-pair.
+Let $X_1:= S_1$ and we say that each $x\in X_1$ covers itself.
 
@@
     \caption{The proof of \cref{make_disjoint}.
-Disks are points in $S$ and crosses are points the resulting elements of $X$.
+Disks are vertices in $S$ and crosses are the resulting elements of $X$.
 The large yellow boxes are $B_{r-1}(v)$ for some $v\in S$ and the small mauve and green boxes are $B_{p+1}(x)$ and $B_{p}(x)$, respectively, for some $x\in X$.}
@@
   Let $S_2$ be the subset of $S\setminus S_1$ containing those vertices that take part in a $2p$-pair.
-For each $2p$-pair $v,w\in S_2$, there exists a $x\in V(\interior_r(G))$ such that $v,w\in \tilde{B}_{p}(x)$.
-We include $x$ in $X_2$ and say that $x$ covers $v$ and $w$.
+For each $2p$-pair $v,w\in S_2$, there exists  $x\in V(\interior_r(G))$ such that $v,w\in \tilde{B}_{p}(x)$.
+We include $x$ in $X_2$ and say that $x$ \defin{covers} $v$ and $w$.
 
   All that remains are the vertices in $S_3:=S\setminus (S_1\cup S_2)$ which take part in some $(r-1)$-pair but do not take part in any $2p$ pair.
-For each $(r-1)$-pair $v,w\in S_3$, there exists two vertices $x,y\in V(G)$ such that $v\in \tilde{B}_{p}(x)$, $w\in\tilde{B}_{p}(y)$ and $B_{p+1}(x)\cap B_{p+1}(y)=\emptyset$.
+For each $(r-1)$-pair $v,w\in S_3$, there exists two vertices $x,y\in V(G)$ such that $v\in \tilde{B}_{p}(x)$, $w\in\tilde{B}_{p}(y)$ and $\tilde{B}_{p+1}(x)\cap \tilde{B}_{p+1}(y)=\emptyset$.
 We include $x$ and $y$ in $X_3$ and say that $x$ covers $v$ and $y$ covers $w$.
@@
 By construction, $X$ satisfies (\ref{covers_s}), so it only remains to show that $X$ satisfies (\ref{two_per_box}) and (\ref{disjoint_boxes}).
-That $X$ satisfies (\ref{two_per_box}) follows easily from (\ref{disjoint_boxes}), the fact that each element of $S$ is covered by some $x\in X$, and the fact that each $x\in X$ covers at most two elements of $S$.
+That $X$ satisfies (\ref{two_per_box}) follows from (\ref{disjoint_boxes}), the fact that each element of $S$ is covered by some $x\in X$, and the fact that each $x\in X$ covers at most two elements of $S$.
 
-  To see that $X$ satisifies (\ref{disjoint_boxes}), observe that if $x\in X$ covers $v\in S$ then $B_{p+1}(x)\subseteq B_{2p+2}(v)$.
-Therefore, if $x\in X$ covers $v\in S$ and $y\in X$ covers $w\in S$ and $B_{p+1}(x)\cap B_{p+1}(y)\neq\emptyset$ then $w\in B_{4p+4}(v)$ or $v\in B_{4p+4}(w)$.
+  To see that $X$ satisifies (\ref{disjoint_boxes}), observe that if $x\in X$ covers $v\in S$ then $\tilde{B}_{p+1}(x)\subseteq \tilde{B}_{2p+2}(v)$.
+Therefore, if $x\in X$ covers $v\in S$ and $y\in X$ covers $w\in S$ and $\tilde{B}_{p+1}(x)\cap \tilde{B}_{p+1}(y)\neq\emptyset$ then $w\in \tilde{B}_{4p+4}(v)$ or $v\in \tilde{B}_{4p+4}(w)$.
 Since $r-1 \ge 4p+4$, this implies that $v=w$ or $v$ and $w$ are an $(r-1)$-pair.
@@
 If $v,w\in S_2$ then, again $x=y$ and there is nothing to prove.
-If $v,w\in S_3$ then $x$ and $y$ are specifically chosen so that $B_{p+1}(x)$ and $B_{p+1}(y)$ are disjoint.
+If $v,w\in S_3$ then $x$ and $y$ are specifically chosen so that $\tilde{B}_{p+1}(x)$ and $\tilde{B}_{p+1}(y)$ are disjoint.
 \end{proof}
@@
 \begin{lem}\label{pick_up_everything}
-  Let $r\ge 9$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$ for each $v\in S$.
+  Let $r\ge 9$, let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=\{P_\mu:\mu\in\VE(G_k)\}$ and let $S\subseteq V(\interior_r(G))$ be such that $|\tilde{B}_{r}(v)\cap S|\le 2$ for each $v\in S$.
 Then $G$ contains a path that contains every vertex in $S$.
@@
 Now, consider a `snake-like' path $P_0$ that contains every row $i$ with $i\equiv 1\pmod{2p+1}$ and, aside from those rows contains only vertices of columns $1$ and $k$.
-For each $x\in X$, this path intersects exactly one row of $B_{p+1}(x)$.
-For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{p+1}(x)$ with a path that is contained in $B_{p+1}(x)$ and contains the (at most two) vertices of $B_{p+1}(x)\cap S$.
+For each $x\in X$, this path intersects exactly one row of $\tilde{B}_{p+1}(x)$.
+For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{p+1}(x)$ with a path that is contained in $\tilde{B}_{p+1}(x)$ and contains the (at most two) vertices of $\tilde{B}_{p+1}(x)\cap S$.
 After doing this for each $x\in X$ we obtain a path $P$ that contains every vertex in $S$.
 \end{proof}
-
-\todo[inline]{A bunch of $B$'s should be $\tilde{B}$'s}
 
@@
 \begin{proof}
-  For simplicity, assume that $G_k$ is embedded in $G_{k+2r}$ with the elements of $Q$ all contained in $\VE(\interior_{2r}(G_{k+2r})$.
-In particular, for each $\mu_1\in Q$, this implies that $|B_{\rho}(\nu)| =\vol{\rho}$ for any $\rho\in\{0,\ldots,2r\}$.
-(This assumption is not critical, but eliminates the need to discuss boundary conditions that have no effect on the final result.)
-
   The packing condition that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $\VE(G_{r/2-1}(\mu_1))\cap \VE(G_{r/2-1}(\mu_2))=\emptyset$ for each $\mu_1,\mu_2\in Q$.
@@
 %     # Sage Code
-%     g(r) = f(5*r/2)/f(r/2)
-%     show("g: ", g)
-%     show("g.full_simplify(): ", g.full_simplify())
-%     show("limit(g, r=oo): ", limit(g, r=oo))
-%     show(plot(g, r, 1, 200))
-%     show(solve(g(r)==25, r))
+%   g(r)=2*(r+1)*(2*r+1)  # |B_r(mu)|
+%   show(g)               
+%   show(g.factor())
+%   show(g.full_simplify())
+%   f(r)=2*(r+1)*(2*r+1 + 2*r) + (2*r+1)^2  # |VE(G[B_r(mu)])
+%   show(f)
+%   show(f.factor())
+%   vol = f
+%   ratio(r) = vol(5*r/2) / vol(r/2)  # ratio that appears in packing lemma
+%   show(plot(ratio, r, 1, 10))
+%   show(limit(ratio(r), r=oo))
 %   \end{lstlisting}
-  which implies that $|Q'|\le \vol{5r/2}/\vol{r/2}$.  For $r > 2$, this bound is increasing in $r$ and is less than $25$ for any even $r\ge 2$.
+  which implies that $|Q'|\le \vol{5r/2}/\vol{r/2}$.  For $r > 2$, this bound is increasing in $r$ and is less than $25$ for any $r\ge 2$.
 \end{proof}
@@
 \begin{lem}\label{doubled_colour_set}
-  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
+  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.\newline
   Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)|\ge d|A|$.
@@
 \begin{proof}
-  We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi$ of $\VE(G_k)$.
+  We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi:\VE(G_k)\to\varphi(V(G))$.
 Observe that it is now sufficient to find a $2|\varphi(V(G))|$-element subset $Q\subseteq\VE(\interior_r(G_k))$ such that

### 2022-05-09 lin-vs-cen.tex
Cool

@@
 and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$ where $P_{\mu_v}$ is the unique part in $\mathcal{P}$ that contains $v$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never be any ambiguity as to which partition is being used.}
-For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.\todo{Use $B$, $G$, $\tilde{B}$, and $\tilde{G}$ consistently!}
+For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
 
@@
     \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\varphi^{-1}(\alpha)\cap S|= 2$; and
-    \item \label{spread_out} for each $v\in S$, $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
+    \item \label{spread_out} for each $v\in S$, $|\tilde{B}_r(v)\cap S|\le 2$.
   \end{compactenum}
@@
     \item $\phi(\mu_1)=\phi(\mu_2)=\alpha$; and\label{hits_both_q}
-    \item for each $b\in\{1,2\}$, $\mu_b\not\in \VE(G_r(\mu_{3-b})) \cup \bigcup_{\mu\in Q_1} \VE(G_r(\mu))$,\label{spread_out_q}
+    \item for each $b\in\{1,2\}$, $\mu_b\not\in \VE(G_{2r+2}(\mu_{3-b})) \cup \bigcup_{\mu\in Q_1} \VE(G_{2r+2}(\mu))$,\label{spread_out_q}
   \end{compactenum}
@@
   \[
-    \VE(G_r(\mu_i))\setminus\left(\bigcup_{j=1}^{i-1}\VE(G_{2r}(\mu_j))\right)
+    \VE(G_{2r+2}(\mu_i))\setminus\left(\bigcup_{j=1}^{i-1}\VE(G_{2r+2}(\mu_j))\right)
   \]
   contains some object $\nu$ of colour $\phi(\nu)=\alpha$.
-Note the asymmetry here: In the first round, each element $\mu_j$ that takes part in $Q_1$ prevents elements in $\VE(G_r(\mu))$ from also being part of $Q_1$.
-In the second round, $\mu_j$ defines a larger box $\VE(G_{2r}(\mu_j))$ that must be avoided when choosing a second element from $\VE(G_r(\mu_i))$, for each $i>j$.
+Note the asymmetry here: In the first round, each element $\mu_j$ that takes part in $Q_1$ prevents any elements in $\VE(G_{2r+2}(\mu))$ from also being part of $Q_1$.
+In the second round, $\VE(G_{2r+2}(\mu_j))$ need only be avoided when choosing a second element from $\VE(G_r(\mu_i))$, for $i>j$.
 
-  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{r}$ for each $\alpha\in X$.
-  For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.
-Furthermore all but at most $\vol{r}$ elements of $\Phi_\alpha$ are contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.\todo{Explain more or redefine $\Phi_\alpha$ as $\Phi_\alpha\cap (\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu)))$)}
+  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{2r+2}$ for each $\alpha\in X$.
+For each $\alpha\in X$, let $\Phi^0_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi^0_\alpha| \ge d$.
+Furthermore all but at most $\vol{2r+2}$ elements of $\Phi^0_\alpha$ are contained in $\bigcup_{\mu\in Y}\VE(G_{2r+2}(\mu))$ since otherwise the first round would have been an $\alpha$-success.
+Therefore, the set $\Phi_\alpha:=\Phi^0_\alpha\cap (\bigcup_{\mu\in Y}\VE(G_{2r+2}(\mu)))$ has size $|\Phi_\alpha|\ge d-\vol{2r+2}$.
+
+  % \todo{Continue here...}\todo{Explain more or redefine $\Phi_\alpha$ as $\Phi_\alpha\cap (\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu)))$)}
   % For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$. Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{10r}},1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{r})$.
-(We make no attempt to optimize constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$.
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{5r}},1/4)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+2})$.
+(We make no attempt to optimize constants here.)  Suppose that, for some $\mu\in Y$, $B_{2r+2}(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$.
 Then, by \cref{packing_lemma}, the set
   \[
-     Y_\nu := \{\xi\in Y:\nu\in\VE(G_{2r}(\xi))\}
+     Y_\nu := \{\xi\in Y:\nu\in\VE(G_{2r+2}(\xi))\}
   \]
-  has size at most $25$.
-Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.
-This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
-This already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.
+  has size at most $4$.\todo{Update \cref{packing_lemma}} Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.
+This happens with probability exactly $1/|Y_\nu|\ge 1/4$.
+This already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.
+To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+2}(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus\VE(G_{5r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $4r+O(1)$ would be sufficient.}  This process continues until $\VE(G_{2r+2}(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
+Since $|\VE(G_{5r}(\mu))|\le \vol{5r}$, each step in this process eliminates at most $\vol{5r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{5r}}$
 
-  In order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.
-To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $\VE(G_r(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
-Since $|\VE(G_{10r}(\mu))|\le \vol{10r}$, each step in this process eliminates at most $\vol{10r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{10r}}$
-
-  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be the corresponding elements of $\Phi_\alpha$.
+  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_t\}\subseteq\Phi_\alpha$ be the corresponding elements of $\Phi_\alpha$.
 The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are disjoint.
-Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_k$ are mutually independent.
+Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_t$ are mutually independent.
 Indeed, each $U_i$ depends only on the relative order of $Y_{\nu_i}$ in the  permutation $\pi$.
-For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.
-Therefore, $\sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/25)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
+For each $i\in\{1,\ldots,t\}$, $\Pr(U_i)\ge 1/4$.
+Therefore, $\deg_H(\alpha)\ge \sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/4)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
   \begin{align}
-    \Pr\left(\deg_H(\alpha)< 2\vol{r}\right)
-    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 2\vol{r}\right) \notag \\
-    & \le \Pr(\binomial(t,1/25)\le 2\vol{r}) \notag \\
-    & \le (2t)^{2\vol{r}}(1/25)^{r^2}(1-1/25)^{t-2\vol{r}} \notag \\
+    \Pr\left(\deg_H(\alpha)< 2\vol{2r+2}\right)
+    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 2\vol{2r+2}\right) \notag \\
+    & \le \Pr(\binomial(t,1/4)\le 2\vol{2r+2}) \notag \\
+    & = \sum_{x=0}^{2\vol{2r+2}-1} \binom{t}{x}\cdot (1/4)^x \cdot (3/4)^{t-x} \\
+    & \le (2t)^{2\vol{2r+2}}(1/4)^{2\vol{2r+2}}(3/4)^{t-2\vol{2r+2}} \notag \\
     % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
-    & \le \exp({2\vol{r}}(1+\log(2t))-t/25) \notag \\
-    & \le \exp({2\vol{r}}(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/25\vol{10r}) \label{probability}
+    & \le \exp({2\vol{2r+2}}(1+\log(2t))-t/4) \notag \\
+    & \le \exp({2\vol{2r+2}}(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/4\vol{5r}) \label{probability}
     % & \le \exp(-6(2r+1)^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\

### 2022-05-09 lin-vs-cen.tex
I'm about to chuck all these changes

@@
   \caption{Some examples of $3$- and $4$-element vertex subsets that cannot all be contained in a single path (that must also contain other vertices not pictured).}
+  \label{bad_examples}
 \end{figure}
@@
 In \cref{path_finding} we explain how, given a well-separated set $S$ of vertices in $G$, to find a path that contains every vertex in $S$.
-In \cref{packing_lemma_sec} we introduce a fairly standard-looking packing lemma that is needed in several places later.
+% In \cref{packing_lemma_sec} we introduce a fairly standard-looking packing lemma that is needed in several places later.
 In \cref{well_separated} we explain how find a well-separated set $S$ that contains two vertices of each colour.
@@
 
-We extend these definitions to vertices of a $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in\VE(G_k)$, define
+Later, we will make use of the following observation about sets of boxes, none of which contains the center of any other.
+(Intuitively, each of the four quadrants surrounding $\nu$ can contain at most one element $\mu\in Q$ with $\nu\in \VE(G_r(\mu))$.)
+\begin{obs}\label{packing_lemma}
+  Let $r,k\ge 1$ be integers and let $Q\subseteq\VE(G_k)$ be such that $\VE(G_r(\mu))\cap Q=\{\mu\}$ for each $\mu\in Q$.  Then, for each $\nu\in\VE(G_k)$, $|\{\mu\in Q:\nu\in \VE(G_r(\mu))\}|\le 4$.
+\end{obs}
+
+We extend the definition of boxes to vertices of a $k\times k$ pseudogrid $G$ with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ as follows. For any $\mu\in\VE(G_k)$, define
 \[
@@
 
-\subsection{A Packing Lemma}
-\label{packing_lemma_sec}
-
-We will make use of the following fairly standard looking packing lemma.
-
-\begin{lem}\label{packing_lemma}
-  Let $r\ge 2$ be an even integer and let $Q\subseteq \VE(\interior_r(G_k))$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
-  $|\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}| \le 25$.
-\end{lem}
-
-\begin{proof}
-  The packing condition that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $\VE(G_{r/2-1}(\mu_1))\cap \VE(G_{r/2-1}(\mu_2))=\emptyset$ for each $\mu_1,\mu_2\in Q$.
+% \begin{proof}
+%   The packing condition that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $\VE(G_{r/2-1}(\mu_1))\cap \VE(G_{r/2-1}(\mu_2))=\emptyset$ for each $\mu_1,\mu_2\in Q$.
 Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu_1))\}$ so that our task is to show that $|Q'|\le 25$.
 For any $\mu_1\in Q'$,
-  $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.
-  Since this is true for all $\mu_1\in Q$ we have
-  \[  \vol{5r/2} \ge |\VE(G_{5r/2}(\mu))| \ge \sum_{\mu_1\in Q'} |\VE(G_{r/2}(\mu_1))|
-  = |Q'|\cdot \vol{r/2} \enspace .
-  \]
-%   /* \begin{lstlisting}[frame=single,language=python]
-%     # Sage Code
-%   g(r)=2*(r+1)*(2*r+1)  # |B_r(mu)|
-%   show(g)
-%   show(g.factor())
-%   show(g.full_simplify())
-%   f(r)=2*(r+1)*(2*r+1 + 2*r) + (2*r+1)^2  # |VE(G[B_r(mu)])
-%   show(f)
-%   show(f.factor())
-%   vol = f
-%   ratio(r) = vol(5*r/2) / vol(r/2)  # ratio that appears in packing lemma
-%   show(plot(ratio, r, 1, 10))
-%   show(limit(ratio(r), r=oo))
-%   \end{lstlisting}
-  which implies that $|Q'|\le \vol{5r/2}/\vol{r/2}$.  For $r > 2$, this bound is increasing in $r$ and is less than $25$ for any $r\ge 2$.
-\end{proof}
+%   $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.
+%   Since this is true for all $\mu_1\in Q$ we have
+%   \[  \vol{5r/2} \ge |\VE(G_{5r/2}(\mu))| \ge \sum_{\mu_1\in Q'} |\VE(G_{r/2}(\mu_1))|
+%   = |Q'|\cdot \vol{r/2} \enspace .
+%   \]
+% %   /* \begin{lstlisting}[frame=single,language=python]
+% %     # Sage Code
+% %   g(r)=2*(r+1)*(2*r+1)  # |B_r(mu)|
+% %   show(g)
+% %   show(g.factor())
+% %   show(g.full_simplify())
+% %   f(r)=2*(r+1)*(2*r+1 + 2*r) + (2*r+1)^2  # |VE(G[B_r(mu)])
+% %   show(f)
+% %   show(f.factor())
+% %   vol = f
+% %   ratio(r) = vol(5*r/2) / vol(r/2)  # ratio that appears in packing lemma
+% %   show(plot(ratio, r, 1, 10))
+% %   show(limit(ratio(r), r=oo))
+% %   \end{lstlisting}
+%   which implies that $|Q'|\le \vol{5r/2}/\vol{r/2}$.  For $r > 2$, this bound is increasing in $r$ and is less than $25$ for any $r\ge 2$.
+% \end{proof}
+%
 
@@
 \begin{lem}\label{doubled_colour_set}
-  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.\newline
+  For every integer $r\ge 1$ there exists an integer $d\in O(r^4\log r)$ such that the following is true, for every integer $k\ge 1$.\newline
   Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)|\ge d|A|$.
@@
   Define a bipartite graph $H$ with parts $L(H):=X:=\varphi(V(G))\setminus \overline{X}$ and $R(H):=Y:=Q_1$.
-Let $\pi:=\mu_1,\ldots,\mu_{2|K|}$ be a random permutation of $Y$.
+Let $\pi:=\mu_1,\ldots,\mu_{|Y|}$ be a random permutation of $Y$.
 We include an edge $\alpha\mu_i$ in $H$ if and only if
@@
 This already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.
-To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+2}(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus\VE(G_{5r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $4r+O(1)$ would be sufficient.}  This process continues until $\VE(G_{2r+2}(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
+To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+2}(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus\VE(G_{5r}(\mu))$.\footnote{The constant $5$ is overkill here and is only used for simplicity; $4r+O(1)$ would be sufficient.}  This process continues until $\VE(G_{2r+2}(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
 Since $|\VE(G_{5r}(\mu))|\le \vol{5r}$, each step in this process eliminates at most $\vol{5r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{5r}}$
@@
 For each $i\in\{1,\ldots,t\}$, $\Pr(U_i)\ge 1/4$.
-Therefore, $\deg_H(\alpha)\ge \sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/4)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
+Therefore, $\deg_H(\alpha)\ge \sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/4)$ random variable.\footnote{A random variable $X$ \defin{dominates} a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
   \begin{align}
@@
     & \le \Pr(\binomial(t,1/4)\le 2\vol{2r+2}) \notag \\
-    & = \sum_{x=0}^{2\vol{2r+2}-1} \binom{t}{x}\cdot (1/4)^x \cdot (3/4)^{t-x} \\
-    & \le (2t)^{2\vol{2r+2}}(1/4)^{2\vol{2r+2}}(3/4)^{t-2\vol{2r+2}} \notag \\
+    & = \sum_{x=0}^{2\vol{2r+2}} \binom{t}{x}\cdot (1/4)^x \cdot (3/4)^{t-x} \notag \\
+    & \le \sum_{x=0}^{2\vol{2r+2}} \binom{t}{x}\cdot (3/4)^{t} \notag  \\
+    & \le (2t)^{2\vol{2r+2}}\cdot (3/4)^{t} \notag \\
     % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
-    & \le \exp({2\vol{2r+2}}(1+\log(2t))-t/4) \notag \\
-    & \le \exp({2\vol{2r+2}}(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/4\vol{5r}) \label{probability}
+    & < \exp\left({2\vol{2r+2}}\ln(2t)- t/4\right)
+      & \text{(since $\ln(3/4) < -1/4$)}\notag \\
+    & \le \exp\left({2\vol{2r+2}}\ln(2|\Phi_\alpha|) -|\Phi_\alpha|/4\vol{5r}\right)
+      & \text{(since $|\Phi_\alpha|/\vol{5r}\le t\le |\Phi_\alpha|$)}\label{probability}
     % & \le \exp(-6(2r+1)^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
@@
   We are now ready for an application of \cref{weighted_lovasz}.
-For each $\alpha\in X$, let Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{r}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
+For each $\alpha\in X$, let Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{2r+2}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
 For each $\alpha\in X$, define
   \[
-     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\neq\emptyset\right\}
+     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{5r}(\nu))\neq\emptyset\right\}
   \]
-  and observe that $|\Gamma_\alpha|\le \vol{10r}\cdot|\Phi_\alpha|$.
-  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
+  and observe that $|\Gamma_\alpha|\le \vol{5r}\cdot|\Phi_\alpha|$.
+  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r+2}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r+2}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
 
-  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>0$ to be determined shortly.  Then
+  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>1$ to be determined shortly.  Then
   \begin{align}
@@
     & \ge \exp\left(-\log\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
-    & \ge \exp\left(-\log\tau - \vol{10r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
+    & \ge \exp\left(-\log\tau - \vol{5r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
   \end{align}
@@
   Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
-  Taking $\tau = {50\vol{10r}}+1$, and using \cref{weight_product,probability}, we find that
+  Taking $\tau = {8\vol{5r}}+1$, and using \cref{weight_product,probability}, we find that
   \[
@@
   provided that
-  \[   |\Phi_\alpha| \ge {50\vol{10r}}\left(\log({50\vol{10r}}+1) + {2\vol{r}}(1+\log(2|\Phi_\alpha|))\right) \enspace .
+  \[   |\Phi_\alpha| \ge {8\vol{5r}}\big(\log({8\vol{5r}}+1) + {2\vol{2r+2}}\log(2|\Phi_\alpha|)\big) \enspace .
   \]
-  The right hand side of this last equation is of the form $O(r^4+r^2\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4)$.
+  The right hand side of this last equation is of the form $O(r^2\log r+r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d-\vol{2r+2}$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
 
@@
 
-  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
-On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 < {\vol{r}}$ for each $y\in Y$.
-Therefore, for any $A\subseteq X$, $|N_H(A)|\ge {2\vol{r}}|A|/{\vol{r}} = 2|A|$.
+  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\pi$ of $Y$ that produces a bipartite graph $H$ such that $\deg_H(\alpha)\ge {2\vol{2r+2}}$ for each $\alpha\in X$.
+On the other hand, $\deg_{H}(y)\le |\VE(G_{2r+2}(y))|-1 < {\vol{2r+2}}$ for each $y\in Y$.
+Therefore, for any $A\subseteq X$, $|N_H(A)|\ge {2\vol{2r+2}}|A|/{\vol{2r+2}} = 2|A|$.
 Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$.
@@
 Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)\cap\interior_r(G_{k'})|\ge d|A|$, for any $d\le 1/2\epsilon-2$.
-In particular, for some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
+In particular, for some $\epsilon \in \Omega(1/r^4\log r)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
 By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.

### 2022-05-10 lin-vs-cen.tex
Got rid of stupid letterpaper option

--- 2022-05-05 lin-vs-cen.tex
@@
     & \le \Pr(\binomial(t,1/25)\le 2\vol{r}) \notag \\
-    & \le (2t)^{2\vol{r}}(1/25)^{r^2}(1-1/25)^{t-2\vol{r}} \notag \\
+    & = \sum_{x=0}^{2\vol{r}} \binom{t}{x}\left(\frac{1}{25}\right)^x\left(\frac{24}{25}\right)^{t-x} \notag \\
+    & < \sum_{x=0}^{2\vol{r}} \binom{t}{x}\left(\frac{24}{25}\right)^{t} \notag \\
+    & \le (2t)^{2\vol{r}}\cdot\left(\frac{24}{25}\right)^{t} \notag \\
     % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
-    & \le \exp({2\vol{r}}(1+\log(2t))-t/25) \notag \\
-    & \le \exp({2\vol{r}}(1+\log(2|\Phi_\alpha|))-|\Phi_\alpha|/25\vol{10r}) \label{probability}
+    & = \exp\big({2\vol{r}}\ln(2t)-t\ln(25/24)\big) \notag \\
+    & \le \exp\left({2\vol{r}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|\ln(25/24)}{\vol{10r}}\right)
+      & \text{(since $|\Phi_\alpha|/\vol{10r} \le t \le |\Phi_\alpha|$)} \label{probability}
     % & \le \exp(-6(2r+1)^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
@@
 
-  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\alpha_1,\ldots,\alpha_{2|K|}$ such that $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
+  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\pi$ of $Y$ that produces a bipartition graph $H$ with $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
 On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 < {\vol{r}}$ for each $y\in Y$.

### 2022-05-10 lin-vs-cen.tex
Cleaned up some calculations

@@
   We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{r}$ for each $\alpha\in X$.
-  For each $\alpha\in X$, let $\Phi_\alpha:=\{\nu\in\VE(G_k):\phi(\nu)=\alpha\}$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\Phi_\alpha| \ge d$.
-Furthermore all but at most $\vol{r}$ elements of $\Phi_\alpha$ are contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.\todo{Explain more or redefine $\Phi_\alpha$ as $\Phi_\alpha\cap (\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu)))$)}
+  For each $\alpha\in X$, consider the set $\phi^{-1}_{\mathcal{P}}(\alpha)\subseteq \VE(G_k)$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\phi^{-1}_{\mathcal{P}}(\alpha)| \ge d$.
+Furthermore all but at most $\vol{r}$ elements of $\phi^{-1}_{\mathcal{P}}(\alpha)$ are contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.
+Let $\Phi_\alpha:=\phi^{-1}_{\mathcal{P}}(\alpha)\cap \left(\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))\right)$, so $|\Phi_\alpha|\ge d-\vol{2r+2}$.
+In words, $\Phi_\alpha$ consists of all objects of colour $\alpha$ that could not be chosen in the first round because they are too close to an element of $Y$.
+
+  % \todo{Explain more or redefine $\Phi_\alpha$ as $\Phi_\alpha\cap (\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu)))$)}
   % For each $\mu\in Y$, $|\VE(G_{r+1}(\mu))|\in O(r^2)$. Therefore, there are $\Omega(|\Phi_\alpha|/r^2)$ objects $\mu\in Y$ such that $\VE(G_{r+1}(\mu))\cap \Phi_\alpha\neq\emptyset$.
@@
 This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
-This already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$.
-
-  In order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.
-To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $\VE(G_r(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
-Since $|\VE(G_{10r}(\mu))|\le \vol{10r}$, each step in this process eliminates at most $\vol{10r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{10r}}$
+This is already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.\footnote{There are many ways to obtain a concentration result here, we use a simple method with the goal of being self-contained.
+For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $\VE(G_r(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
+Since $|\VE(G_{10r}(\mu))|\le \vol{10r}$, each step in this process eliminates at most $\vol{10r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{10r}}$.
 
@@
     & < \sum_{x=0}^{2\vol{r}} \binom{t}{x}\left(\frac{24}{25}\right)^{t} \notag \\
-    & \le (2t)^{2\vol{r}}\cdot\left(\frac{24}{25}\right)^{t} \notag \\
+    & < (2t)^{2\vol{r}}\cdot\left(\frac{24}{25}\right)^{t} \notag \\
     % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
@@
     & \le \exp\left({2\vol{r}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|\ln(25/24)}{\vol{10r}}\right)
-      & \text{(since $|\Phi_\alpha|/\vol{10r} \le t \le |\Phi_\alpha|$)} \label{probability}
+      & \text{(since $|\Phi_\alpha|/\vol{10r} \le t \le |\Phi_\alpha|$)} \notag \\
+    & \le \exp\left({2\vol{r}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|}{25\vol{10r}}\right)
+      & \text{(since $\ln(25/24) > 1/25$).}
+       \label{probability}
     % & \le \exp(-6(2r+1)^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
@@
   We are now ready for an application of \cref{weighted_lovasz}.
-For each $\alpha\in X$, let Let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{r}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
+For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{r}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
 For each $\alpha\in X$, define
   \[
-     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\neq\emptyset\right\}
+     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \left(\bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\right)\neq\emptyset\right\}
   \]
@@
     & = \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|-1} \notag \\
-    & \ge \exp\left(-\log\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
-    & \ge \exp\left(-\log\tau - \vol{10r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
+    & \ge \exp\left(-\ln\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
+    & \ge \exp\left(-\ln\tau - \vol{10r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
   \end{align}
@@
   provided that
-  \[   |\Phi_\alpha| \ge {50\vol{10r}}\left(\log({50\vol{10r}}+1) + {2\vol{r}}(1+\log(2|\Phi_\alpha|))\right) \enspace .
+  \[   |\Phi_\alpha| \ge {50\vol{10r}}\big(\ln({50\vol{10r}}+1) + {2\vol{r}}\ln(2|\Phi_\alpha|)\big) \enspace .
   \]
-  The right hand side of this last equation is of the form $O(r^4+r^2\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4)$.
+  The right hand side of this last equation is of the form $O(r^2+r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
 
@@
 
-  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\pi$ of $Y$ that produces a bipartition graph $H$ with $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
+  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\pi$ of $Y$ that produces a bipartite graph $H$ with $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
 On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 < {\vol{r}}$ for each $y\in Y$.

### 2022-05-11 lin-vs-cen.tex
Pfffft

@@
 \begin{lem}\label{packing_lemma}
-  Let $r\ge 2$ be an even integer and let $Q\subseteq \VE(\interior_r(G_k))$ be such that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
-  $|\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu))\}| \le 25$.
+  Let $r\ge 10$ be an even integer and let $Q\subseteq \VE(G_k)$ be such that $\mu_1\not\in \VE(G_{2r+1}(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$.  Then, for any $\mu\in \VE(G_k)$,
+  $|\{\mu_1\in Q: \mu\in \VE(G_{3r+1}(\mu_1))\}| \le 16$.
 \end{lem}
@@
 \begin{proof}
-  The packing condition that $\mu_1\not\in \VE(G_r(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $\VE(G_{r/2-1}(\mu_1))\cap \VE(G_{r/2-1}(\mu_2))=\emptyset$ for each $\mu_1,\mu_2\in Q$.
-Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{2r}(\mu_1))\}$ so that our task is to show that $|Q'|\le 25$.
+  To avoid the need to discuss boundary conditions that have no effect on the result, it is simpler to consider point that are sufficiently far from the boundary of $G_k$.
+More precisely, we may assume that $Q\cup\{\mu\}\subseteq\interior_{4r+2}(G_{k+8r+4})$.
+
+  The packing condition that $\mu_1\not\in \VE(G_{2r+1}(\mu_2))$ for each distinct $\mu_1,\mu_2\in Q$ implies that $\VE(G_{r}(\mu_1))\cap \VE(G_{r}(\mu_2))=\emptyset$.
+Let $Q':=\{\mu_1\in Q: \mu\in \VE(G_{3r+1}(\mu_1))\}$ so that our task is to show that $|Q'|\le 16$.
 For any $\mu_1\in Q'$,
-  $\mu_1\in\VE(G_{2r+1}(\mu))$ and consequently, $\VE(G_{r/2-1}(\mu_1))\subseteq \VE(G_{5r/2}(\mu))$.
-  Since this is true for all $\mu_1\in Q$ we have
-  \[  \vol{5r/2} \ge |\VE(G_{5r/2}(\mu))| \ge \sum_{\mu_1\in Q'} |\VE(G_{r/2}(\mu_1))|
-  = |Q'|\cdot \vol{r/2} \enspace .
+  $\mu_1\in\VE(G_{3r+2}(\mu))$ and consequently, $\VE(G_{r}(\mu_1))\subseteq \VE(G_{4r+2}(\mu))$.  For each $\mu_1\in Q'$, $|\VE(G_r(\mu_1))|= 12r^2+8r+1$.  Therefore,
+  \[  |Q'|\le \frac{\vol{4r+2}}{12r^2+8r+1} < 17 \enspace ,
   \]
-%   /* \begin{lstlisting}[frame=single,language=python]
-%     # Sage Code
-%   g(r)=2*(r+1)*(2*r+1)  # |B_r(mu)|
-%   show(g)
-%   show(g.factor())
-%   show(g.full_simplify())
-%   f(r)=2*(r+1)*(2*r+1 + 2*r) + (2*r+1)^2  # |VE(G[B_r(mu)])
-%   show(f)
-%   show(f.factor())
-%   vol = f
-%   ratio(r) = vol(5*r/2) / vol(r/2)  # ratio that appears in packing lemma
-%   show(plot(ratio, r, 1, 10))
-%   show(limit(ratio(r), r=oo))
-%   \end{lstlisting}
-  which implies that $|Q'|\le \vol{5r/2}/\vol{r/2}$.  For $r > 2$, this bound is increasing in $r$ and is less than $25$ for any $r\ge 2$.
+  for all $r\ge 10$.
 \end{proof}
@@
 \begin{lem}\label{doubled_colour_set}
-  For every integer $r\ge 1$ there exists an integer $d\in O(r^4\log r)$ such that the following is true, for every integer $k\ge 1$.\newline
+  For every integer $r\ge 10$ there exists an integer $d\in O(r^4\log r)$ such that the following is true, for every integer $k\ge 1$.\newline
   Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)|\ge d|A|$.
@@
     \item $\phi(\mu_1)=\phi(\mu_2)=\alpha$; and\label{hits_both_q}
-    \item for each $b\in\{1,2\}$, $\mu_b\not\in \VE(G_r(\mu_{3-b})) \cup \bigcup_{\mu\in Q_1} \VE(G_r(\mu))$,\label{spread_out_q}
+    \item for each $b\in\{1,2\}$, $\mu_b\not\in \VE(G_{2r+1}(\mu_{3-b})) \cup \bigcup_{\mu\in Q_1} \VE(G_{2r+1}(\mu))$,\label{spread_out_q}
   \end{compactenum}
@@
   Let $\overline{X}\subseteq\varphi(V(G))$ be the set of colours $\alpha$ for which the first round was an $\alpha$-success.
-At the end of this process, the set $Q_1$ certainly satisfies (\ref{spread_out}) since (\ref{spread_out_q}) implies that $\VE(G_r(\mu))\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$.
+At the end of this process, the set $Q_1$ certainly satisfies (\ref{spread_out}).
+In fact, it implies an even stronger property:  $|B_r(\nu)\cap Q_1|\le 1$ for any $\nu\in\VE(G_k)$.
 However, $Q_1$ only satisfies (\ref{hits_both}) for the colours in $\overline{X}$.
-We now use the second round to create a set $Q_2$ to fix this.
+We now use the second round to create a set $Q_2$ to handle the colours in $X:=\varphi(V(G))\setminus \overline{X}$.
   % Our strategy is to use \cref{d_hall} to find two matchings between the colours in $X$ and the boxes in $\{\tilde{B}_r(\mu):\mu\in Q_1\}$ that contain those colours.
 
-  Define a bipartite graph $H$ with parts $L(H):=X:=\varphi(V(G))\setminus \overline{X}$ and $R(H):=Y:=Q_1$.
-Let $\pi:=\mu_1,\ldots,\mu_{|Y|}$ be a random permutation of $Y$.
-We include an edge $\alpha\mu_i$ in $H$ if and only if
+  Define a bipartite graph $H$ with parts $L(H):=X$ and $R(H):=Y:=Q_1$.
+Let $\pi:=\mu_1,\ldots,\mu_{|Y|}$ be a random permutation of $Y$.\footnote{Here random permutation means a permutation chosen uniformly over all $|Y|!$ permutations of $Y$.}  We include an edge $\alpha\mu_i$ in $H$ if and only if
   \[
-    \VE(G_r(\mu_i))\setminus\left(\bigcup_{j=1}^{i-1}\VE(G_{2r}(\mu_j))\right)
+    \VE(G_r(\mu_i))\setminus\left(\bigcup_{j=1}^{i-1}\VE(G_{3r+1}(\mu_j))\right)
   \]
@@
 Note the asymmetry here: In the first round, each element $\mu_j$ that takes part in $Q_1$ prevents elements in $\VE(G_r(\mu))$ from also being part of $Q_1$.
-In the second round, $\mu_j$ defines a larger box $\VE(G_{2r}(\mu_j))$ that must be avoided when choosing a second element from $\VE(G_r(\mu_i))$, for each $i>j$.
+In the second round, $\mu_j$ defines a larger box $\VE(G_{3r+1}(\mu_j))$ that must be avoided when choosing a second element from $\VE(G_{2r+1}(\mu_i))$, for each $i>j$.
 
-  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{r}$ for each $\alpha\in X$.
-  For each $\alpha\in X$, consider the set $\phi^{-1}_{\mathcal{P}}(\alpha)\subseteq \VE(G_k)$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\phi^{-1}_{\mathcal{P}}(\alpha)| \ge d$.
-Furthermore all but at most $\vol{r}$ elements of $\phi^{-1}_{\mathcal{P}}(\alpha)$ are contained in $\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.
-Let $\Phi_\alpha:=\phi^{-1}_{\mathcal{P}}(\alpha)\cap \left(\bigcup_{\mu\in Y}\VE(G_{r+1}(\mu))\right)$, so $|\Phi_\alpha|\ge d-\vol{2r+2}$.
-In words, $\Phi_\alpha$ consists of all objects of colour $\alpha$ that could not be chosen in the first round because they are too close to an element of $Y$.
+  We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{2r+1}$ for each $\alpha\in X$.
+  For each $\alpha\in X$, consider the set $\phi^{-1}_{\mathcal{P}}(\alpha)\subseteq \VE(G_k)$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\phi^{-1}(\alpha)| \ge d$.
+Furthermore all but at most $\vol{2r+1}$ elements of $\phi^{-1}(\alpha)$ are contained in $\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.
+Let $\Phi_\alpha:=\phi^{-1}(\alpha)\cap \left(\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))\right)$, so $|\Phi_\alpha|\ge d-\vol{2r+2}$.
+In words, $\Phi_\alpha$ consists of all objects of colour $\alpha$ that were eliminated from consideration because they are too close to an element of $Y$.
 
@@
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{10r}},1/25)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{r})$.
-(We make no attempt to optimize constants here.)  Suppose that, for some $\mu\in Y$, $B_r(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$.
-Then, by \cref{packing_lemma}, the set
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{10r}},1/16)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{r})$.
+(We make no attempt to optimize constants here.)  Suppose that, for some $\alpha\in X$ and $\mu\in Y$, $B_{2r+1}(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and consider the set
   \[
-     Y_\nu := \{\xi\in Y:\nu\in\VE(G_{2r}(\xi))\}
+     Y_\nu := \{\xi\in Y:\nu\in\VE(G_{3r+1}(\xi))\} \enspace .
   \]
-  has size at most $25$.
   Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.
-This happens with probability exactly $1/|Y_\nu|\ge 1/25$.
+By \cref{packing_lemma}, $|Y_\nu|\le 16$, so the probability that the edge $\alpha\mu$ appears in $H$ is at least $1/|Y_\nu|\ge 1/16$.
 This is already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.\footnote{There are many ways to obtain a concentration result here, we use a simple method with the goal of being self-contained.
-For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J-\VE(G_{10r}(\mu))$.\footnote{The constant $10$ is overkill here and is only used for simplicity; $8r+O(1)$ would be sufficient.}  This process continues until $\VE(G_r(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
-Since $|\VE(G_{10r}(\mu))|\le \vol{10r}$, each step in this process eliminates at most $\vol{10r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{10r}}$.
+For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus \VE(G_{7r}(\nu))$.\footnote{The constant $7$ is overkill here and is only used for simplicity; $6r+O(1)$ would be sufficient.}  This process continues until $\VE(G_r(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
+Since $|\VE(G_{7r}(\mu))|\le \vol{7r}$, each step in this process eliminates at most $\vol{7r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{7r}}$.
 
-  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be the corresponding elements of $\Phi_\alpha$.
-The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are disjoint.
-Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_k$ are mutually independent.
-Indeed, each $U_i$ depends only on the relative order of $Y_{\nu_i}$ in the  permutation $\pi$.
-For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.
-Therefore, $\sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/25)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
+  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_t\}\subseteq\Phi_\alpha$ be the corresponding elements of $\Phi_\alpha$.
+The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are pairwise disjoint.
+Indeed, $Y_{\nu_i}\subseteq B_{3r+2}(\nu_i)$, $Y_{\nu_j}\subseteq B_{3r+2}(\nu_j)$ and the fact that $\nu_i\not\in B_{7r}(\nu_j)$ implies that $B_{3r+2}(\nu_i)\cap B_{3r+2}(\nu_j)=\emptyset$.
+Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_t$ are mutually independent.
+Indeed, each $U_i$ depends only on the relative order of the subset $Y_{\nu_i}$ in the permutation $\pi$.
+For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/16$.
+Therefore, $\deg_H(\alpha)\ge \sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/16)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
   \begin{align}
@@
     & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 2\vol{r}\right) \notag \\
-    & \le \Pr(\binomial(t,1/25)\le 2\vol{r}) \notag \\
-    & = \sum_{x=0}^{2\vol{r}} \binom{t}{x}\left(\frac{1}{25}\right)^x\left(\frac{24}{25}\right)^{t-x} \notag \\
-    & < \sum_{x=0}^{2\vol{r}} \binom{t}{x}\left(\frac{24}{25}\right)^{t} \notag \\
-    & < (2t)^{2\vol{r}}\cdot\left(\frac{24}{25}\right)^{t} \notag \\
+    & \le \Pr(\binomial(t,1/16)\le 2\vol{r}) \notag \\
+    & = \sum_{x=0}^{2\vol{2r+1}} \binom{t}{x}\left(\frac{1}{16}\right)^x\left(\frac{15}{16}\right)^{t-x} \notag \\
+    & < \sum_{x=0}^{2\vol{2r+1}} \binom{t}{x}\left(\frac{15}{16}\right)^{t} \notag \\
+    & < (2t)^{2\vol{2r+1}}\cdot\left(\frac{15}{16}\right)^{t} \notag \\
     % & \le \exp(6(2r+1)^2\log(2t)+(t-6(2r+1)^2)\log(1-1/25)) \\
-    & = \exp\big({2\vol{r}}\ln(2t)-t\ln(25/24)\big) \notag \\
-    & \le \exp\left({2\vol{r}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|\ln(25/24)}{\vol{10r}}\right)
+    & = \exp\big({2\vol{2r+1}}\ln(2t)-t\ln(16/15)\big) \notag \\
+    & \le \exp\left({2\vol{2r+1}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|\ln(16/15)}{\vol{10r}}\right)
       & \text{(since $|\Phi_\alpha|/\vol{10r} \le t \le |\Phi_\alpha|$)} \notag \\
-    & \le \exp\left({2\vol{r}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|}{25\vol{10r}}\right)
-      & \text{(since $\ln(25/24) > 1/25$).}
+    & \le \exp\left({2\vol{2r+1}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|}{16\vol{10r}}\right)
+      & \text{(since $\ln(16/15) > 1/16$).}
        \label{probability}
@@
   We are now ready for an application of \cref{weighted_lovasz}.
-For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{r}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
+For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{2r+1}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
 For each $\alpha\in X$, define
   \[
-     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \left(\bigcup_{\nu\in \Phi_\alpha} \VE(G_{10r}(\nu))\right)\neq\emptyset\right\}
+     \Gamma_\alpha := \left\{E_\beta \in \mathcal{E}: \Phi_\beta\cap \left(\bigcup_{\nu\in \Phi_\alpha} \VE(G_{7r}(\nu))\right)\neq\emptyset\right\}
   \]
-  and observe that $|\Gamma_\alpha|\le \vol{10r}\cdot|\Phi_\alpha|$.
-  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
+  and observe that $|\Gamma_\alpha|\le \vol{7r}\cdot|\Phi_\alpha|$.
+  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{3r+1}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{3r+1}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each entirely determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
 
-  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>0$ to be determined shortly.  Then
+  For each $\alpha\in X$, let $w(E_\alpha):=1/\tau$, for some $\tau>1$ to be defined shortly.  Then
   \begin{align}
@@
     & \ge \exp\left(-\ln\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
-    & \ge \exp\left(-\ln\tau - \vol{10r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
+    & \ge \exp\left(-\ln\tau - \vol{7r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
   \end{align}
@@
   Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
-  Taking $\tau = {50\vol{10r}}+1$, and using \cref{weight_product,probability}, we find that
+  Taking $\tau = {32\vol{7r}}+1$, and using \cref{weight_product,probability}, we find that
   \[
@@
   provided that
-  \[   |\Phi_\alpha| \ge {50\vol{10r}}\big(\ln({50\vol{10r}}+1) + {2\vol{r}}\ln(2|\Phi_\alpha|)\big) \enspace .
+  \[   |\Phi_\alpha| \ge {32\vol{7r}}\big(\ln({32\vol{7r}}+1) + {2\vol{2r+1}}\ln(2|\Phi_\alpha|)\big) \enspace .
   \]
   The right hand side of this last equation is of the form $O(r^2+r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
-
-  % Taking the ratio of \cref{ratioo} and $\Pr(E_\alpha)$ we obtain
-  % \begin{align*}
-  %   \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{\Pr(E_\alpha)}
-  %   & \ge \frac{e^{-C|\Phi_\alpha|/2r^2}-e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
-  %   & \ge e^{C|\Phi_\alpha|/2r^2} - \frac{e^{-C|(\Gamma_\alpha|+1)/50r^2}}{e^{-C|\Phi_\alpha|/r^2}} \\
-  % \end{align*}
-  %
-  % \todo[inline]{Hmmmmm. Not working.  Try something like $w(E_\alpha)=1-e^{-1/50}$}
-  %
-  % Finally, by \cref{packing_lemma} $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.
-  %
-  % \begin{align*}
-  %     \Pr(E_\alpha) \le \exp(-|\Phi_\alpha|\cdot C/r^2) \enspace .
-  % \end{align*}
-  %
-  % For each $\alpha\in X$, define $w(E_\alpha)=C/2r^2$ and observe that
-  % \begin{align*}
-  %   p(1-p)^{|\Gamma_\alpha|}
-  %   & \ge  p(1-p)^{25|\Phi_\alpha|} = p\exp(|\Phi_\alpha|\cdot 25\log(1-c/2r^2))
-  % \end{align*}
-  % \todo[inline]{Sort out these calculations}
 
@@
 Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)\cap\interior_r(G_{k'})|\ge d|A|$, for any $d\le 1/2\epsilon-2$.
-In particular, for some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
+In particular, for some $\epsilon \in \Omega(1/r^4\log r)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
 By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.
-Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
+Since $P$ is contained in $G'$, $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
 \end{proof}

### 2022-05-13 lin-vs-cen.tex
Ready to go?

@@
 Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.
+The fundamental problem on linear colouring is the question of upper bounding $\chicen(G)$ in terms of $\chilin(G)$.
 
@@
 
-\noindent The proof of \cref{kun_obrien_general} has three main ingredients:
+In the other direction, \citet{kun.obrien.ea:polynomial} show that, for any $\epsilon >0$, there exists a graph $G$ with $\chicen(G) \ge (2-\epsilon)\chilin(G)$.
+The same authors make the following bold conjecture:
+
+\begin{conj}[\citet{kun.obrien.ea:polynomial}]\label{crazy_conjecture}
+  For any graph $G$, $\chicen(G)\le 2\chilin(G)$.
+\end{conj}
+
+Currently, the only class of graphs for which centred chromatic number is known to be bounded by a linear function of linear chromatic number is the class of bounded degree trees \cite[Theorem~4]{kun.obrien.ea:polynomial}.
+To place the current contribution into context it is useful to understand the proof of \cref{kun_obrien_general}, which has three main ingredients:
 \begin{enumerate}
@@
 
-% \todo[inline]{Open problem here: Is there anything special about the relationship between treedepth and grid minors in planar graphs or product structured graphs?
-It's probably worth reading \citet{czerwinski.nadara.ea:improved} to see how their proof works.}
-
-% In addition to the improvement to \cref{kun_obrien_general}, this results in considerably better bounds for some specific graph classes.
-For example, the Excluded Grid Theorem for planar graphs asserts that any planar graph of treewidth $k$ contains a $\Omega(k)\times \Omega(k)$ grid minor.
-Using this result instead of Point~2, above, implies that for any planar graph $G$, $\chicen(G)\in O(\chilin^2(G))$.
-
-\citet{kun.obrien.ea:polynomial} boldly conjecture that, for any graph $G$, $\chicen(G)\le 2\chilin(G)$.
-However, before the current work, the only class of graphs for which centred chromatic number was known to be bounded by a linear function of linear chromatic number was bounded degree trees \cite[Theorem~4]{kun.obrien.ea:polynomial}.
-The current result adds further evidence in support of this conjecture by establishing that, when $G$ is a $k\times k$ grid (or even a pseudogrid), then $\chicen(G)\in\Theta(\chilin(G))$.
+In addition to the (admittedly modest) improvement on the exponent in \cref{kun_obrien_general}, the current result adds further evidence in support of \cref{crazy_conjecture} by establishing that, when $G$ is a $k\times k$ pseudogrid, then $\chicen(G)\in\Theta(\chilin(G))$.
 
@@
 
-  \item For these failed colours, we use \cref{d_hall} (the Polygamous Marriage Theorem) to find a $d$-to-$1$ matching $M$ that associates each failed colour $\alpha$ with $d$ vertices in $S_1$ that each contain a vertex of colour $\alpha$ nearby.
-In $M$, no vertex of $S_1$ is matched with more than one failed colour.
-
-  \item Using $M$ and \cref{weighted_lovasz} (the Lovász Local Lemma) we identify a set $S_2$ that contains two vertices of each of the missing colours and such that no vertex of $S_2$ is close to any other vertex of $S_2$.
+  \item For these failed colours, we use \cref{d_hall} (the Polygamous Marriage Theorem) and \cref{weighted_lovasz} (the Lovász Local Lemma) to identify a set $S_2$ that contains two vertices of each of the missing colours and such that no vertex of $S_2$ is close to any other vertex of $S_2$ and each vertex of $S_1$ is close to at most one vertex of $S_2$.
 
@@
 These vertices of $S_1$ are well-separated, but choosing one point from each of the three boxes to take part in $S_2$ could result in three vertices of $S_2$ being very close to each other.
-In particular, these three points could be vertices of $P_{\mu_1}$, $P_{\mu_2}$ and $P_{\mu_2}$ where $\mu_1$, $\mu_2$, and $\mu_3$ are objects in $\VE(G_{k\times k})$ that all contain a common vertex, making it impossible to find a path that contains all three (see \cref{bad_examples}).
-The local lemma is used to show that, in situations like this, we can still obtain two vertices from each missing colour class to put into $S_2$ without having to take more than one vertex from these three boxes.
+In particular, these three points could be vertices of $P_{\mu_1}$, $P_{\mu_2}$ and $P_{\mu_2}$ where $\mu_1$, $\mu_2$, and $\mu_3$ are objects in $\VE(G_{k\times k})$ that all contain a common vertex, making it difficult or impossible to find a single path that contains all three (see \cref{bad_examples}).
 
@@
 
+To avoid this, we take a random permutation $\pi$ of the vertices in $S_1$ and process them in order.
+When a vertex is processed it \defin{claims} all the unmarked vertices in its box and then marks all the vertices in a larger box so that no subsequent vertex in $S_1$ is able to claim them.
+The Local Lemma is then used to show that with positive probability, for each missing colour $\alpha$, there are many vertices of $S_1$ that claim a vertex of colour $\alpha$.
+Finally, the Marriage Theorem is then used to show that we can find a matching between vertices in $S_1$ and claimed vertices so that each vertex of $S_1$ that takes part in the matching is matched with a vertex it claims and, for each missing colour $\alpha$, there are two vertices of colour $\alpha$ that take part in the matching.
+
+The rest of this section expands the preceding sketch into a complete proof of \cref{pseudogrid_lower_bound}.
 In \cref{only_frequent_sec} we explain how to remove rows and columns of $G$ to obtain a sub-pseudogrid in which every colour appears frequently.
@@
 
-The following lemma gives conditions that allow us to delete rows and columns from $G$ to obtain a sub-pseudogrid in which every colour occurs frequently in the its interior.
+The following lemma gives conditions that allow us to delete rows and columns from $G$ to obtain a sub-pseudogrid in which every colour occurs frequently in the interior.
+% \todo{Do we really need $\interior_r(G)$ or can we use $\interior_2(G)$?  We could change it, but let's not.}
 
@@
 
-Having each vertex or edge $\mu$ of $G_k$ associated with a set $\varphi_\mathcal{P}(\mu)$ of colours rather than a single colour is problematic for what we want to to do next.
+Having each object $\mu\in\VE(G_k)$ associated with a set $\varphi_\mathcal{P}(\mu)$ of colours rather than a single colour is problematic for what we want to to do next.
 The following lemma allows us to choose one representative colour $\phi(\mu)$ from $\varphi_{\mathcal{P}}(\mu)$ for each $\mu\in \VE(G_k)$ while still ensuring that each colour appears frequently.
@@
 \begin{lem}\label{one_colour_per_object}
-  Let $d,r>1$ be integers, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)| \ge d|A|$.
-Then there exists a colouring $\phi:\VE(G_{k})\to\{1,\ldots,c,\perp\}$ with the following properties:
+  Let $d,r>1$ be integers, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)| \ge d|A|$.
+Then there exists a colouring $\phi:\VE(G_{k})\to \varphi(V(G))\cup\{\perp\}$ with the following properties:
   \begin{compactenum}[(i)]
-    \item For each $\mu\not\in\interior_r(G_{k})$, $\phi(\mu)=\perp$.
-    \item For each $\mu\in\VE(G_{k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi_\mathcal{P}(\mu)$.
-    \item For each $\alpha\in\varphi(V(G))$,
-    $|\phi^{-1}(\alpha)|\ge d$.
+    \item $\phi(\mu)=\perp$ for each For each $\mu\not\in\VE(\interior_r(G_{k}))$;
+    \item $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi_\mathcal{P}(\mu)$ for each $\mu\in\VE(G_{k})$; and
+    \item $|\phi^{-1}(\alpha)|\ge d$ for each $\mu\in\VE(G_{k})$.
     % $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
@@
 and, for any $v\in V(G)$, let $\tilde{B}_r(v):=\tilde{B}_r(\mu_v)$ where $P_{\mu_v}$ is the unique part in $\mathcal{P}$ that contains $v$.\footnote{Technically, the notation for $\tilde{B}_r(v)$ should include the partition $\mathcal{P}$, but we omit this since there will never be any ambiguity as to which partition is being used.}
-For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.\todo{Use $B$, $G$, $\tilde{B}$, and $\tilde{G}$ consistently!}
+For any $v\in V(G)$, define $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.
 
-The following lemma, whose proof is a (somewhat tedious) case analysis that appears in \cref{pick_up_two_proof}, is the main tool we use to build a path that contains a set of vertices that can be paired off in such a way that each pair is far from all other vertices.
+The following lemma, whose proof is a case analysis that appears in \cref{pick_up_two_proof}, is the main tool we use to build a path that contains a set of vertices that can be paired off in such a way that each pair is far from all other vertices.
 
@@
 
-The next lemma shows how to take a well separated collection of pairs and cover them with disjoint boxes that can be used with \cref{pick_up_two}.
+The next lemma shows how to take a well separated collection of pairs of vertices and cover them with disjoint boxes, each of which is compatible with \cref{pick_up_two}.
 
@@
 \begin{proof}
-  For any integer $q$, say that two elements $v,w\in S$ are a \defin{$q$-pair} if $v\in B_{q}(w)$ or $w\in B_q(v)$.
+  For any integer $q$, say that two elements $v,w\in S$ are a \defin{$q$-pair} if $v\in \tilde{B}_{q}(w)$ or $w\in \tilde{B}_q(v)$.
 First observe that any $v\in S$ takes part in at most one $(r-1)$-pair since, otherwise, $|\tilde{B}_{r}(v)\cap S|\ge 3$.
@@
 Now, consider a `snake-like' path $P_0$ that contains every row $i$ with $i\equiv 1\pmod{2p+1}$ and, aside from those rows contains only vertices of columns $1$ and $k$.
-For each $x\in X$, this path intersects exactly one row of $\tilde{B}_{p+1}(x)$.
+For each $x\in X$, this path intersects exactly one row of $G[\tilde{B}_{p+1}(x)]$.
 For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{p+1}(x)$ with a path that is contained in $\tilde{B}_{p+1}(x)$ and contains the (at most two) vertices of $\tilde{B}_{p+1}(x)\cap S$.
@@
 \begin{proof}
-  To avoid the need to discuss boundary conditions that have no effect on the result, it is simpler to consider point that are sufficiently far from the boundary of $G_k$.
+  To avoid the need to discuss boundary conditions that have no effect on the result, it is simpler to consider points that are sufficiently far from the boundary of $G_k$.
 More precisely, we may assume that $Q\cup\{\mu\}\subseteq\interior_{4r+2}(G_{k+8r+4})$.
@@
   \begin{compactenum}[(i)]
-    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\varphi^{-1}(\alpha)\cap S|= 2$; and
-    \item \label{spread_out} for each $v\in S$, $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
+    \item \label{hits_both} $|\varphi^{-1}(\alpha)\cap S|= 2$ for each $\alpha\in\varphi(V(G))$; and
+    \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
   \end{compactenum}
@@
 \begin{proof}
-  We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi:\VE(G_k)\to\varphi(V(G))$.
+  We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi:\VE(G_k)\to\varphi(V(G))\cup\{\perp\}$.
 Observe that it is now sufficient to find a $2|\varphi(V(G))|$-element subset $Q\subseteq\VE(\interior_r(G_k))$ such that
@@
 At the end of this process, the set $Q_1$ certainly satisfies (\ref{spread_out}).
-In fact, it implies an even stronger property:  $|B_r(\nu)\cap Q_1|\le 1$ for any $\nu\in\VE(G_k)$.
+In fact, it implies an even stronger property: $B_{2r+1}(\mu)\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$, which implies that $|B_r(\nu)\cap Q_1|\le 1$ for each $\nu\in\VE(G_k)$.
 However, $Q_1$ only satisfies (\ref{hits_both}) for the colours in $\overline{X}$.
 We now use the second round to create a set $Q_2$ to handle the colours in $X:=\varphi(V(G))\setminus \overline{X}$.
-  % Our strategy is to use \cref{d_hall} to find two matchings between the colours in $X$ and the boxes in $\{\tilde{B}_r(\mu):\mu\in Q_1\}$ that contain those colours.
 
   Define a bipartite graph $H$ with parts $L(H):=X$ and $R(H):=Y:=Q_1$.
-Let $\pi:=\mu_1,\ldots,\mu_{|Y|}$ be a random permutation of $Y$.\footnote{Here random permutation means a permutation chosen uniformly over all $|Y|!$ permutations of $Y$.}  We include an edge $\alpha\mu_i$ in $H$ if and only if
+Let $\pi:=\mu_1,\ldots,\mu_{|Y|}$ be a random permutation of $Y$.\footnote{Here random permutation means a permutation chosen uniformly from the set of all $|Y|!$ permutations of $Y$.}  We include an edge $\alpha\mu_i$ in $H$ if and only if
   \[
-    \VE(G_r(\mu_i))\setminus\left(\bigcup_{j=1}^{i-1}\VE(G_{3r+1}(\mu_j))\right)
+    \VE(G_{2r+1}(\mu_i))\setminus\left(\bigcup_{j=1}^{i-1}\VE(G_{3r+1}(\mu_j))\right)
   \]
   contains some object $\nu$ of colour $\phi(\nu)=\alpha$.
-Note the asymmetry here: In the first round, each element $\mu_j$ that takes part in $Q_1$ prevents elements in $\VE(G_r(\mu))$ from also being part of $Q_1$.
-In the second round, $\mu_j$ defines a larger box $\VE(G_{3r+1}(\mu_j))$ that must be avoided when choosing a second element from $\VE(G_{2r+1}(\mu_i))$, for each $i>j$.
+A helpful way to view this process is as follows: Initially, every vertex of $G_k$ is \defin{unmarked}.
+For $i=1,\ldots,|Y|$, $\mu_i$ claims every unmarked object in $\VE(G_{2r+1}(\mu_i))$ and then marks every object in $\VE(G_{3r+1}(\mu_i))$.
+Then $H$ contains the edge $\alpha\mu_i$ if and only if $\mu_i$ claims at least one object of colour $\alpha$.
 
   We now want to use the Local Lemma (\cref{weighted_lovasz}) to show that, with probability greater than zero, $\deg_H(\alpha)\ge 2\vol{2r+1}$ for each $\alpha\in X$.
-  For each $\alpha\in X$, consider the set $\phi^{-1}_{\mathcal{P}}(\alpha)\subseteq \VE(G_k)$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\phi^{-1}(\alpha)| \ge d$.
-Furthermore all but at most $\vol{2r+1}$ elements of $\phi^{-1}(\alpha)$ are contained in $\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.
-Let $\Phi_\alpha:=\phi^{-1}(\alpha)\cap \left(\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))\right)$, so $|\Phi_\alpha|\ge d-\vol{2r+2}$.
+For each $\alpha\in X$, consider the set $\phi^{-1}(\alpha)\subseteq \VE(G_k)$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\phi^{-1}(\alpha)| \ge d$.
+Furthermore all but at most $\vol{r+1}$ elements of $\phi^{-1}(\alpha)$ are contained in $\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.
+Let $\Phi_\alpha:=\phi^{-1}(\alpha)\cap \left(\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))\right)$, so $|\Phi_\alpha|\ge d-\vol{r+1}$.
 In words, $\Phi_\alpha$ consists of all objects of colour $\alpha$ that were eliminated from consideration because they are too close to an element of $Y$.
@@
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{10r}},1/16)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{r})$.
-(We make no attempt to optimize constants here.)  Suppose that, for some $\alpha\in X$ and $\mu\in Y$, $B_{2r+1}(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and consider the set
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{7r}},1/16)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+1})$.
+Suppose that, for some $\alpha\in X$ and $\mu\in Y$, $B_{2r+1}(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and consider the set
   \[
@@
   \]
-  Observe that the edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.
+  The edge $\alpha\mu$ will appear in $H$ if, in our random permutation $\pi$ of $Y$, $\nu$ appears before any other element of $Y_\nu$.
 By \cref{packing_lemma}, $|Y_\nu|\le 16$, so the probability that the edge $\alpha\mu$ appears in $H$ is at least $1/|Y_\nu|\ge 1/16$.
-This is already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.\footnote{There are many ways to obtain a concentration result here, we use a simple method with the goal of being self-contained.
-For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_r(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus \VE(G_{7r}(\nu))$.\footnote{The constant $7$ is overkill here and is only used for simplicity; $6r+O(1)$ would be sufficient.}  This process continues until $\VE(G_r(\mu))\cap J=\emptyset$ for each $\mu\in Y$.
-Since $|\VE(G_{7r}(\mu))|\le \vol{7r}$, each step in this process eliminates at most $\vol{7r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{7r}}$.
+This is already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.\footnote{There are many ways to obtain a concentration result here and we use a simple method with the goal of being self-contained.
+For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+1}(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus \VE(G_{7r}(\nu))$.\footnote{The constant $7$ is overkill here and is only used for simplicity; $6r+O(1)$ would be sufficient.}  This process continues until $J$ is empty.
+Since $|\VE(G_{7r}(\mu))|\le \vol{7r}$, each iteration in this process eliminates at most $\vol{7r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{7r}}$.
 
@@
 The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are pairwise disjoint.
-Indeed, $Y_{\nu_i}\subseteq B_{3r+2}(\nu_i)$, $Y_{\nu_j}\subseteq B_{3r+2}(\nu_j)$ and the fact that $\nu_i\not\in B_{7r}(\nu_j)$ implies that $B_{3r+2}(\nu_i)\cap B_{3r+2}(\nu_j)=\emptyset$.
+Indeed, $Y_{\nu_i}\subseteq B_{3r+2}(\nu_i)$, $Y_{\nu_j}\subseteq B_{3r+2}(\nu_j)$ and the fact that $\nu_i\not\in B_{7r}(\nu_j)$ implies that $G_{3r+2}(\nu_i)\cap G_{3r+2}(\nu_j)=\emptyset$.
 Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_t$ are mutually independent.
@@
   \begin{align}
-    \Pr\left(\deg_H(\alpha)< 2\vol{r}\right)
-    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 2\vol{r}\right) \notag \\
-    & \le \Pr(\binomial(t,1/16)\le 2\vol{r}) \notag \\
+    \Pr\left(\deg_H(\alpha)< 2\vol{2r+1}\right)
+    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 2\vol{2r+1}\right) \notag \\
+    & \le \Pr(\binomial(t,1/16)\le 2\vol{2r+1}) \notag \\
     & = \sum_{x=0}^{2\vol{2r+1}} \binom{t}{x}\left(\frac{1}{16}\right)^x\left(\frac{15}{16}\right)^{t-x} \notag \\
@@
     & = \exp\big({2\vol{2r+1}}\ln(2t)-t\ln(16/15)\big) \notag \\
-    & \le \exp\left({2\vol{2r+1}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|\ln(16/15)}{\vol{10r}}\right)
-      & \text{(since $|\Phi_\alpha|/\vol{10r} \le t \le |\Phi_\alpha|$)} \notag \\
-    & \le \exp\left({2\vol{2r+1}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|}{16\vol{10r}}\right)
+    & \le \exp\left({2\vol{2r+1}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|\ln(16/15)}{\vol{7r}}\right)
+      & \text{(since $|\Phi_\alpha|/\vol{7r} \le t \le |\Phi_\alpha|$)} \notag \\
+    & \le \exp\left({2\vol{2r+1}}\ln(2|\Phi_\alpha|)-\frac{|\Phi_\alpha|}{16\vol{7r}}\right)
       & \text{(since $\ln(16/15) > 1/16$).}
@@
   We are now ready for an application of \cref{weighted_lovasz}.
-For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 2\vol{2r+1}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
+For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)< 2\vol{2r+1}$'' and let $\mathcal{E}:=\{E_\alpha:\alpha\in X\}$.
 For each $\alpha\in X$, define
@@
   and observe that $|\Gamma_\alpha|\le \vol{7r}\cdot|\Phi_\alpha|$.
-  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{3r+1}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{3r+1}(\mu)\cap \Phi_\beta\neq\emptyset\}$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each entirely determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
+  The event $E_\alpha$ is mutually independent of $\mathcal{E}\setminus\Gamma_\alpha$ since, for any $E_\beta\in\mathcal{E}\setminus\Gamma_\alpha$ the sets $Y_\alpha:=\bigcup_{\nu\in\Phi_\alpha} Y_\nu$ and $Y_\beta:=\bigcup_{\nu\in \Phi_\beta} Y_\nu$ are disjoint and $\deg_H(\alpha)$ and $\deg_H(\beta)$ are each entirely determined by the relative orders of $Y_\alpha$ and $Y_\beta$ in $\pi$, respectively.
 
@@
     & \ge \exp\left(-\ln\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
-    & \ge \exp\left(-\ln\tau - \vol{7r}\cdot|\Phi_\alpha|/(\tau-1)\right) \enspace . \label{weight_product}
+    & \ge \exp\left(-\ln\tau - \vol{7r}\cdot|\Phi_\alpha|/(\tau-1)\right)
+    & \text{(since $|\Gamma_\alpha|\le \vol{7r}\cdot\Phi_\alpha$).}  \label{weight_product}
   \end{align}
@@
   Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
-  Taking $\tau = {32\vol{7r}}+1$, and using \cref{weight_product,probability}, we find that
+  Taking $\tau = {32\vol{7r}^2}+1$, and using \cref{weight_product,probability}, we find that
   \[
-     \Pr\left(E_\alpha\right) \le w(E_\alpha)\prod_{\beta\in\Gamma_\alpha}(1-w(\Gamma_\alpha))
+     \Pr\left(E_\alpha\right) \le w(E_\alpha)\prod_{\beta\in\Gamma_\alpha}(1-w(E_\beta))
   \]
@@
   \]
-  The right hand side of this last equation is of the form $O(r^2+r^4\log|\Phi_\alpha|)$.  Since $|\Phi_\alpha|\ge d$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
+  The right hand side of this last equation is of the form $O(r^2\log r+r^4\log|\Phi_\alpha|)$.
+Since $|\Phi_\alpha|\ge d-\vol{r+1}\in d- O(r^2)$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
 
-  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\pi$ of $Y$ that produces a bipartite graph $H$ with $\deg_H(\alpha)\ge {2\vol{r}}$ for each $\alpha\in X$.
-On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 < {\vol{r}}$ for each $y\in Y$.
-Therefore, for any $A\subseteq X$, $|N_H(A)|\ge {2\vol{r}}|A|/{\vol{r}} = 2|A|$.
+  Therefore, by \cref{weighted_lovasz}, there exists a permutation $\pi$ of $Y$ that produces a bipartite graph $H$ with $\deg_H(\alpha)\ge {2\vol{2r+1}}$ for each $\alpha\in X$.
+On the other hand, $\deg_{H}(y)\le |\VE(G_r(y))|-1 < {\vol{2r+1}}$ for each $y\in Y$.
+Therefore, for any $A\subseteq X$, $|N_H(A)|> {2\vol{2r+1}}|A|/{\vol{2r+1}} = 2|A|$.
 Therefore, by \cref{d_hall}, there is a subgraph $M$ of $H$ in which $\deg_M(\alpha)=2$ for each $\alpha\in X$ and $\deg_M(\mu)\le 1$ for each $\mu\in Y$.
-Each edge $\alpha\mu$ of $M$ corresponds to some $\nu\in X_\alpha\cap\VE(G_r(\mu))$, which we place in $Q_2$.
+Each edge $\alpha\mu$ of $M$ corresponds to some $\nu\in\Phi_\alpha\cap\VE(G_r(\mu))$, which we place in $Q_2$.
 
@@
 \begin{proof}[Proof of \cref{pick_up_two}]
-  Let $\mathcal{P}:\{P_\mu:\mu\in \VE(G_k)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_a)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.
+  Let $\mathcal{P}:=\{P_\mu:\mu\in \VE(G_a)\}$ be a grid partition of $G$ and let $\mu_v,\mu_w\in \VE(G_a)$ be such that $v\in V(P_{\mu_v})$ and $w\in V(P_{\mu_w})$.
 Define $\mu_s$ and $\mu_t$ similarly, with respect to $s$ and $t$.
@@
       \item If $\mu_v$ is straight then we replace $\mu_v$ with the other vertical edge incident on $\mu_v$.
-      \item If $\mu_v$ is bent then we replace $\mu_v$ with the horizontal edge joining $\mu_v$ to a vertex in column $i+1$.
+      \item If $\mu_v$ is bent then we replace $\mu_v$ with the horizontal edge joining $\mu_v$ to a vertex in column $i-1$.
     \end{compactitem}
@@
     \end{compactitem}
-    If $\mu_w$ is an edge of $G_a$ then there is nothing else to. If $\mu_w$ is a vertex of $G_a$ then there are two cases to consider:
+    If $\mu_w$ is an edge of $G_a$ then there is nothing else to do. If $\mu_w$ is a vertex of $G_a$ then there are two cases to consider:
     \begin{compactitem}

### 2022-05-27 lin-vs-cen.tex
Last touch-ups.  Submit this weekend, maybe.

@@
 For a vertex $v\in V(G)$,  $N_G(v):=\{w\in V(G):vw\in E(G)\}$  denotes the open neighbourhood of $v$ in $G$ and for any set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$.
-We use $\deg_G(v):=|N_G(v)|$ to denote the degree of the vertex $v$ in the graph $G$.
+We use $\deg_G(v):=|N_G(v)|$ to denote the \defin{degree} of the vertex $v$ in the graph $G$.
 
@@
 
-\cref{d_hall} can be deduced by adding $d-1$ twins $x_2,\ldots,x_d$ for each vertex $x\in L(H)$, i.e., by applying \cref{hall} on the graph $H':=H\cup\{x_iy:x\in L(H),\, i\in\{2,\ldots,d\},\, y\in N_H(x)\}$.
+\cref{d_hall} can be deduced from \cref{hall} by adding $d-1$ \defin{twins} $x_2,\ldots,x_d$ for each vertex $x\in L(H)$, i.e., by applying \cref{hall} on the graph $H':=H\cup\{x_iy:x\in L(H),\, i\in\{2,\ldots,d\},\, y\in N_H(x)\}$.
 
@@
 For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the \defin{$j$th row} is the vertex set $\{(1,j),\ldots,(a,j)\}$.
-For any integer $0\le r\le\min\{a,b\}/2$, the \defin{r-interior} of $G_{a\times b}$ defined as $\interior_r(G_{a\times b}):=G_{a\times b}[\{1+r,\ldots,a-r\}\times\{1+r,\ldots,b-r\}]$.
+For any integer $0\le r<\min\{a,b\}/2$, the \defin{r-interior} of $G_{a\times b}$ defined as $\interior_r(G_{a\times b}):=G_{a\times b}[\{1+r,\ldots,a-r\}\times\{1+r,\ldots,b-r\}]$.
 For $r\ge\min\{a,b\}/2$, $\interior_r(G_{a\times b})$ is the empty graph.
@@
 (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
-  \item Replace each degree-$4$ vertex $(i,j)$ of $G_{a\times b}$ with a non-empty path $P_v$.
-If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each of $(i\pm 1,j\pm 1)$ and this has no effect on the underlying graph.
-Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $(i\pm 1,j\pm 1)$.
-It is useful to consider three possible cases:
+  \item Replace each degree-$4$ vertex $v:=(i,j)$ of $G_{a\times b}$ with a non-empty path $P_v$.
+If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each element of $N_{G_{a\times b}}(v)$ and this has no effect on the underlying graph.
+Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $N_{G_{a\times b}}(v)$.
+It is useful to consider three possible cases, each of which appears at least once in \cref{pseudogrid_fig}:
   \begin{compactenum}[(Q1)]
@@
 Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path $C$  in $G$ that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
-This correspondence allows us to talk about the rows and columns of $G$, which we will do immediately.
+This correspondence allows us to talk about the \defin{rows} and \defin{columns} of $G$, which we will do immediately.
 
@@
 We will prove \cref{pseudogrid_lower_bound} by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
-We will prove the existence of $P$ in several steps (see \cref{outline}):
+We will prove the existence of $P$ in several steps; see \cref{outline}:
 
@@
   % This makes it easy to find a single path in $G$ that contains each of these vertices.
-  Unfortunately, this process will fail for some subset of the colours that appear in $G$.
+  Unfortunately, this process can fail for some subset of the colours that appear in $G$.
 
@@
 In \cref{packing_lemma_sec} we introduce a fairly standard-looking packing lemma that is needed in several places later.
-In \cref{well_separated} we explain how find a well-separated set $S$ that contains two vertices of each colour.
-Finally \cref{wrapping_up} shows how to assemble these various pieces to prove \cref{pseudogrid_lower_bound}.
+In \cref{well_separated} we explain how to find a well-separated set $S$ that contains two vertices of each colour.
+Finally, \cref{wrapping_up} shows how to assemble these various pieces to prove \cref{pseudogrid_lower_bound}.
 
@@
 
-For convenience, let $G_k:=G_{k\times k}$.
+For the sake of compactness, let $G_k:=G_{k\times k}$.
 Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=\{P_\mu:\mu\in\VE(G_{k})\}$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
@@
 For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-1}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-1}(\alpha)$.
-(Throughout this section, we will use the idiom $\varphi(V(G))$ to denote the set of all colours used by $\varphi$ to colour the vertices of $G$.)
+Throughout this section, we will use the idiom $\varphi(V(G))$ to denote the set of all colours used by $\varphi$ to colour the vertices of $G$.
 
@@
 \begin{lem}\label{pick_up_two}
-  Let $G$ be a $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in\interior_1(G)$, and $t$ in column $a$ of $G$.
+  Let $G$ be a $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.
 Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
@@
   Refer to \cref{make_disjoint_fig}.
-We will define the set $X$ so that it satisfies (\ref{covers_s}) and for each $x\in X$ we will choose $1$ or $2$ elements of $\tilde{B}_{p}(x)\cap S$ and say that $x$ \defin{covers} those elements.
+We will define the set $X$ so that it satisfies (\ref{covers_s}) and for each $x\in X$ we will choose one or two elements of $\tilde{B}_{p}(x)\cap S$ and say that $x$ \defin{covers} those elements.
 Let $S_1$ be the subset of $S$ containing only those elements that do not take part in any $(r-1)$-pair.
@@
   To see that $X$ satisifies (\ref{disjoint_boxes}), observe that if $x\in X$ covers $v\in S$ then $\tilde{B}_{p+1}(x)\subseteq \tilde{B}_{2p+2}(v)$.
-Therefore, if $x\in X$ covers $v\in S$ and $y\in X$ covers $w\in S$ and $\tilde{B}_{p+1}(x)\cap \tilde{B}_{p+1}(y)\neq\emptyset$ then $w\in \tilde{B}_{4p+4}(v)$ or $v\in \tilde{B}_{4p+4}(w)$.
+Therefore, if $x\in X$ covers $v\in S$,  $y\in X$ covers $w\in S$, and $\tilde{B}_{p+1}(x)\cap \tilde{B}_{p+1}(y)\neq\emptyset$ then $w\in \tilde{B}_{4p+4}(v)$ or $v\in \tilde{B}_{4p+4}(w)$.
 Since $r-1 \ge 4p+4$, this implies that $v=w$ or $v$ and $w$ are an $(r-1)$-pair.
@@
 For each $x\in X$, this path intersects exactly one row of $G[\tilde{B}_{p+1}(x)]$.
-For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{p+1}(x)$ with a path that is contained in $\tilde{B}_{p+1}(x)$ and contains the (at most two) vertices of $\tilde{B}_{p+1}(x)\cap S$.
+For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap \tilde{B}_{p+1}(x)$ with a path that is contained in $\tilde{B}_{p+1}(x)$ and contains the (at most two) vertices of $\tilde{B}_{p+1}(x)\cap S$.
 After doing this for each $x\in X$ we obtain a path $P$ that contains every vertex in $S$.
@@
 At the end of this process, the set $Q_1$ certainly satisfies (\ref{spread_out}).
-In fact, it implies an even stronger property: $B_{2r+1}(\mu)\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$, which implies that $|B_r(\nu)\cap Q_1|\le 1$ for each $\nu\in\VE(G_k)$.
+In fact, it satisfies an even stronger property: $B_{2r+1}(\mu)\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$, which implies that $|B_r(\nu)\cap Q_1|\le 1$ for each $\nu\in\VE(G_k)$.
 However, $Q_1$ only satisfies (\ref{hits_both}) for the colours in $\overline{X}$.
@@
 This is already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.\footnote{There are many ways to obtain a concentration result here and we use a simple method with the goal of being self-contained.
-For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+1}(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus \VE(G_{7r}(\nu))$.\footnote{The constant $7$ is overkill here and is only used for simplicity; $6r+O(1)$ would be sufficient.}  This process continues until $J$ is empty.
+For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+1}(\mu))\cap J$ contains at least one element $\nu$ and then setting $J:=J\setminus \VE(G_{7r}(\nu))$.\footnote{The constant $7$ is overkill here and is only used for simplicity; $6r+O(1)$ would be sufficient.}  This process continues until $J$ is empty.
 Since $|\VE(G_{7r}(\mu))|\le \vol{7r}$, each iteration in this process eliminates at most $\vol{7r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{7r}}$.
@@
 The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are pairwise disjoint.
-Indeed, $Y_{\nu_i}\subseteq B_{3r+2}(\nu_i)$, $Y_{\nu_j}\subseteq B_{3r+2}(\nu_j)$ and the fact that $\nu_i\not\in B_{7r}(\nu_j)$ implies that $G_{3r+2}(\nu_i)\cap G_{3r+2}(\nu_j)=\emptyset$.
+Indeed, $Y_{\nu_i}\subseteq B_{3r+2}(\nu_i)$, $Y_{\nu_j}\subseteq B_{3r+2}(\nu_j)$ and the fact that $\nu_i\not\in B_{7r}(\nu_j)$ implies that $\VE(G_{3r+2}(\nu_i))\cap \VE(G_{3r+2}(\nu_j))=\emptyset$.
 Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_t$ are mutually independent.
@@
 We are grateful to the organizers and participants of both workshops for providing a stimulating research environment.
-Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
+Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace and to Omer Angel for pointing out that the problem on $k\times k$ grid has a simple solution.
 

### 2022-05-30 lin-vs-cen.tex
Fixed sentence in Ack

@@
 We are grateful to the organizers and participants of both workshops for providing a stimulating research environment.
-Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace and to Omer Angel for pointing out that the problem on $k\times k$ grid has a simple solution.
+Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
+We are also grateful to Omer Angel for pointing out that the problem on $k\times k$ grid (as opposed to pseudogrid) has a simple solution.
 

### 2022-05-30 lin-vs-cen.tex
Added affiliations

@@
 \title{\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}}
-\author{Prosenjit~Bose, Vida~Dujmović, Hussein~Houdrouge, Mehrnoosh~Javarsineh, and Pat~Morin}
+\author{Prosenjit~Bose%
+  \thanks{School of Computer Science, Carleton University},\,
+  Vida~Dujmović%
+  \thanks{Department of Computer Science and Electrical Engineering, University of Ottawa},\,\,
+  Hussein~Houdrouge\footnotemark[2],\,
+  Mehrnoosh~Javarsineh\footnotemark[2],\,
+  and Pat~Morin\footnotemark[2]}
 

### 2022-07-20 lin-vs-cen.tex
Vida's new abstract

--- 2022-05-30 lin-vs-cen.tex
@@
 \begin{abstract}
-  We prove that the linear chromatic number of any $k\times k$ pseudogrid is $\Omega(k)$.
+  A \emph{centered colouring} of a graph is one in which every connected subgraph contains a vertex whose colour is unique and a \emph{linear colouring} is one in which every path contains a vertex whose colour is unique.
+For a graph $G$ the \emph{centered colouring number} $\chicen(G)$ and the \emph{linear colouring number} $\chilin(G)$ denote the minimum number distinct colours required for a centered, respectively, linear colouring of $G$.
+From these definitions, it follows immediately that $\chilin(G)\le \chicen(G)$ for every graph $G$.
+The centered colouring number is equivalent to treedepth and has been studied extensively, but much less is known about linear colouring.
+Kun \etal\ [\textit{Algorithmica} \textbf{83}(1)] prove that $\chicen(G) \le \tilde{O}(\chilin(G)^{190})$ for any graph $G$ and conjecture that $\chicen(G)\le 2\chilin(G)$.
+Their upper bound was subsequently improved by Czerwinski \etal\ [\textit{SIDMA} \textbf{35}(2)] to $\chicen(G)\le\tilde{O}(\chilin(G)^{19})$.
+The proof of both upper bounds relies on establishing a lower bound on the linear coloring number of pseudogrids, which appear in the proof due to their critical relationship to treewidth.
+Specifically, Kun \etal\ prove that $k\times k$ pseudogrids have linear colouring $\Omega(\sqrt{k})$.
+Our main contribution is establishing a tight
+  bound on the linear coloring number of pseudogrids, specifically $\chilin(G)\ge \Omega(k)$
+  for every $k\times k$ pseudogrid $G$. As a consequence we improve the general bound
+  for all graphs to $\chicen(G)\le \tilde{O}(\chilin(G)^{9})$.
+In addition this tight bound gives further evidence in support of Kun \etal's conjecture (above) that the centered colouring number (i.e.
+treedepth) of any graph is upper bounded by a linear function of its linear chromatic number.
+%   entered coloring (def). linear col definition. centered col same as
+% well studied tree depth. x_cen <= x_lin. There are plethora of results
+% on centred colouring much less is known about linear col. Bla proved
+% that x_cen <= BLA and conjectured that x_cen < 2 x_lin, for all
+% graphs. Their upper bound proof relies on establishing a lower bound
+% on the linear coloring number of pseudo grids. Pseudo grids appear in
+% the proof due to their critical relationship to tree width.
+% Specifically they prove that k x k pseud grids have linear colouring
+% number Omega(sqrt(k)). Our main contribution is establishing a tight
+% bound on linear coloring number of pseudo grids, specifically Theta(k)
+% for k x k pseudogrids. As a consequence we improved the general bound
+% for all graphs to BLA. In addition this tight bound gives further
+% evidence in support to the BLA's conjecture above that centered
+% colouring's (i.e. treedepth) of any graph is upper bounded by a linear
+% function of its linear chromatic number.
+
+  % We prove that the linear chromatic number of any $k\times k$ pseudogrid is $\Omega(k)$.
 By an argument of Kun et al (Algorithmica, 2021), this result gives a tighter upper bound on the treedepth of a graph as a function of its linear chromatic number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear chromatic number.
@@
 For a vertex $v\in V(G)$,  $N_G(v):=\{w\in V(G):vw\in E(G)\}$  denotes the open neighbourhood of $v$ in $G$ and for any set $S\subseteq V(G)$, $N_G(S):=\{w\in V(G):vw\in E(G),\, v\in S,\, w\not\in S\}$.
-We use $\deg_G(v):=|N_G(v)|$ to denote the \defin{degree} of the vertex $v$ in the graph $G$.
+We use $\deg_G(v):=|N_G(v)|$ to denote the degree of the vertex $v$ in the graph $G$.
 
@@
 
-\cref{d_hall} can be deduced from \cref{hall} by adding $d-1$ \defin{twins} $x_2,\ldots,x_d$ for each vertex $x\in L(H)$, i.e., by applying \cref{hall} on the graph $H':=H\cup\{x_iy:x\in L(H),\, i\in\{2,\ldots,d\},\, y\in N_H(x)\}$.
+\cref{d_hall} can be deduced by adding $d-1$ twins $x_2,\ldots,x_d$ for each vertex $x\in L(H)$, i.e., by applying \cref{hall} on the graph $H':=H\cup\{x_iy:x\in L(H),\, i\in\{2,\ldots,d\},\, y\in N_H(x)\}$.
 
@@
 For each $i\in\{1,\ldots,a\}$, the \defin{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the \defin{$j$th row} is the vertex set $\{(1,j),\ldots,(a,j)\}$.
-For any integer $0\le r<\min\{a,b\}/2$, the \defin{r-interior} of $G_{a\times b}$ defined as $\interior_r(G_{a\times b}):=G_{a\times b}[\{1+r,\ldots,a-r\}\times\{1+r,\ldots,b-r\}]$.
+For any integer $0\le r\le\min\{a,b\}/2$, the \defin{r-interior} of $G_{a\times b}$ defined as $\interior_r(G_{a\times b}):=G_{a\times b}[\{1+r,\ldots,a-r\}\times\{1+r,\ldots,b-r\}]$.
 For $r\ge\min\{a,b\}/2$, $\interior_r(G_{a\times b})$ is the empty graph.
@@
 (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
-  \item Replace each degree-$4$ vertex $v:=(i,j)$ of $G_{a\times b}$ with a non-empty path $P_v$.
-If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each element of $N_{G_{a\times b}}(v)$ and this has no effect on the underlying graph.
-Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $N_{G_{a\times b}}(v)$.
-It is useful to consider three possible cases, each of which appears at least once in \cref{pseudogrid_fig}:
+  \item Replace each degree-$4$ vertex $(i,j)$ of $G_{a\times b}$ with a non-empty path $P_v$.
+If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each of $(i\pm 1,j\pm 1)$ and this has no effect on the underlying graph.
+Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $(i\pm 1,j\pm 1)$.
+It is useful to consider three possible cases:
   \begin{compactenum}[(Q1)]
@@
 Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path $C$  in $G$ that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$.
-This correspondence allows us to talk about the \defin{rows} and \defin{columns} of $G$, which we will do immediately.
+This correspondence allows us to talk about the rows and columns of $G$, which we will do immediately.
 
@@
 We will prove \cref{pseudogrid_lower_bound} by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
-We will prove the existence of $P$ in several steps; see \cref{outline}:
+We will prove the existence of $P$ in several steps (see \cref{outline}):
 
@@
   % This makes it easy to find a single path in $G$ that contains each of these vertices.
-  Unfortunately, this process can fail for some subset of the colours that appear in $G$.
+  Unfortunately, this process will fail for some subset of the colours that appear in $G$.
 
@@
 In \cref{packing_lemma_sec} we introduce a fairly standard-looking packing lemma that is needed in several places later.
-In \cref{well_separated} we explain how to find a well-separated set $S$ that contains two vertices of each colour.
-Finally, \cref{wrapping_up} shows how to assemble these various pieces to prove \cref{pseudogrid_lower_bound}.
+In \cref{well_separated} we explain how find a well-separated set $S$ that contains two vertices of each colour.
+Finally \cref{wrapping_up} shows how to assemble these various pieces to prove \cref{pseudogrid_lower_bound}.
 
@@
 
-For the sake of compactness, let $G_k:=G_{k\times k}$.
+For convenience, let $G_k:=G_{k\times k}$.
 Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=\{P_\mu:\mu\in\VE(G_{k})\}$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.
@@
 For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-1}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-1}(\alpha)$.
-Throughout this section, we will use the idiom $\varphi(V(G))$ to denote the set of all colours used by $\varphi$ to colour the vertices of $G$.
+(Throughout this section, we will use the idiom $\varphi(V(G))$ to denote the set of all colours used by $\varphi$ to colour the vertices of $G$.)
 
@@
 \begin{lem}\label{pick_up_two}
-  Let $G$ be a $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.
+  Let $G$ be a $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in\interior_1(G)$, and $t$ in column $a$ of $G$.
 Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
@@
   Refer to \cref{make_disjoint_fig}.
-We will define the set $X$ so that it satisfies (\ref{covers_s}) and for each $x\in X$ we will choose one or two elements of $\tilde{B}_{p}(x)\cap S$ and say that $x$ \defin{covers} those elements.
+We will define the set $X$ so that it satisfies (\ref{covers_s}) and for each $x\in X$ we will choose $1$ or $2$ elements of $\tilde{B}_{p}(x)\cap S$ and say that $x$ \defin{covers} those elements.
 Let $S_1$ be the subset of $S$ containing only those elements that do not take part in any $(r-1)$-pair.
@@
   To see that $X$ satisifies (\ref{disjoint_boxes}), observe that if $x\in X$ covers $v\in S$ then $\tilde{B}_{p+1}(x)\subseteq \tilde{B}_{2p+2}(v)$.
-Therefore, if $x\in X$ covers $v\in S$,  $y\in X$ covers $w\in S$, and $\tilde{B}_{p+1}(x)\cap \tilde{B}_{p+1}(y)\neq\emptyset$ then $w\in \tilde{B}_{4p+4}(v)$ or $v\in \tilde{B}_{4p+4}(w)$.
+Therefore, if $x\in X$ covers $v\in S$ and $y\in X$ covers $w\in S$ and $\tilde{B}_{p+1}(x)\cap \tilde{B}_{p+1}(y)\neq\emptyset$ then $w\in \tilde{B}_{4p+4}(v)$ or $v\in \tilde{B}_{4p+4}(w)$.
 Since $r-1 \ge 4p+4$, this implies that $v=w$ or $v$ and $w$ are an $(r-1)$-pair.
@@
 For each $x\in X$, this path intersects exactly one row of $G[\tilde{B}_{p+1}(x)]$.
-For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap \tilde{B}_{p+1}(x)$ with a path that is contained in $\tilde{B}_{p+1}(x)$ and contains the (at most two) vertices of $\tilde{B}_{p+1}(x)\cap S$.
+For each $x\in X$ we use \cref{pick_up_two} to replace $V(P_0)\cap B_{p+1}(x)$ with a path that is contained in $\tilde{B}_{p+1}(x)$ and contains the (at most two) vertices of $\tilde{B}_{p+1}(x)\cap S$.
 After doing this for each $x\in X$ we obtain a path $P$ that contains every vertex in $S$.
@@
 At the end of this process, the set $Q_1$ certainly satisfies (\ref{spread_out}).
-In fact, it satisfies an even stronger property: $B_{2r+1}(\mu)\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$, which implies that $|B_r(\nu)\cap Q_1|\le 1$ for each $\nu\in\VE(G_k)$.
+In fact, it implies an even stronger property: $B_{2r+1}(\mu)\cap Q_1=\{\mu\}$ for each $\mu\in Q_1$, which implies that $|B_r(\nu)\cap Q_1|\le 1$ for each $\nu\in\VE(G_k)$.
 However, $Q_1$ only satisfies (\ref{hits_both}) for the colours in $\overline{X}$.
@@
 This is already enough to establish that $\E(\deg_H(\alpha)) \in \Omega(|\Phi_\alpha|/r^2)$, but in order to obtain a sufficiently strong concentration result for $\deg_H(\alpha)$ we need to find some independence.\footnote{There are many ways to obtain a concentration result here and we use a simple method with the goal of being self-contained.
-For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+1}(\mu))\cap J$ contains at least one element $\nu$ and then setting $J:=J\setminus \VE(G_{7r}(\nu))$.\footnote{The constant $7$ is overkill here and is only used for simplicity; $6r+O(1)$ would be sufficient.}  This process continues until $J$ is empty.
+For a tighter result, use the method of bounded differences \cite{mcdiarmid:on}.}  To do this, set $J:=\Phi_\alpha$ and consider the greedy process of repeatedly choosing some $\mu\in Y$ such that $\VE(G_{2r+1}(\mu))\cap J$ contains at least one element $\nu$ and then set $J:=J\setminus \VE(G_{7r}(\nu))$.\footnote{The constant $7$ is overkill here and is only used for simplicity; $6r+O(1)$ would be sufficient.}  This process continues until $J$ is empty.
 Since $|\VE(G_{7r}(\mu))|\le \vol{7r}$, each iteration in this process eliminates at most $\vol{7r}$ elements from $J$ so the number, $t$, of iterations is at least $\ceil{|\Phi_\alpha|/\vol{7r}}$.
@@
 The important observation now is that the sets $Y_{\nu_1},\ldots,Y_{\nu_t}$ are pairwise disjoint.
-Indeed, $Y_{\nu_i}\subseteq B_{3r+2}(\nu_i)$, $Y_{\nu_j}\subseteq B_{3r+2}(\nu_j)$ and the fact that $\nu_i\not\in B_{7r}(\nu_j)$ implies that $\VE(G_{3r+2}(\nu_i))\cap \VE(G_{3r+2}(\nu_j))=\emptyset$.
+Indeed, $Y_{\nu_i}\subseteq B_{3r+2}(\nu_i)$, $Y_{\nu_j}\subseteq B_{3r+2}(\nu_j)$ and the fact that $\nu_i\not\in B_{7r}(\nu_j)$ implies that $G_{3r+2}(\nu_i)\cap G_{3r+2}(\nu_j)=\emptyset$.
 Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears in $\pi$ before any other element of $Y_{\nu_i}$'' then the events $U_1,\ldots,U_t$ are mutually independent.
@@
 Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
-We are also grateful to Omer Angel for pointing out that the problem on $k\times k$ grid (as opposed to pseudogrid) has a simple solution.
 

### 2022-07-21 lin-vs-cen.tex
Pffft

--- 2022-05-30 lin-vs-cen.tex
@@
 \title{\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}}
-\author{Prosenjit~Bose%
-  \thanks{School of Computer Science, Carleton University},\,
-  Vida~Dujmović%
-  \thanks{Department of Computer Science and Electrical Engineering, University of Ottawa},\,\,
-  Hussein~Houdrouge\footnotemark[2],\,
-  Mehrnoosh~Javarsineh\footnotemark[2],\,
-  and Pat~Morin\footnotemark[2]}
+\author{Prosenjit~Bose, Vida~Dujmović, Hussein~Houdrouge, Mehrnoosh~Javarsineh, and Pat~Morin}
 
@@
 \begin{abstract}
-  We prove that the linear chromatic number of any $k\times k$ pseudogrid is $\Omega(k)$.
-By an argument of Kun et al (Algorithmica, 2021), this result gives a tighter upper bound on the treedepth of a graph as a function of its linear chromatic number and gives further evidence in support of their conjecture that the treedepth of any graph is upper bounded by a linear function of its linear chromatic number.
+  A \emph{centred colouring} of a graph is a vertex colouring in which every connected subgraph contains a vertex whose colour is unique and a \emph{linear colouring} is a vertex colouring in which every path contains a vertex whose colour is unique.
+For a graph $G$ the \emph{centred chromatic number} $\chicen(G)$ and the \emph{linear chromatic number} $\chilin(G)$ denote the minimum number distinct colours required for a centred, respectively, linear colouring of $G$.
+From these definitions, it follows immediately that $\chilin(G)\le \chicen(G)$ for every graph $G$.
+The centred chromatic number is equivalent to treedepth and has been studied extensively.
+Much less is known about linear colouring.
+Kun \etal\ [\textit{Algorithmica} \textbf{83}(1)] prove that $\chicen(G) \le \tilde{O}(\chilin(G)^{190})$ for any graph $G$ and conjecture that $\chicen(G)\le 2\chilin(G)$.
+Their upper bound was subsequently improved by Czerwinski \etal\ [\textit{SIDMA} \textbf{35}(2)] to $\chicen(G)\le\tilde{O}(\chilin(G)^{19})$.
+The proof of both upper bounds relies on establishing a lower bound on the linear coloring number of pseudogrids, which appear in the proof due to their critical relationship to treewidth.
+Specifically, Kun \etal\ prove that $k\times k$ pseudogrids have linear colouring $\Omega(\sqrt{k})$.
+Our main contribution is establishing a tight bound on the linear coloring number of pseudogrids, specifically $\chilin(G)\ge \Omega(k)$ for every $k\times k$ pseudogrid $G$.
+As a consequence we improve the general bound for all graphs to $\chicen(G)\le \tilde{O}(\chilin(G)^{9})$.
+In addition this tight bound gives further evidence in support of Kun \etal's conjecture (above) that the centred chromatic number (i.e.
+treedepth) of any graph is upper bounded by a linear function of its linear chromatic number.
 \end{abstract}
@@
 A colouring $\varphi$ of $G$ is \defin{linear} if every simple path in $G$ has a centre with respect to $\varphi$.
+The \defin{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring and the \defin{linear chromatic number} $\chilin(G)$ is the minimum integer $c$ such $G$ has a linear $c$-colouring.
 
-The \defin{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring.
-An easy argument shows that centred chromatic number and treedepth are equivalent, i.e., $\td(G)=\chicen(G)$ for every graph $G$ \cite{nesetril.ossona:sparsity}.
-The \defin{linear chromatic number} of $G$ is the minimum integer $\ell$ such that $G$ has a linear $\ell$-colouring.
+The centered chromatic number of any graph $G$ is equal to the treedepth of $G$ and has been studied extensively \cite{nesetril.ossona:sparsity,X,Y}.
+Linear chromatic number was introduced by \citet{kun.obrien.ea:polynomial} who were motivated by finding efficiently-computable approximations of treedepth in bounded expansion classes.
+Much less is known about linear chromatic number.
 Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.
-The fundamental problem on linear colouring is the question of upper bounding $\chicen(G)$ in terms of $\chilin(G)$.
-
-Motivated by finding efficiently-computable approximations of treedepth in bounded expansion classes, linear colourings were introduced by \citet{kun.obrien.ea:polynomial} who were able to establish a polynomial upper bound on $\chicen(G)$ as a function of $\chilin(G)$.
-By combining their arguments with recent improvements by \citet{czerwinski.nadara.ea:improved} on the relationship between treedepth and treewidth, the following theorem is established:
+In the other direction, \citet{kun.obrien.ea:polynomial} were able to establish that $\chicen(G)\le \chilin(G)^{190}\log^{O(1)}\chilin(G)$.
+This upper bounded was subsequently improved by \citet{czerwinski.nadara.ea:improved}, who reduced the exponent to $19$.
 
@@
 
-In the other direction, \citet{kun.obrien.ea:polynomial} show that, for any $\epsilon >0$, there exists a graph $G$ with $\chicen(G) \ge (2-\epsilon)\chilin(G)$.
-The same authors make the following bold conjecture:
+\citet{kun.obrien.ea:polynomial} give a family of graphs that contains, for every $\epsilon > 0$, a graph $G$ with $\chicen(G)\ge (2-\epsilon)\chilin(G)$. They conjecture that this bound is tight:
 
-\begin{conj}[\citet{kun.obrien.ea:polynomial}]\label{crazy_conjecture}
-  For any graph $G$, $\chicen(G)\le 2\chilin(G)$.
+\begin{conj}[\cite{kun.obrien.ea:polynomial}]\label{crazy_conjecture}
+  For every graph $G$, $\chicen(G)\le 2\chilin(G)$.
 \end{conj}
 
-Currently, the only class of graphs for which centred chromatic number is known to be bounded by a linear function of linear chromatic number is the class of bounded degree trees \cite[Theorem~4]{kun.obrien.ea:polynomial}.
-To place the current contribution into context it is useful to understand the proof of \cref{kun_obrien_general}, which has three main ingredients:
+This is a very bold conjecture since until now the only class of graphs for which a linear bound is known is the class of bounded degree trees.
+Specifically, for any tree $T$ of maximum-degree $\Delta\ge 3$, $\chicen(T) \le (\log_2(\Delta))\chilin(G)$ \cite[Theorem~4]{kun.obrien.ea:polynomial}.
+
+To put our results into context, we summarize the proof of \cref{kun_obrien_general}:
+
 \begin{enumerate}
@@
 This is just an application of their result with $a=k^{18}\log^q k$ and $b=k$.}  In the former case, an asymptotically optimal result of \citet{kun.obrien.ea:polynomial} on subcubic trees completes the proof, so we are left with the case where $\tw(G)\in\Omega(k^{18}\log^q k)$.
+
   \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy.tan:towards} shows that, if $\tw(G)\in\Omega(k^{18}\log^q k)$ (for a particular fixed positive $q$), then $G$ contains an $\Omega(k^2)\times \Omega(k^2)$ grid minor.
+
   \item A technical lemma \cite[Lemma~5]{kun.obrien.ea:polynomial} shows that, if $G$ contains a $\Omega(k^2)\times \Omega(k^2)$ grid minor, then $\chilin(G)\in\Omega(k)$.
@@
 
-In the current work, we spend considerable effort to improve the technical lemma in Point~3 to show that a $\Omega(k)\times\Omega(k)$ grid minor is sufficient to guarantee that $\chilin(G)\in\Omega(k)$.
+In the current work, we spend considerable effort to improve the technical lemma in Point~3 and prove the following:
+
+\begin{lem}\label{pseudogrid_lower_bound}
+  For any $k\times k$ pseudogrid $G$, $\chilin(G)\in\Omega(k)$.
+\end{lem}
+
 This improves the exponent in \cref{kun_obrien_general} from $19$ to $10$, yielding the following improvement to \cref{kun_obrien_general}:
+
 \begin{thm}\label{kun_obrien_general2}
@@
 
-In addition to the (admittedly modest) improvement on the exponent in \cref{kun_obrien_general}, the current result adds further evidence in support of \cref{crazy_conjecture} by establishing that, when $G$ is a $k\times k$ pseudogrid, then $\chicen(G)\in\Theta(\chilin(G))$.
+Any further improvement to \cref{kun_obrien_general2} will either require an improved Excluded Grid Theorem or an entirely new approach.
+
+In addition to the improvement on the exponent in \cref{kun_obrien_general}, \cref{pseudogrid_lower_bound} adds further evidence in support of \cref{crazy_conjecture} by establishing that, when $G$ is a $k\times k$ pseudogrid, then $\chicen(G)\in\Theta(\chilin(G))$.
 
@@
 If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph \cite{kun.obrien.ea:polynomial}.
-Therefore to prove \cref{kun_obrien_general2}, it suffices to establish:
-
-\begin{lem}\label{pseudogrid_lower_bound}
-  For any $k\times k$ pseudogrid $G$, $\chilin(G)\in\Omega(k)$.
-\end{lem}
-
-We will prove \cref{pseudogrid_lower_bound} by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
-We will prove the existence of $P$ in several steps; see \cref{outline}:
+Therefore to prove \cref{kun_obrien_general2}, it suffices to establish \cref{pseudogrid_lower_bound}.
+We do this by showing that, for sufficiently small $\epsilon >0$, any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$.
+We prove the existence of $P$ in several steps; see \cref{outline}:
 
@@
   \item We greedily choose a set $S_1$ of vertices in $G$ that contains two vertices of each colour and that is `well-separated' in the sense that the corresponding set of vertices/edges in $G_{k\times k}$ have a minimum distance between them.
-In \cref{outline} this minimum distance corresponds to the fact that the box drawn centered at each vertex in $S_1$ contains no other vertices of $S_1$.
+In \cref{outline} this minimum distance corresponds to the fact that the box drawn centred at each vertex in $S_1$ contains no other vertices of $S_1$.
   % This makes it easy to find a single path in $G$ that contains each of these vertices.

### 2022-08-15 lin-vs-cen.tex
Addressed Vida's comments

@@
 Their upper bound was subsequently improved by Czerwinski \etal\ [\textit{SIDMA} \textbf{35}(2)] to $\chicen(G)\le\tilde{O}(\chilin(G)^{19})$.
-The proof of both upper bounds relies on establishing a lower bound on the linear coloring number of pseudogrids, which appear in the proof due to their critical relationship to treewidth.
-Specifically, Kun \etal\ prove that $k\times k$ pseudogrids have linear colouring $\Omega(\sqrt{k})$.
-Our main contribution is establishing a tight bound on the linear coloring number of pseudogrids, specifically $\chilin(G)\ge \Omega(k)$ for every $k\times k$ pseudogrid $G$.
+The proof of both upper bounds relies on establishing a lower bound on the linear chromatic number of pseudogrids, which appear in the proof due to their critical relationship to treewidth.
+Specifically, Kun \etal\ prove that $k\times k$ pseudogrids have linear chromatic number $\Omega(\sqrt{k})$.
+Our main contribution is establishing a tight bound on the linear chromatic number of pseudogrids, specifically $\chilin(G)\ge \Omega(k)$ for every $k\times k$ pseudogrid $G$.
 As a consequence we improve the general bound for all graphs to $\chicen(G)\le \tilde{O}(\chilin(G)^{9})$.
@@
 
-The centered chromatic number of any graph $G$ is equal to the treedepth of $G$ and has been studied extensively \cite{nesetril.ossona:sparsity,X,Y}.
+The centered chromatic number of any graph $G$ is equal to the treedepth of $G$ and has been studied extensively \cite{nesetril.ossona:sparsity,nesetril.ossona:on,nesetril.ossona:grad,nesetril.ossona:tree-depth,pilipczuk.siebertz:polynomial,dereniowski.nadolski:vertex,dereniowski.kubale:cholesky,deogun.kloks.ea:on,bodlaender.gilbert.ea:approximating,bodlaender.deogun.ea:rankings}.
+Much less is known about linear chromatic number.
 Linear chromatic number was introduced by \citet{kun.obrien.ea:polynomial} who were motivated by finding efficiently-computable approximations of treedepth in bounded expansion classes.
-Much less is known about linear chromatic number.
 Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.
@@
 
-To put our results into context, we summarize the proof of \cref{kun_obrien_general}:
+To prove \cref{kun_obrien_general}, \citet{kun.obrien.ea:polynomial} establish the critical role that lower bounds on the linear linear chromatic number of pseudogrids and subcubic trees play in establishing an upper bound on $\chicen(G)$ as a function of $\chilin(G)$.
+In their work, they establish (asymptotically) tight lower bounds for the linear chromatic number of subcubic trees, but their lower bounds for pseudogrids are not tight.
+
+With the goal of making progress on \cref{crazy_conjecture}, our objective in this paper is to establish a tight lower bound on the linear chromatic number of pseudogrids.
+To put our results into context and to be more precise, we first summarize the proof of \cref{kun_obrien_general}:
 
@@
 
-  \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy.tan:towards} shows that, if $\tw(G)\in\Omega(k^{18}\log^q k)$ (for a particular fixed positive $q$), then $G$ contains an $\Omega(k^2)\times \Omega(k^2)$ grid minor.
+  \item The current-best version of the Excluded Grid Theorem due to \citet{chuzhoy.tan:towards} shows that, if $\tw(G)\in\Omega(k^{18}\log^q k)$ (for a particular fixed positive $q$), then $G$ contains an $\Omega(k^2)\times \Omega(k^2)$ grid minor (equivalently, $G$ contains a $\Omega(k^2)\times\Omega(k^2)$ pseudogrid as a subgraph).
 
-  \item A technical lemma \cite[Lemma~5]{kun.obrien.ea:polynomial} shows that, if $G$ contains a $\Omega(k^2)\times \Omega(k^2)$ grid minor, then $\chilin(G)\in\Omega(k)$.
+  \item Points~1 and 2 demonstrate that in order to establish \cref{kun_obrien_general}, a lower bound on the linear chromatic number of a $\Omega(k^2)\times\Omega(k^2)$ grid minor (i.e., pseudogrid) is needed.
+\citet[Lemma~5]{kun.obrien.ea:polynomial} establish such a lower bound.
+Specifically, they show that, if $G$ contains a $\Omega(k^2)\times \Omega(k^2)$ grid minor, then $\chilin(G)\in\Omega(k)$.
+  In the current work, we spend considerable effort to prove the following tight bound:
 \end{enumerate}
-
-In the current work, we spend considerable effort to improve the technical lemma in Point~3 and prove the following:
 
@@
 
+In addition to the improvement on the exponent in \cref{kun_obrien_general}, \cref{pseudogrid_lower_bound} adds further evidence in support of \cref{crazy_conjecture} by establishing that, when $G$ is a $k\times k$ pseudogrid, then $\chicen(G)\in\Theta(\chilin(G))$.
+
 Any further improvement to \cref{kun_obrien_general2} will either require an improved Excluded Grid Theorem or an entirely new approach.
-
-In addition to the improvement on the exponent in \cref{kun_obrien_general}, \cref{pseudogrid_lower_bound} adds further evidence in support of \cref{crazy_conjecture} by establishing that, when $G$ is a $k\times k$ pseudogrid, then $\chicen(G)\in\Theta(\chilin(G))$.
 

### 2022-08-18 lin-vs-cen.tex
Vida's comments
Small typo fix

@@
 
-To prove \cref{kun_obrien_general}, \citet{kun.obrien.ea:polynomial} establish the critical role that lower bounds on the linear linear chromatic number of pseudogrids and subcubic trees play in establishing an upper bound on $\chicen(G)$ as a function of $\chilin(G)$.
+To prove \cref{kun_obrien_general}, \citet{kun.obrien.ea:polynomial} establish the critical role that lower bounds on the linear chromatic number of pseudogrids and subcubic trees play in establishing an upper bound on $\chicen(G)$ as a function of $\chilin(G)$.
 In their work, they establish (asymptotically) tight lower bounds for the linear chromatic number of subcubic trees, but their lower bounds for pseudogrids are not tight.
@@
 \citet[Lemma~5]{kun.obrien.ea:polynomial} establish such a lower bound.
-Specifically, they show that, if $G$ contains a $\Omega(k^2)\times \Omega(k^2)$ grid minor, then $\chilin(G)\in\Omega(k)$.
+Specifically, they show that, for any $\Omega(k^2)\times \Omega(k^2)$ pseudogrid $G$, $\chilin(G)\in\Omega(k)$.
   In the current work, we spend considerable effort to prove the following tight bound:

### 2024-04-03 lin-vs-cen.tex
Pffft
Fixed calculation error pointed out by referee.

@@
     & = \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|-1} \notag \\
-    & \ge \exp\left(-\ln\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
-    & \ge \exp\left(-\ln\tau - \vol{7r}\cdot|\Phi_\alpha|/(\tau-1)\right)
+    & > \tfrac{1}{\tau}\left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|}
+      & \text{(since $1-1/\tau<1$)}\notag \\
+    % & = \exp(-\ln\tau)\cdot \left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|} \notag \\
+    % & = \exp(-\ln\tau)\cdot \left(1-\tfrac{1}{\tau}\right)^{|\Gamma_\alpha|} \notag \\
+    & > \exp(-\ln\tau-|\Gamma_\alpha|/\tau)
+      & \text{(since $e^{-1/\tau}> 1-1/\tau$)}\notag \\
+    % & \ge \exp\left(-\ln\tau - |\Gamma_\alpha|/(\tau-1)\right) & \text{(Since $(1-1/\tau)^{\tau-1} > 1/e$)} \notag \\
+    & \ge \exp\left(-\ln\tau - \vol{7r}\cdot|\Phi_\alpha|/\tau\right)
       & \text{(since $|\Gamma_\alpha|\le \vol{7r}\cdot\Phi_\alpha$).}  \label{weight_product}
@@
   Comparing \cref{weight_product,probability} we see that both quantities decrease exponentially in $|\Phi_\alpha|$ but \cref{weight_product} does so at a rate that can be controlled with $\tau$.
-  Taking $\tau = {32\vol{7r}^2}+1$, and using \cref{weight_product,probability}, we find that
+  Taking $\tau = {32\vol{7r}^2}$, and using \cref{weight_product,probability}, we find that
   \[
@@
 Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
-We are also grateful to Omer Angel for pointing out that the problem on $k\times k$ grid (as opposed to pseudogrid) has a simple solution.
+We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution.
 

### 2024-04-04 referee-response.md
Responding to referee

The referee had 4 broad comments:

1. The paper's title is a bit ambitious since, as the authors point out, the machinery for relating the linear chromatic number to the centred one was already introduced previously (and this manuscript does not develop on it at all). A more honest title would be something along the lines of "The linear chromatic number of pseudogrids".  
We have chosen to leave the title as is.  In our opinion, the current title doesn't promise much.  The proposed title, on the other hand, suggests that the results apply only to pseudogrids.  Consider, for example, the similar-sounding title "The linear chromatic number of cactus graphs."  
With the current title a reader will still learn, within the first two pages that the improvement comes only by studying the linear chromatic number pseudogrids and decide for themselves whether it's worthwhile reading the rest of the argument.  (These first two pages also have the benefit of catching-up a reader on the current state of the art and the limitations of Kun et al's argument.)

2. There are a few places (e.g. in section 3.6 -- see PDF) where the authors cut a few corners in their proofs; this should be addressed, but I don't think it is anything serious.

We have gone through the attached PDF carefully and corrected the minor issues pointed out. Some larger changes in response to the referee's comments are discussed below.

3. A lot of the arguments in the middle sections, such as the packing lemma, seem to use the underlying intuition that grids (and by extension pseudogrids) look like the plane with the taxicab metric and the Lebesgue measure. I imagine the authors have already attempted this, but is there really no way to make that intuition rigorous by defining a metric and/or a measure on appropriate sets, and express the lemmas in terms of those? Something like that would make arguments such as the one in Lemma 13 much cleaner. In this direction, have the authors considered the idea of using a grid with every edge subdivided once as the index set for the path partition of the pseudogrid (the subdivision vertices symbolising the edges), with the purpose of making definitions such as the r-boxes more uniform (and of having a symmetric relation, whereby if \mu is in the r-box of \nu, then \nu should be in the r-box of \mu)?

During the initial writing phase of this result, we tried several such strategies.  The main annoying issue is that mapping the vertices of a subdivided grid onto objects in the pseudogrid sometimes maps a vertex onto a single edge (but not the endpoints of that edge).  In terms of colouring, this means that the subdivision vertex has an empty colour set.  It is possible to define things carefully enough for this to work, but then the reader is constantly having to verify that simple-sounding statements are correct under these careful definitions.

4. In the acknowledgements, the authors mention that it was pointed out to them the problem on a grid (as opposed to pseudogrid) has a simple solution. Assuming the person who pointed that out is happy with it, it might be interesting to include the solution.

TODO

---

**Referee Comment:** Bottom of Page 2: A tiny bit misleading (since as far as I can tell, the bottleneck for getting conjecture 2 is the grid theorem, not this).

Top of Page 3: This somewhat addresses the previous comment.

**Response:** We have added further explanation of the limits of this argument here. In particular, even with an Excluded Grid Theorem that matches the currently known lower bound, this would only establish that the centred chromatic number is at most cubic in the linear chromatic number.  Even for planar graphs, the argument can only show that the centred chromatic number is at most quadratic in the linear chromatic number.

**Referee Comment:** I do not understand how this works when removing edges in $P_{v_i}$ -- doesn't $P_{v_i}$ belong to both a row and a column (and thus deleting it leaves us with both fewer rows and columns?)

**Response:** The referee is correct.  The description of deletion has been corrected.  Now we only delete edges of $G$ that correspond to edges of the grid (and then repeatedly remove vertices of degree at most $1$).

**Referee Comment:** $d$ has not been defined. The last part of the proof (where it is shown that a certain value of $d$ would work) should come first -- or at the very least, $d$ should be chosen bigger than some bound given here (and the last part of the proof would show this bound works). This is important, since I am having some trouble keeping track of the dependencies between the variables, so that should be made more transparent.

**Response:** The variable $d$ is defined in the statement of the lemma ("there exists a value $d\in O(r^4\log r)$ such that....").  In the first line of the proof, we use this value of $d$ to apply Lemma 9 to obtain an assignment of colours of objects so that each colour is assigned to at least $d$ objects.  We have expanded the first line of the proof to point this out, so that it is familiar when it is mentioned again at this point in the proof.

**Referee Comment:** From what I can tell, we are saying that if there are enough elements of $\phi^{-1}(\alpha)$ far from the elements of $Q_1$, there will be two such elements far from each other. I agree with this in principle, but where does the actual bound given here come from? The intuition, I presume, is that a set of that size is "maximal" of diameter at most $2r + 1$, but how is that made rigorous? Presumably some argument involving convexity,  and it would be good to see it written.

**Response:** It is much simpler than this.  If more than $\boxplus_{r+1}$ objects of $\phi^{-1}(\alpha)$ are left uncovered by the radius $2r+1$ boxes around the objects in $Y$ then this was also the case when the colour $\alpha$ was considered in the first round.  At that point, we could have chosen any of these uncovered objects $\mu_1$ of colour $\alpha$.  Then $B_{2r+1}(\mu_1)$ contains at most $\boxplus_r$ objects (by definition), so there is still another uncovered object $\mu_2$ of colour $\alpha$.  In this case, we would have put $\mu_1$ and $\mu_2$ into $Q_1$ and round 1 would have succeeded for $\alpha$.

### 2024-04-04 lin-vs-cen.tex
Fixed dumb typo in abstract
Responding to referee

@@
 \begin{abstract}
-  A \emph{centred colouring} of a graph is a vertex colouring in which every connected subgraph contains a vertex whose colour is unique and a \emph{linear colouring} is a vertex colouring in which every path contains a vertex whose colour is unique.
-For a graph $G$ the \emph{centred chromatic number} $\chicen(G)$ and the \emph{linear chromatic number} $\chilin(G)$ denote the minimum number distinct colours required for a centred, respectively, linear colouring of $G$.
+  A \emph{centred colouring} of a graph is a vertex colouring in which every connected subgraph contains a vertex whose colour is unique and a \emph{linear colouring} is a vertex colouring in which every (not-necessarily induced) path contains a vertex whose colour is unique.
+For a graph $G$, the \emph{centred chromatic number} $\chicen(G)$ and the \emph{linear chromatic number} $\chilin(G)$ denote the minimum number of distinct colours required for a centred, respectively, linear colouring of $G$.
 From these definitions, it follows immediately that $\chilin(G)\le \chicen(G)$ for every graph $G$.
@@
 Our main contribution is establishing a tight bound on the linear chromatic number of pseudogrids, specifically $\chilin(G)\ge \Omega(k)$ for every $k\times k$ pseudogrid $G$.
-As a consequence we improve the general bound for all graphs to $\chicen(G)\le \tilde{O}(\chilin(G)^{9})$.
-In addition this tight bound gives further evidence in support of Kun \etal's conjecture (above) that the centred chromatic number (i.e.
-treedepth) of any graph is upper bounded by a linear function of its linear chromatic number.
+As a consequence we improve the general bound for all graphs to $\chicen(G)\le \tilde{O}(\chilin(G)^{10})$.
+In addition this tight bound gives further evidence in support of Kun \etal's conjecture (above) that the centred chromatic number (i.e., the treedepth) of any graph is upper bounded by a linear function of its linear chromatic number.
 \end{abstract}
@@
 A colouring $\varphi$ of $G$ is \defin{centred} if every connected subgraph of $G$ has a centre with respect to $\varphi$.
-A colouring $\varphi$ of $G$ is \defin{linear} if every simple path in $G$ has a centre with respect to $\varphi$.
+A colouring $\varphi$ of $G$ is \defin{linear} if every path\footnote{A \defin{path} in a graph $G$ is a sequence of distinct vertices $v_0,\ldots,v_r$ such that $v_{i-1}v_i$ is an edge of $G$ for each $i\in\{1,\ldots,r\}$.} in $G$ has a centre with respect to $\varphi$.
 The \defin{centred chromatic number} $\chicen(G)$ of $G$ is the minimum integer $c$ such that $G$ has a centred $c$-colouring and the \defin{linear chromatic number} $\chilin(G)$ is the minimum integer $c$ such $G$ has a linear $c$-colouring.
+Since each edge of $G$ is both a connected subgraph and a path in $G$, any centred colouring and any linear colouring is a proper colouring of $G$.
 
 The centered chromatic number of any graph $G$ is equal to the treedepth of $G$ and has been studied extensively \cite{nesetril.ossona:sparsity,nesetril.ossona:on,nesetril.ossona:grad,nesetril.ossona:tree-depth,pilipczuk.siebertz:polynomial,dereniowski.nadolski:vertex,dereniowski.kubale:cholesky,deogun.kloks.ea:on,bodlaender.gilbert.ea:approximating,bodlaender.deogun.ea:rankings}.
-Much less is known about linear chromatic number.
+Much less is known about the linear chromatic number.
 Linear chromatic number was introduced by \citet{kun.obrien.ea:polynomial} who were motivated by finding efficiently-computable approximations of treedepth in bounded expansion classes.
 Since every path in $G$ is a connected subgraph of $G$, every centred colouring of $G$ is also a linear colouring of $G$, so $\chilin(G)\le\chicen(G)$.
-In the other direction, \citet{kun.obrien.ea:polynomial} were able to establish that $\chicen(G)\le \chilin(G)^{190}\log^{O(1)}\chilin(G)$.
-This upper bounded was subsequently improved by \citet{czerwinski.nadara.ea:improved}, who reduced the exponent to $19$.
+In the other direction, \citet{kun.obrien.ea:polynomial} were able to establish that $\chicen(G)\le \chilin(G)^{190}\cdot(\log(\chilin(G)))^{O(1)}$.
+This upper bound was subsequently improved by \citet{czerwinski.nadara.ea:improved}, who reduced the exponent to $19$.
 
@@
 
-With the goal of making progress on \cref{crazy_conjecture}, our objective in this paper is to establish a tight lower bound on the linear chromatic number of pseudogrids.
+With the goal of better understanding the difficulty of \cref{crazy_conjecture}, our objective in this paper is to establish a tight lower bound on the linear chromatic number of pseudogrids.
 To put our results into context and to be more precise, we first summarize the proof of \cref{kun_obrien_general}:
@@
 
-In addition to the improvement on the exponent in \cref{kun_obrien_general}, \cref{pseudogrid_lower_bound} adds further evidence in support of \cref{crazy_conjecture} by establishing that, when $G$ is a $k\times k$ pseudogrid, then $\chicen(G)\in\Theta(\chilin(G))$.
+In addition to the improvement on the exponent in \cref{kun_obrien_general}, \cref{pseudogrid_lower_bound} adds further evidence in support of \cref{crazy_conjecture} by establishing that, when $G$ is a $k\times k$ pseudogrid, $\chicen(G)\in\Theta(\chilin(G))$.
 
 Any further improvement to \cref{kun_obrien_general2} will either require an improved Excluded Grid Theorem or an entirely new approach.
+However, no improvement to the Excluded Grid Theorem will sufficient to establish a linear relationship between the centred and the linear chromatic number.
+Indeed, the best possible Excluded Grid Theorem would state that any graph of treewidth $k^2\log k$ contains an $\Omega(k)\times\Omega(k)$ grid minor, and the preceding argument would only show that $\chicen(G)\in O((\chilin(G))^3\log(\chilin(G)))$.
+Even the Excluded Grid Theorem for Planar Graphs states that any planar graph of treewidth $k$ contains an $\Omega(k)\times\Omega(k)$ grid minor.
+Combining this with the argument above and \cref{pseudogrid_lower_bound} shows only that, for any planar graph $G$, $\chicen(G)\in\Theta((\chilin(G))^2)$.
 
@@
 
-For positive integers $a$ and $b$, the \defin{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(i_1,j_1)$ and $(i_2,j_2)$ if and only if $|i_1-i_2|+|j_1-j_2|=1$.
+For positive integers $a$ and $b$, the \defin{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contains an edge with endpoints $(i_1,j_1)$ and $(i_2,j_2)$ if and only if $|i_1-i_2|+|j_1-j_2|=1$.
 Such an edge is \defin{vertical} if $i_1=i_2$ and \defin{horizontal} if $j_1=j_2$.
@@
   \end{center}
-  \caption{A $6\times 4$ pseudogrid.  Black vertices are contained in $P_v$ for some vertex $v$ of $G_{6\times 4}$ and white vertices are contained in $P_{vw}$ for some edge of $G_{6\times 4}$.}
+  \caption{A $6\times 4$ pseudogrid.
+Black vertices are contained in $P_v$ for some vertex $v$ of $G_{6\times 4}$ and white vertices are contained in $P_{vw}$ for some edge of $G_{6\times 4}$.
+Colour are used to distinguish between cases (Q1) (blue), (Q2) (pink), and (Q3) (gold).}
   \label{pseudogrid_fig}
@@
 \begin{compactitem}
-  \item Replace each edge $vw$ of $G_{a\times b}$ with a path $\overline{P}_{vw}$ whose endpoints are $v$ and $w$.
-(In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
-  \item Replace each degree-$4$ vertex $v:=(i,j)$ of $G_{a\times b}$ with a non-empty path $P_v$.
-If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each element of $N_{G_{a\times b}}(v)$ and this has no effect on the underlying graph.
-Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $N_{G_{a\times b}}(v)$.
+  \item Replace each degree-$4$ vertex $v$ whose neighbours (in counterclockwise order, starting with the neighbour $w$ above $v$) are $w,x,y,z$ with a non-empty path $P_v$.
+If $P_v$ has only one vertex then the unique vertex of $P_v$ is adjacent to each of $w,x,y,z$ and this has no effect on the underlying graph.
+Otherwise, $P_v$ has two endpoints $p$ and $q$, each of which is adjacent to two vertices among $w,x,y,z$.
 It is useful to consider three possible cases, each of which appears at least once in \cref{pseudogrid_fig}:
   \begin{compactenum}[(Q1)]
-    \item \label{q_i} $p$ is adjacent to $\{(i-1,j), (i,j-1)\}$ and $q$ is adjacent to $(i+1,j),(i,j+1)$;
-    \item \label{q_ii} $p$ is adjacent to $\{(i,j+1), (i-1,j)\}$ and $q$ is adjacent to $(i,j-1),(i+1,j)$; or
-    \item \label{q_iii} $p$ is adjacent to $\{(i-1,j),(i+1,j)\}$ and $q$ is adjacent to $\{(i,j-1),(i,j+1)\}$.
+    \item \label{q_i} $p$ is adjacent to $\{w,x\}$ and $q$ is adjacent to $\{y,z\}$;
+    \item \label{q_ii} $p$ is adjacent to $\{x,y\}$ and $q$ is adjacent to $\{w,z\}$; or
+    \item \label{q_iii} $p$ is adjacent to $\{w,y\}$ and $q$ is adjacent to $\{x,z\}$.
   \end{compactenum}
+  \item At this point, each edge $vw$ of $G_{a\times b}$ has a corresponding edge $v'w'$ in the modified graph, and we replace $v'w'$ with a path $\overline{P}_{vw}$ whose endpoints are $v'$ and $w'$.
+(In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $v'w'$ zero or more times.)
 \end{compactitem}
@@
 
-As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.
+As part of our proof, we use the operation of \defin{deleting} a row (or column) of $G$.
+To delete a row $R$ of $G$ that corresponds to the row $R':=v_1,\ldots,v_a$, we remove the edges of $\overline{P}_{v_{i}v_{i+1}}$ for each $i\in\{2,\ldots,a\}$.
 If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_r$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove vertices of degree at most $1$ until none remain.
@@
   Let $d,k,r\ge 1$ be integers, let $G$ be a $k\times k$ pseudogrid and let $\varphi$ be a vertex colouring of $G$ that uses $|\varphi(V(G))|\le k/(d+2r)$ colours.
-  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - (d+2r)|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k})\}$ such that
+  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k - (d+2r)|\varphi(V(G))|$ that has a grid-partition $\mathcal{P}':=\{V(P'_\mu):\mu\in \VE(G_{k'})\}$ such that
   for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-1}(A)\cap \interior_r(G_{k'})| \ge d|A|$.
@@
 If there exists no $A\subseteq \varphi(V(G))$ with $|\varphi_{\mathcal{P}}^{-1}(A)\cap\interior_r(G_k)| < d|A|$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma.
-Otherwise, we will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|+2r \le (d+2r)|A|$ to eliminate all vertices with colours in $A$, as follows:
+Otherwise, fix such a set $A$.
+We will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le d|A|+2r \le (d+2r)|A|$ to eliminate all vertices with colours in $A$, as follows:
   \begin{compactitem}
     \item For each vertex $v:=(i,j)\in V(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.
-    \item For each horizontal edge $vw\in E(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
-    \item For each vertical edge $vw\in E(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row of $G$ that contains $P_{vw}$ in $R$.
+    \item For each horizontal edge $vw\in E(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the row $R$ of $G$ that contains $P_{vw}$.
+    \item For each vertical edge $vw\in E(\interior_r(G_{k}))$ with $A\cap\varphi_\mathcal{P}(vw)\neq\emptyset$, we include the column $C$ of $G$ that contains $P_{vw}$.
     \item We add the first and last $r$ rows to $R$ and the first and last $r$ columns to $C$.
@@
   \begin{compactenum}[(i)]
-    \item $\phi(\mu)=\perp$ for each For each $\mu\not\in\VE(\interior_r(G_{k}))$;
+    \item $\phi(\mu)=\perp$ for each $\mu\not\in\VE(\interior_r(G_{k}))$;
     \item $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi_\mathcal{P}(\mu)$ for each $\mu\in\VE(G_{k})$; and
-    \item $|\phi^{-1}(\alpha)|\ge d$ for each $\mu\in\VE(G_{k})$.
+    \item $|\phi^{-1}(\alpha)|\ge d$ for each $\alpha\in\varphi(V(G))$.
+    % $\mu\in\VE(G_{k})$.
     % $|\{\mu\in\VE(G_{k}):\phi(\mu)=\alpha\}|\ge d$.
@@
 \begin{proof}
-  Consider the bipartite graph $H$ with parts $X:=\{1,\ldots,c\}$ and $Y:=\VE(\interior_r(G_k))$ and edge set
+  Consider the bipartite graph $H$ with parts $X:=\varphi(V(G))$ and $Y:=\VE(\interior_r(G_k))$ and edge set
   \[
@@
 \begin{lem}\label{pick_up_two}
-  Let $G$ be a $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.
+  Let $G$ be an $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.
 Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
@@
 \begin{proof}
-  We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi:\VE(G_k)\to\varphi(V(G))\cup\{\perp\}$.
+  We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi:\VE(G_k)\to\varphi(V(G))\cup\{\perp\}$ in which $\phi^{-1}(\alpha)\ge d$ for each $\alpha\in\varphi(V(G))$.
 Observe that it is now sufficient to find a $2|\varphi(V(G))|$-element subset $Q\subseteq\VE(\interior_r(G_k))$ such that
@@
   Define a bipartite graph $H$ with parts $L(H):=X$ and $R(H):=Y:=Q_1$.
-Let $\pi:=\mu_1,\ldots,\mu_{|Y|}$ be a random permutation of $Y$.\footnote{Here random permutation means a permutation chosen uniformly from the set of all $|Y|!$ permutations of $Y$.}  We include an edge $\alpha\mu_i$ in $H$ if and only if
+Let $\pi:=\mu_1,\ldots,\mu_{|Y|}$ be a random permutation of $Y$.\footnote{A \defin{random permutation} of $Y$ is a permutation chosen uniformly from the set of all $|Y|!$ permutations of $Y$.}  We include an edge $\alpha\mu_i$ in $H$ if and only if
   \[
@@
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates a $\binomial(\ceil{|\Phi_\alpha|/\vol{7r}},1/16)$ random variable, which allows us to establish en exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+1})$.
+  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.} a $\binomial(\ceil{|\Phi_\alpha|/\vol{7r}},1/16)$ random variable, which allows us to establish an exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+1})$.
 Suppose that, for some $\alpha\in X$ and $\mu\in Y$, $B_{2r+1}(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and consider the set
@@
 For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/16$.
-Therefore, $\deg_H(\alpha)\ge \sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/16)$ random variable.\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.}  Therefore,
+Therefore, $\deg_H(\alpha)\ge \sum_{i=1}^t \mathbbm{1}_{U_i}$ dominates a $\binomial(t,1/16)$ random variable.
+Therefore,
   \begin{align}
@@
 We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution.
+(By partitioning the rows into consecutive pairs and the columns  into consecutive pairs, and then repeatedly deleting a pair of rows and columns whose intersection contains a vertex whose colour is unique, one is left with a graph that has a Hamiltonian path in which each colour apppears on at least two vertices.)
 

### 2024-04-04 referee-response.md
More changes suggested by the referee.

@@
 
-1.
-The paper's title is a bit ambitious since, as the authors point out, the machinery for relating the linear chromatic number to the centred one was already introduced previously (and this manuscript does not develop on it at all).
+**Referee Comment 1:** The paper's title is a bit ambitious since, as the authors point out, the machinery for relating the linear chromatic number to the centred one was already introduced previously (and this manuscript does not develop on it at all).
 A more honest title would be something along the lines of "The linear chromatic number of pseudogrids".
-We have chosen to leave the title as is.
+
+**Response:** We have chosen to leave the title as is.
 In our opinion, the current title doesn't promise much.
@@
 Consider, for example, the similar-sounding title "The linear chromatic number of cactus graphs."
+
 With the current title a reader will still learn, within the first two pages that the improvement comes only by studying the linear chromatic number pseudogrids and decide for themselves whether it's worthwhile reading the rest of the argument.
@@
 
-2. There are a few places (e.g. in section 3.6 -- see PDF) where the authors cut a few corners in their proofs; this should be addressed, but I don't think it is anything serious.
+**Referee Comment 2:** There are a few places (e.g. in section 3.6 -- see PDF) where the authors cut a few corners in their proofs; this should be addressed, but I don't think it is anything serious.
 
@@
 
-3.
-A lot of the arguments in the middle sections, such as the packing lemma, seem to use the underlying intuition that grids (and by extension pseudogrids) look like the plane with the taxicab metric and the Lebesgue measure.
+**Referee Comment 3:** A lot of the arguments in the middle sections, such as the packing lemma, seem to use the underlying intuition that grids (and by extension pseudogrids) look like the plane with the taxicab metric and the Lebesgue measure.
 I imagine the authors have already attempted this, but is there really no way to make that intuition rigorous by defining a metric and/or a measure on appropriate sets, and express the lemmas in terms of those?
@@
 During the initial writing phase of this result, we tried several such strategies.
-The main annoying issue is that mapping the vertices of a subdivided grid onto objects in the pseudogrid sometimes maps a vertex onto a single edge (but not the endpoints of that edge).
+The main annoying issue is that mapping the vertices of a subdivided grid onto objects in the pseudogrid sometimes maps a vertex onto a single edge of the pseudogrid (but not the endpoints of that edge).
 In terms of colouring, this means that the subdivision vertex has an empty colour set.
@@
 
-4.
-In the acknowledgements, the authors mention that it was pointed out to them the problem on a grid (as opposed to pseudogrid) has a simple solution.
+**Referee Comment 4:** In the acknowledgements, the authors mention that it was pointed out to them the problem on a grid (as opposed to pseudogrid) has a simple solution.
 Assuming the person who pointed that out is happy with it, it might be interesting to include the solution.
@@
 
-Top of Page 3: This somewhat addresses the previous comment.
+**Referee Comment:** Top of Page 3: This somewhat addresses the previous comment.
 
-**Response:** We have added further explanation of the limits of this argument here.
+**Response:** We have added further explanation (at the location of the second comment) of the limits of this argument here.
 In particular, even with an Excluded Grid Theorem that matches the currently known lower bound, this would only establish that the centred chromatic number is at most cubic in the linear chromatic number.
@@
 In the first line of the proof, we use this value of $d$ to apply Lemma 9 to obtain an assignment of colours of objects so that each colour is assigned to at least $d$ objects.
-We have expanded the first line of the proof to point this out, so that it is familiar when it is mentioned again at this point in the proof.
+We have expanded the first line of the proof to point this out, so that it sounds familiar when it is mentioned again at this point in the proof.
 
@@
 **Response:** It is much simpler than this.
-If more than $\boxplus_{r+1}$ objects of $\phi^{-1}(\alpha)$ are left uncovered by the radius $2r+1$ boxes around the objects in $Y$ then this was also the case when the colour $\alpha$ was considered in the first round.
-At that point, we could have chosen any of these uncovered objects $\mu_1$ of colour $\alpha$.
-Then $B_{2r+1}(\mu_1)$ contains at most $\boxplus_r$ objects (by definition), so there is still another uncovered object $\mu_2$ of colour $\alpha$.
-In this case, we would have put $\mu_1$ and $\mu_2$ into $Q_1$ and round 1 would have succeeded for $\alpha$.
+During Round 1, when an object is added to $Q_1$, it *covers* all the nearby objects.
+Round 1 succeeds for $\alpha$ if, when the colour $\alpha$ is considered, there exists two uncovered vertices $\mu_1$ and $\mu_2$ of colour $\alpha$ that are "far" from each other.
+For a single object $\mu_1$, the number of vertices that "close" to $\mu_1$ is $\boxplus_r$ (by definition).
+So if there are $\boxplus_{r+1}>\boxplus_r$ uncovered objects of colour $\alpha$, we can choose any of them to be $\mu_1$, eliminate the at most $\boxplus_{r}$ objects that are close to $\mu_1$, and then take $\mu_2$ to be any of the remaining objects of colour $\alpha$.
+We have added a paragraph describing this argument.

### 2024-04-04 lin-vs-cen.tex
More changes suggested by the referee.

@@
 For each $\alpha\in X$, consider the set $\phi^{-1}(\alpha)\subseteq \VE(G_k)$ and recall that, since $\phi$ comes from \cref{one_colour_per_object}, $|\phi^{-1}(\alpha)| \ge d$.
-Furthermore all but at most $\vol{r+1}$ elements of $\phi^{-1}(\alpha)$ are contained in $\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))$ since otherwise the first round would have been an $\alpha$-success.
-Let $\Phi_\alpha:=\phi^{-1}(\alpha)\cap \left(\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))\right)$, so $|\Phi_\alpha|\ge d-\vol{r+1}$.
+
+  For each $\alpha\in X$, let $\Phi_\alpha:=\phi^{-1}(\alpha)\cap \left(\bigcup_{\mu\in Y}\VE(G_{2r+1}(\mu))\right)$.
 In words, $\Phi_\alpha$ consists of all objects of colour $\alpha$ that were eliminated from consideration because they are too close to an element of $Y$.
+We claim that, for each $\alpha\in X$, $|\Phi_\alpha|\ge d-\vol{r}$.
+Otherwise $|\phi^{-1}(\alpha)\setminus\Phi_\alpha| > \vol{r}$.
+Then, when the colour $\alpha$ was considered during the first round, we could have taken $\mu_1$ to be any object in $\phi^{-1}(\alpha)\setminus\Phi_\alpha$ and taken $\mu_2$ to be any object in $\phi^{-1}(\alpha)\setminus\Phi_\alpha\setminus B_{2r+1}(\mu_1)$.
+(Recall that $\vol{r}\ge |B_{2r+1}(\mu_1)|$, by definition.)  Thus, the first round would have been an $\alpha$-success, contradicting the fact that $\alpha\in X$.
 
@@
 
-  We begin by showing that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.} a $\binomial(\ceil{|\Phi_\alpha|/\vol{7r}},1/16)$ random variable, which allows us to establish an exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+1})$.
+  We now show that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.} a $\binomial(\ceil{|\Phi_\alpha|/\vol{7r}},1/16)$ random variable, which allows us to establish an exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+1})$.
 Suppose that, for some $\alpha\in X$ and $\mu\in Y$, $B_{2r+1}(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and consider the set
@@
   The right hand side of this last equation is of the form $O(r^2\log r+r^4\log|\Phi_\alpha|)$.
-Since $|\Phi_\alpha|\ge d-\vol{r+1}\in d- O(r^2)$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
+Since $|\Phi_\alpha|\ge d-\vol{r}\in d- O(r^2)$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
 
@@
 \begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
-  Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.
-Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)\cap\interior_r(G_{k'})|\ge d|A|$, for any $d\le 1/2\epsilon-2$.
-In particular, for some $\epsilon \in \Omega(1/r^4\log r)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
+  Suppose, for the sake of contradiction, that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours where $\epsilon:=1/(d+2r)$, $r:= 13$ and $d\in O(r^4\log r)=O(1)$ is some large integer constant.
+Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)\cap\interior_r(G_{k'})|\ge d|A|$.
+\Cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
 By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.

### 2024-04-04 referee-response.md
added math delimiter

@@
 Something like that would make arguments such as the one in Lemma 13 much cleaner.
-In this direction, have the authors considered the idea of using a grid with every edge subdivided once as the index set for the path partition of the pseudogrid (the subdivision vertices symbolising the edges), with the purpose of making definitions such as the r-boxes more uniform (and of having a symmetric relation, whereby if \mu is in the r-box of \nu, then \nu should be in the r-box of \mu)?
+In this direction, have the authors considered the idea of using a grid with every edge subdivided once as the index set for the path partition of the pseudogrid (the subdivision vertices symbolising the edges), with the purpose of making definitions such as the r-boxes more uniform (and of having a symmetric relation, whereby if $\mu$ is in the r-box of $\nu$, then $\nu$ should be in the r-box of $\mu$)?
 

### 2024-04-04 referee-response.md
Addressed everything the referee asked. Just need to reread.

@@
-The referee had 4 broad comments:
+We thank the referee for carefully reading the manuscript and giving thoughtful and helpful feedback.  The referee had 4 broad comments:
 
@@
 
-With the current title a reader will still learn, within the first two pages that the improvement comes only by studying the linear chromatic number pseudogrids and decide for themselves whether it's worthwhile reading the rest of the argument.
+With the current title a reader will still learn, within the first two pages, that the improvement comes only by studying the linear chromatic number pseudogrids and decide for themselves whether it's worthwhile reading the rest of the argument.
 (These first two pages also have the benefit of catching-up a reader on the current state of the art and the limitations of Kun et al's argument.)
@@
 
-TODO
+**Response:** We have added the solution along with the acknowledgement.
 
@@
 Now we only delete edges of $G$ that correspond to edges of the grid (and then repeatedly remove vertices of degree at most $1$).
+This avoids disconnecting a column when deleting a row, and vice versa.
 

### 2024-04-04 lin-vs-cen.tex
Addressed everything the referee asked. Just need to reread.

@@
 \usepackage{dsfont}  % for \mathds{A}
+\usepackage[OT1]{fontenc}
 \usepackage[utf8x]{inputenc}
@@
 Pat~Morin especially enjoyed early-morning discussions with Stefan~Langerman on the Seabourne Terrace.
-We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution.
-(By partitioning the rows into consecutive pairs and the columns  into consecutive pairs, and then repeatedly deleting a pair of rows and columns whose intersection contains a vertex whose colour is unique, one is left with a graph that has a Hamiltonian path in which each colour apppears on at least two vertices.)
+
+We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution: Partition the grid into $2\times 2$ grids and repeatedly delete a subgrid if it contains a vertex whose colour is unique.
+If this causes the grid to become disconnected, keep only the largest remaining component and continue.
+This process deletes at most $4\epsilon k$ vertices.
+For $\epsilon < 1/4$, this implies that the graph that remains at the end has at least $k^2/2$ vertices.
+In particular, the graph that remains is non-empty and every colour that appears in this graph appears in at least two vertices.
+Finally, this graph has a Hamiltonian path (which can be constructed by traversing the boundary of  a spanning-tree of the dual graph).
 

### 2024-04-04 lin-vs-cen.tex
Fucking annoying

@@
 \usepackage{listings}
+\usepackage{thmtools}
+\usepackage{thm-restate}
+
+\newtheorem{zeb}[theorem]{Lemma}
 
@@
 As a consequence we improve the general bound for all graphs to $\chicen(G)\le \tilde{O}(\chilin(G)^{10})$.
-In addition this tight bound gives further evidence in support of Kun \etal's conjecture (above) that the centred chromatic number (i.e., the treedepth) of any graph is upper bounded by a linear function of its linear chromatic number.
+In addition, this tight bound gives further evidence in support of Kun \etal's conjecture (above) that the centred chromatic number (i.e., the treedepth) of any graph is upper bounded by a linear function of its linear chromatic number.
 \end{abstract}
@@
 
-\citet{kun.obrien.ea:polynomial} give a family of graphs that contains, for every $\epsilon > 0$, a graph $G$ with $\chicen(G)\ge (2-\epsilon)\chilin(G)$. They conjecture that this bound is tight:
+\citet{kun.obrien.ea:polynomial} construct a family of graphs that contains, for every $\epsilon > 0$, a graph $G$ with $\chicen(G)\ge (2-\epsilon)\chilin(G)$.
+They conjecture that this bound is tight:
 
@@
 However, no improvement to the Excluded Grid Theorem will sufficient to establish a linear relationship between the centred and the linear chromatic number.
-Indeed, the best possible Excluded Grid Theorem would state that any graph of treewidth $k^2\log k$ contains an $\Omega(k)\times\Omega(k)$ grid minor, and the preceding argument would only show that $\chicen(G)\in O((\chilin(G))^3\log(\chilin(G)))$.
-Even the Excluded Grid Theorem for Planar Graphs states that any planar graph of treewidth $k$ contains an $\Omega(k)\times\Omega(k)$ grid minor.
+Indeed, the best possible Excluded Grid Theorem would state that any graph of treewidth $k^2\log k$ contains an $\Omega(k)\times\Omega(k)$ grid minor \cite{robertson.seymour.ea:quickly}, and the preceding argument would only show that $\chicen(G)\in O((\chilin(G))^3\log(\chilin(G)))$.
+Even the Excluded Grid Theorem for Planar Graphs states that any planar graph of treewidth $k$ contains an $\Omega(k)\times\Omega(k)$ grid minor
+% \cite[(6.2)]{robertson.seymour.ea:quickly}.
+\cite{robertson.seymour.ea:quickly}.
 Combining this with the argument above and \cref{pseudogrid_lower_bound} shows only that, for any planar graph $G$, $\chicen(G)\in\Theta((\chilin(G))^2)$.
@@
 As part of our proof, we use the operation of \defin{deleting} a row (or column) of $G$.
-To delete a row $R$ of $G$ that corresponds to the row $R':=v_1,\ldots,v_a$, we remove the edges of $\overline{P}_{v_{i}v_{i+1}}$ for each $i\in\{2,\ldots,a\}$.
+To delete a row $R$ of $G$ that corresponds to the row $R':=v_1,\ldots,v_a$ in $G_{a\times b}$, we remove the edges of $\overline{P}_{v_{i}v_{i+1}}$ for each $i\in\{2,\ldots,a\}$.
 If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$ or when $R=v_1,\ldots,v_r$ does not contains $P_{v_i}$ for some $i\in\{1,\ldots,r\}$) then we repeatedly remove vertices of degree at most $1$ until none remain.
@@
 
-\begin{lem}\label{pick_up_two}
+\begin{restatable}{zeb}{blechblech}\label{pick_up_two}
   Let $G$ be an $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.
 Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
-\end{lem}
+\end{restatable}
 
@@
 
-  We now show that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.} a $\binomial(\ceil{|\Phi_\alpha|/\vol{7r}},1/16)$ random variable, which allows us to establish an exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+1})$.
+  Our next step is to show that, for each $\alpha\in X$, the random variable $\deg_H(\alpha)$ dominates\footnote{A random variable $X$ dominates a random variable $Y$ if $\Pr(X\ge x)\ge \Pr(Y\ge x)$ for all $x\in\R$.} a $\binomial(\ceil{|\Phi_\alpha|/\vol{7r}},1/16)$ random variable, which allows us to establish an exponential inequality for $\Pr(\deg_H(\alpha) < 2\vol{2r+1})$.
 Suppose that, for some $\alpha\in X$ and $\mu\in Y$, $B_{2r+1}(\mu)\cap \Phi_\alpha$ contains at least one element $\nu$ and consider the set
@@
 \begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
-  Suppose, for the sake of contradiction, that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours where $\epsilon:=1/(d+2r)$, $r:= 13$ and $d\in O(r^4\log r)=O(1)$ is some large integer constant.
+  Let $G$ be a $k\times k$ pseudogrid and suppose, for the sake of contradiction, that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours where $\epsilon:=1/(d+2r)$, $r:= 13$ and $d\in O(r^4\log r)=O(1)$ is some large integer constant.
 Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)\cap\interior_r(G_{k'})|\ge d|A|$.
-\Cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
+For a sufficiently large constant $d$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
 By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.
@@
 This process deletes at most $4\epsilon k$ vertices.
-For $\epsilon < 1/4$, this implies that the graph that remains at the end has at least $k^2/2$ vertices.
-In particular, the graph that remains is non-empty and every colour that appears in this graph appears in at least two vertices.
+For $\epsilon < 1/4$, this implies that the graph that remains at the end of thie process has at least $k^2/2$ vertices.
+In particular, the graph that remains is non-empty and every colour that appears in this graph is used on at least two vertices.
 Finally, this graph has a Hamiltonian path (which can be constructed by traversing the boundary of  a spanning-tree of the dual graph).
@@
 \label{pick_up_two_proof}
+
+\blechblech
 

### 2024-04-05 lin-vs-cen.tex
Pffft

@@
 \usepackage{paralist}
-\usepackage{dsfont}  % for \mathds{A}
 \usepackage[OT1]{fontenc}
 \usepackage[utf8x]{inputenc}
-\usepackage{skull}
 \usepackage{paralist}
-\usepackage{graphicx}
-\usepackage[noend]{algorithmic}
 \usepackage{bbm}  % needed for \mathbbm{1}
-\usepackage{listings}
-\usepackage{thmtools}
-\usepackage{thm-restate}
-
-\newtheorem{zeb}[theorem]{Lemma}
-
-\usepackage[normalem]{ulem}
-\usepackage{cancel}
-%\usepackage{enumitem}
 
@@
 
-\begin{restatable}{zeb}{blechblech}\label{pick_up_two}
+\begin{lem}[restate=pickuptwo,label=pick_up_two]
   Let $G$ be an $a\times a$ pseudogrid with $a\ge 5$,  let $s,v,w,t$ be vertices of $G$ with $s$ in column $1$ of $G$, $v,w\in \interior_1(G)$, and $t$ in column $a$ of $G$.
 Then $G$ contains a path $P$ with endpoints $s$ and $t$ that contains $v$ and $w$.
-\end{restatable}
+\end{lem}
 
@@
 
-We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution: Partition the grid into $2\times 2$ grids and repeatedly delete a subgrid if it contains a vertex whose colour is unique.
+We are also grateful to Omer Angel for pointing out that the problem on a $k\times k$ grid (as opposed to pseudogrid) has a simple solution, illustrated in \cref{omer}: Partition the grid into $2\times 2$ grids and repeatedly delete a subgrid if it contains a vertex whose colour is unique.
 If this causes the grid to become disconnected, keep only the largest remaining component and continue.
@@
 In particular, the graph that remains is non-empty and every colour that appears in this graph is used on at least two vertices.
-Finally, this graph has a Hamiltonian path (which can be constructed by traversing the boundary of  a spanning-tree of the dual graph).
+Finally, this graph has a Hamiltonian path (which can be constructed by traversing the boundary of a certain tree contained in the dual graph).
+Unfortunately, this solution breaks down immediately for pseudogrids since a pseudogrid is not necessarily Hamiltonian, even before removing some of its vertices and edges.
+
+\begin{figure}
+  \centering
+  \begin{tabular}{cccc}
+    \includegraphics[page=1]{figs/omer} &
+    \includegraphics[page=2]{figs/omer} &
+    % \includegraphics[page=3]{figs/omer} &
+    \includegraphics[page=4]{figs/omer}
+  \end{tabular}
+  \caption{A simpler solution for colourings of the grid.
+The (orange) tree used to create the (purple) Hamiltonian path is created so that its edges and leaves use only odd-numbered rows and columns of the dual graph.}
+  \label{omer}
+\end{figure}
 
@@
 
-\blechblech
+\pickuptwo*
 

### 2024-04-10 lin-vs-cen.tex
WTF

@@
 \title{\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}}
-\author{Prosenjit~Bose, Vida~Dujmović, Hussein~Houdrouge, Mehrnoosh~Javarsineh, and Pat~Morin}
+\author{Prosenjit~Bose\thanks{School of Computer Science, Carleton University},\quad Vida~Dujmović\thanks{Department of Computer Science and EE, University of Ottawa},\quad Hussein~Houdrouge\footnotemark[2],\quad Mehrnoosh~Javarsineh\footnotemark[2],\quad and Pat~Morin\footnotemark[2]}
 

### 2024-04-10 lin-vs-cen.tex
Added contact email as requested by publisher

--- 2024-04-05 lin-vs-cen.tex
@@
 \title{\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}}
-\author{Prosenjit~Bose, Vida~Dujmović, Hussein~Houdrouge, Mehrnoosh~Javarsineh, and Pat~Morin}
+\author{Prosenjit~Bose\thanks{School of Computer Science, Carleton University},\quad Vida~Dujmović\thanks{Department of Computer Science and EE, University of Ottawa},\quad Hussein~Houdrouge\footnotemark[2],\quad Mehrnoosh~Javarsineh\footnotemark[2],\quad and Pat~Morin\thanks{School of Computer Science, Carleton University, \email{morin@scs.carleton.ca}}}
 

### 2025-09-19 lin-vs-cen.tex
Added keywords as requested by Wiley
Switched to PDF figures
Pffft

@@
 
-\title{\MakeUppercase{Linear versus centred chromatic numbers}\thanks{This research was partly funded by NSERC.}}
+\title{\MakeUppercase{Linear Versus Centred Colouring via Pseudogrids}\thanks{This research was partly funded by NSERC.}}
 \author{Prosenjit~Bose\thanks{School of Computer Science, Carleton University},\quad Vida~Dujmović\thanks{Department of Computer Science and EE, University of Ottawa},\quad Hussein~Houdrouge\footnotemark[2],\quad Mehrnoosh~Javarsineh\footnotemark[2],\quad and Pat~Morin\thanks{School of Computer Science, Carleton University, \email{morin@scs.carleton.ca}}}
@@
 \end{abstract}
+
+\keywords{treedepth; graph colouring; centered colouring; linear colouring}
 
@@
 Any further improvement to \cref{kun_obrien_general2} will either require an improved Excluded Grid Theorem or an entirely new approach.
-However, no improvement to the Excluded Grid Theorem will sufficient to establish a linear relationship between the centred and the linear chromatic number.
+However, no improvement to the Excluded Grid Theorem will be sufficient to establish a linear relationship between the centred and the linear chromatic number.
 Indeed, the best possible Excluded Grid Theorem would state that any graph of treewidth $k^2\log k$ contains an $\Omega(k)\times\Omega(k)$ grid minor \cite{robertson.seymour.ea:quickly}, and the preceding argument would only show that $\chicen(G)\in O((\chilin(G))^3\log(\chilin(G)))$.
@@
   \begin{center}
-    \includegraphics{figs/pseudogrid-2}
+    \includegraphics[page=2]{figs/pseudogrid}
   \end{center}
@@
     \begin{tabular}{ccc}
-      \includegraphics{figs/outline-2} &
-      \includegraphics{figs/outline-3} &
-      \includegraphics{figs/outline-4} \\
+      \includegraphics[page=2]{figs/outline} &
+      \includegraphics[page=3]{figs/outline} &
+      \includegraphics[page=4]{figs/outline} \\
       $S_1$ & $S_1\cup S_2$ & $P$
@@
     \begin{tabular}{ccc}
-      \includegraphics{figs/bad-example-1} &
-      \includegraphics{figs/bad-example-2} &
-      \includegraphics{figs/bad-example-3}
+      \includegraphics[page=1]{figs/bad-example} &
+      \includegraphics[page=2]{figs/bad-example} &
+      \includegraphics[page=3]{figs/bad-example}
     \end{tabular}
@@
       \begin{tabular}{ccc}
-        \includegraphics[width=.3\textwidth]{figs/make_disjoint-1} &
-        \includegraphics[width=.3\textwidth]{figs/make_disjoint-2} &
-        \includegraphics[width=.3\textwidth]{figs/make_disjoint-3} \\
+        \includegraphics[page=1,width=.3\textwidth]{figs/make_disjoint} &
+        \includegraphics[page=2,width=.3\textwidth]{figs/make_disjoint} &
+        \includegraphics[page=3,width=.3\textwidth]{figs/make_disjoint} \\
         $X_1$ & $X_2$ & $X_3$
@@
 \begin{proof}
+  Fix a real number $c\ge 1$ such that $d=cr^4\log r$ is an integer.
   We begin by applying \cref{one_colour_per_object} to obtain the colouring $\phi:\VE(G_k)\to\varphi(V(G))\cup\{\perp\}$ in which $\phi^{-1}(\alpha)\ge d$ for each $\alpha\in\varphi(V(G))$.
@@
 In words, $\Phi_\alpha$ consists of all objects of colour $\alpha$ that were eliminated from consideration because they are too close to an element of $Y$.
-We claim that, for each $\alpha\in X$, $|\Phi_\alpha|\ge d-\vol{r}$.
-Otherwise $|\phi^{-1}(\alpha)\setminus\Phi_\alpha| > \vol{r}$.
+We claim that, for each $\alpha\in X$, $|\Phi_\alpha|\ge d-\vol{2r+1}$.
+Otherwise $|\phi^{-1}(\alpha)\setminus\Phi_\alpha| > \vol{2r+1}$.
 Then, when the colour $\alpha$ was considered during the first round, we could have taken $\mu_1$ to be any object in $\phi^{-1}(\alpha)\setminus\Phi_\alpha$ and taken $\mu_2$ to be any object in $\phi^{-1}(\alpha)\setminus\Phi_\alpha\setminus B_{2r+1}(\mu_1)$.
-(Recall that $\vol{r}\ge |B_{2r+1}(\mu_1)|$, by definition.)  Thus, the first round would have been an $\alpha$-success, contradicting the fact that $\alpha\in X$.
+(Recall that $\vol{2r+1}\ge |B_{2r+1}(\mu_1)|$, by definition.)  Thus, the first round would have been an $\alpha$-success, contradicting the fact that $\alpha\in X$.
 
@@
   provided that
-  \[   |\Phi_\alpha| \ge {32\vol{7r}}\big(\ln({32\vol{7r}}+1) + {2\vol{2r+1}}\ln(2|\Phi_\alpha|)\big) \enspace .
+  \begin{equation}
+    |\Phi_\alpha| \ge {32\vol{7r}}\big(\ln({32\vol{7r}}+1) + {2\vol{2r+1}}\ln(2|\Phi_\alpha|)\big) \enspace . \label{critical}
+  \end{equation}
+  The right hand side of this last equation is at most $c'(r^2\log r+r^4\log|\Phi_\alpha|) \le c''r^4\log|\Phi_\alpha|$ for some absolute constants $c'$ and $c''$.
+Thus, it suffices to choose $c$ so that
+  \[
+    \frac{|\Phi_\alpha|}{\log|\Phi_\alpha|}
+    \ge \frac{(d-\vol{r})}{\log(d-\vol{r})}
+    \ge \frac{(d-\vol{r})}{\log d}
+    = \frac{cr^4\log r-\vol{r}}{\log(cr^4\log r)}
+    \ge \frac{cr^4\log r-\vol{r}}{\log(cr^5)}
+    \ge \frac{cr^4\log r}{2\log(cr^5)}
+    \ge c''r^4 \enspace .
   \]
-  The right hand side of this last equation is of the form $O(r^2\log r+r^4\log|\Phi_\alpha|)$.
-Since $|\Phi_\alpha|\ge d-\vol{r}\in d- O(r^2)$, this can therefore be satisfied for some $d\in O(r^4\log r)$.
+  Rewriting the last inequality gives $c-2\log_r c\ge 10c''$.  Since $r\ge 10$, $c-2\log_r c\ge c/2$, for all $c\ge 1$.  So \eqref{critical} holds for all $c\ge 20c''$.
 
@@
 This process deletes at most $4\epsilon k$ vertices.
-For $\epsilon < 1/4$, this implies that the graph that remains at the end of thie process has at least $k^2/2$ vertices.
+For $\epsilon < 1/4$, this implies that the graph that remains at the end of this process has at least $k^2/2$ vertices.
 In particular, the graph that remains is non-empty and every colour that appears in this graph is used on at least two vertices.
@@
       \begin{tabular}{cc}
-        \includegraphics{figs/generic_path-1} &
-        \includegraphics{figs/generic_path-2}
+        \includegraphics[page=1]{figs/generic_path} &
+        \includegraphics[page=2]{figs/generic_path}
       \end{tabular}
